### 2016-11-11 build/paper.tex
Initial commit

\section{Ausgangssituation}\label{ausgangssituation}

\section{Zielsetzung}\label{zielsetzung}

\section{Vorgehensweise}\label{vorgehensweise}

### 2016-11-14 paper.tex
pdf/tex build

\section{Ausgangssituation}\label{ausgangssituation}

Softwareentwicklung für den Browser ist ohne Zweifel eine komplexe
Angelegenheit. Die Gründe für diesen Umstand sind sehr vielfältig.
Einige Probleme sind ``hausgemacht'' wie beispielsweise Designschwächen
der de facto einzigen Zielsprache aller Browser JavaScript. Andere
Probleme sind einfach dem Umstand der rasanten Evolution im Internet von
statischen Seiten hin zu dynamischen Web Apps geschuldet.

Das Resultat dieser Entwicklungen sind gravierend. Die Anzahl der
Frameworks, Werkzeuge und Bibliotheken ist schier unüberblickbar
geworden und wandelt sich in einer Geschwindigkeit, die Entwicklern in
der JavaScript Welt Schwierigkeiten bereitet.{[}\^{}1{]} Aber auch die
andere Seite, die Konsumenten der Webseiten, bekommen diese Probleme
spüren. Frederic Filloux zeigte in einem Blogpost, dass sich in einem
4667 Buchstaben langen Zeitungsartikel der britischen Zeitung ``The
Guardian'' 485527 Buchstaben an Quelltext verstecken.{[}1{]}

\section{Zielsetzung}\label{zielsetzung}

Ziel dieser Arbeit ist die systematische Erfassung der neuen
Technologien sowie eine praktische Erläuterung möglicher
Architekturmodelle.

\section{Vorgehensweise}\label{vorgehensweise}

{[}\^{}1{]}
https://hackernoon.com/how-it-feels-to-learn-javascript-in-2016-d3a717dd577f

\hypertarget{refs}{}
\hypertarget{ref-Filloux2016}{}
{[}1{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016.

### 2016-11-14 paper.tex
pdf/tex build

@@
 geworden und wandelt sich in einer Geschwindigkeit, die Entwicklern in
-der JavaScript Welt Schwierigkeiten bereitet.{[}\^{}1{]} Aber auch die
+der JavaScript Welt Schwierigkeiten bereitet.{[}fn\^{}1{]} Aber auch die
 andere Seite, die Konsumenten der Webseiten, bekommen diese Probleme
@@
 
-{[}\^{}1{]}
+{[}fn\^{}1{]}
 https://hackernoon.com/how-it-feels-to-learn-javascript-in-2016-d3a717dd577f

### 2016-11-14 paper.tex
pdf/tex build

--- 2016-11-14 paper.tex
@@
 geworden und wandelt sich in einer Geschwindigkeit, die Entwicklern in
-der JavaScript Welt Schwierigkeiten bereitet.{[}\^{}1{]} Aber auch die
-andere Seite, die Konsumenten der Webseiten, bekommen diese Probleme
-spüren. Frederic Filloux zeigte in einem Blogpost, dass sich in einem
-4667 Buchstaben langen Zeitungsartikel der britischen Zeitung ``The
-Guardian'' 485527 Buchstaben an Quelltext verstecken.{[}1{]}
+der JavaScript Welt Schwierigkeiten bereitet.\footnote{https://hackernoon.com/how-it-feels-to-learn-javascript-in-2016-d3a717dd577f}
+Aber auch die andere Seite, die Konsumenten der Webseiten, bekommen
+diese Probleme spüren. Frederic Filloux zeigte in einem Blogpost, dass
+sich in einem 4667 Buchstaben langen Zeitungsartikel der britischen
+Zeitung ``The Guardian'' 485527 Buchstaben an Quelltext
+verstecken.{[}1{]}
 
@@
 
-\section{Vorgehensweise}\label{vorgehensweise}
-
-{[}\^{}1{]}
-https://hackernoon.com/how-it-feels-to-learn-javascript-in-2016-d3a717dd577f
+\section*{Vorgehensweise}\label{vorgehensweise}
+\addcontentsline{toc}{section}{Vorgehensweise}
 

### 2016-11-14 paper.tex
pdf/tex build

--- 2016-11-14 paper.tex
@@
 Softwareentwicklung für den Browser ist ohne Zweifel eine komplexe
-Angelegenheit. Die Gründe für diesen Umstand sind sehr vielfältig.
-Einige Probleme sind ``hausgemacht'' wie beispielsweise Designschwächen
-der de facto einzigen Zielsprache aller Browser JavaScript. Andere
-Probleme sind einfach dem Umstand der rasanten Evolution im Internet von
-statischen Seiten hin zu dynamischen Web Apps geschuldet.
+Angelegenheit. Die Gründe für diesen Umstand sind so vielfältig, dass
+darüber eine ganze Forschungsarbeit geschrieben werden könnte. Je nach
+Blickwinkel könnte man von der technischen Seite argumentieren und
+beispielsweise Designschwächen von JavaScript oder die unflexiblen
+Browserengines als Bremsklotz der Webentwicklung benennen.{[}1{]} Oder
+man könnte die organisatorische Seite betrachten und die vielen
+beteiligten Organisationen bzw. Unternehmen und die daraus
+resultierenden langsamen Innovationszyklen als Problem benennen.{[}2{]}
+Oder aber man betrachtet die historische Entwicklung, die dazu geführt
+hat, dass die technische Entwicklung der rasanten Evolution im Internet
+nicht Schritt gehalten hat. Die Zeiten statischer HTML/CSS Seiten auf
+einfachen Desktopgeräten ist mithin noch gar nicht so lange her.
 
-Das Resultat dieser Entwicklungen sind gravierend. Die Anzahl der
-Frameworks, Werkzeuge und Bibliotheken ist schier unüberblickbar
-geworden und wandelt sich in einer Geschwindigkeit, die Entwicklern in
-der JavaScript Welt Schwierigkeiten bereitet.{[}\^{}1{]} Aber auch die
-andere Seite, die Konsumenten der Webseiten, bekommen diese Probleme
-spüren. Frederic Filloux zeigte in einem Blogpost, dass sich in einem
-4667 Buchstaben langen Zeitungsartikel der britischen Zeitung ``The
-Guardian'' 485527 Buchstaben an Quelltext verstecken.{[}1{]}
+Dennoch tragen alle diese Facetten dazu bei, dass die Entwicklung und
+Wartung von komplexen Webapplikationen, wie sie heute Standard sind, mit
+enormen Zeit- und Geldaufwand verbunden sind. Die Anzahl der Frameworks,
+Werkzeuge und Bibliotheken mit Javascript als Zielsprache ist
+unüberblickbar geworden und wandelt sich in einer Geschwindigkeit, die
+vielen Entwicklern Schwierigkeiten bereitet.\footnote{https://hackernoon.com/how-it-feels-to-learn-javascript-in-2016-d3a717dd577f}
+Aber auch die andere Seite, die Nutzer der Webdienste, bekommen diese
+Probleme spüren. Frederic Filloux zeigte in einem Blogpost, dass sich in
+einem 4667 Buchstaben langen Zeitungsartikel der britischen Zeitung
+``The Guardian'' 485527 Buchstaben an Quelltext verstecken.{[}3{]}
+
+Um diesen Entwicklungen, die viel Zeit und Geld kosten, etwas
+entgegenzusetzen,
+
+https://extensiblewebmanifesto.org/
 
@@
 
-{[}\^{}1{]}
-https://hackernoon.com/how-it-feels-to-learn-javascript-in-2016-d3a717dd577f
+Den Anfang dieser Arbeit soll eine Analyse der aktuellen Situation der
+Frontend Entwicklung aufzeigen. In ihr sollen aktuelle
+Architekturmodelle analysiert werden. Außerdem soll dieser Teil vor
+Augen führen, warum es bisher nur mit zusätzlichen Schichten der
+Abstraktion möglich war einen modularen Aufbau von Web Apps zu
+ermöglichen. In diesem Teil sollen auch die Probleme die der Entwicklung
+und Wartung dieser Systeme aufzeigen.
+
+Im nächsten Abschnitt der Arbeit sollen die
 
 \hypertarget{refs}{}
+\hypertarget{ref-Katz2013}{}
+{[}1{]} Y. Katz, ``Extend the web forward,'' 2013.
+
+\hypertarget{ref-Walton2016}{}
+{[}2{]} P. Walton, ``Houdini: Maybe the most exciting development in css
+you've never heard of,'' 2016.
+
 \hypertarget{ref-Filloux2016}{}
-{[}1{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016.
+{[}3{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016.

### 2016-11-18 paper.tex
pdf/tex build

@@
 beispielsweise Designschwächen von JavaScript oder die unflexiblen
-Browserengines als Bremsklotz der Webentwicklung benennen.{[}1{]} Oder
-man könnte die organisatorische Seite betrachten und die vielen
-beteiligten Organisationen bzw. Unternehmen und die daraus
-resultierenden langsamen Innovationszyklen als Problem benennen.{[}2{]}
-Oder aber man betrachtet die historische Entwicklung, die dazu geführt
-hat, dass die technische Entwicklung der rasanten Evolution im Internet
-nicht Schritt gehalten hat. Die Zeiten statischer HTML/CSS Seiten auf
-einfachen Desktopgeräten ist mithin noch gar nicht so lange her.
+Browserplattformen als Bremsklotz der Webentwicklung benennen.{[}1{]}
+Oder man könnte die organisatorische Seite betrachten und die vielen
+beteiligten Organisationen benennen, die für Standardisierungsprozesse
+im Internet zuständig sind.{[}2{]} Oder aber man nimmt die historische
+Entwicklung in Augenschein, die dazu geführt hat, dass die technische
+Entwicklung der rasanten Evolution im Internet nicht Schritt gehalten
+hat. Die Zeiten statischer HTML/CSS Seiten auf einfachen Desktopgeräten
+ist mithin noch gar nicht so lange her.
 
@@
 enormen Zeit- und Geldaufwand verbunden sind. Die Anzahl der Frameworks,
-Werkzeuge und Bibliotheken mit Javascript als Zielsprache ist
-unüberblickbar geworden und wandelt sich in einer Geschwindigkeit, die
-vielen Entwicklern Schwierigkeiten bereitet.\footnote{https://hackernoon.com/how-it-feels-to-learn-javascript-in-2016-d3a717dd577f}
-Aber auch die andere Seite, die Nutzer der Webdienste, bekommen diese
-Probleme spüren. Frederic Filloux zeigte in einem Blogpost, dass sich in
-einem 4667 Buchstaben langen Zeitungsartikel der britischen Zeitung
-``The Guardian'' 485527 Buchstaben an Quelltext verstecken.{[}3{]}
+Werkzeuge und Bibliotheken mit Javascript als Zielsprache ist nahezu
+unüberblickbar und wandelt sich in einer Geschwindigkeit, die vielen
+Entwicklern Schwierigkeiten bereitet.\footnote{Ein fiktives Gespräch mit
+  einem JavaScript Neuling
+  https://hackernoon.com/how-it-feels-to-learn-javascript-in-2016-d3a717dd577f}
+Auch die Konsumenten und Nutzern der Webdiensten bekommen diese Probleme
+zumindest teilweise spüren. Einerseits spürbar in Form von ``Downtimes''
+der Services. Andererseits subtil wie etwa durch lange Ladezeiten wegen
+aufgeblähtem Quellcode. Frederic Filloux zeigte in einem Blogpost
+anschaulich, dass sich in einem Zeitungsartikel der britischen Zeitung
+``The Guardian'' zu jedem lesbaren Buchstaben des Artikels über 100
+Zeichen Code addieren.{[}3{]}
 
-Um diesen Entwicklungen, die viel Zeit und Geld kosten, etwas
-entgegenzusetzen,
-
-https://extensiblewebmanifesto.org/
+Aus diesen Grund wurde schon 2013 das ``Extensible Web Manifesto''
+proklamiert. Darin wird, kurz gefasst, eine Öffnung der Webplattformen
+avisiert, um Webprogrammierung teilweise von Browserstandards zu
+entkoppeln.{[}\^{}fn2{]} Drei Jahre später sind diese Ziele in W3C
+Spezifikationen konkretisiert worden. Einige dieser neuen Standards
+werden bereits nativ in (einigen) Browsern unterstützt.
 
@@
 
-Ziel dieser Arbeit ist die systematische Erfassung der neuen
-Technologien sowie eine praktische Erläuterung möglicher
-Architekturmodelle.
+Viele populäre Frameworks für den Browser als Zielplattform weisen
+heutzutage eine monolithische Architektur auf. Der Grund dafür liegt
+unter anderem darin, dass es bisher nicht möglich war einzelne
+Browserelemente in ihrem Aussehen und Verhalten zu isolieren. Mit dem
+neuen W3C Spezifikation ``Web Components'', die mehrere Substandards
+unter sich vereint , Ziel dieser Arbeit soll ein In jüngster Zeit haben
+einige Vorschläge für eine ``low-level API'' Ziel dieser Arbeit ist die
+systematische Erfassung der neuen Technologien sowie eine praktische
+Erläuterung möglicher Architekturmodelle.
+
+houdini https://www.w3.org/standards/techs/components
 
@@
 
+\subsection[:
+https://extensiblewebmanifesto.org/]{\texorpdfstring{\footnote{Ein
+  fiktives Gespräch mit einem JavaScript Neuling
+  https://hackernoon.com/how-it-feels-to-learn-javascript-in-2016-d3a717dd577f}:
+https://extensiblewebmanifesto.org/}{: https://extensiblewebmanifesto.org/}}\label{fn1-httpsextensiblewebmanifesto.org}
+
+title: Microservices im Browser durch Web Components subtitle:
+Kombinierte W3C Spezifikationen für plattformübergreifende Browsermodule
+author: Jan Peteler, FHWS Würzburg-Schweinfurt,
+jan.peteler@student.fhws.de date: November 2016 abstract: \ldots{}
+
 \hypertarget{refs}{}

### 2016-11-18 build/paper.tex
pdf/tex build

@@
 \section{Ausgangssituation}\label{ausgangssituation}
+
+Softwareentwicklung für den Browser ist ohne Zweifel eine komplexe
+Angelegenheit. Die Gründe für diesen Umstand sind so vielfältig, dass
+darüber eine ganze Forschungsarbeit geschrieben werden könnte. Je nach
+Blickwinkel könnte man von der technischen Seite argumentieren und
+beispielsweise Designschwächen von JavaScript oder die unflexiblen
+Browserplattformen als Bremsklotz der Webentwicklung benennen.{[}1{]}
+Oder man könnte die organisatorische Seite betrachten und die vielen
+beteiligten Organisationen benennen, die für Standardisierungsprozesse
+im Internet zuständig sind.{[}2{]} Oder aber man nimmt die historische
+Entwicklung in Augenschein, die dazu geführt hat, dass die technische
+Entwicklung der rasanten Evolution im Internet nicht Schritt gehalten
+hat. Die Zeiten statischer HTML/CSS Seiten auf einfachen Desktopgeräten
+ist mithin noch gar nicht so lange her.
+
+Dennoch tragen alle diese Facetten dazu bei, dass die Entwicklung und
+Wartung von komplexen Webapplikationen, wie sie heute Standard sind, mit
+enormen Zeit- und Geldaufwand verbunden sind. Die Anzahl der Frameworks,
+Werkzeuge und Bibliotheken mit Javascript als Zielsprache ist nahezu
+unüberblickbar und wandelt sich in einer Geschwindigkeit, die vielen
+Entwicklern Schwierigkeiten bereitet.\footnote{Ein fiktives Gespräch mit
+  einem JavaScript Neuling
+  https://hackernoon.com/how-it-feels-to-learn-javascript-in-2016-d3a717dd577f}
+Auch die Konsumenten und Nutzern der Webdiensten bekommen diese Probleme
+zumindest teilweise spüren. Einerseits spürbar in Form von ``Downtimes''
+der Services. Andererseits subtil wie etwa durch lange Ladezeiten wegen
+aufgeblähtem Quellcode. Frederic Filloux zeigte in einem Blogpost
+anschaulich, dass sich in einem Zeitungsartikel der britischen Zeitung
+``The Guardian'' zu jedem lesbaren Buchstaben des Artikels über 100
+Zeichen Code addieren.{[}3{]}
+
+Aus diesen Grund wurde schon 2013 das ``Extensible Web Manifesto''
+proklamiert. Darin wird, kurz gefasst, eine Öffnung der Webplattformen
+avisiert, um Webprogrammierung teilweise von Browserstandards zu
+entkoppeln.\footnote{https://extensiblewebmanifesto.org/} Drei Jahre
+später sind diese Ziele in W3C Spezifikationen konkretisiert worden.
+Einige dieser neuen Standards werden bereits nativ in (einigen) Browsern
+unterstützt.
 
@@
 
+Viele populäre Frameworks für den Browser als Zielplattform versprechen
+dem Entwickler eine bessere Kontrolle über die Webseite. Der Grund dafür
+liegt unter anderem darin, dass es bisher nicht möglich war einzelne
+Browserelemente in ihrem Aussehen und Verhalten zu isolieren. Selbst
+kleinste Änderungen können daher die Funktionalität des komplexen
+Gesamtsystems beeinträchtigen. Mit der neuen W3C Spezifikation ``Web
+Components'',{[}\^{}fn3{]} die mehrere Substandards unter sich vereint,
+soll die Modularität nun auch nativ im Browser unterstützt werden.
+Modularität als Designpattern ist in der IT schon länger Standard um
+komplexe Systeme beherrschbar zu gestalten und auch zukünftige
+Unsicherheiten abzuwägen.{[}4{]} Die Unsicherheit, die Frameworks immer
+anhaftet ist die Lebensdauer derselben.
+
+Web Components ist der Entwickler in der Lage, eigene HTML Elemente zu
+definieren und diese per Templates zu exportieren oder zu importieren.
+Darüber hinaus können sie JavaScript und CSS Funktionalität enthalten
+und diese vom Rest der Webseite abzukapseln. Ziel dieser Arbeit ist die
+systematische Erfassung und Risikoabwägung der neuen Technologien sowie
+eine praktische Erläuterung möglicher Architekturmodelle. Bisher gibt es
+keine einheitlichen ``best practices'' wie denn die Komponenten
+umzusetzen sind, obwohl diese Technologien bereits länger durch
+Polyfills genutzt werden können. Manch ein Entwickler fürchtet schon
+eine Flut von schlecht gebauten Komponenten, die wiederum neue Probleme
+schaffen könnten.{[}5{]}
+
 \section{Vorgehensweise}\label{vorgehensweise}
+
+Den Anfang dieser Arbeit soll eine Analyse der aktuellen Situation der
+Frontend Entwicklung aufzeigen. In ihr sollen aktuelle
+Architekturmodelle analysiert werden. Außerdem soll dieser Teil vor
+Augen führen, warum es bisher nur mit zusätzlichen Schichten der
+Abstraktion möglich war einen modularen Aufbau von Web Apps zu
+ermöglichen. In diesem Teil sollen auch die Probleme die der Entwicklung
+und Wartung dieser Systeme aufzeigen.
+
+Im nächsten Abschnitt der Arbeit sollen die
+
+\subsection{{[}\^{}fn3{]}:
+https://www.w3.org/standards/techs/components}\label{fn3-httpswww.w3.orgstandardstechscomponents}
+
+title: Microservices im Browser durch Web Components subtitle:
+Kombinierte W3C Spezifikationen für plattformübergreifende Browsermodule
+author: Jan Peteler, FHWS Würzburg-Schweinfurt,
+jan.peteler@student.fhws.de date: November 2016 abstract: \ldots{}
+
+\hypertarget{refs}{}
+\hypertarget{ref-Katz2013}{}
+{[}1{]} Y. Katz, ``Extend the web forward,'' 2013.
+
+\hypertarget{ref-Walton2016}{}
+{[}2{]} P. Walton, ``Houdini: Maybe the most exciting development in css
+you've never heard of,'' 2016.
+
+\hypertarget{ref-Filloux2016}{}
+{[}3{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016.
+
+\hypertarget{ref-Baldwin2006}{}
+{[}4{]} C. Y. Baldwin and K. B. Clark, ``Modularity in the design of
+complex engineering systems,'' in \emph{Complex engineered systems:
+Science meets technology}, D. Braha, A. A. Minai, and Y. Bar-Yam, Eds.
+Berlin, Heidelberg: Springer Berlin Heidelberg, 2006, pp. 175--205.
+
+\hypertarget{ref-Keith2014}{}
+{[}5{]} J. Keith, ``Web components,'' 2014.

### 2016-11-18 paper.tex
pdf/tex build

@@
 avisiert, um Webprogrammierung teilweise von Browserstandards zu
-entkoppeln.{[}\^{}fn2{]} Drei Jahre später sind diese Ziele in W3C
-Spezifikationen konkretisiert worden. Einige dieser neuen Standards
-werden bereits nativ in (einigen) Browsern unterstützt.
+entkoppeln.\footnote{https://extensiblewebmanifesto.org/} Drei Jahre
+später sind diese Ziele in W3C Spezifikationen konkretisiert worden.
+Einige dieser neuen Standards werden bereits nativ in (einigen) Browsern
+unterstützt.
 
@@
 
-Viele populäre Frameworks für den Browser als Zielplattform weisen
-heutzutage eine monolithische Architektur auf. Der Grund dafür liegt
-unter anderem darin, dass es bisher nicht möglich war einzelne
-Browserelemente in ihrem Aussehen und Verhalten zu isolieren. Mit dem
-neuen W3C Spezifikation ``Web Components'', die mehrere Substandards
-unter sich vereint , Ziel dieser Arbeit soll ein In jüngster Zeit haben
-einige Vorschläge für eine ``low-level API'' Ziel dieser Arbeit ist die
-systematische Erfassung der neuen Technologien sowie eine praktische
-Erläuterung möglicher Architekturmodelle.
+Viele populäre Frameworks für den Browser als Zielplattform versprechen
+dem Entwickler eine bessere Kontrolle über die Webseite. Der Grund dafür
+liegt unter anderem darin, dass es bisher nicht möglich war einzelne
+Browserelemente in ihrem Aussehen und Verhalten zu isolieren. Selbst
+kleinste Änderungen können daher die Funktionalität des komplexen
+Gesamtsystems beeinträchtigen. Mit der neuen W3C Spezifikation ``Web
+Components''{[}\^{}fn3{]}, die mehrere Substandards unter sich vereint,
+soll die Modularität nun auch nativ im Browser unterstützt werden.
+Modularität als Designpattern ist in der IT schon länger Standard um
+komplexe Systeme beherrschbar zu gestalten und auch zukünftige
+Unsicherheiten abzuwägen.{[}4{]} Die Unsicherheit, die Frameworks immer
+anhaftet ist die Lebensdauer derselben.
 
-houdini https://www.w3.org/standards/techs/components
+Web Components ist der Entwickler in der Lage, eigene HTML Elemente zu
+definieren und diese per Templates zu exportieren oder zu importieren.
+Darüber hinaus können sie JavaScript und CSS Funktionalität enthalten
+und diese vom Rest der Webseite abzukapseln. Ziel dieser Arbeit ist die
+systematische Erfassung und Risikoabwägung der neuen Technologien sowie
+eine praktische Erläuterung möglicher Architekturmodelle. Bisher gibt es
+keine einheitlichen ``best practices'' wie denn die Komponenten
+umzusetzen sind, obwohl diese Technologien bereits länger durch
+Polyfills genutzt werden können. Manch ein Entwickler fürchtet schon
+eine Flut von schlecht gebauten Komponenten, die wiederum neue Probleme
+schaffen könnten.{[}5{]}
 
@@
 
-\subsection[:
-https://extensiblewebmanifesto.org/]{\texorpdfstring{\footnote{Ein
-  fiktives Gespräch mit einem JavaScript Neuling
-  https://hackernoon.com/how-it-feels-to-learn-javascript-in-2016-d3a717dd577f}:
-https://extensiblewebmanifesto.org/}{: https://extensiblewebmanifesto.org/}}\label{fn1-httpsextensiblewebmanifesto.org}
+\subsection{{[}\^{}fn3{]}:
+https://www.w3.org/standards/techs/components}\label{fn3-httpswww.w3.orgstandardstechscomponents}
 
@@
 {[}3{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016.
+
+\hypertarget{ref-Baldwin2006}{}
+{[}4{]} C. Y. Baldwin and K. B. Clark, ``Modularity in the design of
+complex engineering systems,'' in \emph{Complex engineered systems:
+Science meets technology}, D. Braha, A. A. Minai, and Y. Bar-Yam, Eds.
+Berlin, Heidelberg: Springer Berlin Heidelberg, 2006, pp. 175--205.
+
+\hypertarget{ref-Keith2014}{}
+{[}5{]} J. Keith, ``Web components,'' 2014.

### 2016-11-18 build/paper.tex
pdf/tex build

@@
 unüberblickbar und wandelt sich in einer Geschwindigkeit, die vielen
-Entwicklern Schwierigkeiten bereitet.\footnote{Ein fiktives Gespräch mit
-  einem JavaScript Neuling
-  https://hackernoon.com/how-it-feels-to-learn-javascript-in-2016-d3a717dd577f}
-Auch die Konsumenten und Nutzern der Webdiensten bekommen diese Probleme
-zumindest teilweise spüren. Einerseits spürbar in Form von ``Downtimes''
-der Services. Andererseits subtil wie etwa durch lange Ladezeiten wegen
-aufgeblähtem Quellcode. Frederic Filloux zeigte in einem Blogpost
-anschaulich, dass sich in einem Zeitungsartikel der britischen Zeitung
-``The Guardian'' zu jedem lesbaren Buchstaben des Artikels über 100
-Zeichen Code addieren.{[}3{]}
+Entwicklern Schwierigkeiten bereitet.{[}3{]} Auch die Konsumenten und
+Nutzern der Webdiensten bekommen diese Probleme zumindest teilweise
+spüren. Einerseits spürbar in Form von ``Downtimes'' der Services.
+Andererseits subtil wie etwa durch lange Ladezeiten wegen aufgeblähtem
+Quellcode. Frederic Filloux zeigte in einem Blogpost anschaulich, dass
+sich in einem Zeitungsartikel der britischen Zeitung ``The Guardian'' zu
+jedem lesbaren Buchstaben des Artikels über 100 Zeichen Code
+addieren.{[}4{]}
 
@@
 Gesamtsystems beeinträchtigen. Mit der neuen W3C Spezifikation ``Web
-Components'',{[}\^{}fn3{]} die mehrere Substandards unter sich vereint,
-soll die Modularität nun auch nativ im Browser unterstützt werden.
-Modularität als Designpattern ist in der IT schon länger Standard um
-komplexe Systeme beherrschbar zu gestalten und auch zukünftige
-Unsicherheiten abzuwägen.{[}4{]} Die Unsicherheit, die Frameworks immer
-anhaftet ist die Lebensdauer derselben.
+Components'', die mehrere Substandards unter sich vereint, soll die
+Modularität nun auch nativ im Browser unterstützt
+werden.{[}\^{}w3cWebComponent{]} Modularität als Designpattern ist in
+der IT schon länger Standard um komplexe Systeme beherrschbar zu
+gestalten und auch zukünftige Unsicherheiten abzuwägen.{[}5, p. 1{]} Die
+Unsicherheit, die Frameworks immer anhaftet, ist die Lebensdauer
+derselben.
 
@@
 eine Flut von schlecht gebauten Komponenten, die wiederum neue Probleme
-schaffen könnten.{[}5{]}
+schaffen könnten.{[}6{]}
 
@@
 Den Anfang dieser Arbeit soll eine Analyse der aktuellen Situation der
-Frontend Entwicklung aufzeigen. In ihr sollen aktuelle
-Architekturmodelle analysiert werden. Außerdem soll dieser Teil vor
-Augen führen, warum es bisher nur mit zusätzlichen Schichten der
-Abstraktion möglich war einen modularen Aufbau von Web Apps zu
-ermöglichen. In diesem Teil sollen auch die Probleme die der Entwicklung
-und Wartung dieser Systeme aufzeigen.
+Frontend Entwicklung aufzeigen. Dort soll auf die Problematik der
+bisherigen Architekturmodelle und die Probleme mit monolithischen
+Frameworks eingegangen werden. Außerdem soll dieser Teil vor Augen
+führen, warum es bisher nur mit zusätzlicher Abstraktionsebenen möglich
+war einen modularen Aufbau von Web Applikationen zu ermöglichen.
 
-Im nächsten Abschnitt der Arbeit sollen die
+Im nächsten Abschnitt der Arbeit sollen die neuen Standards
+aufgeschlüsselt, zugänglich gemacht und auf Anwendungsmöglichkeiten
+untersucht werden. Explorativ sollen gute Bespiele offengelegt werden
+und mögliche Einbettungen ins Gesamtsystem diskutiert werden. Auch die
+Probleme, wie sie beispielsweise bei alten Browsern auftreten können,
+sollen hier behandelt werden.
 
-\subsection{{[}\^{}fn3{]}:
-https://www.w3.org/standards/techs/components}\label{fn3-httpswww.w3.orgstandardstechscomponents}
+Der letzte Abschnitt soll das Thema noch mit einer Metaperspektive
+abrunden, in der auch die neuen low-level CSS Standards Houdini als Teil
+des Gesamtsystems betrachtet werden sollen.{[}\^{}w3cHoudini{]}
+
+\subsection{{[}\^{}w3cHoudini{]}:
+https://drafts.css-houdini.org/}\label{w3choudini-httpsdrafts.css-houdini.org}
 
@@
 
+\hypertarget{ref-Berner2016}{}
+{[}3{]} D. Berner, ``Not an imposter: Fighting front-end fatigue,''
+2016.
+
 \hypertarget{ref-Filloux2016}{}
-{[}3{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016.
+{[}4{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016.
 
 \hypertarget{ref-Baldwin2006}{}
-{[}4{]} C. Y. Baldwin and K. B. Clark, ``Modularity in the design of
+{[}5{]} C. Y. Baldwin and K. B. Clark, ``Modularity in the design of
 complex engineering systems,'' in \emph{Complex engineered systems:
@@
 \hypertarget{ref-Keith2014}{}
-{[}5{]} J. Keith, ``Web components,'' 2014.
+{[}6{]} J. Keith, ``Web components,'' 2014.

### 2016-11-18 build/paper.tex
pdf/tex build

@@
 Components'', die mehrere Substandards unter sich vereint, soll die
-Modularität nun auch nativ im Browser unterstützt
-werden.{[}\^{}w3cWebComponent{]} Modularität als Designpattern ist in
-der IT schon länger Standard um komplexe Systeme beherrschbar zu
-gestalten und auch zukünftige Unsicherheiten abzuwägen.{[}5, p. 1{]} Die
-Unsicherheit, die Frameworks immer anhaftet, ist die Lebensdauer
-derselben.
+Modularität nun auch nativ im Browser unterstützt werden.\footnote{https://www.w3.org/standards/techs/components}
+Modularität als Designpattern ist in der IT schon länger Standard um
+komplexe Systeme beherrschbar zu gestalten und auch zukünftige
+Unsicherheiten abzuwägen.{[}5, p. 1{]} Die Unsicherheit, die Frameworks
+immer anhaftet, ist die Lebensdauer derselben.
 
@@
 abrunden, in der auch die neuen low-level CSS Standards Houdini als Teil
-des Gesamtsystems betrachtet werden sollen.{[}\^{}w3cHoudini{]}
+des Gesamtsystems betrachtet werden sollen.{[}\^{}h{]}
 
-\subsection{{[}\^{}w3cHoudini{]}:
-https://drafts.css-houdini.org/}\label{w3choudini-httpsdrafts.css-houdini.org}
+\subsection{{[}\^{}h{]}:
+https://drafts.css-houdini.org/}\label{h-httpsdrafts.css-houdini.org}
 

### 2016-11-18 paper.tex
pdf/tex build

@@
 Gesamtsystems beeinträchtigen. Mit der neuen W3C Spezifikation ``Web
-Components''{[}\^{}fn3{]}, die mehrere Substandards unter sich vereint,
-soll die Modularität nun auch nativ im Browser unterstützt werden.
+Components'', die mehrere Substandards unter sich vereint, soll die
+Modularität nun auch nativ im Browser unterstützt werden.\footnote{https://www.w3.org/standards/techs/components}
 Modularität als Designpattern ist in der IT schon länger Standard um
 komplexe Systeme beherrschbar zu gestalten und auch zukünftige
-Unsicherheiten abzuwägen.{[}4{]} Die Unsicherheit, die Frameworks immer
-anhaftet ist die Lebensdauer derselben.
+Unsicherheiten abzuwägen.{[}5, p. 1{]} Die Unsicherheit, die Frameworks
+immer anhaftet, ist die Lebensdauer derselben.
 
@@
 Den Anfang dieser Arbeit soll eine Analyse der aktuellen Situation der
-Frontend Entwicklung aufzeigen. In ihr sollen aktuelle
-Architekturmodelle analysiert werden. Außerdem soll dieser Teil vor
-Augen führen, warum es bisher nur mit zusätzlichen Schichten der
-Abstraktion möglich war einen modularen Aufbau von Web Apps zu
-ermöglichen. In diesem Teil sollen auch die Probleme die der Entwicklung
-und Wartung dieser Systeme aufzeigen.
+Frontend Entwicklung aufzeigen. Dort soll auf die Problematik der
+bisherigen Architekturmodelle und die Probleme mit monolithischen
+Frameworks eingegangen werden. Außerdem soll dieser Teil vor Augen
+führen, warum es bisher nur mit zusätzlicher Abstraktionsebenen möglich
+war einen modularen Aufbau von Web Applikationen zu ermöglichen.
 
-Im nächsten Abschnitt der Arbeit sollen die
+Im nächsten Abschnitt der Arbeit sollen die neuen Standards
+aufgeschlüsselt, zugänglich gemacht und auf Anwendungsmöglichkeiten
+untersucht werden. Explorativ sollen gute Bespiele offengelegt werden
+und mögliche Einbettungen ins Gesamtsystem diskutiert werden. Auch die
+Probleme, wie sie beispielsweise bei alten Browsern auftreten können,
+sollen hier behandelt werden.
 
-\subsection{{[}\^{}fn3{]}:
-https://www.w3.org/standards/techs/components}\label{fn3-httpswww.w3.orgstandardstechscomponents}
-
-title: Microservices im Browser durch Web Components subtitle:
-Kombinierte W3C Spezifikationen für plattformübergreifende Browsermodule
-author: Jan Peteler, FHWS Würzburg-Schweinfurt,
-jan.peteler@student.fhws.de date: November 2016 abstract: \ldots{}
+Der letzte Abschnitt soll das Thema noch mit einer Metaperspektive
+abrunden, in der auch die neuen low-level CSS Standards Houdini als Teil
+des Gesamtsystems betrachtet werden sollen.\footnote{https://drafts.css-houdini.org/}
 

### 2016-11-18 build/paper.tex
pdf/tex build

--- 2016-11-18 build/paper.tex
@@
 
-Softwareentwicklung für den Browser ist ohne Zweifel eine komplexe
-Angelegenheit. Die Gründe für diesen Umstand sind so vielfältig, dass
-darüber eine ganze Forschungsarbeit geschrieben werden könnte. Je nach
-Blickwinkel könnte man von der technischen Seite argumentieren und
-beispielsweise Designschwächen von JavaScript oder die unflexiblen
-Browserplattformen als Bremsklotz der Webentwicklung benennen.{[}1{]}
-Oder man könnte die organisatorische Seite betrachten und die vielen
-beteiligten Organisationen benennen, die für Standardisierungsprozesse
-im Internet zuständig sind.{[}2{]} Oder aber man nimmt die historische
-Entwicklung in Augenschein, die dazu geführt hat, dass die technische
-Entwicklung der rasanten Evolution im Internet nicht Schritt gehalten
-hat. Die Zeiten statischer HTML/CSS Seiten auf einfachen Desktopgeräten
-ist mithin noch gar nicht so lange her.
+Softwareentwicklung für den Browser ist ohne Zweifel eine komplexe und
+fehleranfällige Angelegenheit. Die Gründe für diesen Umstand sind so
+vielfältig, dass darüber eine ganze Forschungsarbeit geschrieben werden
+könnte. Je nach Blickwinkel könnte man von der technischen Seite
+argumentieren und die Designschwächen der Browserplattform
+erläutern.{[}1{]} Oder man könnte die organisatorische Seite betrachten
+und die Schwerfälligkeit von Standardisierungsprozessen
+anprangern.{[}2{]} Oder aber man nimmt die historische Entwicklung in
+Augenschein. Die Zeiten statischer HTML/CSS Seiten auf einfachen
+Desktopgeräten ist mithin noch gar nicht so lange her.
 
 Dennoch tragen alle diese Facetten dazu bei, dass die Entwicklung und
-Wartung von komplexen Webapplikationen, wie sie heute Standard sind, mit
-enormen Zeit- und Geldaufwand verbunden sind. Die Anzahl der Frameworks,
-Werkzeuge und Bibliotheken mit Javascript als Zielsprache ist nahezu
-unüberblickbar und wandelt sich in einer Geschwindigkeit, die vielen
-Entwicklern Schwierigkeiten bereitet.{[}3{]} Auch die Konsumenten und
-Nutzern der Webdiensten bekommen diese Probleme zumindest teilweise
-spüren. Einerseits spürbar in Form von ``Downtimes'' der Services.
-Andererseits subtil wie etwa durch lange Ladezeiten wegen aufgeblähtem
-Quellcode. Frederic Filloux zeigte in einem Blogpost anschaulich, dass
-sich in einem Zeitungsartikel der britischen Zeitung ``The Guardian'' zu
-jedem lesbaren Buchstaben des Artikels über 100 Zeichen Code
-addieren.{[}4{]}
+Wartung von komplexen Webapplikationen mit enormen Zeit- und Geldaufwand
+verbunden sind. Die Anzahl der Frameworks, Werkzeuge und Bibliotheken
+mit Javascript als Zielsprache wandelt sich in einer Geschwindigkeit,
+die vielen Entwicklern Schwierigkeiten bereitet.{[}3{]} Auch die
+Konsumenten und Nutzern der Webdiensten bekommen diese Probleme
+zumindest teilweise spüren. Einerseits spürbar in Form von ``Downtimes''
+der Dienste. Andererseits subtil wie etwa durch lange Ladezeiten wegen
+aufgeblähtem Quellcode. Frederic Filloux zeigte in einem Blogpost
+anschaulich, dass sich in einem Zeitungsartikel der britischen Zeitung
+``The Guardian'' zu jedem lesbaren Buchstaben des Artikels über 100
+Zeichen Code addieren.{[}4{]}
 
@@
 Gesamtsystems beeinträchtigen. Mit der neuen W3C Spezifikation ``Web
-Components'', die mehrere Substandards unter sich vereint, soll die
-Modularität nun auch nativ im Browser unterstützt
-werden.{[}\^{}w3cWebComponent{]} Modularität als Designpattern ist in
-der IT schon länger Standard um komplexe Systeme beherrschbar zu
+Components,''\footnote{https://www.w3.org/standards/techs/components}
+die mehrere Substandards unter sich vereint, soll die Modularität nun
+auch nativ im Browser unterstützt werden. Modularität als Designpattern
+ist in der IT schon länger Standard um komplexe Systeme beherrschbar zu
 gestalten und auch zukünftige Unsicherheiten abzuwägen.{[}5, p. 1{]} Die
@@
 Web Components ist der Entwickler in der Lage, eigene HTML Elemente zu
-definieren und diese per Templates zu exportieren oder zu importieren.
-Darüber hinaus können sie JavaScript und CSS Funktionalität enthalten
-und diese vom Rest der Webseite abzukapseln. Ziel dieser Arbeit ist die
-systematische Erfassung und Risikoabwägung der neuen Technologien sowie
-eine praktische Erläuterung möglicher Architekturmodelle. Bisher gibt es
+definieren und diese per Templates zu ex- \& importieren. Darüber hinaus
+können sie JavaScript und CSS Funktionalität enthalten und diese vom
+Rest der Webseite abzukapseln. Ziel dieser Arbeit ist die systematische
+Erfassung und Risikoabwägung der neuen Technologien sowie eine
+praktische Erläuterung möglicher Architekturmodelle. Bisher gibt es
 keine einheitlichen ``best practices'' wie denn die Komponenten
@@
 abrunden, in der auch die neuen low-level CSS Standards Houdini als Teil
-des Gesamtsystems betrachtet werden sollen.{[}\^{}w3cHoudini{]}
-
-\subsection{{[}\^{}w3cHoudini{]}:
-https://drafts.css-houdini.org/}\label{w3choudini-httpsdrafts.css-houdini.org}
-
-title: Microservices im Browser durch Web Components subtitle:
-Kombinierte W3C Spezifikationen für plattformübergreifende Browsermodule
-author: Jan Peteler, FHWS Würzburg-Schweinfurt,
-jan.peteler@student.fhws.de date: November 2016 abstract: \ldots{}
+des Gesamtsystems betrachtet werden sollen.\footnote{https://drafts.css-houdini.org/}
 

### 2016-11-18 build/paper.tex
pdf/tex build

@@
 anprangern.{[}2{]} Oder aber man nimmt die historische Entwicklung in
-Augenschein. Die Zeiten statischer HTML/CSS Seiten auf einfachen
-Desktopgeräten ist mithin noch gar nicht so lange her.
+Augenschein. Die Zeiten statischer HTML/CSS Seiten auf Desktopgeräten
+ist mithin noch nicht lange her.
 
@@
 Wartung von komplexen Webapplikationen mit enormen Zeit- und Geldaufwand
-verbunden sind. Die Anzahl der Frameworks, Werkzeuge und Bibliotheken
-mit Javascript als Zielsprache wandelt sich in einer Geschwindigkeit,
-die vielen Entwicklern Schwierigkeiten bereitet.{[}3{]} Auch die
-Konsumenten und Nutzern der Webdiensten bekommen diese Probleme
-zumindest teilweise spüren. Einerseits spürbar in Form von ``Downtimes''
-der Dienste. Andererseits subtil wie etwa durch lange Ladezeiten wegen
-aufgeblähtem Quellcode. Frederic Filloux zeigte in einem Blogpost
-anschaulich, dass sich in einem Zeitungsartikel der britischen Zeitung
-``The Guardian'' zu jedem lesbaren Buchstaben des Artikels über 100
-Zeichen Code addieren.{[}4{]}
+verbunden ist. Offensichtlich wird das Problem, wenn man die schiere
+Anzahl der Frameworks und Bibliotheken mit Javascript als Zielsprache
+betrachtet, die vielen Entwicklern Schwierigkeiten bereitet.{[}3{]}
 
-Aus diesen Grund wurde schon 2013 das ``Extensible Web Manifesto''
-proklamiert. Darin wird, kurz gefasst, eine Öffnung der Webplattformen
-avisiert, um Webprogrammierung teilweise von Browserstandards zu
-entkoppeln.\footnote{https://extensiblewebmanifesto.org/} Drei Jahre
-später sind diese Ziele in W3C Spezifikationen konkretisiert worden.
-Einige dieser neuen Standards werden bereits nativ in (einigen) Browsern
-unterstützt.
+Um die inhärenten Designschwächen von Browsern aus der Welt zu schaffen,
+wurde schon im Jahr 2013 das ``Extensible Web Manifesto''
+proklamiert.\footnote{https://extensiblewebmanifesto.org/} Darin wird,
+kurz gefasst, eine Öffnung der Webplattformen avisiert, um
+Webprogrammierung teilweise von Browserstandards zu entkoppeln. Drei
+Jahre später sind diese Ziele in W3C Standards konkretisiert worden und
+sind teilweise schon umgesetzt worden.
 
@@
 ist in der IT schon länger Standard um komplexe Systeme beherrschbar zu
-gestalten und auch zukünftige Unsicherheiten abzuwägen.{[}5, p. 1{]} Die
+gestalten und auch zukünftige Unsicherheiten abzuwägen.{[}4, p. 1{]} Die
 Unsicherheit, die Frameworks immer anhaftet, ist die Lebensdauer
@@
 eine Flut von schlecht gebauten Komponenten, die wiederum neue Probleme
-schaffen könnten.{[}6{]}
+schaffen könnten.{[}5{]}
 
@@
 
-\hypertarget{ref-Filloux2016}{}
-{[}4{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016.
-
 \hypertarget{ref-Baldwin2006}{}
-{[}5{]} C. Y. Baldwin and K. B. Clark, ``Modularity in the design of
+{[}4{]} C. Y. Baldwin and K. B. Clark, ``Modularity in the design of
 complex engineering systems,'' in \emph{Complex engineered systems:
@@
 \hypertarget{ref-Keith2014}{}
-{[}6{]} J. Keith, ``Web components,'' 2014.
+{[}5{]} J. Keith, ``Web components,'' 2014.

### 2016-11-18 build/paper.tex
pdf/tex build

@@
 ist in der IT schon länger Standard um komplexe Systeme beherrschbar zu
-gestalten und auch zukünftige Unsicherheiten abzuwägen.{[}4, p. 1{]} Die
+gestalten und auch zukünftige Unsicherheiten abzuwägen.{[}4{]} Die
 Unsicherheit, die Frameworks immer anhaftet, ist die Lebensdauer

### 2016-11-18 build/paper.tex
pdf/tex build

@@
 Um die inhärenten Designschwächen von Browsern aus der Welt zu schaffen,
-wurde schon im Jahr 2013 das ``Extensible Web Manifesto''
+wurde schon im Jahr 2013 das \emph{Extensible Web Manifesto}
 proklamiert.\footnote{https://extensiblewebmanifesto.org/} Darin wird,
@@
 Jahre später sind diese Ziele in W3C Standards konkretisiert worden und
-sind teilweise schon umgesetzt worden.
+teilweise schon umgesetzt worden.
 
@@
 Viele populäre Frameworks für den Browser als Zielplattform versprechen
-dem Entwickler eine bessere Kontrolle über die Webseite. Der Grund dafür
-liegt unter anderem darin, dass es bisher nicht möglich war einzelne
-Browserelemente in ihrem Aussehen und Verhalten zu isolieren. Selbst
-kleinste Änderungen können daher die Funktionalität des komplexen
-Gesamtsystems beeinträchtigen. Mit der neuen W3C Spezifikation ``Web
-Components,''\footnote{https://www.w3.org/standards/techs/components}
+dem Entwickler eine bessere Kontrolle über seinen Webservice. Die
+Notwendigkeit rührt unter anderem daher, dass es bisher nicht möglich
+war einzelne Browserelemente in ihrem Aussehen und Verhalten zu
+isolieren. Selbst kleinste Änderungen können daher die Funktionalität
+des Gesamtsystems beeinträchtigen. Mit der neuen W3C Spezifikation
+\emph{Web Components},\footnote{https://www.w3.org/standards/techs/components}
 die mehrere Substandards unter sich vereint, soll die Modularität nun
@@
 
-Web Components ist der Entwickler in der Lage, eigene HTML Elemente zu
-definieren und diese per Templates zu ex- \& importieren. Darüber hinaus
-können sie JavaScript und CSS Funktionalität enthalten und diese vom
-Rest der Webseite abzukapseln. Ziel dieser Arbeit ist die systematische
-Erfassung und Risikoabwägung der neuen Technologien sowie eine
-praktische Erläuterung möglicher Architekturmodelle. Bisher gibt es
-keine einheitlichen ``best practices'' wie denn die Komponenten
-umzusetzen sind, obwohl diese Technologien bereits länger durch
-Polyfills genutzt werden können. Manch ein Entwickler fürchtet schon
-eine Flut von schlecht gebauten Komponenten, die wiederum neue Probleme
-schaffen könnten.{[}5{]}
+Mit \emph{Web Components} ist der Entwickler in der Lage, eigene HTML
+Elemente zu definieren und diese per Templates zu ex- \& importieren.
+Darüber hinaus können sie JavaScript und CSS Funktionalität enthalten
+und diese vom Rest der Webseite abzukapseln. Ziel dieser Arbeit ist die
+systematische Erfassung und Risikoabwägung der neuen Technologien sowie
+eine praktische Erläuterung möglicher Architekturmodelle. Bisher gibt es
+keine einheitliche Bestpractice, wie denn die Komponenten umzusetzen
+sind, obwohl diese Technologien bereits länger durch Polyfills genutzt
+werden können. Manch ein Entwickler fürchtet schon eine Flut von
+schlecht gebauten Komponenten, die wiederum neue Probleme schaffen
+könnten.{[}5{]}
 
@@
 
-Der letzte Abschnitt soll das Thema noch mit einer Metaperspektive
-abrunden, in der auch die neuen low-level CSS Standards Houdini als Teil
+Der letzte Abschnitt soll das Thema mit einer Metaperspektive abrunden,
+in der auch die neuen low-level CSS Standards \emph{Houdini} als Teil
 des Gesamtsystems betrachtet werden sollen.\footnote{https://drafts.css-houdini.org/}

### 2016-11-18 build/paper.tex
pdf/tex build

@@
 ist in der IT schon länger Standard um komplexe Systeme beherrschbar zu
-gestalten und auch zukünftige Unsicherheiten abzuwägen.{[}4{]} Die
+gestalten und auch zukünftige Unsicherheiten abzuwägen.{[}4, p. 1{]} Die
 Unsicherheit, die Frameworks immer anhaftet, ist die Lebensdauer
@@
 aufgeschlüsselt, zugänglich gemacht und auf Anwendungsmöglichkeiten
-untersucht werden. Explorativ sollen gute Bespiele offengelegt werden
+untersucht werden. Explorativ sollen Bestpractices offengelegt werden
 und mögliche Einbettungen ins Gesamtsystem diskutiert werden. Auch die
@@
 
-Der letzte Abschnitt soll das Thema mit einer Metaperspektive abrunden,
-in der auch die neuen low-level CSS Standards \emph{Houdini} als Teil
-des Gesamtsystems betrachtet werden sollen.\footnote{https://drafts.css-houdini.org/}
+Der letzte Abschnitt soll das Thema mit einer Metaebene abrunden, in der
+auch die neuen CSS Standards \emph{Houdini} als Teil des Gesamtsystems
+betrachtet werden sollen.\footnote{https://drafts.css-houdini.org/}
 

### 2016-11-18 build/paper.tex
pdf/tex build

@@
 \hypertarget{ref-Katz2013}{}
-{[}1{]} Y. Katz, ``Extend the web forward,'' 2013.
+{[}1{]} Y. Katz, ``Extend the web forward,'' 2013 {[}Online{]}.
+Available:
+\url{http://yehudakatz.com/2013/05/21/extend-the-web-forward/}
 
@@
 {[}2{]} P. Walton, ``Houdini: Maybe the most exciting development in css
-you've never heard of,'' 2016.
+you've never heard of,'' 2016 {[}Online{]}. Available:
+\url{https://www.smashingmagazine.com/2016/03/houdini-maybe-the-most-exciting-development-in-css-youve-never-heard-of}
 
 \hypertarget{ref-Berner2016}{}
-{[}3{]} D. Berner, ``Not an imposter: Fighting front-end fatigue,''
-2016.
+{[}3{]} D. Berner, ``Not an imposter: Fighting front-end fatigue,'' 2016
+{[}Online{]}. Available:
+\url{https://www.smashingmagazine.com/2016/11/not-an-imposter-fighting-front-end-fatigue/}
 
@@
 Science meets technology}, D. Braha, A. A. Minai, and Y. Bar-Yam, Eds.
-Berlin, Heidelberg: Springer Berlin Heidelberg, 2006, pp. 175--205.
+Berlin, Heidelberg: Springer Berlin Heidelberg, 2006, pp. 175--205
+{[}Online{]}. Available: \url{http://dx.doi.org/10.1007/3-540-32834-3_9}
 
 \hypertarget{ref-Keith2014}{}
-{[}5{]} J. Keith, ``Web components,'' 2014.
+{[}5{]} J. Keith, ``Web components,'' 2014 {[}Online{]}. Available:
+\url{https://adactio.com/journal/7431}

### 2016-12-07 paper.tex
pdf/tex build

@@
 
-Softwareentwicklung für den Browser ist ohne Zweifel eine komplexe
-Angelegenheit. Die Gründe für diesen Umstand sind so vielfältig, dass
-darüber eine ganze Forschungsarbeit geschrieben werden könnte. Je nach
-Blickwinkel könnte man von der technischen Seite argumentieren und
-beispielsweise Designschwächen von JavaScript oder die unflexiblen
-Browserplattformen als Bremsklotz der Webentwicklung benennen.{[}1{]}
-Oder man könnte die organisatorische Seite betrachten und die vielen
-beteiligten Organisationen benennen, die für Standardisierungsprozesse
-im Internet zuständig sind.{[}2{]} Oder aber man nimmt die historische
-Entwicklung in Augenschein, die dazu geführt hat, dass die technische
-Entwicklung der rasanten Evolution im Internet nicht Schritt gehalten
-hat. Die Zeiten statischer HTML/CSS Seiten auf einfachen Desktopgeräten
-ist mithin noch gar nicht so lange her.
+Softwareentwicklung für den Browser ist ohne Zweifel eine komplexe und
+fehleranfällige Angelegenheit. Die Gründe für diesen Umstand sind so
+vielfältig, dass darüber eine ganze Forschungsarbeit geschrieben werden
+könnte. Je nach Blickwinkel könnte man von der technischen Seite
+argumentieren und die Designschwächen der Browserplattform
+erläutern.{[}1{]} Oder man könnte die organisatorische Seite betrachten
+und die Schwerfälligkeit von Standardisierungsprozessen
+anprangern.{[}2{]} Oder aber man nimmt die historische Entwicklung in
+Augenschein. Die Zeiten statischer HTML/CSS Seiten auf Desktopgeräten
+ist mithin noch nicht lange her.
 
 Dennoch tragen alle diese Facetten dazu bei, dass die Entwicklung und
-Wartung von komplexen Webapplikationen, wie sie heute Standard sind, mit
-enormen Zeit- und Geldaufwand verbunden sind. Die Anzahl der Frameworks,
-Werkzeuge und Bibliotheken mit Javascript als Zielsprache ist nahezu
-unüberblickbar und wandelt sich in einer Geschwindigkeit, die vielen
-Entwicklern Schwierigkeiten bereitet.{[}3{]} Auch die Konsumenten und
-Nutzern der Webdiensten bekommen diese Probleme zumindest teilweise
-spüren. Einerseits spürbar in Form von ``Downtimes'' der Services.
-Andererseits subtil wie etwa durch lange Ladezeiten wegen aufgeblähtem
-Quellcode. Frederic Filloux zeigte in einem Blogpost anschaulich, dass
-sich in einem Zeitungsartikel der britischen Zeitung ``The Guardian'' zu
-jedem lesbaren Buchstaben des Artikels über 100 Zeichen Code
-addieren.{[}4{]}
+Wartung von komplexen Webapplikationen mit enormen Zeit- und Geldaufwand
+verbunden ist. Offensichtlich wird das Problem, wenn man die schiere
+Anzahl der Frameworks und Bibliotheken mit Javascript als Zielsprache
+betrachtet, die vielen Entwicklern Schwierigkeiten bereitet.{[}3{]}
 
-Aus diesen Grund wurde schon 2013 das ``Extensible Web Manifesto''
-proklamiert. Darin wird, kurz gefasst, eine Öffnung der Webplattformen
-avisiert, um Webprogrammierung teilweise von Browserstandards zu
-entkoppeln.\footnote{https://extensiblewebmanifesto.org/} Drei Jahre
-später sind diese Ziele in W3C Spezifikationen konkretisiert worden.
-Einige dieser neuen Standards werden bereits nativ in (einigen) Browsern
-unterstützt.
+Um die inhärenten Designschwächen von Browsern aus der Welt zu schaffen,
+wurde schon im Jahr 2013 das \emph{Extensible Web Manifesto}
+proklamiert.\footnote{https://extensiblewebmanifesto.org/} Darin wird,
+kurz gefasst, eine Öffnung der Webplattformen avisiert, um
+Webprogrammierung teilweise von Browserstandards zu entkoppeln. Drei
+Jahre später sind diese Ziele in W3C Standards konkretisiert worden und
+teilweise schon umgesetzt worden.
 
@@
 Viele populäre Frameworks für den Browser als Zielplattform versprechen
-dem Entwickler eine bessere Kontrolle über die Webseite. Der Grund dafür
-liegt unter anderem darin, dass es bisher nicht möglich war einzelne
-Browserelemente in ihrem Aussehen und Verhalten zu isolieren. Selbst
-kleinste Änderungen können daher die Funktionalität des komplexen
-Gesamtsystems beeinträchtigen. Mit der neuen W3C Spezifikation ``Web
-Components'', die mehrere Substandards unter sich vereint, soll die
-Modularität nun auch nativ im Browser unterstützt werden.\footnote{https://www.w3.org/standards/techs/components}
-Modularität als Designpattern ist in der IT schon länger Standard um
-komplexe Systeme beherrschbar zu gestalten und auch zukünftige
-Unsicherheiten abzuwägen.{[}5, p. 1{]} Die Unsicherheit, die Frameworks
-immer anhaftet, ist die Lebensdauer derselben.
+dem Entwickler eine bessere Kontrolle über seinen Webservice. Die
+Notwendigkeit rührt unter anderem daher, dass es bisher nicht möglich
+war einzelne Browserelemente in ihrem Aussehen und Verhalten zu
+isolieren. Selbst kleinste Änderungen können daher die Funktionalität
+des Gesamtsystems beeinträchtigen. Mit der neuen W3C Spezifikation
+\emph{Web Components},\footnote{https://www.w3.org/standards/techs/components}
+die mehrere Substandards unter sich vereint, soll die Modularität nun
+auch nativ im Browser unterstützt werden. Modularität als Designpattern
+ist in der IT schon länger Standard um komplexe Systeme beherrschbar zu
+gestalten und auch zukünftige Unsicherheiten abzuwägen.{[}4, S. 1{]} Die
+Unsicherheit, die Frameworks immer anhaftet, ist die Lebensdauer
+derselben.
 
-Web Components ist der Entwickler in der Lage, eigene HTML Elemente zu
-definieren und diese per Templates zu exportieren oder zu importieren.
+Mit \emph{Web Components} ist der Entwickler in der Lage, eigene HTML
+Elemente zu definieren und diese per Templates zu ex- \& importieren.
 Darüber hinaus können sie JavaScript und CSS Funktionalität enthalten
@@
 eine praktische Erläuterung möglicher Architekturmodelle. Bisher gibt es
-keine einheitlichen ``best practices'' wie denn die Komponenten
-umzusetzen sind, obwohl diese Technologien bereits länger durch
-Polyfills genutzt werden können. Manch ein Entwickler fürchtet schon
-eine Flut von schlecht gebauten Komponenten, die wiederum neue Probleme
-schaffen könnten.{[}6{]}
+keine einheitliche Bestpractice, wie denn die Komponenten umzusetzen
+sind, obwohl diese Technologien bereits länger durch Polyfills genutzt
+werden können. Manch ein Entwickler fürchtet schon eine Flut von
+schlecht gebauten Komponenten, die wiederum neue Probleme schaffen
+könnten.{[}5{]}
 
@@
 
-Der letzte Abschnitt soll das Thema noch mit einer Metaperspektive
-abrunden, in der auch die neuen low-level CSS Standards Houdini als Teil
-des Gesamtsystems betrachtet werden sollen.\footnote{https://drafts.css-houdini.org/}
+Der letzte Abschnitt soll das Thema mit einer Metaebene abrunden, in der
+auch die neuen CSS Standards \emph{Houdini} als Teil des Gesamtsystems
+betrachtet werden sollen.\footnote{https://drafts.css-houdini.org/}
 
@@
 \hypertarget{ref-Katz2013}{}
-{[}1{]} Y. Katz, ``Extend the web forward,'' 2013.
+{[}1{]} Y. Katz, „Extend the Web Forward``, 2013 {[}Online{]}. Verfügbar
+unter: \url{http://yehudakatz.com/2013/05/21/extend-the-web-forward/}
 
 \hypertarget{ref-Walton2016}{}
-{[}2{]} P. Walton, ``Houdini: Maybe the most exciting development in css
-you've never heard of,'' 2016.
+{[}2{]} P. Walton, „Houdini: Maybe The Most Exciting Development In CSS
+You've Never Heard Of``, 2016 {[}Online{]}. Verfügbar unter:
+\url{https://www.smashingmagazine.com/2016/03/houdini-maybe-the-most-exciting-development-in-css-youve-never-heard-of}
 
 \hypertarget{ref-Berner2016}{}
-{[}3{]} D. Berner, ``Not an imposter: Fighting front-end fatigue,''
-2016.
-
-\hypertarget{ref-Filloux2016}{}
-{[}4{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016.
+{[}3{]} D. Berner, „Not An Imposter: Fighting Front-End Fatigue``, 2016
+{[}Online{]}. Verfügbar unter:
+\url{https://www.smashingmagazine.com/2016/11/not-an-imposter-fighting-front-end-fatigue/}
 
 \hypertarget{ref-Baldwin2006}{}
-{[}5{]} C. Y. Baldwin and K. B. Clark, ``Modularity in the design of
-complex engineering systems,'' in \emph{Complex engineered systems:
-Science meets technology}, D. Braha, A. A. Minai, and Y. Bar-Yam, Eds.
-Berlin, Heidelberg: Springer Berlin Heidelberg, 2006, pp. 175--205.
+{[}4{]} C. Y. Baldwin und K. B. Clark, „Modularity in the Design of
+Complex Engineering Systems``, in \emph{Complex Engineered Systems:
+Science Meets Technology}, D. Braha, A. A. Minai, und Y. Bar-Yam, Hrsg.
+Berlin, Heidelberg: Springer Berlin Heidelberg, 2006, S. 175--205
+{[}Online{]}. Verfügbar unter:
+\url{http://dx.doi.org/10.1007/3-540-32834-3_9}
 
 \hypertarget{ref-Keith2014}{}
-{[}6{]} J. Keith, ``Web components,'' 2014.
+{[}5{]} J. Keith, „Web Components``, 2014 {[}Online{]}. Verfügbar unter:
+\url{https://adactio.com/journal/7431}

### 2016-12-07 paper.tex
pdf/tex build

@@
+\documentclass[ngerman,]{assets/latex/ieee}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[unicode=true]{hyperref}
+\hypersetup{
+            pdftitle={Browsernative Microservices},
+            pdfauthor={Jan Peteler, FH Würzburg-Schweinfurt, jan.peteler@student.fhws.de},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[shorthands=off,main=ngerman]{babel}
+\else
+  \usepackage{polyglossia}
+  \setmainlanguage[]{german}
+\fi
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+% set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+
+\title{Browsernative Microservices}
+\providecommand{\subtitle}[1]{}
+\subtitle{Modulare Webarchitekturen durch neue Browserstandards entwickeln}
+\author{Jan Peteler, FH Würzburg-Schweinfurt, jan.peteler@student.fhws.de}
+\date{Januar 2017}
+
+\begin{document}
+\maketitle
+
 \section{Ausgangssituation}\label{ausgangssituation}
@@
 \url{https://adactio.com/journal/7431}
+
+\end{document}

### 2016-12-07 build/paper.tex
update

@@
+\documentclass[ngerman,]{../assets/latex/ieee}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[unicode=true]{hyperref}
+\hypersetup{
+            pdftitle={Browsernative Microservices},
+            pdfauthor={Jan Peteler, FH Würzburg-Schweinfurt, jan.peteler@student.fhws.de},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[shorthands=off,main=ngerman]{babel}
+\else
+  \usepackage{polyglossia}
+  \setmainlanguage[]{german}
+\fi
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\newenvironment{Shaded}{}{}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{{#1}}}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{{#1}}}
+\newcommand{\ImportTok}[1]{{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{{#1}}}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{{#1}}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{{#1}}}
+\newcommand{\BuiltInTok}[1]{{#1}}
+\newcommand{\ExtensionTok}[1]{{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{{#1}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{{#1}}}
+\newcommand{\RegionMarkerTok}[1]{{#1}}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
+\newcommand{\NormalTok}[1]{{#1}}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+% set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+
+\title{Browsernative Microservices}
+\providecommand{\subtitle}[1]{}
+\subtitle{Modulare Webarchitekturen durch neue Browserstandards entwickeln}
+\author{Jan Peteler, FH Würzburg-Schweinfurt, jan.peteler@student.fhws.de}
+\date{Januar 2017}
+
+\begin{document}
+\maketitle
+
 \section{Ausgangssituation}\label{ausgangssituation}
 
-Softwareentwicklung für den Browser ist ohne Zweifel eine komplexe und
-fehleranfällige Angelegenheit. Die Gründe für diesen Umstand sind so
-vielfältig, dass darüber eine ganze Forschungsarbeit geschrieben werden
-könnte. Je nach Blickwinkel könnte man von der technischen Seite
-argumentieren und die Designschwächen der Browserplattform
-erläutern.{[}1{]} Oder man könnte die organisatorische Seite betrachten
-und die Schwerfälligkeit von Standardisierungsprozessen
-anprangern.{[}2{]} Oder aber man nimmt die historische Entwicklung in
-Augenschein. Die Zeiten statischer HTML/CSS Seiten auf Desktopgeräten
-ist mithin noch nicht lange her.
+en der Browserplattform erläutern.{[}1{]} Oder man könnte die
+organisatorische Seite betrachten und die Schwerfälligkeit von
+Standardisierungsprozessen anprangern.{[}2{]} Oder aber man nimmt die
+historische Entwicklung in Augenschein. Die Zeiten statischer HTML/CSS
+Seiten auf Desktopgeräten ist mithin noch nicht lange her.
 
-Dennoch tragen alle diese Facetten dazu bei, dass die Entwicklung und
-Wartung von komplexen Webapplikationen mit enormen Zeit- und Geldaufwand
-verbunden ist. Offensichtlich wird das Problem, wenn man die schiere
-Anzahl der Frameworks und Bibliotheken mit Javascript als Zielsprache
-betrachtet, die vielen Entwicklern Schwierigkeiten bereitet.{[}3{]}
+cript als Zielsprache betrachtet, die vielen Entwicklern Schwierigkeiten
+bereitet.{[}3{]}
 
-Um die inhärenten Designschwächen von Browsern aus der Welt zu schaffen,
-wurde schon im Jahr 2013 das \emph{Extensible Web Manifesto}
-proklamiert.\footnote{https://extensiblewebmanifesto.org/} Darin wird,
-kurz gefasst, eine Öffnung der Webplattformen avisiert, um
-Webprogrammierung teilweise von Browserstandards zu entkoppeln. Drei
-Jahre später sind diese Ziele in W3C Standards konkretisiert worden und
-teilweise schon umgesetzt worden.
-
-\section{Zielsetzung}\label{zielsetzung}
-
-Viele populäre Frameworks für den Browser als Zielplattform versprechen
-dem Entwickler eine bessere Kontrolle über seinen Webservice. Die
-Notwendigkeit rührt unter anderem daher, dass es bisher nicht möglich
-war einzelne Browserelemente in ihrem Aussehen und Verhalten zu
-isolieren. Selbst kleinste Änderungen können daher die Funktionalität
-des Gesamtsystems beeinträchtigen. Mit der neuen W3C Spezifikation
-\emph{Web Components},\footnote{https://www.w3.org/standards/techs/components}
-die mehrere Substandards unter sich vereint, soll die Modularität nun
-auch nativ im Browser unterstützt werden. Modularität als Designpattern
-ist in der IT schon länger Standard um komplexe Systeme beherrschbar zu
-gestalten und auch zukünftige Unsicherheiten abzuwägen.{[}4, p. 1{]} Die
-Unsicherheit, die Frameworks immer anhaftet, ist die Lebensdauer
-derselben.
+nftige Unsicherheiten abzuwägen.{[}4, S. 1--3{]} Die Unsicherheit, die
+Frameworks immer anhaftet, ist die Lebensdauer derselben.
 
@@
 Probleme, wie sie beispielsweise bei alten Browsern auftreten können,
-sollen hier behandelt werden.
+sollen hier behandelt werden.\\
+So we are
+
+\begin{quote}
+This is a blockquote
+\end{quote}
+
+´hello world´
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{Javascript baby}
+\end{Highlighting}
+\end{Shaded}
+
+\texttt{so\ this\ is\ the\ js\ world}
+
+one more last test
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  This is the list
+\end{enumerate}
+
+This is a sub
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  Anothoter list
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  {[} {]} oohhhhhhh
+\end{itemize}
 
@@
 \hypertarget{ref-Katz2013}{}
-{[}1{]} Y. Katz, ``Extend the web forward,'' 2013 {[}Online{]}.
-Available:
-\url{http://yehudakatz.com/2013/05/21/extend-the-web-forward/}
+{[}1{]} Y. Katz, „Extend the Web Forward``, 2013 {[}Online{]}. Verfügbar
+unter: \url{http://yehudakatz.com/2013/05/21/extend-the-web-forward/}
 
 \hypertarget{ref-Walton2016}{}
-{[}2{]} P. Walton, ``Houdini: Maybe the most exciting development in css
-you've never heard of,'' 2016 {[}Online{]}. Available:
+{[}2{]} P. Walton, „Houdini: Maybe The Most Exciting Development In CSS
+You've Never Heard Of``, 2016 {[}Online{]}. Verfügbar unter:
 \url{https://www.smashingmagazine.com/2016/03/houdini-maybe-the-most-exciting-development-in-css-youve-never-heard-of}
@@
 \hypertarget{ref-Berner2016}{}
-{[}3{]} D. Berner, ``Not an imposter: Fighting front-end fatigue,'' 2016
-{[}Online{]}. Available:
+{[}3{]} D. Berner, „Not An Imposter: Fighting Front-End Fatigue``, 2016
+{[}Online{]}. Verfügbar unter:
 \url{https://www.smashingmagazine.com/2016/11/not-an-imposter-fighting-front-end-fatigue/}
 
-\hypertarget{ref-Baldwin2006}{}
-{[}4{]} C. Y. Baldwin and K. B. Clark, ``Modularity in the design of
-complex engineering systems,'' in \emph{Complex engineered systems:
-Science meets technology}, D. Braha, A. A. Minai, and Y. Bar-Yam, Eds.
-Berlin, Heidelberg: Springer Berlin Heidelberg, 2006, pp. 175--205
-{[}Online{]}. Available: \url{http://dx.doi.org/10.1007/3-540-32834-3_9}
+\hypertarget{ref-Dodson2016}{}
+{[}4{]} R. Dodson, „The Case for Custom Elements: Part 2 -- Dev Channel
+-- Medium``. 2016 {[}Online{]}. Verfügbar unter:
+\href{https://medium.com/dev-channel/the-case-for-custom-elements-part-2-2efe42ce9133\%7B/\#\%7D.gl5nuty2r}{https://medium.com/dev-channel/the-case-for-custom-elements-part-2-2efe42ce9133\{\textbackslash{}\#\}.gl5nuty2r}.
+{[}Zugegriffen: 01-Dez-2016{]}
 
 \hypertarget{ref-Keith2014}{}
-{[}5{]} J. Keith, ``Web components,'' 2014 {[}Online{]}. Available:
+{[}5{]} J. Keith, „Web Components``, 2014 {[}Online{]}. Verfügbar unter:
 \url{https://adactio.com/journal/7431}
+
+\end{document}

### 2017-01-03 build/paper.tex
pdf/tex build

--- 2016-11-18 build/paper.tex
@@
-\section{Ausgangssituation}\label{ausgangssituation}
+\documentclass[]{assets/latex/ieee}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[unicode=true]{hyperref}
+\hypersetup{
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-Softwareentwicklung für den Browser ist ohne Zweifel eine komplexe und
-fehleranfällige Angelegenheit. Die Gründe für diesen Umstand sind so
-vielfältig, dass darüber eine ganze Forschungsarbeit geschrieben werden
-könnte. Je nach Blickwinkel könnte man von der technischen Seite
-argumentieren und die Designschwächen der Browserplattform
-erläutern.{[}1{]} Oder man könnte die organisatorische Seite betrachten
-und die Schwerfälligkeit von Standardisierungsprozessen
-anprangern.{[}2{]} Oder aber man nimmt die historische Entwicklung in
-Augenschein. Die Zeiten statischer HTML/CSS Seiten auf Desktopgeräten
-ist mithin noch nicht lange her.
+% set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 
-Dennoch tragen alle diese Facetten dazu bei, dass die Entwicklung und
-Wartung von komplexen Webapplikationen mit enormen Zeit- und Geldaufwand
-verbunden ist. Offensichtlich wird das Problem, wenn man die schiere
-Anzahl der Frameworks und Bibliotheken mit Javascript als Zielsprache
-betrachtet, die vielen Entwicklern Schwierigkeiten bereitet.{[}3{]}
+\date{}
 
-Um die inhärenten Designschwächen von Browsern aus der Welt zu schaffen,
-wurde schon im Jahr 2013 das \emph{Extensible Web Manifesto}
-proklamiert.\footnote{https://extensiblewebmanifesto.org/} Darin wird,
-kurz gefasst, eine Öffnung der Webplattformen avisiert, um
-Webprogrammierung teilweise von Browserstandards zu entkoppeln. Drei
-Jahre später sind diese Ziele in W3C Standards konkretisiert worden und
-teilweise schon umgesetzt worden.
+\begin{document}
 
-\section{Zielsetzung}\label{zielsetzung}
+\section{Introduction}\label{introduction}
 
-Viele populäre Frameworks für den Browser als Zielplattform versprechen
-dem Entwickler eine bessere Kontrolle über seinen Webservice. Die
-Notwendigkeit rührt unter anderem daher, dass es bisher nicht möglich
-war einzelne Browserelemente in ihrem Aussehen und Verhalten zu
-isolieren. Selbst kleinste Änderungen können daher die Funktionalität
-des Gesamtsystems beeinträchtigen. Mit der neuen W3C Spezifikation
-\emph{Web Components},\footnote{https://www.w3.org/standards/techs/components}
-die mehrere Substandards unter sich vereint, soll die Modularität nun
-auch nativ im Browser unterstützt werden. Modularität als Designpattern
-ist in der IT schon länger Standard um komplexe Systeme beherrschbar zu
-gestalten und auch zukünftige Unsicherheiten abzuwägen.{[}4, p. 1{]} Die
-Unsicherheit, die Frameworks immer anhaftet, ist die Lebensdauer
-derselben.
+\begin{quote}
+Simplicity is prerequisite for reliability. - Edsger W. Dijkstra
+\end{quote}
 
-Mit \emph{Web Components} ist der Entwickler in der Lage, eigene HTML
-Elemente zu definieren und diese per Templates zu ex- \& importieren.
-Darüber hinaus können sie JavaScript und CSS Funktionalität enthalten
-und diese vom Rest der Webseite abzukapseln. Ziel dieser Arbeit ist die
-systematische Erfassung und Risikoabwägung der neuen Technologien sowie
-eine praktische Erläuterung möglicher Architekturmodelle. Bisher gibt es
-keine einheitliche Bestpractice, wie denn die Komponenten umzusetzen
-sind, obwohl diese Technologien bereits länger durch Polyfills genutzt
-werden können. Manch ein Entwickler fürchtet schon eine Flut von
-schlecht gebauten Komponenten, die wiederum neue Probleme schaffen
-könnten.{[}5{]}
+Der Erfinder der Programmiersprache Clojure, Rich Hickey, ist ohne
+Zweifel eine Koryphäe auf seinem Gebiet, der Strukturierung von
+komplexen Systemen. In einer vielbeachteten Keynote aus dem Jahr 2012
+geht er auf etymologisch-philosophische Spurensuche nach dem Wort
+\textbf{simplicity } aus Sicht eines Softwareentwicklers.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=48s}{Rails
+  Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} Das Adjektiv
+\emph{simple} hat demnach seinen Ursprung im lateinischen Wort
+\emph{simplex}, was soviel wie \emph{einfach} oder \emph{einzeln}
+bedeutet. In Gegensatz dazu stehen Eigenschaften wie \emph{complex} oder
+\emph{multiplex}. Qualitative Software ist, so Hickey, vor allem simpel
+- im Prozess, im Design, in der Struktur und in der Entwicklung.
 
-\section{Vorgehensweise}\label{vorgehensweise}
+Bezeichnend für diese Keynote ist, dass sie von Rich Hickey im Rahmen
+einer Webentwicklerkonferenz gehalten wurde. Software für den Browser
+war (und ist) seit langer Zeit maßgeblich geprägt von
+\textbf{Komplexität} auf mehreren Ebenen. Im Laufe dieser Bachelorarbeit
+werden diese Stru
 
-Den Anfang dieser Arbeit soll eine Analyse der aktuellen Situation der
-Frontend Entwicklung aufzeigen. Dort soll auf die Problematik der
-bisherigen Architekturmodelle und die Probleme mit monolithischen
-Frameworks eingegangen werden. Außerdem soll dieser Teil vor Augen
-führen, warum es bisher nur mit zusätzlicher Abstraktionsebenen möglich
-war einen modularen Aufbau von Web Applikationen zu ermöglichen.
+\begin{itemize}
+\tightlist
+\item
+  Das \emph{Document Object Model} als Rückgrad jeder Webapplikation ist
+  hierarchisch strukturiert und deren Elemente damit keinesfalls
+  unabhängig in ihrer Darstellung und Reihenfolge.
+\item
+  JavaScript als single-threaded Skriptsprache lässt sich schlecht ihn
+  ihrem Verhalten isolieren, ist Fehleranfällig und hatte lange Zeit nur
+  sehr wenig idiomatische Lösungsansätze für komplexe Probleme, wie
+  beispielsweise Asynchronität
+\item
+  CSS ist geprägt von stetigem Überschreiben vorher definierter Regeln
+  und verletzt damit Simplizität in ihrer Struktur, Design und dem
+  Entwicklungsprozesses auf bester Art und Weise.
+\end{itemize}
 
-Im nächsten Abschnitt der Arbeit sollen die neuen Standards
-aufgeschlüsselt, zugänglich gemacht und auf Anwendungsmöglichkeiten
-untersucht werden. Explorativ sollen Bestpractices offengelegt werden
-und mögliche Einbettungen ins Gesamtsystem diskutiert werden. Auch die
-Probleme, wie sie beispielsweise bei alten Browsern auftreten können,
-sollen hier behandelt werden.
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
 
-Der letzte Abschnitt soll das Thema mit einer Metaebene abrunden, in der
-auch die neuen CSS Standards \emph{Houdini} als Teil des Gesamtsystems
-betrachtet werden sollen.\footnote{https://drafts.css-houdini.org/}
+Betrachtet man andere populäre Webframeworks dieser Zeit, wie
+beispielsweise React oder Angular, lässt sich ziemlich schnell ein
+gemeinsames Designpattern ausmachen, wie Komponenten intern ihre
+Zuständen verwalten. Der Facebook Entwickler Dan Abramov hat dieses
+dichotome Pattern systematisch erfasst und in zwei Kategorien
+eingeteilt.
+
+Nach Abramov existieren zum einen Komponenten, die alleine für das
+\textbf{Darstellen von Information} zuständig sind. Diese Komponenten
+nehmen keinerlei Einfluss auf Informationen oder anderen Komponenten um
+sich herum. Sie sind im wahrsten Sinne passiv und fremdgesteuert über
+klar definierte Schnittstellen. Diese Komponenten sind häufig flexibel
+einsetzbar und hochgradig wiederwerwendbar. Abramov nennt diese
+Komponenten ``Presentational Components''.{[}1{]} Verortet man diese Art
+von Komponenten innerhalb des \emph{MVC Pattern}, sind die Komponenten
+reine \emph{View-Elemente}.
+
+Im Gegensatz dazu stehen Komponenten, die für die \textbf{Verarbeitung
+von Informationen} zuständig sind. Diese so genannten ``Container
+Components'' haben oft einen internen Zustand, den sie verändern können
+und an ihre Kindkomponenten weiterreichen können.{[}1{]} Im \emph{MVC
+Pattern} handelt es sich um die \emph{Controller}. In der funktionalen
+Programmiersprache Elm werden diese Elemente \emph{Updater} genannt, was
+ihre Funktion noch besser umschreibt.
+
+Ein üblicher eventgesteuerter Webservice setzt sich aus
+unterschiedlichsten Komponenten zusammen, die wiederum
+unterschiedlichste Eventlistener \& -emitter in sich subsummieren. Diese
+inhärente Komplexität verlangt geradezu nach einer klaren,
+deterministischen Struktur des Webservices, die das Zusammenspiel
+orchestriert. In der Analogie des Orchesters gesprochen, benötigt der
+Webservice (oder sogar die gesamte Webapplikation) einen Dirigenten, der
+für die Steuerung verantwortlich ist.
+
+\section{Store}\label{store}
+
+\begin{verbatim}
+   +-----------------------+
+   |         Store         |
+   +-----------------------+
+       /\|  |
+
+\end{verbatim}
+
+die Änderungen des Zustands vorhersehbar aber auch zurückverfolgbar
+macht.
+
+\section{Microservices}\label{microservices}
+
+In a nutshell a microservice is a small, autonomous service that works
+together with other services seamlessly.{[}2, p. 2{]} While the concept
+of microservices lacks a formal standardization there are certain
+defacto standardized ideas emergine from this pattern. The purpose of
+this section is to match those ideas against the manifestations of web
+components.
+
+First of all, from a technical perspective, a microservice reinforces
+the \emph{Single Responsibility Principle} defined by Robert C. Martin:
+``Gather together those things that change for the same reason and
+separate those things that change for different reasons.''{[}3{]} An a
+way this principle tackles another often cited design principle of the
+\emph{seperation of concerns}. Web Components incorporate this principle
+in multiple ways while still remaining flexible.
+
+Most obvious ist the gathering of all related code under the umbrellar
+of a single HTML tag. Grouping together HTML, JS and CSS Code in a safe,
+sandboxed environment exposes the possibility to build more cohesive and
+understandable services. In the typical global nature of web development
+those three pillars are separated. This circumstance left the developer
+switching back and forth between code bases trying to find a tricky way
+to glue related parts together.
+
+Secondly, the sub-standard \emph{custom elements} introduces so called
+lifecycle methods and a getter/setter interface exposing the
+functionality to the developer. Event handling, for example, can be
+registered in place which is much more declarative than assigning event
+listeners from the outside. Of course, this events can be pushed down to
+nested tags, allowing an increasingly granular system design. This
+approach will be explained further in the upcoming sections.
+
+The concept of microservices incooperates not only a technical
+perspective. Microservice patterns are a product of real-world
+usage.{[}2, p. 1{]} In a real world we typical have to deal with the so
+called \emph{Conway's Law}:
+
+\begin{quote}
+``organizations which design systems \ldots{} are constrained to produce
+designs which are copies of the communication~structures~of these
+organizations''. {[}4{]}
+\end{quote}
+
+Following this logic any company, whether its web-related or not, should
+be devided in units grouped around a destinct business service to
+optimise the workflow. Fowler and Lewis outlines this approach as an
+``alignment of business capabilities''{[}5, p. 2{]} While this kind of
+structure may be true for companies like Google or Amazon, there is a
+vast majority of companies developing for the web which are grouped
+around tasks.{[}6{]} A very common structure is formed by the technology
+stack: UX Designers, Frontend- \& Backend Developers.
+
+The microservice approach models all those functions into
+business-related units. Web components are one (but important) way to
+tie up those diciplines as one component can host a single independent
+business service. Combined with a flexible backend service these
+components can be huge gain over the cumbersome functional
+organizational approach.
+
+Macroperspektive / Composition
+
+http://alistair.cockburn.us/Hexagonal+architecture
+
+MVC Pattern
+
+\section{Progressive Enhancement}\label{progressive-enhancement}
+
+\subsection{Chapter about progressive
+enhancement}\label{chapter-about-progressive-enhancement}
+
+title: Browsernative Microservices subtitle: Modulare Webarchitekturen
+durch neue W3C-Spezifikationen author: Jan Peteler, FH
+Würzburg-Schweinfurt, jan.peteler@student.fhws.de date: Januar 2017
+lang: de-DE abstract: Webapplikationen lassen sich oft nur mit
+zusätzlichen Abstraktionsebenen wie beispielsweise ``Frameworks''
+verwirklichen. Neue Browserstandards ermöglichen eine native
+Servicearchitektur, die direkt auf der Browser-Engine aufsetzt. Diese
+Charakteristiken sollen systematisch erfasst werden. \ldots{}
 
 \hypertarget{refs}{}
-\hypertarget{ref-Katz2013}{}
-{[}1{]} Y. Katz, ``Extend the web forward,'' 2013.
+\hypertarget{ref-Abramov2015}{}
+{[}1{]} D. Abramov, ``Presentational and Container Components --
+Medium.'' 2015 {[}Online{]}. Available:
+\url{https://medium.com/@dan_abramov/smart-and-dumb-components-7ca2f9a7c7d0}.
+{[}Accessed: 01-Dec-2016{]}
 
-\hypertarget{ref-Walton2016}{}
-{[}2{]} P. Walton, ``Houdini: Maybe the most exciting development in css
-you've never heard of,'' 2016.
+\hypertarget{ref-Newman2015}{}
+{[}2{]} S. Newman, \emph{Building microservices}. O'Reilly Media, Inc.,
+2016 {[}Online{]}. Available:
+\url{http://www.ebook.de/de/product/22539693/sam_newmann_building_microservices.html}
 
-\hypertarget{ref-Berner2016}{}
-{[}3{]} D. Berner, ``Not an imposter: Fighting front-end fatigue,''
-2016.
+\hypertarget{ref-Martin}{}
+{[}3{]} R. C. Martin, ``The single responsibility principle.''
+{[}Online{]}. Available:
+\url{http://programmer.97things.oreilly.com/wiki/index.php/The_Single_Responsibility_Principle}
 
-\hypertarget{ref-Baldwin2006}{}
-{[}4{]} C. Y. Baldwin and K. B. Clark, ``Modularity in the design of
-complex engineering systems,'' in \emph{Complex engineered systems:
-Science meets technology}, D. Braha, A. A. Minai, and Y. Bar-Yam, Eds.
-Berlin, Heidelberg: Springer Berlin Heidelberg, 2006, pp. 175--205.
+\hypertarget{ref-Conway1968}{}
+{[}4{]} M. E. Conway, ``How do committees invent?'' 1968 {[}Online{]}.
+Available: \url{http://www.melconway.com/Home/Committees_Paper.html}
 
-\hypertarget{ref-Keith2014}{}
-{[}5{]} J. Keith, ``Web components,'' 2014.
+\hypertarget{ref-Fowler2015}{}
+{[}5{]} M. Fowler and J. Lewis, ``Microservices: Nur ein weiteres
+konzept in der softwarearchitektur oder mehr?'' \emph{OBJEKTspektrum},
+Jan. 2015.
+
+\hypertarget{ref-Issa2016}{}
+{[}6{]} B. Issa, ``The way of the web.'' Polymer Summit 2016, Oct-2016
+{[}Online{]}. Available:
+\url{https://www.youtube.com/watch?v=8ZTFEhPBJEE}
+
+\end{document}

### 2017-01-04 paper.tex
pdf/tex build

@@
-\documentclass[ngerman,]{assets/latex/ieee}
+\documentclass[]{assets/latex/ieee}
 \usepackage{lmodern}
@@
 \urlstyle{same}  % don't use monospace font for urls
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[shorthands=off,main=ngerman]{babel}
-\else
-  \usepackage{polyglossia}
-  \setmainlanguage[]{german}
-\fi
 \IfFileExists{parskip.sty}{%
@@
 \providecommand{\subtitle}[1]{}
-\subtitle{Modulare Webarchitekturen durch neue Browserstandards entwickeln}
+\subtitle{Modular web architecture through new W3C specifications}
 \author{Jan Peteler, FH Würzburg-Schweinfurt, jan.peteler@student.fhws.de}
@@
 \maketitle
+\begin{abstract}
+Building complex web applications nowadays require additional layers of
+abstraction and often heavily depend on proprietary frameworks. New
+specifications build right into the browserengine provide a native
+service API to overcome tricky abstraction constraints.
+\end{abstract}
 
-\section{Ausgangssituation}\label{ausgangssituation}
+\section{Introduction}\label{introduction}
 
-Softwareentwicklung für den Browser ist ohne Zweifel eine komplexe und
-fehleranfällige Angelegenheit. Die Gründe für diesen Umstand sind so
-vielfältig, dass darüber eine ganze Forschungsarbeit geschrieben werden
-könnte. Je nach Blickwinkel könnte man von der technischen Seite
-argumentieren und die Designschwächen der Browserplattform
-erläutern.{[}1{]} Oder man könnte die organisatorische Seite betrachten
-und die Schwerfälligkeit von Standardisierungsprozessen
-anprangern.{[}2{]} Oder aber man nimmt die historische Entwicklung in
-Augenschein. Die Zeiten statischer HTML/CSS Seiten auf Desktopgeräten
-ist mithin noch nicht lange her.
+\begin{quote}
+Simplicity is prerequisite for reliability. - Edsger W. Dijkstra
+\end{quote}
 
-Dennoch tragen alle diese Facetten dazu bei, dass die Entwicklung und
-Wartung von komplexen Webapplikationen mit enormen Zeit- und Geldaufwand
-verbunden ist. Offensichtlich wird das Problem, wenn man die schiere
-Anzahl der Frameworks und Bibliotheken mit Javascript als Zielsprache
-betrachtet, die vielen Entwicklern Schwierigkeiten bereitet.{[}3{]}
+Introduction: from simplicity to Microservices to w3c specifications
+\ldots{} and beyond
 
-Um die inhärenten Designschwächen von Browsern aus der Welt zu schaffen,
-wurde schon im Jahr 2013 das \emph{Extensible Web Manifesto}
-proklamiert.\footnote{https://extensiblewebmanifesto.org/} Darin wird,
-kurz gefasst, eine Öffnung der Webplattformen avisiert, um
-Webprogrammierung teilweise von Browserstandards zu entkoppeln. Drei
-Jahre später sind diese Ziele in W3C Standards konkretisiert worden und
-teilweise schon umgesetzt worden.
+Der Erfinder der Programmiersprache Clojure, Rich Hickey, ist ohne
+Zweifel eine Koryphäe auf seinem Gebiet, der Strukturierung von
+komplexen Systemen. In einer vielbeachteten Keynote aus dem Jahr 2012
+geht er auf etymologisch-philosophische Spurensuche nach dem Wort
+\textbf{simplicity } aus Sicht eines Softwareentwicklers.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=48s}{Rails
+  Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} Das Adjektiv
+\emph{simple} hat demnach seinen Ursprung im lateinischen Wort
+\emph{simplex}, was soviel wie \emph{einfach} oder \emph{einzeln}
+bedeutet. In Gegensatz dazu stehen Eigenschaften wie \emph{complex} oder
+\emph{multiplex}. Qualitative Software ist, so Hickey, vor allem simpel
+- im Prozess, im Design, in der Struktur und in der Entwicklung.
 
-\section{Zielsetzung}\label{zielsetzung}
+Bezeichnend für diese Keynote ist, dass sie von Rich Hickey im Rahmen
+einer Webentwicklerkonferenz gehalten wurde. Software für den Browser
+war (und ist) seit langer Zeit maßgeblich geprägt von
+\textbf{Komplexität} auf mehreren Ebenen. Im Laufe dieser Bachelorarbeit
+werden diese Stru
 
-Viele populäre Frameworks für den Browser als Zielplattform versprechen
-dem Entwickler eine bessere Kontrolle über seinen Webservice. Die
-Notwendigkeit rührt unter anderem daher, dass es bisher nicht möglich
-war einzelne Browserelemente in ihrem Aussehen und Verhalten zu
-isolieren. Selbst kleinste Änderungen können daher die Funktionalität
-des Gesamtsystems beeinträchtigen. Mit der neuen W3C Spezifikation
-\emph{Web Components},\footnote{https://www.w3.org/standards/techs/components}
-die mehrere Substandards unter sich vereint, soll die Modularität nun
-auch nativ im Browser unterstützt werden. Modularität als Designpattern
-ist in der IT schon länger Standard um komplexe Systeme beherrschbar zu
-gestalten und auch zukünftige Unsicherheiten abzuwägen.{[}4, S. 1{]} Die
-Unsicherheit, die Frameworks immer anhaftet, ist die Lebensdauer
-derselben.
+\begin{itemize}
+\tightlist
+\item
+  Das \emph{Document Object Model} als Rückgrad jeder Webapplikation ist
+  hierarchisch strukturiert und deren Elemente damit keinesfalls
+  unabhängig in ihrer Darstellung und Reihenfolge.
+\item
+  JavaScript als single-threaded Skriptsprache lässt sich schlecht ihn
+  ihrem Verhalten isolieren, ist Fehleranfällig und hatte lange Zeit nur
+  sehr wenig idiomatische Lösungsansätze für komplexe Probleme, wie
+  beispielsweise Asynchronität
+\item
+  CSS ist geprägt von stetigem Überschreiben vorher definierter Regeln
+  und verletzt damit Simplizität in ihrer Struktur, Design und dem
+  Entwicklungsprozesses auf bester Art und Weise.
+\end{itemize}
 
-Mit \emph{Web Components} ist der Entwickler in der Lage, eigene HTML
-Elemente zu definieren und diese per Templates zu ex- \& importieren.
-Darüber hinaus können sie JavaScript und CSS Funktionalität enthalten
-und diese vom Rest der Webseite abzukapseln. Ziel dieser Arbeit ist die
-systematische Erfassung und Risikoabwägung der neuen Technologien sowie
-eine praktische Erläuterung möglicher Architekturmodelle. Bisher gibt es
-keine einheitliche Bestpractice, wie denn die Komponenten umzusetzen
-sind, obwohl diese Technologien bereits länger durch Polyfills genutzt
-werden können. Manch ein Entwickler fürchtet schon eine Flut von
-schlecht gebauten Komponenten, die wiederum neue Probleme schaffen
-könnten.{[}5{]}
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
 
-\section{Vorgehensweise}\label{vorgehensweise}
+Betrachtet man andere populäre Webframeworks dieser Zeit, wie
+beispielsweise React oder Angular, lässt sich ziemlich schnell ein
+gemeinsames Designpattern ausmachen, wie Komponenten intern ihre
+Zuständen verwalten. Der Facebook Entwickler Dan Abramov hat dieses
+dichotome Pattern systematisch erfasst und in zwei Kategorien
+eingeteilt.
 
-Den Anfang dieser Arbeit soll eine Analyse der aktuellen Situation der
-Frontend Entwicklung aufzeigen. Dort soll auf die Problematik der
-bisherigen Architekturmodelle und die Probleme mit monolithischen
-Frameworks eingegangen werden. Außerdem soll dieser Teil vor Augen
-führen, warum es bisher nur mit zusätzlicher Abstraktionsebenen möglich
-war einen modularen Aufbau von Web Applikationen zu ermöglichen.
+Nach Abramov existieren zum einen Komponenten, die alleine für das
+\textbf{Darstellen von Information} zuständig sind. Diese Komponenten
+nehmen keinerlei Einfluss auf Informationen oder anderen Komponenten um
+sich herum. Sie sind im wahrsten Sinne passiv und fremdgesteuert über
+klar definierte Schnittstellen. Diese Komponenten sind häufig flexibel
+einsetzbar und hochgradig wiederwerwendbar. Abramov nennt diese
+Komponenten ``Presentational Components''.{[}1{]} Verortet man diese Art
+von Komponenten innerhalb des \emph{MVC Pattern}, sind die Komponenten
+reine \emph{View-Elemente}.
 
-Im nächsten Abschnitt der Arbeit sollen die neuen Standards
-aufgeschlüsselt, zugänglich gemacht und auf Anwendungsmöglichkeiten
-untersucht werden. Explorativ sollen Bestpractices offengelegt werden
-und mögliche Einbettungen ins Gesamtsystem diskutiert werden. Auch die
-Probleme, wie sie beispielsweise bei alten Browsern auftreten können,
-sollen hier behandelt werden.
+Im Gegensatz dazu stehen Komponenten, die für die \textbf{Verarbeitung
+von Informationen} zuständig sind. Diese so genannten ``Container
+Components'' haben oft einen internen Zustand, den sie verändern können
+und an ihre Kindkomponenten weiterreichen können.{[}1{]} Im \emph{MVC
+Pattern} handelt es sich um die \emph{Controller}. In der funktionalen
+Programmiersprache Elm werden diese Elemente \emph{Updater} genannt, was
+ihre Funktion noch besser umschreibt.
 
-Der letzte Abschnitt soll das Thema mit einer Metaebene abrunden, in der
-auch die neuen CSS Standards \emph{Houdini} als Teil des Gesamtsystems
-betrachtet werden sollen.\footnote{https://drafts.css-houdini.org/}
+Ein üblicher eventgesteuerter Webservice setzt sich aus
+unterschiedlichsten Komponenten zusammen, die wiederum
+unterschiedlichste Eventlistener \& -emitter in sich subsummieren. Diese
+inhärente Komplexität verlangt geradezu nach einer klaren,
+deterministischen Struktur des Webservices, die das Zusammenspiel
+orchestriert. In der Analogie des Orchesters gesprochen, benötigt der
+Webservice (oder sogar die gesamte Webapplikation) einen Dirigenten, der
+für die Steuerung verantwortlich ist.
+
+\section{Microservices}\label{microservices}
+
+Opening up the case for \emph{Browsernative Microservices} brings up the
+question about the concept microservices in general. In fact the concept
+of microservices has many facets, stretching beyond disciplines and
+technical boundaries. It lacks a formal standardization but there are
+certain ideas emergine from this pattern. As a primary source of truth
+this articles relies on the work of Sam Newman, who has written a
+comprehensive guide in \emph{Building Microservices}. The purpose of
+this section is to match those ideas against the manifestations of web
+components.
+
+In a nutshell a microservice is a small, autonomous service that works
+together with other services seamlessly.{[}2, p. 2{]} or with the words
+of Fowler and Lewis: ``It is an approach to developing a single
+application as a suite of small services, each running in its own
+process and communicating with lightweight mechanisms,\ldots{}''{[}3{]}
+Microservices incorporate many ideas, like \emph{domain-driven design}
+where we try to represent the real world in our code.{\textbf{???}} Or
+making use of \emph{continuous delivery} for pushing software rapidly
+through \emph{automated deployment} mechanisms in production.{[}3{]}
+And, last but not least, microservices utilizes the idea of small teams
+with a lot of product knowledge working mostly autonomous on their very
+own service with their very own set of tools and techniques.
+
+\subsection{Technical perspective}\label{technical-perspective}
+
+Shift of paradigms / BFF / Platform agnostic / changes in infrastructure
+like APIs Databanks / Deployment
+
+Talking about microservices in a browsernative context isn't that far
+fetched as microservices themselves incorporate many ideas from the web.
+Exemplary, microservices often communicate via an HTTP request-response
+with resource API's and lightweight messaging.{[}3{]} Nevertheless,
+while (server-side) microservices offer wide ranges of technical
+possibilities, we must take into account that (client-side) browsers
+come with certain constraints and limitations.
+
+Fowler and Lewis issue a call for using services as components. A
+component~is regarded as a unit of software that is independently
+replaceable and upgradeable. The main advantage of a component in
+contrast of library is the possibility of an independent deployment. It
+aligns perfectly with the main goal of a microservice architecture to
+strip away most of the dependencies in favour of clean
+interfaces.{[}3{]}
+
+Components
+
+First of all, from a technical perspective, a microservice reinforces
+the \emph{Single Responsibility Principle} defined by Robert C. Martin:
+``Gather together those things that change for the same reason and
+separate those things that change for different reasons.''{[}4{]} An a
+way this principle tackles another often cited design principle of the
+\emph{seperation of concerns}. Web Components incorporate this principle
+in multiple ways while still remaining flexible.
+
+Most obvious ist the gathering of all related code under the umbrellar
+of a single HTML tag. Grouping together HTML, JS and CSS Code in a safe,
+sandboxed environment exposes the possibility to build more cohesive and
+understandable services. In the typical global nature of web development
+those three pillars are separated. This circumstance left the developer
+switching back and forth between code bases developing a tricky (and
+sometimes biased) way to glue related parts together.
+
+Secondly, the sub-standard \emph{custom elements} introduces so called
+lifecycle methods and a getter/setter interface exposing the
+functionality to the developer. Event handling, for example, can be
+registered in place which is much more declarative than assigning event
+listeners from the outside. Of course, this events can be pushed down to
+nested tags, allowing an increasingly granular system design. This
+approach will be explained further in the upcoming sections.
+
+The concept of microservices incooperates not only a technical
+perspective. Microservice patterns are a product of real-world
+usage.{[}2, p. 1{]} In a real world we typical have to deal with the so
+called \emph{Conway's Law}:
+
+\begin{quote}
+``organizations which design systems \ldots{} are constrained to produce
+designs which are copies of the communication~structures~of these
+organizations''. {[}5{]}
+\end{quote}
+
+Following this logic any company, whether it is web-related or not,
+should be devided in units grouped around a destinct business service to
+optimise the workflow. Fowler and Lewis outlines this approach as an
+``alignment of business capabilities''{[}3{]} While this kind of
+structure may be true for companies like Google or Amazon, there is a
+vast majority of companies developing for the web which are grouped
+around tasks.{[}6{]} A very common structure is formed by the technology
+stack (UX Designers, Frontend- \& Backend Developers) or by separating
+teams along the product lifecycle (development, testing, deployment).
+
+Advocators from the microservice approach propose a different model.
+best described by . Web components are one (but important) way to tie up
+those diciplines as one component can host a single independent business
+service. Combined with a flexible backend service these components can
+be huge gain over the cumbersome functional organizational approach.
+
+The ideas transcending from the microservice approach offers plenty of
+choices and decisions how to proceed with designing a program or to
+structure a process.
+
+Macroperspektive / Composition
+
+http://alistair.cockburn.us/Hexagonal+architecture
+
+MVC Pattern
+
+Pure frontend vs heavy backend
+
+\section{Progressive Enhancement}\label{progressive-enhancement}
+
+Chapter about progressive enhancement
 
 \hypertarget{refs}{}
-\hypertarget{ref-Katz2013}{}
-{[}1{]} Y. Katz, „Extend the Web Forward``, 2013 {[}Online{]}. Verfügbar
-unter: \url{http://yehudakatz.com/2013/05/21/extend-the-web-forward/}
+\hypertarget{ref-Abramov2015}{}
+{[}1{]} D. Abramov, ``Presentational and Container Components --
+Medium.'' 2015 {[}Online{]}. Available:
+\url{https://medium.com/@dan_abramov/smart-and-dumb-components-7ca2f9a7c7d0}.
+{[}Accessed: 01-Dec-2016{]}
 
-\hypertarget{ref-Walton2016}{}
-{[}2{]} P. Walton, „Houdini: Maybe The Most Exciting Development In CSS
-You've Never Heard Of``, 2016 {[}Online{]}. Verfügbar unter:
-\url{https://www.smashingmagazine.com/2016/03/houdini-maybe-the-most-exciting-development-in-css-youve-never-heard-of}
+\hypertarget{ref-Newman2015}{}
+{[}2{]} S. Newman, \emph{Building microservices}. O'Reilly Media, Inc.,
+2016 {[}Online{]}. Available:
+\url{http://www.ebook.de/de/product/22539693/sam_newmann_building_microservices.html}
 
-\hypertarget{ref-Berner2016}{}
-{[}3{]} D. Berner, „Not An Imposter: Fighting Front-End Fatigue``, 2016
-{[}Online{]}. Verfügbar unter:
-\url{https://www.smashingmagazine.com/2016/11/not-an-imposter-fighting-front-end-fatigue/}
+\hypertarget{ref-Fowler2014}{}
+{[}3{]} M. Fowler and J. Lewis, ``Microservices: A definition of this
+new architectural term,'' Jan. 2014 {[}Online{]}. Available:
+\url{http://www.martinfowler.com/articles/microservices.html}
 
-\hypertarget{ref-Baldwin2006}{}
-{[}4{]} C. Y. Baldwin und K. B. Clark, „Modularity in the Design of
-Complex Engineering Systems``, in \emph{Complex Engineered Systems:
-Science Meets Technology}, D. Braha, A. A. Minai, und Y. Bar-Yam, Hrsg.
-Berlin, Heidelberg: Springer Berlin Heidelberg, 2006, S. 175--205
-{[}Online{]}. Verfügbar unter:
-\url{http://dx.doi.org/10.1007/3-540-32834-3_9}
+\hypertarget{ref-Martin}{}
+{[}4{]} R. C. Martin, ``The single responsibility principle.''
+{[}Online{]}. Available:
+\url{http://programmer.97things.oreilly.com/wiki/index.php/The_Single_Responsibility_Principle}
 
-\hypertarget{ref-Keith2014}{}
-{[}5{]} J. Keith, „Web Components``, 2014 {[}Online{]}. Verfügbar unter:
-\url{https://adactio.com/journal/7431}
+\hypertarget{ref-Conway1968}{}
+{[}5{]} M. E. Conway, ``How do committees invent?'' 1968 {[}Online{]}.
+Available: \url{http://www.melconway.com/Home/Committees_Paper.html}
+
+\hypertarget{ref-Issa2016}{}
+{[}6{]} B. Issa, ``The way of the web.'' Polymer Summit 2016, Oct-2016
+{[}Online{]}. Available:
+\url{https://www.youtube.com/watch?v=8ZTFEhPBJEE}
 

### 2017-01-04 paper.tex
pdf/tex build

@@
 Microservices incorporate many ideas, like \emph{domain-driven design}
-where we try to represent the real world in our code.{\textbf{???}} Or
+where we try to represent the real world in our code.{[}2, p. 2{]} Or
 making use of \emph{continuous delivery} for pushing software rapidly

### 2017-01-05 paper.tex
pdf/tex build

@@
 \urlstyle{same}  % don't use monospace font for urls
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\newenvironment{Shaded}{}{}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{{#1}}}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{{#1}}}
+\newcommand{\ImportTok}[1]{{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{{#1}}}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{{#1}}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{{#1}}}
+\newcommand{\BuiltInTok}[1]{{#1}}
+\newcommand{\ExtensionTok}[1]{{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{{#1}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{{#1}}}
+\newcommand{\RegionMarkerTok}[1]{{#1}}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
+\newcommand{\NormalTok}[1]{{#1}}
+\usepackage[normalem]{ulem}
+% avoid problems with \sout in headers with hyperref:
+\pdfstringdefDisableCommands{\renewcommand{\sout}{}}
 \IfFileExists{parskip.sty}{%
@@
 
-Introduction: from simplicity to Microservices to w3c specifications
-\ldots{} and beyond
+Introduction: from simplicity to Microservices to extensible web
+manifesto to w3c specifications \ldots{} and beyond
 
@@
 
+\section{W3C specifications}\label{w3c-specifications}
+
+For building a native microservice running on the ``bare-metal'' browser
+engine requires a bunch of new specifications and assumptions. Starting
+right into it, the first and foremost the quasi specification \emph{Web
+Components} is needed. \emph{Web Components} is not a real standard.
+It's an amalgam of APIs from multiple w3c specs which can be used
+independently, too. Some people argue for only two specs{[}7{]}, some
+people go for three specs {[}8{]}, but the majority advocating the four
+specs variant, which can looked up on \sout{the official}
+\href{http://webcomponents.org}{webcomponents.org} website. For the
+purpose of this article, those four specs will be discussed briefly to
+provide a rough understanding. It is not meant to cover all bits and
+pieces.
+
+\subsection{\texorpdfstring{Custom Elements
+\href{http://w3c.github.io/webcomponents/spec/custom/}{(w3c)}}{Custom Elements (w3c)}}\label{custom-elements-w3c}
+
+\emph{Custom Elements} are the fundamental building block for
+browsernative microservices. They provide a way to create customized
+HTML tags. An obligatory \textbf{HelloWorld} will help to understand the
+spec quickly:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
+\KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+    \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
+        \KeywordTok{super}\NormalTok{()}\OperatorTok{;} \CommentTok{// mandatory!}
+        \KeywordTok{this}\NormalTok{.}\AttributeTok{innerHTML} \OperatorTok{=} \StringTok{"hello world"}\OperatorTok{;}
+        \KeywordTok{this}\NormalTok{.}\AttributeTok{onclick} \OperatorTok{=} \NormalTok{e }\OperatorTok{=>} \AttributeTok{alert}\NormalTok{(}\KeywordTok{this}\NormalTok{.}\AttributeTok{innerHTML}\NormalTok{)}\OperatorTok{;}
+    \OperatorTok{\}}
+\OperatorTok{\}}
+\VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> index.html}
+\KeywordTok{<hello-world></hello-world>}
+\end{Highlighting}
+\end{Shaded}
+
+Most obvious, this spec relies on the new \emph{ES6 Class Syntax} in
+favor of the original prototype-based inheritance model.
+``Extending~\texttt{HTMLElement}~ensures the custom element inherits the
+entire DOM API and means any properties/methods that you add to the
+class become part of the element's DOM interface.''{[}9{]} The keyword
+\texttt{this} points to the element itself and let us easily define and
+access properties of the element. The so called \emph{fat-arrow}
+(\texttt{=\textgreater{}}) is just a new feature of ES6 and nothing more
+than a \texttt{function()}.
+
+After defining the element it needs to be registered in the new global
+build-in \texttt{customElements} with an DOM alias like
+\texttt{hello-world}. Note that there must be a dash inside the name to
+be valid. Finally, the new element can go live inside the HTML Document
+\texttt{index.html}.
+
 Macroperspektive / Composition
@@
 
+\hypertarget{ref-Buchner2016}{}
+{[}7{]} D. Buchner, ``Demythstifying web components,'' 2016
+{[}Online{]}. Available:
+\url{http://www.backalleycoder.com/2016/08/26/demythstifying-web-components/}
+
+\hypertarget{ref-vanKesteren2014}{}
+{[}8{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
+{[}Online{]}. Available:
+\url{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}
+
+\hypertarget{ref-Bidelman2016}{}
+{[}9{]} E. Bidelman, ``Custom elements v1: reusable web components.''
+2016 {[}Online{]}. Available:
+\url{https://developers.google.com/web/fundamentals/primers/customelements/}.
+{[}Accessed: 01-Dec-2016{]}
+
 \end{document}

### 2017-01-11 paper.tex
pdf/tex build

@@
 For building a native microservice running on the ``bare-metal'' browser
-engine requires a bunch of new specifications and assumptions. Starting
-right into it, the first and foremost the quasi specification \emph{Web
-Components} is needed. \emph{Web Components} is not a real standard.
-It's an amalgam of APIs from multiple w3c specs which can be used
-independently, too. Some people argue for only two specs{[}7{]}, some
-people go for three specs {[}8{]}, but the majority advocating the four
-specs variant, which can looked up on \sout{the official}
+engine requires a bunch of new specifications and assumptions. Most
+importantly the quasi specification \textbf{Web Components} is needed.
+\emph{Web Components} is not a real standard. It's an amalgam of APIs
+from multiple w3c specs which can be used independently, too. A
+webdeveloper may choose one spec and embrace the freedom in architecture
+which can be combined with other frameworks/libraries.
+
+Depending on the context, some people argue for only two specs which
+essentially make it possible to create a scoped component but not caring
+too much on it's distribution{[}7{]}. Some people prefer the three specs
+{[}8{]}, but the majority advocating the four specs variant, which is
+listed on \sout{the official}
 \href{http://webcomponents.org}{webcomponents.org} website. For the
-purpose of this article, those four specs will be discussed briefly to
+purpose of this article, all four specs will be discussed briefly to
 provide a rough understanding. It is not meant to cover all bits and
@@
 
-\emph{Custom Elements} are the fundamental building block for
-browsernative microservices. They provide a way to create customized
-HTML tags. An obligatory \textbf{HelloWorld} will help to understand the
-spec quickly:
+\emph{Custom elements} are the fundamental building blocks for web
+components. Essentially, they provide a way to create custom HTML tags
+enriched with behavior, design and functionality. An obligatory
+\textbf{HelloWorld} will help to grasp the spec:
 
@@
     \KeywordTok{super}\NormalTok{()}\OperatorTok{;} \CommentTok{// mandatory!}
-        \KeywordTok{this}\NormalTok{.}\AttributeTok{innerHTML} \OperatorTok{=} \StringTok{"hello world"}\OperatorTok{;}
-        \KeywordTok{this}\NormalTok{.}\AttributeTok{onclick} \OperatorTok{=} \NormalTok{e }\OperatorTok{=>} \AttributeTok{alert}\NormalTok{(}\KeywordTok{this}\NormalTok{.}\AttributeTok{innerHTML}\NormalTok{)}\OperatorTok{;}
+    \KeywordTok{this}\NormalTok{.}\AttributeTok{onclick} \OperatorTok{=} \NormalTok{e }\OperatorTok{=>} \AttributeTok{alert}\NormalTok{(}\StringTok{"hello"}\NormalTok{)}\OperatorTok{;}
+    \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(...)}\OperatorTok{;}
   \OperatorTok{\}}
@@
 \NormalTok{> index.html}
-\KeywordTok{<hello-world></hello-world>}
+\KeywordTok{<hello-world>}\NormalTok{say hello}\KeywordTok{</hello-world>}
 \end{Highlighting}
@@
 entire DOM API and means any properties/methods that you add to the
-class become part of the element's DOM interface.''{[}9{]} The keyword
-\texttt{this} points to the element itself and let us easily define and
-access properties of the element. The so called \emph{fat-arrow}
+class become part of the element's DOM interface.''{[}9{]} Like any
+other \emph{ES6 class}, \emph{custom element}s can be sub-classed
+further on with the typical \texttt{extends} inheritance.
+
+The beauty of \emph{custom elements} comes with the keyword
+\texttt{this} which points to the element itself. Instead of querying
+and assigning behavior AFTER creation of the node in question
+\texttt{this} functionality allows a \textbf{declarative programming
+style}. Assigning functionality happens right in place BEFORE creation
+or insertion of the DOM. The so called \emph{fat-arrow}
 (\texttt{=\textgreater{}}) is just a new feature of ES6 and nothing more
-than a \texttt{function()}.
+than an anonymous \texttt{function()}.
 
-After defining the element it needs to be registered in the new global
-build-in \texttt{customElements} with an DOM alias like
-\texttt{hello-world}. Note that there must be a dash inside the name to
-be valid. Finally, the new element can go live inside the HTML Document
-\texttt{index.html}.
+After definition, the element needs to be registered in the new global
+build-in \texttt{customElements} with an tag name like
+\texttt{\textless{}hello-world\textgreater{}}. Mind the dash inside the
+tag name to conform the spec. Finally, the new element can go live
+inside the HTML Document \texttt{index.html}.
+
+\subsubsection{Lifecycle methods}\label{lifecycle-methods}
+
+In addition to the \texttt{constructor()}, the spec defines so called
+\emph{lifecycle callbacks} for controlling the \textbf{behaviour in the
+DOM}. Many popular frameworks like ReactJS or AngularJS rely on similar
+approaches:
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{connectedCallback()}\\
+  Called upon the time of \emph{connecting or upgrading the node} which
+  means the moment the node is inserted and rendered inside the DOM.
+  Typically this block of code contains setup code, such as fetching
+  resources or rendering.{[}9{]} For performance issues it's highly
+  preferable to put much code in here.
+\item
+  \texttt{disconnectedCallback()}\\
+  Called upon the time of \emph{node removal}. Cleanup code like
+  removing eventListeners or disconnecting websockets can be put here.
+\item
+  \texttt{attributeChangedCallback(attrName,\ oldVal,\ newVal)}\\
+  This method provides an \emph{Onchange handler} that runs for certain
+  attributes called with three values as defined in the signature. It is
+  meant to control an elements' transition from on \texttt{oldVal} to a
+  \texttt{newVal}. Due to performance issues, this callback is only
+  triggered for attributes registered in an \emph{observedAttributes}
+  array.
+\item
+  \texttt{adoptedCallback()}\\
+  Called when moving the node \emph{between documents}.
+\end{itemize}
+
+\subsubsection{Custom attributes}\label{custom-attributes}
+
+As previously mentioned, the custom elements must \texttt{extend} the
+\texttt{HTMLElement}. Consequently, the new element inherits properties
+and methods from it (and it's parent \texttt{Element}) and things like
+\texttt{id,\ class,\ addEventListner,\ ...} run out-of-the-box.
+Additionally, it is possible to define custom attributes using the
+\emph{custom elements'} \textbf{getter / setter interface} to steer the
+behavior of the element.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
+\KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+  \NormalTok{...}
+  \NormalTok{set }\AttributeTok{sayhello}\NormalTok{(val) }\OperatorTok{\{}
+    \KeywordTok{this}\NormalTok{.}\AttributeTok{_hello} \OperatorTok{=} \NormalTok{val}\OperatorTok{;}
+    \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\KeywordTok{this}\NormalTok{.}\AttributeTok{_hello}\NormalTok{)}\OperatorTok{;}
+  \OperatorTok{\}}
+  \NormalTok{get }\AttributeTok{sayhello}\NormalTok{() }\OperatorTok{\{}
+    \ControlFlowTok{return} \KeywordTok{this}\NormalTok{.}\AttributeTok{_hello}\OperatorTok{;}
+  \OperatorTok{\}}
+\OperatorTok{\}}
+\VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}\OperatorTok{;}
+\KeywordTok{var} \NormalTok{el }\OperatorTok{=} \KeywordTok{new} \AttributeTok{HelloWorld}\NormalTok{()}\OperatorTok{;}
+\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello} \OperatorTok{=} \StringTok{"earth"}\OperatorTok{;}
+\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello}\OperatorTok{;}\CommentTok{//"earth"}
+\end{Highlighting}
+\end{Shaded}
+
+While getters and setters work great in the JS world they fail crossing
+the boundaries to the corresponding HTML node. Declaring
+\texttt{\textless{}hello-world\ sayhello="mars"\textgreater{}\textless{}/hello-world\textgreater{}}
+would't work in the previous setup. A common workaround is archived by
+using the previous mentioned \texttt{attributeChangedCallback} lifecycle
+method to \textbf{reflect changing HTML attributes to JS} and/or map JS
+attributes to HTML with \texttt{this.setAttributes(...)} respectively.
+On \textbf{insertion time} html attributes might raise their hand with
+\texttt{this.hasAttributes(...)} and \texttt{this.getAttributes(...)}.
+Native DOM properties will reflect their values between HTML and JS
+automatically.{[}10, Para. 2.6.1{]}
+
+Concluding this section, a reader might already discover the
+\textbf{mental model} behind \emph{web compontents}. A custom element is
+similar to a named function where attributes treated as \textbf{input
+variables}. In the hierarchical nature of DOM, input can occur either
+top-down via assignments and bottom-up via captured events. The same
+goes true when talking about output. Even though it seems obvious, it
+might be helpful to keep this point in mind.
+
+\subsubsection{Customized build-in
+elements}\label{customized-build-in-elements}
+
+One aspect didn't mentioned yet is the possibility of creating
+sub-classes of \textbf{build-in elements} by extending the native
+Interfaces like the \texttt{HTMLButtonElement} interface. While this
+functionality is perfectly spec'd it is strongly rejected by some
+browser vendors.\footnote{Discussion on the topic:
+  https://github.com/w3c/webcomponents/issues/509} Most likely the spec
+will change in future in one or other way on this issue and therefore
+\textbf{customized build-in elements} left out of this paper
+intentionally.
+
+\subsection{\texorpdfstring{Shadow DOM
+\href{http://w3c.github.io/webcomponents/spec/shadow/}{(w3c)}}{Shadow DOM (w3c)}}\label{shadow-dom-w3c}
+
+The second most important concept of \emph{web components} rewards to
+the \emph{shadow DOM} spec. The concept isn't the easiest to wrap ones
+head around. Basically it's an isolated DOM tree living inside an
+hosting DOM tree. The spec referencing the hosting tree as \textbf{light
+tree} and the attached DOM as \textbf{shadow tree}. Conceptually, the
+Shadow DOM issues a single topic: \textbf{Style Encapsulation}.
+
+With an ever increasing complexity of an single-page application, the
+global nature of the DOM creates a challenging situation for the web
+developer and often ends up in highly fragmented bits of CSS and obscure
+CSS selectors. Of course, this situation lowers code clarity and
+reusability dramatically. The only solution which won't break with the
+existing global paradigm effectively is to allow separate pieces of
+encapsulated code sit on top of the global DOM - introducing the
+shadowed DOM approach!
+
+Enhancing the previous example the new encapsulated \texttt{HelloWorld}
+would like this:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
+\KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
+    \KeywordTok{super}\NormalTok{()}\OperatorTok{;} \CommentTok{// mandatory!}
+    \KeywordTok{const} \NormalTok{shadowRoot }\OperatorTok{=}
+          \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+    \VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{innerHTML} \OperatorTok{=} \StringTok{"hello"}\OperatorTok{;}
+    \KeywordTok{this}\NormalTok{.}\AttributeTok{onclick} \OperatorTok{=} \NormalTok{e }\OperatorTok{=>} \AttributeTok{alert}\NormalTok{(}\StringTok{"hello"}\NormalTok{)}\OperatorTok{;}
+  \OperatorTok{\}}
+\OperatorTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+The new global method \texttt{attachShadow} adds a new document root to
+the \texttt{HelloWorld} which has the same properties as a normal DOM.
+Therefore, it's possible to call the \texttt{innerHTML} method to fill
+the new document (fragment) with some content. Note that shadowRoot is
+attached as \textbf{open} which ensures that some events can bubble out
+and outside JS can reach in the new root. Nested children and other
+content in the light DOM are ``shadowed'' by the new root and must be
+invited in by so called \texttt{slots}.
+
+\subsubsection{Slots}\label{slots}
 
 Macroperspektive / Composition
+
+\subsubsection{Assumptions about custom
+elements}\label{assumptions-about-custom-elements}
+
+Apart from the spec'd perspective there is mental model a webdeveloper
+might
+
+Creating and using webcomponents might require a new mental model how do
+design a
+
+containers
 
@@
 
+\hypertarget{ref-HTML}{}
+{[}10{]} \emph{HTML living standard --- last updated 11 january 2017}.
+{[}Online{]}. Available: \url{https://html.spec.whatwg.org/multipage/}
+
 \end{document}

### 2017-01-16 paper.tex
pdf/tex build

@@
 \KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
-  \NormalTok{...}
+  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}\NormalTok{...}\OperatorTok{\}}
   \NormalTok{set }\AttributeTok{sayhello}\NormalTok{(val) }\OperatorTok{\{}
@@
 The second most important concept of \emph{web components} rewards to
-the \emph{shadow DOM} spec. The concept isn't the easiest to wrap ones
-head around. Basically it's an isolated DOM tree living inside an
-hosting DOM tree. The spec referencing the hosting tree as \textbf{light
-tree} and the attached DOM as \textbf{shadow tree}. Conceptually, the
-Shadow DOM issues a single topic: \textbf{Style Encapsulation}.
+the \emph{shadow DOM} spec. In terms of complexity this spec outpacing
+all others by far. A \emph{shadow DOM} is basically an isolated DOM tree
+living inside an another (hosting) DOM tree. The spec refers the hosting
+tree as \emph{light DOM tree} and the attached DOM as \emph{shadow DOM
+tree}. Conceptually, the \emph{shadow DOM} issues a single important
+topic in software development: \textbf{Encapsulation}. While the first
+spec \emph{custom elements} provides a sufficient way to encapsulate JS
+behavior, \emph{shadow DOM} coined strongly to in the direction of style
+encapsulation.
 
 With an ever increasing complexity of an single-page application, the
-global nature of the DOM creates a challenging situation for the web
-developer and often ends up in highly fragmented bits of CSS and obscure
-CSS selectors. Of course, this situation lowers code clarity and
-reusability dramatically. The only solution which won't break with the
-existing global paradigm effectively is to allow separate pieces of
-encapsulated code sit on top of the global DOM - introducing the
-shadowed DOM approach!
+global nature of the DOM creates a daunting situation for code
+organization and leads over times to highly fragmented bits of CSS and
+obscure CSS selectors or html wrappers. Of course, this situation lowers
+code clarity and reusability dramatically. The only solution which won't
+break with the existing global paradigm effectively is to allow separate
+pieces of encapsulated code sit on top of the global DOM - introducing
+the shadowed DOM approach!
 
@@
   \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
-    \KeywordTok{super}\NormalTok{()}\OperatorTok{;} \CommentTok{// mandatory!}
+    \NormalTok{...}
     \KeywordTok{const} \NormalTok{shadowRoot }\OperatorTok{=}
           \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
-    \VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{innerHTML} \OperatorTok{=} \StringTok{"hello"}\OperatorTok{;}
-    \KeywordTok{this}\NormalTok{.}\AttributeTok{onclick} \OperatorTok{=} \NormalTok{e }\OperatorTok{=>} \AttributeTok{alert}\NormalTok{(}\StringTok{"hello"}\NormalTok{)}\OperatorTok{;}
+    \VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{innerHTML} \OperatorTok{=} \StringTok{'<p>hello</p>'}\OperatorTok{;}
   \OperatorTok{\}}
@@
 the \texttt{HelloWorld} which has the same properties as a normal DOM.
-Therefore, it's possible to call the \texttt{innerHTML} method to fill
-the new document (fragment) with some content. Note that shadowRoot is
-attached as \textbf{open} which ensures that some events can bubble out
-and outside JS can reach in the new root. Nested children and other
-content in the light DOM are ``shadowed'' by the new root and must be
-invited in by so called \texttt{slots}.
+Therefore, invoking \texttt{innerHTML} method would fill the new
+document (fragment) with some arbitrary content. Note that
+\texttt{shadowRoot} is marked as \textbf{open} which ensures that some
+events can bubble out and outside JS can reach in the new root. Nested
+children nodes and other content in the light DOM are ``shadowed'' by
+the new root and must be invited in by so called \texttt{slots}.
 
 \subsubsection{Slots}\label{slots}
+
+Contradicting to the simplified \texttt{HelloWorld} example, a
+\emph{shadow DOM} shouldn't contain any \sout{valuable} content. While
+technical possible any change of an element would require deeply nested
+calls from the \emph{light DOM} to the \emph{shadow DOM} to update the
+element in place. That's why \emph{shadow DOM} should be perceived more
+as \textbf{static HTML template} and provide therefore a kind of
+internal frame for the render engine. \texttt{Slots} are placeholders
+for \emph{light DOM} nodes used to mark the endpoints in question.
+
+Technically, the \emph{light DOM} nodes are not moved inside the
+\emph{shadow DOM}. Their just rendered in place. It's an subtle but
+important difference towards handling a node. JS behaviour and CSS
+styles applied in the \emph{light DOM} will still be valid in the
+\emph{shadow DOM}. The render engine literally taking the nodes and
+putting them inside the \texttt{slot}. This procedure is commonly
+referred as \textbf{flattening} of the DOM trees.
+
+\paragraph{Named slots}\label{named-slots}
+
+A named slot is the preferable way for clear code organization. Taking
+for example
+\texttt{\textless{}slot\ name="hello"\textgreater{}Drop\ me\ a\ "hello"\ node\textless{}/slot\textgreater{}}
+targets all direct \emph{light DOM} child nodes of the hosting node
+matching the slot name like
+\texttt{\textless{}div\ slot="hello"\textgreater{}\textless{}/div\textgreater{}}.
+Writing a little documentation inside the
+\texttt{\textless{}slot\textgreater{}} tag is considered as a good
+practice as it will be rendered only if no matching \emph{light DOM}
+node is available. This functionality makes a \emph{custom element}
+pretty much self-explanatory.
+
+\paragraph{Unnamed slots}\label{unnamed-slots}
+
+Inside a so-called \emph{default slot} which looks like
+\texttt{\textless{}slot\textgreater{}Unnamed\ content\ goes\ here\textless{}/slot\textgreater{}},
+the render engine expands all direct \emph{light DOM} children without a
+\texttt{slot} attribute. In case of multiple default slots, the first
+slot takes it all.
+
+\subsubsection{Styling}\label{styling}
+
+As mentioned in the last section, there is a distinct difference about
+the nature of nodes. Nodes declared and rendered exclusively in the
+\emph{shadow DOM} are not affected by any styling from outside. Nodes
+which are declared outside and distributed via \texttt{slots} will be
+styled in the \emph{light DOM} and can be additionally painted in the
+\emph{shadow DOM} through the new CSS-Selector \texttt{::slotted()}.
+
+Note that styles from the outside have an higher specify than styles
+assigned after distribution. Therefore it is generally a good advice to
+minimize the global stylings to some base styling for uniformity of the
+web site while leaving the specific stylings to the component. Due to
+the cascading nature of CSS, styles will still ``bleed in'' from
+ancestors to the \emph{light DOM} nodes. Therefore it's strongly
+recommended to begin every \emph{shadow DOM} with a \textbf{\emph{CSS
+reset:} \texttt{:host\ \{all:\ initial;\}}}.
+
+Regarding the importance style encapsulation, a couple of new CSS rules
+emerged that are exclusively targeting the \emph{shadow DOM}. The table
+below outlines styling possibilities for the use INSIDE the \emph{shadow
+DOM}:
+
+{[}{[} CSS Selectors {]}{]}
+
+Using the \emph{functional selector} of \texttt{:host()} or even the
+only-functional \texttt{:host-contest()} allows the creation of
+\textbf{context-aware custom elements}. A possible use case would be
+``theming'' a component (example taken from {[}11{]}):
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> index.html}
+\KeywordTok{<body}\OtherTok{ class=}\StringTok{"darktheme"}\KeywordTok{>}
+  \KeywordTok{<fancy-tabs>}
+    \NormalTok{...}
+  \KeywordTok{</fancy-tabs>}
+\KeywordTok{</body>}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> fancy-tabs shadowRoot}
+\NormalTok{<style>}
+\DecValTok{:}\NormalTok{host-context(}\FloatTok{.darktheme}\NormalTok{) }\KeywordTok{\{}
+\ErrorTok{ } \KeywordTok{color:} \DataTypeTok{white}\KeywordTok{;}
+\ErrorTok{ } \KeywordTok{background:} \DataTypeTok{black}\KeywordTok{;}
+\KeywordTok{\}}
+\NormalTok{</style>}
+\end{Highlighting}
+\end{Shaded}
+
+\subsubsection{JS Behavior}\label{js-behavior}
+
+As mentioned earlier any logic applied to \emph{light DOM} nodes stays
+with the node even after redistribution. For the sake of separation of
+concerns the business logic should be part of the \emph{custom element}
+(the \emph{light DOM}) and not the part of the \emph{shadow DOM}. On the
+other hand there are numerous scenarios where JS is just concerned with
+\textbf{styling or animation of an element}.In this case it might be
+more straightforward to apply JS inside the \emph{shadow DOM} to avoid
+mixing with business logic handlers.
+
+Drilling down to a \emph{light DOM} node from an \emph{shadow DOM}
+context is not possible with querying the node directly with
+\texttt{.querySelector()} or \texttt{.getElementById()} as the node is
+not part of the context. To get a distributed node in question it needs
+the way over the slot node and call \texttt{slot.assinedNodes()} to
+receive an array of distributed node(s) which can be accessed and
+manipulated like any other node. Calling \texttt{.assignedNodes()} on an
+empty \texttt{slot} returns an empty array.
+
+Wrapping up this section, \emph{shadow DOM} provides a non-hacky way to
+create uniform looking \emph{custom elements} and even enhance styling
+possibilities without adding much overhead. Still, for smaller
+components with only one or two child nodes, just a little styling
+and/or no structured redistribution a \emph{shadow DOM} might be to hard
+to reason about. Eventually it all depends on the question of ``how hard
+is it to implement it without shadow DOM'' - which can't be answered
+universally. For a more in-depth guide, Google Engineer Eric Bidelman
+wroten a great primer on \emph{shadow DOM}{[}11{]}.
+
+So far, there is still a missing link between \emph{light DOM} and
+\emph{shadow DOM}. The observant reader may have already noticed the
+weak point in the \texttt{HelloWorld} example: how to ``vitalize'' the
+\emph{shadow DOM}. Recapturing the last \texttt{HelloWorld} example a
+string of markup was assigned to the \texttt{shadowRoot.innerHTML}
+property. While it works perfectly fine in this simple case, a string of
+markup is rather cumbersome and error-prone and doesn't scale well. When
+putting quotes inside another quotes things break quickly. It makes the
+life hard for developers to work with it because it requires manual
+indentation and is out of syntax highlighting. That's the time templates
+come into play.
+
+\subsection{\texorpdfstring{HTML Templates
+\href{https://www.w3.org/TR/html5/scripting-1.html\#the-template-element}{(w3c)}}{HTML Templates (w3c)}}\label{html-templates-w3c}
+
+Among all other new standards \emph{HTML templates} are the most mature
+and adopted standard in the browser environment. All major browsers,
+except from Internet Explorer, support this standard.
+
+One core concept in templates is efficiency: Whatever dropped inside a
+\texttt{template} tag \sout{bucket} will be parsed on runtime - but not
+constructed into the \emph{content tree}. It remains plain HTML Markup
+sitting somewhere in the document until the time of activation.
+
+Activation typically takes four steps:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  \textbf{Querying the template node in question}\\
+  \texttt{const\ t\ =\ document.querySelector(\textquotesingle{}\#helloworld\textquotesingle{});}
+\item
+  \textbf{Preparing the templates' content}\\
+  The templates' \texttt{content} property contains all nodes a
+  \emph{DocumentFragment} object. Handling should be straight forward\\
+  \texttt{t.content.querySelector(\textquotesingle{}img\textquotesingle{}).src\ =\ \textquotesingle{}logo.png\textquotesingle{};}
+\item
+  \textbf{Optional: Cloning the \emph{DocumentFragment} for multiple
+  use}\\
+  \texttt{const\ clone\ =\ t.content.cloneNode("deep");}
+\item
+  \textbf{Appending the clone/original to destination}\\
+  \texttt{document.body.appendChild(clone);}
+\end{enumerate}
+
+As easy and minimal \emph{HTML templates} are, they're missing out a
+crucial feature other template implementations usually have. As
+templates are basically just dump containers for HTML Markup, there is
+no way to define some logic as \textbf{placeholders} where content
+should appear. Of course, with heavy use of JS things could be modeled
+this way. But the idiomatic way tends more towards a \emph{Shadow DOM \&
+HTML Templates} symbiosis.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> hello-component.html}
+\KeywordTok{<hello-world>}
+  \KeywordTok{<p}\OtherTok{ id=}\StringTok{"sendto"}\OtherTok{ slot=}\StringTok{"placeholder"}\KeywordTok{>}
+    \NormalTok{Hello World    }
+  \KeywordTok{</p>}
+\KeywordTok{</hello-world>}
+
+\KeywordTok{<template}\OtherTok{ id=}\StringTok{"hello"}\KeywordTok{>}
+  \KeywordTok{<style>}
+    \FloatTok{#stylewrapper} \KeywordTok{\{}
+      \KeywordTok{font-weight:} \DataTypeTok{bold}\KeywordTok{;}
+      \KeywordTok{color:} \NormalTok{orange}\KeywordTok{;}
+    \KeywordTok{\}}
+  \KeywordTok{</style>}
+  \CommentTok{<!-- CONTENT -->}
+  \KeywordTok{<div}\OtherTok{ id=}\StringTok{"stylewrapper"}\KeywordTok{>}
+    \KeywordTok{<slot}\OtherTok{ name=}\StringTok{"placeholder"}\KeywordTok{>}
+      \NormalTok{Named placeholder}
+    \KeywordTok{</slot>}
+  \KeywordTok{</div>}
+\KeywordTok{</template>}
+
+\KeywordTok{<script>}
+  \KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+    \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
+      \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
+      \KeywordTok{const} \NormalTok{shadowRoot }\OperatorTok{=}
+        \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+      \KeywordTok{const} \NormalTok{template }\OperatorTok{=}
+        \VariableTok{document}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'#hello'}\NormalTok{).}\AttributeTok{content}\OperatorTok{;}
+
+      \KeywordTok{this}\NormalTok{.}\VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{appendChild}\NormalTok{(template)}\OperatorTok{;}
+    \OperatorTok{\}}
+  \OperatorTok{\}}
+  \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}\OperatorTok{;}
+\OperatorTok{<}\SpecialStringTok{/script>}
+\end{Highlighting}
+\end{Shaded}
+
+The updated \texttt{HelloWorld} example looks already pretty mature. It
+combines all the previous mentioned standards into one blob of HTML.
+\emph{Custom Elements} serves the logic, \emph{Shadow DOM} scopes the
+styles and \emph{HTML Template} efficiently glues DOM and \emph{Shadow
+DOM} together. The last standard in the row of four is not concerned
+with the internals of a \emph{web component}. \emph{HTML Imports} serves
+the need for an efficient distribution mechanism of components.
+
+\subsection{\texorpdfstring{HTML Imports
+\href{https://www.w3.org/TR/html-imports/}{(w3c)}}{HTML Imports (w3c)}}\label{html-imports-w3c}
 
@@
 
+\hypertarget{ref-Bidelman2016shadow}{}
+{[}11{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
+2016 {[}Online{]}. Available:
+\url{https://developers.google.com/web/fundamentals/getting-started/primers/shadowdom}
+
 \end{document}

### 2017-01-17 build/paper.tex
pdf/tex build

@@
 \hypersetup{
+            pdftitle={Browsernative Microservices},
+            pdfauthor={Jan Peteler, FH Würzburg-Schweinfurt, jan.peteler@student.fhws.de},
             pdfborder={0 0 0},
@@
 
-\date{}
+\title{Browsernative Microservices}
+\providecommand{\subtitle}[1]{}
+\subtitle{Modular web architecture through new W3C specifications}
+\author{Jan Peteler, FH Würzburg-Schweinfurt, jan.peteler@student.fhws.de}
+\date{Januar 2017}
 
 \begin{document}
+\maketitle
+\begin{abstract}
+Building complex web applications nowadays require additional layers of
+abstraction and often heavily depend on proprietary frameworks. New
+specifications build right into the browserengine provide a native
+service API to overcome tricky abstraction constraints.
+\end{abstract}
 
@@
 \end{quote}
+
+Introduction: from simplicity to Microservices to extensible web
+manifesto to w3c specifications \ldots{} and beyond
 
@@
 
-\section{Store}\label{store}
-
-\begin{verbatim}
-   +-----------------------+
-   |         Store         |
-   +-----------------------+
-       /\|  |
-
-\end{verbatim}
-
-die Änderungen des Zustands vorhersehbar aber auch zurückverfolgbar
-macht.
-
 \section{Microservices}\label{microservices}
 
-In a nutshell a microservice is a small, autonomous service that works
-together with other services seamlessly.{[}2, p. 2{]} While the concept
-of microservices lacks a formal standardization there are certain
-defacto standardized ideas emergine from this pattern. The purpose of
+Opening up the case for \emph{Browsernative Microservices} brings up the
+question about the concept microservices in general. In fact the concept
+of microservices has many facets, stretching beyond disciplines and
+technical boundaries. It lacks a formal standardization but there are
+certain ideas emergine from this pattern. As a primary source of truth
+this articles relies on the work of Sam Newman, who has written a
+comprehensive guide in \emph{Building Microservices}. The purpose of
 this section is to match those ideas against the manifestations of web
 components.
+
+In a nutshell a microservice is a small, autonomous service that works
+together with other services seamlessly.{[}2, p. 2{]} or with the words
+of Fowler and Lewis: ``It is an approach to developing a single
+application as a suite of small services, each running in its own
+process and communicating with lightweight mechanisms,\ldots{}''{[}3{]}
+Microservices incorporate many ideas, like \emph{domain-driven design}
+where we try to represent the real world in our code.{[}2, p. 2{]} Or
+making use of \emph{continuous delivery} for pushing software rapidly
+through \emph{automated deployment} mechanisms in production.{[}3{]}
+And, last but not least, microservices utilizes the idea of small teams
+with a lot of product knowledge working mostly autonomous on their very
+own service with their very own set of tools and techniques.
+
+\subsection{Technical perspective}\label{technical-perspective}
+
+Shift of paradigms / BFF / Platform agnostic / changes in infrastructure
+like APIs Databanks / Deployment
+
+Talking about microservices in a browsernative context isn't that far
+fetched as microservices themselves incorporate many ideas from the web.
+Exemplary, microservices often communicate via an HTTP request-response
+with resource API's and lightweight messaging.{[}3{]} Nevertheless,
+while (server-side) microservices offer wide ranges of technical
+possibilities, we must take into account that (client-side) browsers
+come with certain constraints and limitations.
+
+Fowler and Lewis issue a call for using services as components. A
+component~is regarded as a unit of software that is independently
+replaceable and upgradeable. The main advantage of a component in
+contrast of library is the possibility of an independent deployment. It
+aligns perfectly with the main goal of a microservice architecture to
+strip away most of the dependencies in favour of clean
+interfaces.{[}3{]}
+
+Components
 
@@
 ``Gather together those things that change for the same reason and
-separate those things that change for different reasons.''{[}3{]} An a
+separate those things that change for different reasons.''{[}4{]} An a
 way this principle tackles another often cited design principle of the
@@
 those three pillars are separated. This circumstance left the developer
-switching back and forth between code bases trying to find a tricky way
-to glue related parts together.
+switching back and forth between code bases developing a tricky (and
+sometimes biased) way to glue related parts together.
 
@@
 designs which are copies of the communication~structures~of these
-organizations''. {[}4{]}
+organizations''. {[}5{]}
 \end{quote}
 
-Following this logic any company, whether its web-related or not, should
-be devided in units grouped around a destinct business service to
+Following this logic any company, whether it is web-related or not,
+should be devided in units grouped around a destinct business service to
 optimise the workflow. Fowler and Lewis outlines this approach as an
-``alignment of business capabilities''{[}5, p. 2{]} While this kind of
+``alignment of business capabilities''{[}3{]} While this kind of
 structure may be true for companies like Google or Amazon, there is a
@@
 around tasks.{[}6{]} A very common structure is formed by the technology
-stack: UX Designers, Frontend- \& Backend Developers.
+stack (UX Designers, Frontend- \& Backend Developers) or by separating
+teams along the product lifecycle (development, testing, deployment).
 
-The microservice approach models all those functions into
-business-related units. Web components are one (but important) way to
-tie up those diciplines as one component can host a single independent
-business service. Combined with a flexible backend service these
-components can be huge gain over the cumbersome functional
-organizational approach.
+Advocators from the microservice approach propose a different model.
+best described by . Web components are one (but important) way to tie up
+those diciplines as one component can host a single independent business
+service. Combined with a flexible backend service these components can
+be huge gain over the cumbersome functional organizational approach.
+
+The ideas transcending from the microservice approach offers plenty of
+choices and decisions how to proceed with designing a program or to
+structure a process.
+
+\section{W3C specifications}\label{w3c-specifications}
+
+For building a native microservice running on the ``bare-metal'' browser
+engine requires a bunch of new specifications and assumptions. Most
+importantly the quasi specification \textbf{Web Components} is needed.
+\emph{Web Components} is not a real standard. It's an amalgam of APIs
+from multiple w3c specs which can be used independently, too. A
+webdeveloper may choose one spec and embrace the freedom in architecture
+which can be combined with other frameworks/libraries.
+
+Depending on the context, some people argue for only two specs which
+essentially make it possible to create a scoped component but not caring
+too much on it's distribution{[}7{]}. Some people prefer the three specs
+{[}8{]}, but the majority advocating the four specs variant, which is
+listed on \sout{the official}
+\href{http://webcomponents.org}{webcomponents.org} website. For the
+purpose of this article, all four specs will be discussed briefly to
+provide a rough understanding. It is not meant to cover all bits and
+pieces.
+
+\subsection{\texorpdfstring{Custom Elements
+\href{http://w3c.github.io/webcomponents/spec/custom/}{(w3c)}}{Custom Elements (w3c)}}\label{custom-elements-w3c}
+
+\emph{Custom elements} are the fundamental building blocks for web
+components. Essentially, they provide a way to create custom HTML tags
+enriched with behavior, design and functionality. An obligatory
+\textbf{HelloWorld} will help to grasp the spec:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
+\KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
+    \KeywordTok{super}\NormalTok{()}\OperatorTok{;} \CommentTok{// mandatory!}
+    \KeywordTok{this}\NormalTok{.}\AttributeTok{onclick} \OperatorTok{=} \NormalTok{e }\OperatorTok{=>} \AttributeTok{alert}\NormalTok{(}\StringTok{"hello"}\NormalTok{)}\OperatorTok{;}
+    \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(...)}\OperatorTok{;}
+  \OperatorTok{\}}
+\OperatorTok{\}}
+\VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> index.html}
+\KeywordTok{<hello-world>}\NormalTok{say hello}\KeywordTok{</hello-world>}
+\end{Highlighting}
+\end{Shaded}
+
+Most obvious, this spec relies on the new \emph{ES6 Class Syntax} in
+favor of the original prototype-based inheritance model.
+``Extending~\texttt{HTMLElement}~ensures the custom element inherits the
+entire DOM API and means any properties/methods that you add to the
+class become part of the element's DOM interface.''{[}9{]} Like any
+other \emph{ES6 class}, \emph{custom element}s can be sub-classed
+further on with the typical \texttt{extends} inheritance.
+
+The beauty of \emph{custom elements} comes with the keyword
+\texttt{this} which points to the element itself. Instead of querying
+and assigning behavior AFTER creation of the node in question
+\texttt{this} functionality allows a \textbf{declarative programming
+style}. Assigning functionality happens right in place BEFORE creation
+or insertion of the DOM. The so called \emph{fat-arrow}
+(\texttt{=\textgreater{}}) is just a new feature of ES6 and nothing more
+than an anonymous \texttt{function()}.
+
+After definition, the element needs to be registered in the new global
+build-in \texttt{customElements} with an tag name like
+\texttt{\textless{}hello-world\textgreater{}}. Mind the dash inside the
+tag name to conform the spec. Finally, the new element can go live
+inside the HTML Document \texttt{index.html}.
+
+\subsubsection{Lifecycle methods}\label{lifecycle-methods}
+
+In addition to the \texttt{constructor()}, the spec defines so called
+\emph{lifecycle callbacks} for controlling the \textbf{behaviour in the
+DOM}. Many popular frameworks like ReactJS or AngularJS rely on similar
+approaches:
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{connectedCallback()}\\
+  Called upon the time of \emph{connecting or upgrading the node} which
+  means the moment the node is inserted and rendered inside the DOM.
+  Typically this block of code contains setup code, such as fetching
+  resources or rendering.{[}9{]} For performance issues it's highly
+  preferable to put much code in here.
+\item
+  \texttt{disconnectedCallback()}\\
+  Called upon the time of \emph{node removal}. Cleanup code like
+  removing eventListeners or disconnecting websockets can be put here.
+\item
+  \texttt{attributeChangedCallback(attrName,\ oldVal,\ newVal)}\\
+  This method provides an \emph{Onchange handler} that runs for certain
+  attributes called with three values as defined in the signature. It is
+  meant to control an elements' transition from on \texttt{oldVal} to a
+  \texttt{newVal}. Due to performance issues, this callback is only
+  triggered for attributes registered in an \emph{observedAttributes}
+  array.
+\item
+  \texttt{adoptedCallback()}\\
+  Called when moving the node \emph{between documents}.
+\end{itemize}
+
+\subsubsection{Custom attributes}\label{custom-attributes}
+
+As previously mentioned, the custom elements must \texttt{extend} the
+\texttt{HTMLElement}. Consequently, the new element inherits properties
+and methods from it (and it's parent \texttt{Element}) and things like
+\texttt{id,\ class,\ addEventListner,\ ...} run out-of-the-box.
+Additionally, it is possible to define custom attributes using the
+\emph{custom elements'} \textbf{getter / setter interface} to steer the
+behavior of the element.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
+\KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}\NormalTok{...}\OperatorTok{\}}
+  \NormalTok{set }\AttributeTok{sayhello}\NormalTok{(val) }\OperatorTok{\{}
+    \KeywordTok{this}\NormalTok{.}\AttributeTok{_hello} \OperatorTok{=} \NormalTok{val}\OperatorTok{;}
+    \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\KeywordTok{this}\NormalTok{.}\AttributeTok{_hello}\NormalTok{)}\OperatorTok{;}
+  \OperatorTok{\}}
+  \NormalTok{get }\AttributeTok{sayhello}\NormalTok{() }\OperatorTok{\{}
+    \ControlFlowTok{return} \KeywordTok{this}\NormalTok{.}\AttributeTok{_hello}\OperatorTok{;}
+  \OperatorTok{\}}
+\OperatorTok{\}}
+\VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}\OperatorTok{;}
+\KeywordTok{var} \NormalTok{el }\OperatorTok{=} \KeywordTok{new} \AttributeTok{HelloWorld}\NormalTok{()}\OperatorTok{;}
+\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello} \OperatorTok{=} \StringTok{"earth"}\OperatorTok{;}
+\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello}\OperatorTok{;}\CommentTok{//"earth"}
+\end{Highlighting}
+\end{Shaded}
+
+While getters and setters work great in the JS world they fail crossing
+the boundaries to the corresponding HTML node. Declaring
+\texttt{\textless{}hello-world\ sayhello="mars"\textgreater{}\textless{}/hello-world\textgreater{}}
+would't work in the previous setup. A common workaround is archived by
+using the previous mentioned \texttt{attributeChangedCallback} lifecycle
+method to \textbf{reflect changing HTML attributes to JS} and/or map JS
+attributes to HTML with \texttt{this.setAttributes(...)} respectively.
+On \textbf{insertion time} html attributes might raise their hand with
+\texttt{this.hasAttributes(...)} and \texttt{this.getAttributes(...)}.
+Native DOM properties will reflect their values between HTML and JS
+automatically.{[}10, Para. 2.6.1{]}
+
+Concluding this section, a reader might already discover the
+\textbf{mental model} behind \emph{web compontents}. A custom element is
+similar to a named function where attributes treated as \textbf{input
+variables}. In the hierarchical nature of DOM, input can occur either
+top-down via assignments and bottom-up via captured events. The same
+goes true when talking about output. Even though it seems obvious, it
+might be helpful to keep this point in mind.
+
+\subsubsection{Customized build-in
+elements}\label{customized-build-in-elements}
+
+One aspect didn't mentioned yet is the possibility of creating
+sub-classes of \textbf{build-in elements} by extending the native
+Interfaces like the \texttt{HTMLButtonElement} interface. While this
+functionality is perfectly spec'd it is strongly rejected by some
+browser vendors.\footnote{https://github.com/w3c/webcomponents/issues/509}
+Most likely the spec will change in future in one or other way on this
+issue and therefore \textbf{customized build-in elements} left out of
+this paper intentionally.
+
+\subsection{\texorpdfstring{Shadow DOM
+\href{http://w3c.github.io/webcomponents/spec/shadow/}{(w3c)}}{Shadow DOM (w3c)}}\label{shadow-dom-w3c}
+
+The second most important concept of \emph{web components} rewards to
+the \emph{shadow DOM} spec. In terms of complexity this spec outpacing
+all others by far. A \emph{shadow DOM} is basically an isolated DOM tree
+living inside an another (hosting) DOM tree. The spec refers the hosting
+tree as \emph{light DOM tree} and the attached DOM as \emph{shadow DOM
+tree}. Conceptually, the \emph{shadow DOM} issues a single important
+topic in software development: \textbf{Encapsulation}. While the first
+spec \emph{custom elements} provides a sufficient way to encapsulate JS
+behavior, \emph{shadow DOM} coined strongly to in the direction of style
+encapsulation.
+
+With an ever increasing complexity of an single-page application, the
+global nature of the DOM creates a daunting situation for code
+organization and leads over times to highly fragmented bits of CSS and
+obscure CSS selectors or html wrappers. Of course, this situation lowers
+code clarity and reusability dramatically. The only solution which won't
+break with the existing global paradigm effectively is to allow separate
+pieces of encapsulated code sit on top of the global DOM - introducing
+the shadowed DOM approach!
+
+Enhancing the previous example the new encapsulated \texttt{HelloWorld}
+would like this:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
+\KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
+    \NormalTok{...}
+    \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+    \VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{innerHTML} \OperatorTok{=} \StringTok{'<p>hello</p>'}\OperatorTok{;}
+  \OperatorTok{\}}
+\OperatorTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+The new global method \texttt{attachShadow} adds a new document root to
+the \texttt{HelloWorld} which has the same properties as a normal DOM.
+Therefore, invoking \texttt{innerHTML} method would fill the new
+document (fragment) with some arbitrary content. Note that
+\texttt{shadowRoot} is marked as \textbf{open} which ensures that some
+events can bubble out and outside JS can reach in the new root. Nested
+children nodes and other content in the light DOM are ``shadowed'' by
+the new root and must be invited in by so called \texttt{slots}.
+
+\subsubsection{Slots}\label{slots}
+
+Contradicting to the simplified \texttt{HelloWorld} example, a
+\emph{shadow DOM} shouldn't contain any \sout{valuable} content. While
+technical possible any change of an element would require deeply nested
+calls from the \emph{light DOM} to the \emph{shadow DOM} to update the
+element in place. That's why \emph{shadow DOM} should be perceived more
+as \textbf{static HTML template} and provide therefore a kind of
+internal frame for the render engine. \texttt{Slots} are placeholders
+for \emph{light DOM} nodes used to mark the endpoints in question.
+
+Technically, the \emph{light DOM} nodes are not moved inside the
+\emph{shadow DOM}. Their just rendered in place. It's an subtle but
+important difference towards handling a node. JS behaviour and CSS
+styles applied in the \emph{light DOM} will still be valid in the
+\emph{shadow DOM}. The render engine literally taking the nodes and
+putting them inside the \texttt{slot}. This procedure is commonly
+referred as \textbf{flattening} of the DOM trees.
+
+\paragraph{Named slots}\label{named-slots}
+
+A named slot is the preferable way for clear code organization. Taking
+for example
+\texttt{\textless{}slot\ name="hello"\textgreater{}Drop\ me\ a\ "hello"\ node\textless{}/slot\textgreater{}}
+targets all direct \emph{light DOM} child nodes of the hosting node
+matching the slot name like
+\texttt{\textless{}div\ slot="hello"\textgreater{}\textless{}/div\textgreater{}}.
+Writing a little documentation inside the
+\texttt{\textless{}slot\textgreater{}} tag is considered as a good
+practice as it will be rendered only if no matching \emph{light DOM}
+node is available. This functionality makes a \emph{custom element}
+pretty much self-explanatory.
+
+\paragraph{Unnamed slots}\label{unnamed-slots}
+
+Inside a so-called \emph{default slot} which looks like
+\texttt{\textless{}slot\textgreater{}Unnamed\ content\ goes\ here\textless{}/slot\textgreater{}},
+the render engine expands all direct \emph{light DOM} children without a
+\texttt{slot} attribute. In case of multiple default slots, the first
+slot takes it all.
+
+\subsubsection{Styling}\label{styling}
+
+As mentioned in the last section, there is a distinct difference about
+the nature of nodes. Nodes declared and rendered exclusively in the
+\emph{shadow DOM} are not affected by any styling from outside. Nodes
+which are declared outside and distributed via \texttt{slots} will be
+styled in the \emph{light DOM} and can be additionally painted in the
+\emph{shadow DOM} through the new CSS-Selector \texttt{::slotted()}.
+
+Note that styles from the outside have an higher specify than styles
+assigned after distribution. Therefore it is generally a good advice to
+minimize the global stylings to some base styling for uniformity of the
+web site while leaving the specific stylings to the component. Due to
+the cascading nature of CSS, styles will still ``bleed in'' from
+ancestors to the \emph{light DOM} nodes. Therefore it's strongly
+recommended to begin every \emph{shadow DOM} with a \textbf{\emph{CSS
+reset:} \texttt{:host\ \{all:\ initial;\}}}.
+
+Regarding the importance style encapsulation, a couple of new CSS rules
+emerged that are exclusively targeting the \emph{shadow DOM}. The table
+below outlines styling possibilities for the use INSIDE the \emph{shadow
+DOM}:
+
+{[}{[} CSS Selectors {]}{]}
+
+Using the \emph{functional selector} of \texttt{:host()} or even the
+only-functional \texttt{:host-contest()} allows the creation of
+\textbf{context-aware custom elements}. A possible use case would be
+``theming'' a component (example taken from {[}11{]}):
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> index.html}
+\KeywordTok{<body}\OtherTok{ class=}\StringTok{"darktheme"}\KeywordTok{>}
+  \KeywordTok{<fancy-tabs>}
+    \NormalTok{...}
+  \KeywordTok{</fancy-tabs>}
+\KeywordTok{</body>}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> fancy-tabs shadowRoot}
+\NormalTok{<style>}
+\DecValTok{:}\NormalTok{host-context(}\FloatTok{.darktheme}\NormalTok{) }\KeywordTok{\{}
+\ErrorTok{ } \KeywordTok{color:} \DataTypeTok{white}\KeywordTok{;}
+\ErrorTok{ } \KeywordTok{background:} \DataTypeTok{black}\KeywordTok{;}
+\KeywordTok{\}}
+\NormalTok{</style>}
+\end{Highlighting}
+\end{Shaded}
+
+\subsubsection{JS Behavior}\label{js-behavior}
+
+As mentioned earlier any logic applied to \emph{light DOM} nodes stays
+with the node even after redistribution. For the sake of separation of
+concerns the business logic should be part of the \emph{custom element}
+(the \emph{light DOM}) and not the part of the \emph{shadow DOM}. On the
+other hand there are numerous scenarios where JS is just concerned with
+\textbf{styling or animation of an element}.In this case it might be
+more straightforward to apply JS inside the \emph{shadow DOM} to avoid
+mixing with business logic handlers.
+
+Drilling down to a \emph{light DOM} node from an \emph{shadow DOM}
+context is not possible with querying the node directly with
+\texttt{.querySelector()} or \texttt{.getElementById()} as the node is
+not part of the context. To get a distributed node in question it needs
+the way over the slot node and call \texttt{slot.assinedNodes()} to
+receive an array of distributed node(s) which can be accessed and
+manipulated like any other node. Calling \texttt{.assignedNodes()} on an
+empty \texttt{slot} returns an empty array.
+
+Wrapping up this section, \emph{shadow DOM} provides a non-hacky way to
+create uniform looking \emph{custom elements} and even enhance styling
+possibilities without adding much overhead. Still, for smaller
+components with only one or two child nodes, just a little styling
+and/or no structured redistribution a \emph{shadow DOM} might be to hard
+to reason about. Eventually it all depends on the question of ``how hard
+is it to implement it without shadow DOM'' - which can't be answered
+universally. For a more in-depth guide, Google Engineer Eric Bidelman
+wroten a great primer on \emph{shadow DOM}{[}11{]}.
+
+So far, there is still a missing link between \emph{light DOM} and
+\emph{shadow DOM}. The observant reader may have already noticed the
+weak point in the \texttt{HelloWorld} example: how to ``vitalize'' the
+\emph{shadow DOM}. Recapturing the last \texttt{HelloWorld} example a
+string of markup was assigned to the \texttt{shadowRoot.innerHTML}
+property. While it works perfectly fine in this simple case, a string of
+markup is rather cumbersome and error-prone and doesn't scale well. When
+putting quotes inside another quotes things break quickly. It makes the
+life hard for developers to work with it because it requires manual
+indentation and is out of syntax highlighting. That's the time templates
+come into play.
+
+\subsection{\texorpdfstring{HTML Templates
+\href{https://www.w3.org/TR/html5/scripting-1.html\#the-template-element}{(w3c)}}{HTML Templates (w3c)}}\label{html-templates-w3c}
+
+Among all other new standards \emph{HTML templates} are the most mature
+and adopted standard in the browser environment. All major browsers,
+except from Internet Explorer, support this standard.
+
+One core concept in templates is efficiency: Whatever dropped inside a
+\texttt{template} tag \sout{bucket} will be parsed on runtime - but not
+constructed into the \emph{content tree}. It remains plain HTML Markup
+sitting somewhere in the document until the time of activation.
+
+Activation typically takes four steps:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  \textbf{Querying the template node in question}\\
+  const node = document.querySelector(`template');
+\item
+  \textbf{Parsing the content and preparing the templates' content}\\
+  const content = node.content;\\
+  -\textgreater{} Returns a \emph{DocumentFragment} object. Handling is
+  straight forward\\
+  content.querySelector(`img').src = `logo.png';
+\item
+  \textbf{Optional: Cloning the \emph{DocumentFragment} for multiple
+  use}\\
+  const clone = content.cloneNode(``deep'');
+\item
+  \textbf{Appending the clone/original to destination}\\
+  document.body.appendChild(clone);
+\end{enumerate}
+
+As easy and minimal \emph{HTML templates} are, they're missing out a
+crucial feature other template implementations usually have. As
+templates are basically just dump containers for HTML Markup, there is
+no way to define some logic as \textbf{placeholders} where content
+should appear. Of course, with heavy use of JS things could be modeled
+this way. The idiomatic way tends more towards a \emph{Shadow DOM \&
+HTML templates} symbiosis.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> index.html}
+\KeywordTok{<hello-world>}
+  \KeywordTok{<p}\OtherTok{ id=}\StringTok{"sendto"}\OtherTok{ slot=}\StringTok{"placeholder"}\KeywordTok{>}
+    \NormalTok{Hello World Web Component    }
+  \KeywordTok{</p>}
+\KeywordTok{</hello-world>}
+\CommentTok{<!-- COMPONENT STARTS HERE -->}
+\KeywordTok{<template}\OtherTok{ id=}\StringTok{"hello"}\KeywordTok{>}
+  \CommentTok{<!-- STYLES -->}
+  \KeywordTok{<style>}
+    \FloatTok{#stylewrapper} \KeywordTok{\{}
+      \KeywordTok{font-weight:} \DataTypeTok{bold}\KeywordTok{;}
+      \KeywordTok{color:} \NormalTok{orange}\KeywordTok{;}
+    \KeywordTok{\}}
+  \KeywordTok{</style>}
+  \CommentTok{<!-- CONTENT -->}
+  \KeywordTok{<div}\OtherTok{ id=}\StringTok{"stylewrapper"}\KeywordTok{>}
+    \KeywordTok{<slot}\OtherTok{ name=}\StringTok{"placeholder"}\KeywordTok{>}
+      \NormalTok{Named placeholder}
+    \KeywordTok{</slot>}
+  \KeywordTok{</div>}
+\KeywordTok{</template>}
+
+\KeywordTok{<script>}
+  \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,}
+    \KeywordTok{class} \KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+     \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
+       \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
+       \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+       \KeywordTok{const} \NormalTok{template }\OperatorTok{=}
+         \VariableTok{document}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'#hello'}\NormalTok{)}\OperatorTok{;}
+       \KeywordTok{this}\NormalTok{.}\AttributeTok{shadowRoot}
+         \NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{template}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
+     \OperatorTok{\}}
+   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+\OperatorTok{<}\SpecialStringTok{/script>}
+\end{Highlighting}
+\end{Shaded}
+
+The updated \texttt{HelloWorld} component looks already pretty mature.
+It combines all the previous mentioned standards into one blob of HTML.
+\emph{Custom Elements} serves the logic, \emph{Shadow DOM} scopes the
+styles and \emph{HTML Templates} efficiently glues DOM and \emph{Shadow
+DOM} together. This separation of concerns comes with a huge gain in
+flexibility. In a real world scenario \texttt{HelloWorld} would
+contain/reference multiple \emph{HTML Templates} and could switch them
+around without any fuss. A developer might to split up templates into
+\textbf{STYLE} templates and \textbf{CONTENT} templates to increase
+reusability even further.
+
+The last standard in the row of four is not concerned with the internals
+of a \emph{web component}. \emph{HTML Imports} serves the need for an
+efficient distribution mechanism of components.
+
+\subsection{\texorpdfstring{HTML Imports
+\href{https://www.w3.org/TR/html-imports/}{(w3c)}}{HTML Imports (w3c)}}\label{html-imports-w3c}
+
+Importing the \texttt{HelloWorld} component is a one-liner:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"Hello.html"}\OtherTok{ async}\KeywordTok{>}
+\end{Highlighting}
+\end{Shaded}
+
+The \texttt{async} flag is optional but like in any other fetching
+event, strongly recommended. Once the imported HTML document comes into
+scope, activation follows a very similar pattern like the previously
+mentioned \emph{HTML templates}:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  \textbf{Querying the link node}
+\item
+  \textbf{Parsing the content and preparing the render}\\
+  const content = linknode.import; -\textgreater{} Unlike the \emph{HTML
+  template} the content a fully equipped document object.
+\item
+  \textbf{Optional: Cloning some nodes for multiple use}
+\item
+  \textbf{Appending the clone/original to destination}
+\end{enumerate}
+
+This again is the imperative way to handle a generic \emph{HTML Import}.
+In the declarative world of \emph{web components} a component is
+activated, parsed and anchored solely by its' tag name
+\texttt{\textless{}hello-world\textgreater{}\textless{}/hello-world\textgreater{}}.
+Preliminary, the component needs proper configuration as the last
+\texttt{HelloWorld} example wouldn't work like it is currently. The next
+section will provide a better understanding about the right
+configuration and composition of a component to work out-of-the-box.
+
+Despite from being just a practical document importer \emph{HTML
+imports} acts like a fully fledged dependency manager for the browser.
+Multiple resources, ranging from stylesheets, scripts, documents, media
+files and even other \texttt{imports} can be grouped together in a
+logical \texttt{import} statement. Internally, the browser engine keeps
+track for every imported resource so it won't be loaded twice. The
+inherent complexity is in fact a stumbling block for wider browser
+adoption. Currently only Googles blink web engine supports \emph{HTML
+Imports} as they are the driving force behind the \emph{web components}
+spec in general. Mozilla and Apple imposed distaste for \emph{HTML
+Imports} as a whole. One reason for this can be found in the
+incompatibility of the spec with the upcoming \emph{ES6 module
+loader}.\footnote{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}
+
+Despite the discrepancies among browser vendors \emph{HTML Imports}
+should still be part of the paper and future \emph{web components} as no
+other browser technology can bundle up CSS, JS and HTML that efficient.
+As of today, January 2017, only Googles Chrome and related Opera browser
+supporting the full spec and despite from \emph{HTML Imports} all other
+browser vendors most likely will catch up within this year. In the
+meantime, the full \emph{web components} standard can be
+\textbf{polyfilled} and used across all browsers.
+
+\subsection{Appendix: Custom Events}\label{appendix-custom-events}
 
 Macroperspektive / Composition
+
+\subsubsection{Assumptions about custom
+elements}\label{assumptions-about-custom-elements}
+
+Apart from the spec'd perspective there is mental model a webdeveloper
+might
+
+Creating and using webcomponents might require a new mental model how do
+design a
+
+containers
 
@@
 
+Pure frontend vs heavy backend
+
 \section{Progressive Enhancement}\label{progressive-enhancement}
 
-\subsection{Chapter about progressive
-enhancement}\label{chapter-about-progressive-enhancement}
-
-title: Browsernative Microservices subtitle: Modulare Webarchitekturen
-durch neue W3C-Spezifikationen author: Jan Peteler, FH
-Würzburg-Schweinfurt, jan.peteler@student.fhws.de date: Januar 2017
-lang: de-DE abstract: Webapplikationen lassen sich oft nur mit
-zusätzlichen Abstraktionsebenen wie beispielsweise ``Frameworks''
-verwirklichen. Neue Browserstandards ermöglichen eine native
-Servicearchitektur, die direkt auf der Browser-Engine aufsetzt. Diese
-Charakteristiken sollen systematisch erfasst werden. \ldots{}
+Chapter about progressive enhancement
 
@@
 
+\hypertarget{ref-Fowler2014}{}
+{[}3{]} M. Fowler and J. Lewis, ``Microservices: A definition of this
+new architectural term,'' Jan. 2014 {[}Online{]}. Available:
+\url{http://www.martinfowler.com/articles/microservices.html}
+
 \hypertarget{ref-Martin}{}
-{[}3{]} R. C. Martin, ``The single responsibility principle.''
+{[}4{]} R. C. Martin, ``The single responsibility principle.''
 {[}Online{]}. Available:
@@
 \hypertarget{ref-Conway1968}{}
-{[}4{]} M. E. Conway, ``How do committees invent?'' 1968 {[}Online{]}.
+{[}5{]} M. E. Conway, ``How do committees invent?'' 1968 {[}Online{]}.
 Available: \url{http://www.melconway.com/Home/Committees_Paper.html}
-
-\hypertarget{ref-Fowler2015}{}
-{[}5{]} M. Fowler and J. Lewis, ``Microservices: Nur ein weiteres
-konzept in der softwarearchitektur oder mehr?'' \emph{OBJEKTspektrum},
-Jan. 2015.
 
@@
 
+\hypertarget{ref-Buchner2016}{}
+{[}7{]} D. Buchner, ``Demythstifying web components,'' 2016
+{[}Online{]}. Available:
+\url{http://www.backalleycoder.com/2016/08/26/demythstifying-web-components/}
+
+\hypertarget{ref-vanKesteren2014}{}
+{[}8{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
+{[}Online{]}. Available:
+\url{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}
+
+\hypertarget{ref-Bidelman2016}{}
+{[}9{]} E. Bidelman, ``Custom elements v1: reusable web components.''
+2016 {[}Online{]}. Available:
+\url{https://developers.google.com/web/fundamentals/primers/customelements/}.
+{[}Accessed: 01-Dec-2016{]}
+
+\hypertarget{ref-HTML}{}
+{[}10{]} \emph{HTML living standard --- last updated 11 january 2017}.
+{[}Online{]}. Available: \url{https://html.spec.whatwg.org/multipage/}
+
+\hypertarget{ref-Bidelman2016shadow}{}
+{[}11{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
+2016 {[}Online{]}. Available:
+\url{https://developers.google.com/web/fundamentals/getting-started/primers/shadowdom}
+
 \end{document}

### 2017-01-25 paper.tex
pdf/tex build

@@
 
-\section{Introduction}\label{introduction}
+\section{Simplicity and the state of the
+web}\label{simplicity-and-the-state-of-the-web}
 
@@
 
-Introduction: from simplicity to Microservices to extensible web
-manifesto to w3c specifications \ldots{} and beyond
+Computers can scale, humans can't. Ever since, a program or complex
+system made by humans has been constrained by our mental capabilities.
+Like in the analogy of juggling balls, our brain can just ``juggle'' a
+few things at a time. Rich Hickey, the inventor of the programming
+language \emph{Clojure} gave an inspirational keynote on the topic of
+\textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
+  Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} In every sphere
+of a humans life, simplicity aligns perception with our mental
+capacities. Understanding the word \textbf{simple} alone will outline
+the this paper on its own.
 
-Der Erfinder der Programmiersprache Clojure, Rich Hickey, ist ohne
-Zweifel eine Koryphäe auf seinem Gebiet, der Strukturierung von
-komplexen Systemen. In einer vielbeachteten Keynote aus dem Jahr 2012
-geht er auf etymologisch-philosophische Spurensuche nach dem Wort
-\textbf{simplicity } aus Sicht eines Softwareentwicklers.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=48s}{Rails
-  Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} Das Adjektiv
-\emph{simple} hat demnach seinen Ursprung im lateinischen Wort
-\emph{simplex}, was soviel wie \emph{einfach} oder \emph{einzeln}
-bedeutet. In Gegensatz dazu stehen Eigenschaften wie \emph{complex} oder
-\emph{multiplex}. Qualitative Software ist, so Hickey, vor allem simpel
-- im Prozess, im Design, in der Struktur und in der Entwicklung.
+Derived from the ancient Latin word \textbf{simplex}, simple can be
+understood as ``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
+  Dictionary}} which points directly to the unidimensional aspect. While
+complexity describes the multilayered und entangled nature of
+conditions, simplicity empowers the human brain to reason about issues
+in a straightforward manner. It certainly has some overlapping's with
+\emph{easy}, but while \emph{easy} is more of a relative nature, simple
+can be laid out as a objective manner - a goal software development
+should pursuit and praise.
 
-Bezeichnend für diese Keynote ist, dass sie von Rich Hickey im Rahmen
-einer Webentwicklerkonferenz gehalten wurde. Software für den Browser
-war (und ist) seit langer Zeit maßgeblich geprägt von
-\textbf{Komplexität} auf mehreren Ebenen. Im Laufe dieser Bachelorarbeit
-werden diese Stru
+Software development is undoubtedly rich in complexity and full of
+subtle pitfalls for the human brain. In a typical scenario, a piece of
+software evolves over time in one or another \sout{opinionated}
+direction. Layers of new abstractions wrestling with old legacy
+abstractions and mutating objects will get out of control. Subtle bugs
+start to creep in. Eventually the small piece of software ends up in a
+highly complected monolith which will determine future design decisions
+to a painful degree. Any company will be highly determined by the
+current state trying to ``keep the lights on''.
 
-\begin{itemize}
-\tightlist
-\item
-  Das \emph{Document Object Model} als Rückgrad jeder Webapplikation ist
-  hierarchisch strukturiert und deren Elemente damit keinesfalls
-  unabhängig in ihrer Darstellung und Reihenfolge.
-\item
-  JavaScript als single-threaded Skriptsprache lässt sich schlecht ihn
-  ihrem Verhalten isolieren, ist Fehleranfällig und hatte lange Zeit nur
-  sehr wenig idiomatische Lösungsansätze für komplexe Probleme, wie
-  beispielsweise Asynchronität
-\item
-  CSS ist geprägt von stetigem Überschreiben vorher definierter Regeln
-  und verletzt damit Simplizität in ihrer Struktur, Design und dem
-  Entwicklungsprozesses auf bester Art und Weise.
-\end{itemize}
+On the other hand, any truly agile system architecture is laid out in a
+fine-grained manner. As Rich Hickey argues, design decisions should be
+made under the \textbf{impression of extending, substitution, moving,
+combining and repurposing}. The ability to reason about the program at
+any given time is crucial for future decisions and implementations.
+Recalling again the unidimensional nature of simplicity.
+
+The \emph{state of the web} is certainly a different kind of complex
+multidimensional beast. ``The web'' is coined to be meant ``everything
+that runs in the browser''. While simplicity in the backend is mostly a
+matter of principles, any frontend developer is highly restricted on the
+entangled nature of the browser platform. Every call to an webpage
+results in a single \textbf{global DOM tree} managed by the browser
+engine. In the early days of the web a typical static page contained
+just a bunch of text nodes and some styling. Nowadays a typical webapp
+contains text nodes, functionality and styling from different resources
+and is expected to work highly dynamic and frictionless.
+
+In the last four years the average transfer size of a webpage doubled to
+currently around 2.5 MB.{[}\^{}http {]}Subtracting images, fonts or
+other content the size of HTML, CSS and JS sums up to a total average of
+550 kb. One character weights around 1 byte which means an average
+webpage is delivering 550.000 character or around 125 pages of
+single-spaced text. Frederic Filloux calculated the ratio of real
+content on different newspaper websites and came to the conclusion, that
+only round about 5-6 \% of the transferred characters made for
+consumption.{[}1{]}
+
+Having an 95 \% overhead is rather undesirable for both the consumers
+and creators of the website. Since it's a widespread problem without a
+single point of failure one can argue the platform itself is the
+failure. In the recent years many frameworks, libraries and
+methodologies tried to simplify the way of building user interfaces by
+scoping assets and design rules into maintainable components. One of the
+most adopted technique was the \textbf{virtual DOM} approach essential
+ditching the old rule of separated HTML, CSS and JS in favor of
+\textbf{scoped JS components} (containing HTML and CSS). While
+frameworks are a valid approach to solve the problem they remain highly
+opinionated.
+
+In 2013 thinkers, creators and browser vendors joined together to
+propose \emph{The Extensible Web Manifesto}.\footnote{\href{https://extensiblewebmanifesto.org/}{The
+  Extensible Web Manifesto}} The claim of the Manifesto is to enhance
+the current web with new low-level capacities. Those capacities should
+empower creators of the web to write more declarative code and therefore
+overcome known bottlenecks and artificial abstractions. Four years
+later, the new capabilities of JavaScript leapfrogged and many new
+low-level APIs brought to life. With the new APIs at hand a webdeveloper
+can create scoped and highly reusable microservices without additional
+libraries even directly in the browser console.
+
+\emph{Disclaimer:} Currently, many of the new standards proposed
+throughout this paper can only tried and tested inside Google Chrome or
+Opera. As the innovation cycle is pretty short it is expected to see
+those standards go live in all major browsers soon.In the meantime all
+of those functionalities can be safely \textbf{polyfilled}.
+
+\subsection{Microservices}\label{microservices}
+
+Microserve
+
+untangle the web
 
@@
 einsetzbar und hochgradig wiederwerwendbar. Abramov nennt diese
-Komponenten ``Presentational Components''.{[}1{]} Verortet man diese Art
+Komponenten ``Presentational Components''.{[}2{]} Verortet man diese Art
 von Komponenten innerhalb des \emph{MVC Pattern}, sind die Komponenten
@@
 Components'' haben oft einen internen Zustand, den sie verändern können
-und an ihre Kindkomponenten weiterreichen können.{[}1{]} Im \emph{MVC
+und an ihre Kindkomponenten weiterreichen können.{[}2{]} Im \emph{MVC
 Pattern} handelt es sich um die \emph{Controller}. In der funktionalen
@@
 
-\section{Microservices}\label{microservices}
+\section{Microservices}\label{microservices-1}
 
@@
 In a nutshell a microservice is a small, autonomous service that works
-together with other services seamlessly.{[}2, p. 2{]} or with the words
+together with other services seamlessly.{[}3, p. 2{]} or with the words
 of Fowler and Lewis: ``It is an approach to developing a single
 application as a suite of small services, each running in its own
-process and communicating with lightweight mechanisms,\ldots{}''{[}3{]}
+process and communicating with lightweight mechanisms,\ldots{}''{[}4{]}
 Microservices incorporate many ideas, like \emph{domain-driven design}
-where we try to represent the real world in our code.{[}2, p. 2{]} Or
+where we try to represent the real world in our code.{[}3, p. 2{]} Or
 making use of \emph{continuous delivery} for pushing software rapidly
-through \emph{automated deployment} mechanisms in production.{[}3{]}
+through \emph{automated deployment} mechanisms in production.{[}4{]}
 And, last but not least, microservices utilizes the idea of small teams
@@
 Exemplary, microservices often communicate via an HTTP request-response
-with resource API's and lightweight messaging.{[}3{]} Nevertheless,
+with resource API's and lightweight messaging.{[}4{]} Nevertheless,
 while (server-side) microservices offer wide ranges of technical
@@
 strip away most of the dependencies in favour of clean
-interfaces.{[}3{]}
+interfaces.{[}4{]}
 
@@
 ``Gather together those things that change for the same reason and
-separate those things that change for different reasons.''{[}4{]} An a
+separate those things that change for different reasons.''{[}5{]} An a
 way this principle tackles another often cited design principle of the
@@
 perspective. Microservice patterns are a product of real-world
-usage.{[}2, p. 1{]} In a real world we typical have to deal with the so
+usage.{[}3, p. 1{]} In a real world we typical have to deal with the so
 called \emph{Conway's Law}:
@@
 designs which are copies of the communication~structures~of these
-organizations''. {[}5{]}
+organizations''. {[}6{]}
 \end{quote}
@@
 optimise the workflow. Fowler and Lewis outlines this approach as an
-``alignment of business capabilities''{[}3{]} While this kind of
+``alignment of business capabilities''{[}4{]} While this kind of
 structure may be true for companies like Google or Amazon, there is a
 vast majority of companies developing for the web which are grouped
-around tasks.{[}6{]} A very common structure is formed by the technology
+around tasks.{[}7{]} A very common structure is formed by the technology
 stack (UX Designers, Frontend- \& Backend Developers) or by separating
@@
 essentially make it possible to create a scoped component but not caring
-too much on it's distribution{[}7{]}. Some people prefer the three specs
-{[}8{]}, but the majority advocating the four specs variant, which is
+too much on it's distribution{[}8{]}. Some people prefer the three specs
+{[}9{]}, but the majority advocating the four specs variant, which is
 listed on \sout{the official}
@@
 entire DOM API and means any properties/methods that you add to the
-class become part of the element's DOM interface.''{[}9{]} Like any
+class become part of the element's DOM interface.''{[}10{]} Like any
 other \emph{ES6 class}, \emph{custom element}s can be sub-classed
@@
   Called upon the time of \emph{connecting or upgrading the node} which
-  means the moment the node is inserted and rendered inside the DOM.
-  Typically this block of code contains setup code, such as fetching
-  resources or rendering.{[}9{]} For performance issues it's highly
-  preferable to put much code in here.
+  means the moment the node is rendered inside the DOM. Typically this
+  method is called straight after the \texttt{constructor()} on insert.
+  This block of code contains setup code, such as fetching resources or
+  rendering elements according to attributes.{[}10{]} For performance
+  issues it's highly preferable to put much code in here.
 \item
@@
   \OperatorTok{\}}
-\OperatorTok{\}}
+\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
 \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}\OperatorTok{;}
+\CommentTok{// Instantiation}
 \KeywordTok{var} \NormalTok{el }\OperatorTok{=} \KeywordTok{new} \AttributeTok{HelloWorld}\NormalTok{()}\OperatorTok{;}
@@
 Native DOM properties will reflect their values between HTML and JS
-automatically.{[}10, Para. 2.6.1{]}
+automatically.{[}11, Para. 2.6.1{]}
 
@@
 functionality is perfectly spec'd it is strongly rejected by some
-browser vendors.\footnote{Discussion on the topic:
-  https://github.com/w3c/webcomponents/issues/509} Most likely the spec
-will change in future in one or other way on this issue and therefore
-\textbf{customized build-in elements} left out of this paper
-intentionally.
+browser vendors.\footnote{https://github.com/w3c/webcomponents/issues/509}
+Most likely the spec will change in future in one or other way on this
+issue and therefore \textbf{customized build-in elements} left out of
+this paper intentionally.
 
@@
     \NormalTok{...}
-    \KeywordTok{const} \NormalTok{shadowRoot }\OperatorTok{=}
     \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
@@
 \textbf{context-aware custom elements}. A possible use case would be
-``theming'' a component (example taken from {[}11{]}):
+``theming'' a component (example taken from {[}12{]}):
 
@@
 universally. For a more in-depth guide, Google Engineer Eric Bidelman
-wroten a great primer on \emph{shadow DOM}{[}11{]}.
+wroten a great primer on \emph{shadow DOM}{[}12{]}.
 
@@
   \textbf{Querying the template node in question}\\
-  \texttt{const\ t\ =\ document.querySelector(\textquotesingle{}\#helloworld\textquotesingle{});}
+  const node = document.querySelector(`template');
 \item
-  \textbf{Preparing the templates' content}\\
-  The templates' \texttt{content} property contains all nodes a
-  \emph{DocumentFragment} object. Handling should be straight forward\\
-  \texttt{t.content.querySelector(\textquotesingle{}img\textquotesingle{}).src\ =\ \textquotesingle{}logo.png\textquotesingle{};}
+  \textbf{Parsing the content and preparing the templates' content}\\
+  const content = node.content;\\
+  -\textgreater{} Returns a \emph{DocumentFragment} object. Handling is
+  straight forward\\
+  content.querySelector(`img').src = `logo.png';
 \item
@@
   use}\\
-  \texttt{const\ clone\ =\ t.content.cloneNode("deep");}
+  const clone = content.cloneNode(``deep'');
 \item
   \textbf{Appending the clone/original to destination}\\
-  \texttt{document.body.appendChild(clone);}
+  document.body.appendChild(clone);
 \end{enumerate}
@@
 should appear. Of course, with heavy use of JS things could be modeled
-this way. But the idiomatic way tends more towards a \emph{Shadow DOM \&
-HTML Templates} symbiosis.
+this way. The idiomatic way tends more towards a \emph{Shadow DOM \&
+HTML templates} symbiosis.
 
@@
 \begin{Highlighting}[]
-\NormalTok{> hello-component.html}
+\NormalTok{> index.html}
 \KeywordTok{<hello-world>}
   \KeywordTok{<p}\OtherTok{ id=}\StringTok{"sendto"}\OtherTok{ slot=}\StringTok{"placeholder"}\KeywordTok{>}
-    \NormalTok{Hello World    }
+    \NormalTok{Hello World Web Component    }
   \KeywordTok{</p>}
 \KeywordTok{</hello-world>}
-
+\CommentTok{<!-- COMPONENT STARTS HERE -->}
 \KeywordTok{<template}\OtherTok{ id=}\StringTok{"hello"}\KeywordTok{>}
+  \CommentTok{<!-- STYLES -->}
   \KeywordTok{<style>}
@@
 \KeywordTok{<script>}
-  \KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+  \CommentTok{// Switched to anonymous class notation}
+  \CommentTok{// for keeping associated code together.}
+  \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,}
+    \KeywordTok{class} \KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
      \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
        \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
-      \KeywordTok{const} \NormalTok{shadowRoot }\OperatorTok{=}
        \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
        \KeywordTok{const} \NormalTok{template }\OperatorTok{=}
-        \VariableTok{document}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'#hello'}\NormalTok{).}\AttributeTok{content}\OperatorTok{;}
-
-      \KeywordTok{this}\NormalTok{.}\VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{appendChild}\NormalTok{(template)}\OperatorTok{;}
+         \VariableTok{document}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'#hello'}\NormalTok{)}\OperatorTok{;}
+       \KeywordTok{this}\NormalTok{.}\AttributeTok{shadowRoot}
+         \NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{template}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
      \OperatorTok{\}}
-  \OperatorTok{\}}
-  \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}\OperatorTok{;}
+   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
 \OperatorTok{<}\SpecialStringTok{/script>}
@@
 
-The updated \texttt{HelloWorld} example looks already pretty mature. It
-combines all the previous mentioned standards into one blob of HTML.
+The updated \texttt{HelloWorld} component looks already pretty mature.
+It combines all the previous mentioned standards into one blob of HTML.
 \emph{Custom Elements} serves the logic, \emph{Shadow DOM} scopes the
-styles and \emph{HTML Template} efficiently glues DOM and \emph{Shadow
-DOM} together. The last standard in the row of four is not concerned
-with the internals of a \emph{web component}. \emph{HTML Imports} serves
-the need for an efficient distribution mechanism of components.
+styles and \emph{HTML Templates} efficiently glues DOM and \emph{Shadow
+DOM} together. This separation of concerns comes with a huge gain in
+flexibility. In a real world scenario \texttt{HelloWorld} would
+contain/reference multiple \emph{HTML Templates} and could switch them
+around without any fuss. A developer might to split up templates into
+\textbf{STYLE} templates and \textbf{CONTENT} templates to increase
+reusability even further.
+
+The last standard in the row of four is not concerned with the internals
+of a \emph{web component}. \emph{HTML Imports} serves the need for an
+efficient distribution mechanism of components.
 
@@
 \href{https://www.w3.org/TR/html-imports/}{(w3c)}}{HTML Imports (w3c)}}\label{html-imports-w3c}
+
+Importing the \texttt{HelloWorld} component is a one-liner:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"Hello.html"}\OtherTok{ async}\KeywordTok{>}
+\end{Highlighting}
+\end{Shaded}
+
+The \texttt{async} flag is optional but like in any other fetching
+event, strongly recommended. Once the imported HTML document comes into
+scope, activation follows a very similar pattern like the previously
+mentioned \emph{HTML templates}:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  \textbf{Querying the link node}
+\item
+  \textbf{Parsing the content and preparing the render}\\
+  const content = linknode.import; -\textgreater{} Unlike the \emph{HTML
+  template} the content a fully equipped document object.
+\item
+  \textbf{Optional: Cloning some nodes for multiple use}
+\item
+  \textbf{Appending the clone/original to destination}
+\end{enumerate}
+
+This again is the imperative way to handle a generic \emph{HTML Import}.
+In the declarative world of \emph{web components} a component is
+activated, parsed and anchored solely by its' tag name
+\texttt{\textless{}hello-world\textgreater{}\textless{}/hello-world\textgreater{}}.
+Preliminary, the component needs proper configuration as the last
+\texttt{HelloWorld} example wouldn't work like it is currently. The next
+section will elaborate the right configuration and composition of a
+component to work out-of-the-box.
+
+Despite from being just a practical document importer \emph{HTML
+imports} acts like a fully fledged dependency manager for the browser.
+Multiple resources, ranging from stylesheets, scripts, documents, media
+files and even other \texttt{imports} can be grouped together in a
+logical \texttt{import} statement. Internally, the browser engine keeps
+track for every imported resource so it won't be loaded twice. The
+inherent complexity is in fact a stumbling block for wider browser
+adoption. Currently only Googles blink web engine supports \emph{HTML
+Imports} as they are the driving force behind the \emph{web components}
+spec in general. Mozilla and Apple imposed distaste for \emph{HTML
+Imports} as a whole. One reason for this can be found in the
+incompatibility of the spec with the upcoming \emph{ES6 module
+loader}.\footnote{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}
+
+Despite the discrepancies among browser vendors \emph{HTML Imports}
+should still be part of the paper and future \emph{web components} as no
+other native browser technology can bundle up CSS, JS and HTML that
+efficient. As of today, January 2017, only Googles Chrome and related
+Opera browser supporting the full \emph{web components} spec and,
+despite from \emph{HTML Imports}, all other browser vendors most likely
+will catch up with \emph{Custom Elements} and \emph{Shadow DOM} within
+this year. In the meantime, the full \emph{web components} stack can be
+\textbf{polyfilled} and used across all browsers.
+
+\subsection{\texorpdfstring{Appendix A: Custom Events
+\href{https://dom.spec.whatwg.org/\#interface-customevent}{(whatwg)}}{Appendix A: Custom Events (whatwg)}}\label{appendix-a-custom-events-whatwg}
+
+Events are first-class citizens in the browser world and \emph{Custom
+Events} are no exception. The \emph{Custom Elements} interface is part
+of the DOM since years but with the rise of \emph{Custom Elements} they
+will most likely become an indispensable building block of \emph{web
+components}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> index.html}
+\KeywordTok{<hello-world>}
+  \KeywordTok{<button>}\NormalTok{Launch CustomEvent}\KeywordTok{</button>}
+\KeywordTok{</hello-world>}
+\CommentTok{<!-- COMPONENT STARTS HERE -->}
+\KeywordTok{<script>}
+  \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} 
+  \KeywordTok{class} \KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+    \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
+      \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
+      \CommentTok{// Craft a CustomEvent e}
+      \KeywordTok{const} \NormalTok{e }\OperatorTok{=} \KeywordTok{new} \AttributeTok{CustomEvent}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \OperatorTok{\{}
+        \DataTypeTok{bubbles}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,} \CommentTok{//important!}
+        \DataTypeTok{detail}\OperatorTok{:} \StringTok{'Contains string or object'}
+      \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+      \CommentTok{// Launch e on child button click}
+      \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'click'}\OperatorTok{,} \NormalTok{click }\OperatorTok{=>} \OperatorTok{\{}
+        \KeywordTok{this}\NormalTok{.}\AttributeTok{dispatchEvent}\NormalTok{(e)}
+        \VariableTok{click}\NormalTok{.}\AttributeTok{stopPropagation}\NormalTok{()}\OperatorTok{;}
+      \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+      \CommentTok{// Catch e. Typically done by some parent}
+      \CommentTok{// node. Message in detail property}
+      \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>} \OperatorTok{\{}
+        \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\VariableTok{e}\NormalTok{.}\AttributeTok{detail}\NormalTok{)}
+      \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+    \OperatorTok{\}}
+  \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+\OperatorTok{<}\SpecialStringTok{/script>}
+\end{Highlighting}
+\end{Shaded}
+
+Naming events after the emitting tag makes the API almost
+self-explanatory. Fortunately the \texttt{detail} property can transmit
+even objects witch allows the developer to craft almost all
+functionality inside one event. The further sections will elaborate a
+feasible architecture for building a scaling microservice architecture.
+
+Chaining and aggregating events from child nodes should be practiced and
+exercised quiet frequently. As mentioned earlier in the \emph{Custom
+Elements} section, the pattern of \textbf{extending native elements}
+should be somewhat dismissed as it may be implemented outside of Chrome.
+Nevertheless, it is possible to create own kind of quasi native buttons
+when chaining a \emph{CustomEvent} directly after the native click
+event.
+
+Another typical \emph{custom element} use case can be as an actor on
+(native) child elements. In this case, the \emph{Custom Element} catches
+events from children, buffers or rebuild them and eventuall fires an
+event towards the document root.
+
+Unfortunately events only work ``upstream'' towards parent nodes. Still
+the web platform offers plenty of possibilities to talk back to child
+nodes.
+
+\subsection{\texorpdfstring{Appendix B: Web Worker
+\href{https://html.spec.whatwg.org/multipage/workers.html}{(whatwg)}}{Appendix B: Web Worker (whatwg)}}\label{appendix-b-web-worker-whatwg}
+
+Like \emph{Custom Events}, \emph{Web Workers} had been around for a long
+time and therefore embrace full support among major browser vendors.
+They emerged at around 2009 when discussions about browser performance
+was in the early stages of development. Nevertheless, the addressed
+problem of \emph{Web Workers} is a fundamental language problem of JS
+itself.
+
+JS runs in a single-threaded language environment. Every script in the
+browser environment, from handling UI events to query and process larget
+amounts of API data and manipulating the DOM, runs on the same
+thread{[}13{]}. Putting a lot of work to the single main thread can slow
+down the web service significantly. From time to time scripts can block
+or fail to whatever reason which leads to a frozen UI on the users side.
+A worker can overcome the bottleneck of the single-threaded nature with
+spawning a new \textbf{background thread}. Most of the browsers work can
+be leveraged to this new thread. Todays web architectures aims to
+leverage an increasing amout of proccessing to the client to avoid
+time-consuming roundtrips especially in mobiles networks.\footnote{Latency
+  numbers: https://gist.github.com/jboner/2841832} While many native
+APIs like \texttt{fetch} work seamlessly in the new thread, a worker has
+no access to the DOM at all.
+
+A \emph{Web Worker} spawns a new background thread where scripts can run
+concurrent to the main thread. Usually a worker is loaded from a workers
+dedicated file to embrace this kind of separation.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{const} \NormalTok{worker }\OperatorTok{=} \KeywordTok{new} \AttributeTok{Worker}\NormalTok{(}\StringTok{'worker.js'}\NormalTok{)}\OperatorTok{;}
+\end{Highlighting}
+\end{Shaded}
+
+After initialization a worker communicates over a simple \textbf{message
+based interface} with the main thread.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
+\CommentTok{// Send to worker}
+\VariableTok{worker}\NormalTok{.}\AttributeTok{postMessage}\NormalTok{(}\StringTok{'Hello World'}\NormalTok{)}\OperatorTok{;}
+\CommentTok{// Receive msg from worker}
+\VariableTok{worker}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'message'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>}
+  \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\StringTok{'Worker said: '}\OperatorTok{,} \VariableTok{e}\NormalTok{.}\AttributeTok{data}\NormalTok{))}\OperatorTok{;}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\OperatorTok{>} \VariableTok{worker}\NormalTok{.}\AttributeTok{js}
+\CommentTok{// Receive msg and echo back}
+\KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'message'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>}
+  \KeywordTok{this}\NormalTok{.}\AttributeTok{postMessage}\NormalTok{(}\StringTok{"Echo "} \OperatorTok{+} \VariableTok{e}\NormalTok{.}\AttributeTok{data}\NormalTok{))}\OperatorTok{;}
+\end{Highlighting}
+\end{Shaded}
+
+Having no access to the DOM can be seen as hinderance,but i
 
@@
 \hypertarget{refs}{}
+\hypertarget{ref-Filloux2016}{}
+{[}1{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016
+{[}Online{]}. Available:
+\url{https://mondaynote.com/bloated-html-the-best-and-the-worse-cac6eb06496d}
+
 \hypertarget{ref-Abramov2015}{}
-{[}1{]} D. Abramov, ``Presentational and Container Components --
+{[}2{]} D. Abramov, ``Presentational and Container Components --
 Medium.'' 2015 {[}Online{]}. Available:
@@
 \hypertarget{ref-Newman2015}{}
-{[}2{]} S. Newman, \emph{Building microservices}. O'Reilly Media, Inc.,
+{[}3{]} S. Newman, \emph{Building microservices}. O'Reilly Media, Inc.,
 2016 {[}Online{]}. Available:
@@
 \hypertarget{ref-Fowler2014}{}
-{[}3{]} M. Fowler and J. Lewis, ``Microservices: A definition of this
+{[}4{]} M. Fowler and J. Lewis, ``Microservices: A definition of this
 new architectural term,'' Jan. 2014 {[}Online{]}. Available:
@@
 \hypertarget{ref-Martin}{}
-{[}4{]} R. C. Martin, ``The single responsibility principle.''
+{[}5{]} R. C. Martin, ``The single responsibility principle.''
 {[}Online{]}. Available:
@@
 \hypertarget{ref-Conway1968}{}
-{[}5{]} M. E. Conway, ``How do committees invent?'' 1968 {[}Online{]}.
+{[}6{]} M. E. Conway, ``How do committees invent?'' 1968 {[}Online{]}.
 Available: \url{http://www.melconway.com/Home/Committees_Paper.html}
@@
 \hypertarget{ref-Issa2016}{}
-{[}6{]} B. Issa, ``The way of the web.'' Polymer Summit 2016, Oct-2016
+{[}7{]} B. Issa, ``The way of the web.'' Polymer Summit 2016, Oct-2016
 {[}Online{]}. Available:
@@
 \hypertarget{ref-Buchner2016}{}
-{[}7{]} D. Buchner, ``Demythstifying web components,'' 2016
+{[}8{]} D. Buchner, ``Demythstifying web components,'' 2016
 {[}Online{]}. Available:
@@
 \hypertarget{ref-vanKesteren2014}{}
-{[}8{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
+{[}9{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
 {[}Online{]}. Available:
@@
 \hypertarget{ref-Bidelman2016}{}
-{[}9{]} E. Bidelman, ``Custom elements v1: reusable web components.''
+{[}10{]} E. Bidelman, ``Custom elements v1: reusable web components.''
 2016 {[}Online{]}. Available:
@@
 \hypertarget{ref-HTML}{}
-{[}10{]} \emph{HTML living standard --- last updated 11 january 2017}.
+{[}11{]} \emph{HTML living standard --- last updated 11 january 2017}.
 {[}Online{]}. Available: \url{https://html.spec.whatwg.org/multipage/}
@@
 \hypertarget{ref-Bidelman2016shadow}{}
-{[}11{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
+{[}12{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
 2016 {[}Online{]}. Available:
@@
 
+\hypertarget{ref-Bidelman2010}{}
+{[}13{]} E. Bidelman, ``The basics of web workers.'' 2010 {[}Online{]}.
+Available: \url{https://www.html5rocks.com/en/tutorials/workers/basics/}
+
 \end{document}

### 2017-01-25 paper.tex
pdf/tex build

@@
 In the last four years the average transfer size of a webpage doubled to
-currently around 2.5 MB.{[}\^{}http {]}Subtracting images, fonts or
-other content the size of HTML, CSS and JS sums up to a total average of
-550 kb. One character weights around 1 byte which means an average
-webpage is delivering 550.000 character or around 125 pages of
-single-spaced text. Frederic Filloux calculated the ratio of real
-content on different newspaper websites and came to the conclusion, that
-only round about 5-6 \% of the transferred characters made for
-consumption.{[}1{]}
+currently around 2.5 MB.\footnote{\href{http://httparchive.org/trends.php}{HTTPArchive
+  Trends}} Subtracting images, fonts or other content the size of HTML,
+CSS and JS sums up to a total average of 550 kb. One character weights
+around 1 byte which means an average webpage is delivering 550.000
+character or around 125 pages of single-spaced text. Frederic Filloux
+calculated the ratio of real content on different newspaper websites and
+came to the conclusion, that only round about 5-6 \% of the transferred
+characters made for consumption.{[}1{]}
 

### 2017-01-26 build/paper.tex
pdf/tex build

@@
 
-Introduction: from simplicity to Microservices to extensible web
-manifesto to w3c specifications \ldots{} and beyond
+Computers can scale, humans can't. Ever since a program or complex
+system made by humans has been constrained by humans mental
+capabilities. Like in the analogy of juggling balls, our brain can just
+``juggle'' a few things at a time. Rich Hickey, the inventor of the
+programming language \emph{Clojure} gave an inspirational keynote on the
+topic of \textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
+  Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} In every sphere
+of a humans life, simplicity aligns perception with our mental
+capacities. Understanding the word \textbf{simple} alone will outline
+the this paper on its own.
 
-Der Erfinder der Programmiersprache Clojure, Rich Hickey, ist ohne
-Zweifel eine Koryphäe auf seinem Gebiet, der Strukturierung von
-komplexen Systemen. In einer vielbeachteten Keynote aus dem Jahr 2012
-geht er auf etymologisch-philosophische Spurensuche nach dem Wort
-\textbf{simplicity } aus Sicht eines Softwareentwicklers.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=48s}{Rails
-  Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} Das Adjektiv
-\emph{simple} hat demnach seinen Ursprung im lateinischen Wort
-\emph{simplex}, was soviel wie \emph{einfach} oder \emph{einzeln}
-bedeutet. In Gegensatz dazu stehen Eigenschaften wie \emph{complex} oder
-\emph{multiplex}. Qualitative Software ist, so Hickey, vor allem simpel
-- im Prozess, im Design, in der Struktur und in der Entwicklung.
+Derived from the ancient Latin word \textbf{simplex}, simple can be
+understood as ``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
+  Dictionary}} which points directly to the unidimensional aspect. While
+complexity describes the multilayered und entangled nature of
+conditions, simplicity empowers the human brain to reason about issues
+in a straightforward manner. It certainly has some overlapping's with
+\emph{easy}, but while \emph{easy} is more of a relative nature, simple
+can be laid out as a objective manner - a goal software development
+should pursuit and praise.
 
-Bezeichnend für diese Keynote ist, dass sie von Rich Hickey im Rahmen
-einer Webentwicklerkonferenz gehalten wurde. Software für den Browser
-war (und ist) seit langer Zeit maßgeblich geprägt von
-\textbf{Komplexität} auf mehreren Ebenen. Im Laufe dieser Bachelorarbeit
-werden diese Stru
+Software development is undoubtedly rich in complexity and full of
+subtle pitfalls for the human brain. In a typical scenario, a piece of
+software evolves over time in one or another \sout{opinionated}
+direction. Layers of new abstractions wrestling with old legacy
+abstractions and mutating objects will get out of control. Subtle bugs
+start to creep in. Eventually the small piece of software ends up in a
+highly complected monolith which will determine future design decisions
+to a painful degree. Any company will be highly determined by the
+current state trying to ``keep the lights on''.
 
-\begin{itemize}
-\tightlist
-\item
-  Das \emph{Document Object Model} als Rückgrad jeder Webapplikation ist
-  hierarchisch strukturiert und deren Elemente damit keinesfalls
-  unabhängig in ihrer Darstellung und Reihenfolge.
-\item
-  JavaScript als single-threaded Skriptsprache lässt sich schlecht ihn
-  ihrem Verhalten isolieren, ist Fehleranfällig und hatte lange Zeit nur
-  sehr wenig idiomatische Lösungsansätze für komplexe Probleme, wie
-  beispielsweise Asynchronität
-\item
-  CSS ist geprägt von stetigem Überschreiben vorher definierter Regeln
-  und verletzt damit Simplizität in ihrer Struktur, Design und dem
-  Entwicklungsprozesses auf bester Art und Weise.
-\end{itemize}
+On the other hand, any truly agile system architecture is laid out in a
+fine-grained manner. As Rich Hickey argues, design decisions should be
+made under the \textbf{impression of extending, substitution, moving,
+combining and repurposing}. The ability to reason about the program at
+any given time is crucial for future decisions and implementations.
+Recalling again the unidimensional nature of simplicity.
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+The \emph{state of the web} is certainly a different kind of complex
+multidimensional beast. ``The web'' is coined to be meant ``everything
+that runs in the browser''. While simplicity in the backend is mostly a
+matter of principles, any frontend developer is highly restricted on the
+entangled nature of the browser platform.
 
-Betrachtet man andere populäre Webframeworks dieser Zeit, wie
-beispielsweise React oder Angular, lässt sich ziemlich schnell ein
-gemeinsames Designpattern ausmachen, wie Komponenten intern ihre
-Zuständen verwalten. Der Facebook Entwickler Dan Abramov hat dieses
-dichotome Pattern systematisch erfasst und in zwei Kategorien
-eingeteilt.
+In the last four years the average transfer size of a webpage doubled to
+currently around 2.5 MB.\footnote{\href{http://httparchive.org/trends.php}{HTTPArchive
+  Trends}} Subtracting images, fonts or other content the size of HTML,
+CSS and JS sums up to a total average of 550 kb. One character weights
+around 1 byte which means an average webpage is delivering 550.000
+character or around 125 pages of single-spaced text. Frederic Filloux
+calculated the ratio of real content on different newspaper websites and
+came to the conclusion, that only round about 5-6 \% of the transferred
+characters made for consumption.{[}1{]}
 
-Nach Abramov existieren zum einen Komponenten, die alleine für das
-\textbf{Darstellen von Information} zuständig sind. Diese Komponenten
-nehmen keinerlei Einfluss auf Informationen oder anderen Komponenten um
-sich herum. Sie sind im wahrsten Sinne passiv und fremdgesteuert über
-klar definierte Schnittstellen. Diese Komponenten sind häufig flexibel
-einsetzbar und hochgradig wiederwerwendbar. Abramov nennt diese
-Komponenten ``Presentational Components''.{[}1{]} Verortet man diese Art
-von Komponenten innerhalb des \emph{MVC Pattern}, sind die Komponenten
-reine \emph{View-Elemente}.
+Having an 95 \% overhead is rather undesirable for both the consumers
+and creators of the website. Since it's a widespread problem without a
+single point of failure one can argue the platform itself is the
+failure. By design, every pageload results in a \textbf{global DOM tree}
+managed by the browser engine. Regardless of rendering just a bunch of
+static text nodes or an ever changing webapp the underlying global model
+remains the same. \textbf{Encapsulation and modularization} is a typical
+pattern in the computer industry to make complexity manageable and to
+accommodate future uncertainty.{[}2, p. 1{]} Today many websites tend
+towards becoming complex single-page applications (and therefore the
+average JS payload is soaring too). Consequently, the demands to the
+browser platform changed from a static page renderer to a
+\textbf{complexe UI machine}. Without additional layers of abstraction
+modularization would be impossible to archive.
 
-Im Gegensatz dazu stehen Komponenten, die für die \textbf{Verarbeitung
-von Informationen} zuständig sind. Diese so genannten ``Container
-Components'' haben oft einen internen Zustand, den sie verändern können
-und an ihre Kindkomponenten weiterreichen können.{[}1{]} Im \emph{MVC
-Pattern} handelt es sich um die \emph{Controller}. In der funktionalen
-Programmiersprache Elm werden diese Elemente \emph{Updater} genannt, was
-ihre Funktion noch besser umschreibt.
+In the recent years many \textbf{frameworks}, libraries and
+methodologies tried to tackle the global nature of the DOM by scoping
+assets and design rules into maintainable components. While the DOM
+can't be scoped, JS can. Therefore many of those frameworks, like
+ReactJS, AngularJS or VueJS just to name a few, ditched the old rule of
+separated HTML, CSS and JS in favor of an additional layer of
+\textbf{abstracted JS components} (containing content, markup and
+styling). Quiet often those frameworks mimic a MVC pattern on top of the
+browser engine which is a reasonable simple design pattern to build
+graphical user interfaces. This paper will apply the same pattern for
+building a browsernative microservice.While frameworks are a valid
+approach to build scalable web applications they remain highly
+opinionated and can change and break over time. Furthermore, they
+inflate the code with libraries as well. A standardized way for creating
+complex UI interfaces painlessly requires new build-in browser
+capabilities.
 
-Ein üblicher eventgesteuerter Webservice setzt sich aus
-unterschiedlichsten Komponenten zusammen, die wiederum
-unterschiedlichste Eventlistener \& -emitter in sich subsummieren. Diese
-inhärente Komplexität verlangt geradezu nach einer klaren,
-deterministischen Struktur des Webservices, die das Zusammenspiel
-orchestriert. In der Analogie des Orchesters gesprochen, benötigt der
-Webservice (oder sogar die gesamte Webapplikation) einen Dirigenten, der
-für die Steuerung verantwortlich ist.
+In 2013 thinkers, creators and browser vendors joined together to
+propose \emph{The Extensible Web Manifesto}.\footnote{\href{https://extensiblewebmanifesto.org/}{The
+  Extensible Web Manifesto}} The claim of the Manifesto is to enhance
+the current web platforms with new low-level capabilities. Those
+features should empower creators of the web to write more declarative
+code and therefore overcome known bottlenecks and artificial
+abstractions. Four years later, the enhancement of JavaScript
+leapfrogged and many new low-level APIs brought to life. With the new
+APIs at hand a webdeveloper can create scoped and highly reusable
+microservices without additional libraries even directly in the browser
+console.
+
+\emph{Disclaimer:} Currently, all new technologies proposed throughout
+this paper can be tried and tested in Google Chrome, Opera and the
+Safari Technology Preview. Usually the browser adoption time is short in
+time an therefore full support of all standards is expected soon.
+Meanwhile this standards can be used through \textbf{polyfills} on every
+important legacy browser.
 
@@
 
-Opening up the case for \emph{Browsernative Microservices} brings up the
-question about the concept microservices in general. In fact the concept
-of microservices has many facets, stretching beyond disciplines and
-technical boundaries. It lacks a formal standardization but there are
+Calling the need for \emph{Browsernative Microservices} brings up the
+question about the concept of microservices in general. In fact the
+concept of microservices has many facets, stretching beyond disciplines
+and technical boundaries. It lacks a formal standardization but subsumes
 certain ideas emergine from this pattern. As a primary source of truth
-this articles relies on the work of Sam Newman, who has written a
-comprehensive guide in \emph{Building Microservices}. The purpose of
-this section is to match those ideas against the manifestations of web
-components.
+this paper relies on the work of Sam Newman, who has written a
+comprehensive guide in \emph{Building Microservices} and the work of
+Fowler and Lewis. The purpose of this section is to match those ideas
+against the browser platform where different rules of development
+applies.
 
 In a nutshell a microservice is a small, autonomous service that works
-together with other services seamlessly.{[}2, p. 2{]} or with the words
-of Fowler and Lewis: ``It is an approach to developing a single
-application as a suite of small services, each running in its own
-process and communicating with lightweight mechanisms,\ldots{}''{[}3{]}
-Microservices incorporate many ideas, like \emph{domain-driven design}
-where we try to represent the real world in our code.{[}2, p. 2{]} Or
-making use of \emph{continuous delivery} for pushing software rapidly
-through \emph{automated deployment} mechanisms in production.{[}3{]}
-And, last but not least, microservices utilizes the idea of small teams
-with a lot of product knowledge working mostly autonomous on their very
-own service with their very own set of tools and techniques.
+together with other services seamlessly.{[}3, p. 2{]} or with the words
+of Fowler and Lewis: ``In short, the microservice architectural style is
+an approach to developing a single application as a suite of small
+services, each running in its own process and communicating with
+lightweight mechanisms, often an HTTP resource API''{[}4{]}
+Microservices incorporate many other ideas, like \emph{domain-driven
+design} where we try to represent the real world in our code.{[}3, p.
+2{]} Or making use of \emph{continuous delivery} for pushing software
+rapidly through \emph{automated deployment} mechanisms in
+production.{[}4{]} And, last but not least, microservices utilizes the
+idea of small teams with a lot of product knowledge working mostly
+autonomous on their very own service with their very own set of tools
+and techniques.
 
-\subsection{Technical perspective}\label{technical-perspective}
+\subsection{Componentization via
+Services}\label{componentization-via-services}
+
+``A~\textbf{component}~is a unit of software that is independently
+replaceable and upgradeable.''{[}4{]} Compared to libraries a component
+service offers multiple advantages. A library is only loosely coupled to
+the implementation and therefore hard to track in functionality.
+Changing a library often requires an unforeseen amount of time fixing
+implementations. Therefore in the field we often see multiple versions
+of libraries deployed simultaneous (like with JQuery) which adds a huge
+amount of overhead on the page load. Eventually, dead code elimination
+won't work in highly coupled libraries. Components instead incorporate a
+much better evolution mechanism in the service contracts. Changing
+functionality won't break other components. A component can
+progressively enhanced which guarantees functionality throughout
+different versions.
+
+Compared to libraries componentized services offers a more explicit
+interface.{[}4{]} While the functionality of a library needs
+documentation, a component functionality is exposed via the components'
+signature and therefore highly accessible. HTML elements are usually
+controlled via their attributes. A generic HTML service therefore might
+look like this:
+\texttt{\textless{}my-button\ isPrimary\textgreater{}\textless{}/button\textgreater{}}.
+
+\subsection{Organized around Business
+Capabilities}\label{organized-around-business-capabilities}
+
+\subsection{Products not Projects}\label{products-not-projects}
+
+\subsection{Smart endpoints and dumb
+pipes}\label{smart-endpoints-and-dumb-pipes}
+
+\subsection{Decentralized Governance}\label{decentralized-governance}
+
+\subsection{Decentralized Data
+Management}\label{decentralized-data-management}
+
+\subsection{Infrastructure Automation}\label{infrastructure-automation}
+
+\subsection{Design for failure}\label{design-for-failure}
+
+\subsection{Evolutionary Design}\label{evolutionary-design}
 
@@
  \OperatorTok{\}}
-\OperatorTok{\}}
+\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
 \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}\OperatorTok{;}
+\CommentTok{// Instantiation}
 \KeywordTok{var} \NormalTok{el }\OperatorTok{=} \KeywordTok{new} \AttributeTok{HelloWorld}\NormalTok{()}\OperatorTok{;}
@@
   -\textgreater{} Returns a \emph{DocumentFragment} object. Handling is
-  straight forward\\
-  content.querySelector(`img').src = `logo.png';
+  straight forward content.querySelector(`img').src = `logo.png';
 \item
@@
 \KeywordTok{<script>}
+ \CommentTok{// Switched to anonymous class notation}
+ \CommentTok{// for keeping associated code together.}
  \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,}
@@
 \texttt{HelloWorld} example wouldn't work like it is currently. The next
-section will provide a better understanding about the right
-configuration and composition of a component to work out-of-the-box.
+section will elaborate the right configuration and composition of a
+component to work out-of-the-box.
 
@@
 should still be part of the paper and future \emph{web components} as no
-other browser technology can bundle up CSS, JS and HTML that efficient.
-As of today, January 2017, only Googles Chrome and related Opera browser
-supporting the full spec and despite from \emph{HTML Imports} all other
-browser vendors most likely will catch up within this year. In the
-meantime, the full \emph{web components} standard can be
+other native browser technology can bundle up CSS, JS and HTML that
+efficient. As of today, January 2017, only Googles Chrome and related
+Opera browser supporting the full \emph{web components} spec and,
+despite from \emph{HTML Imports}, all other browser vendors most likely
+will catch up with \emph{Custom Elements} and \emph{Shadow DOM} within
+this year. In the meantime, the full \emph{web components} stack can be
 \textbf{polyfilled} and used across all browsers.
 
-\subsection{Appendix: Custom Events}\label{appendix-custom-events}
+\subsection{\texorpdfstring{Appendix A: Custom Events
+\href{https://dom.spec.whatwg.org/\#interface-customevent}{(whatwg)}}{Appendix A: Custom Events (whatwg)}}\label{appendix-a-custom-events-whatwg}
+
+Events are first-class citizens in the browser world and \emph{Custom
+Events} are no exception. The \emph{Custom Elements} interface is part
+of the DOM since years but with the rise of \emph{Custom Elements} they
+will most likely become an indispensable building block of \emph{web
+components}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> index.html}
+\KeywordTok{<hello-world>}
+ \KeywordTok{<button>}\NormalTok{Launch CustomEvent}\KeywordTok{</button>}
+\KeywordTok{</hello-world>}
+\CommentTok{<!-- COMPONENT STARTS HERE -->}
+\KeywordTok{<script>}
+ \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} 
+ \KeywordTok{class} \KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
+     \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
+   \CommentTok{// Craft a CustomEvent e}
+     \KeywordTok{const} \NormalTok{e }\OperatorTok{=} \KeywordTok{new} \AttributeTok{CustomEvent}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \OperatorTok{\{}
+    \DataTypeTok{bubbles}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,} \CommentTok{//important!}
+    \DataTypeTok{detail}\OperatorTok{:} \StringTok{'Contains string or object'}
+   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+   \CommentTok{// Launch e on child button click}
+   \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'click'}\OperatorTok{,} \NormalTok{click }\OperatorTok{=>} \OperatorTok{\{}
+    \KeywordTok{this}\NormalTok{.}\AttributeTok{dispatchEvent}\NormalTok{(e)}
+    \VariableTok{click}\NormalTok{.}\AttributeTok{stopPropagation}\NormalTok{()}\OperatorTok{;}
+   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+   \CommentTok{// Catch e. Typically done by some parent}
+     \CommentTok{// node. Message in detail property}
+   \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>} \OperatorTok{\{}
+    \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\VariableTok{e}\NormalTok{.}\AttributeTok{detail}\NormalTok{)}
+   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+  \OperatorTok{\}}
+ \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+\OperatorTok{<}\SpecialStringTok{/script>}
+\end{Highlighting}
+\end{Shaded}
+
+Naming events after the emitting tag makes the API almost
+self-explanatory. Fortunately the \texttt{detail} property can transmit
+even objects witch allows the developer to craft almost all
+functionality inside one event. The further sections will elaborate a
+feasible architecture for building a scaling microservice architecture.
+
+Chaining and aggregating events from child nodes should be practiced and
+exercised quiet frequently. As mentioned earlier in the \emph{Custom
+Elements} section, the pattern of \textbf{extending native elements}
+should be somewhat dismissed as it may be implemented outside of Chrome.
+Nevertheless, it is possible to create own kind of quasi native buttons
+when chaining a \emph{CustomEvent} directly after the native click
+event.
+
+Another typical \emph{custom element} use case can be as an actor on
+(native) child elements. In this case, the \emph{Custom Element} catches
+events from children, buffers or rebuild them and eventuall fires an
+event towards the document root.
+
+Unfortunately events only work ``upstream'' towards parent nodes. Still
+the web platform offers plenty of possibilities to talk back to child
+nodes.
+
+\subsection{\texorpdfstring{Appendix B: Web Worker
+\href{https://html.spec.whatwg.org/multipage/workers.html}{(whatwg)}}{Appendix B: Web Worker (whatwg)}}\label{appendix-b-web-worker-whatwg}
+
+Like \emph{Custom Events}, \emph{Web Workers} had been around for a long
+time and therefore embrace full support among major browsers. They
+emerged at around 2009 when discussions about browser performance was
+still in the early days. Nevertheless, the addressed problem of
+\emph{Web Workers} is a fundamental language problem of JS itself.
+
+JS runs in a single-threaded language environment. Every script in the
+browser environment, from handling UI events to query and process larget
+amounts of API data and manipulating the DOM, runs on the same
+thread{[}13{]}. Putting a lot of work to the single main thread can slow
+down the web service significantly. From time to time scripts can block
+or fail to whatever reason which leads to a frozen UI on the users side.
+A worker can overcome the bottleneck of the single-threaded nature with
+spawning a new \textbf{background thread}. Most of the browsers work can
+be leveraged to this new thread. Todays web architectures aims to
+leverage an increasing amout of proccessing to the client to avoid
+time-consuming roundtrips especially in mobiles networks.\footnote{Latency
+  numbers: https://gist.github.com/jboner/2841832} While many native
+APIs like \texttt{fetch} work seamlessly in the new thread, a worker has
+no access to the DOM at all.
+
+A \emph{Web Worker} spawns a new background thread where scripts can run
+concurrent to the main thread. Usually a worker is loaded from a workers
+dedicated file to embrace this kind of separation.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{const} \NormalTok{worker }\OperatorTok{=} \KeywordTok{new} \AttributeTok{Worker}\NormalTok{(}\StringTok{'worker.js'}\NormalTok{)}\OperatorTok{;}
+\end{Highlighting}
+\end{Shaded}
+
+After initialization a worker communicates over a simple \textbf{message
+based interface} with the main thread.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
+\CommentTok{// Send to worker}
+\VariableTok{worker}\NormalTok{.}\AttributeTok{postMessage}\NormalTok{(}\StringTok{'Hello World'}\NormalTok{)}\OperatorTok{;}
+\CommentTok{// Receive msg from worker}
+\VariableTok{worker}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'message'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>}
+ \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\StringTok{'Worker said: '}\OperatorTok{,} \VariableTok{e}\NormalTok{.}\AttributeTok{data}\NormalTok{))}\OperatorTok{;}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\OperatorTok{>} \VariableTok{worker}\NormalTok{.}\AttributeTok{js}
+\CommentTok{// Receive msg and echo back}
+\KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'message'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>}
+ \KeywordTok{this}\NormalTok{.}\AttributeTok{postMessage}\NormalTok{(}\StringTok{"Echo "} \OperatorTok{+} \VariableTok{e}\NormalTok{.}\AttributeTok{data}\NormalTok{))}\OperatorTok{;}
+\end{Highlighting}
+\end{Shaded}
+
+Having no access to the DOM can be seen as hinderance,but i
 
@@
 
+dichotome Pattern
+
+Ein üblicher eventgesteuerter Webservice setzt sich aus
+unterschiedlichsten Komponenten zusammen, die wiederum
+unterschiedlichste Eventlistener \& -emitter in sich subsummieren. Diese
+inhärente Komplexität verlangt geradezu nach einer klaren,
+deterministischen Struktur des Webservices, die das Zusammenspiel
+orchestriert. In der Analogie des Orchesters gesprochen, benötigt der
+Webservice (oder sogar die gesamte Webapplikation) einen Dirigenten, der
+für die Steuerung verantwortlich ist.
+
 http://alistair.cockburn.us/Hexagonal+architecture
@@
 \hypertarget{refs}{}
-\hypertarget{ref-Abramov2015}{}
-{[}1{]} D. Abramov, ``Presentational and Container Components --
-Medium.'' 2015 {[}Online{]}. Available:
-\url{https://medium.com/@dan_abramov/smart-and-dumb-components-7ca2f9a7c7d0}.
-{[}Accessed: 01-Dec-2016{]}
+\hypertarget{ref-Filloux2016}{}
+{[}1{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016
+{[}Online{]}. Available:
+\url{https://mondaynote.com/bloated-html-the-best-and-the-worse-cac6eb06496d}
+
+\hypertarget{ref-Baldwin2006}{}
+{[}2{]} C. Y. Baldwin and K. B. Clark, ``Modularity in the Design of
+Complex Engineering Systems,'' in \emph{Complex engineered systems:
+Science meets technology}, D. Braha, A. A. Minai, and Y. Bar-Yam, Eds.
+Berlin, Heidelberg: Springer Berlin Heidelberg, 2006, pp. 175--205
+{[}Online{]}. Available:
+\href{http://dx.doi.org/10.1007/3-540-32834-3\%7B/_\%7D9}{http://dx.doi.org/10.1007/3-540-32834-3\{\textbackslash{}\_\}9}
 
 \hypertarget{ref-Newman2015}{}
-{[}2{]} S. Newman, \emph{Building microservices}. O'Reilly Media, Inc.,
+{[}3{]} S. Newman, \emph{Building microservices}. O'Reilly Media, Inc.,
 2016 {[}Online{]}. Available:

### 2017-01-27 paper.tex
pdf/tex build

@@
 
-Computers can scale, humans can't. Ever since, a program or complex
-system made by humans has been constrained by our mental capabilities.
-Like in the analogy of juggling balls, our brain can just ``juggle'' a
-few things at a time. Rich Hickey, the inventor of the programming
-language \emph{Clojure} gave an inspirational keynote on the topic of
-\textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
+Computers can scale, humans can't. Ever since a program or complex
+system made by humans has been constrained by humans mental
+capabilities. Like in the analogy of juggling balls, our brain can just
+``juggle'' a few things at a time. Rich Hickey, the inventor of the
+programming language \emph{Clojure} gave an inspirational keynote on the
+topic of \textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
   Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} In every sphere
 of a humans life, simplicity aligns perception with our mental
-capacities. Understanding the word \textbf{simple} alone will outline
-the this paper on its own.
+capacities. This paper should be read under the impression of
+simplicity.
 
@@
 in a straightforward manner. It certainly has some overlapping's with
-\emph{easy}, but while \emph{easy} is more of a relative nature, simple
-can be laid out as a objective manner - a goal software development
-should pursuit and praise.
+easy, but while easy is more of a relative nature, simple can be laid
+out as a objective manner - a goal software development should pursuit
+and praise.
 
@@
 
-The \emph{state of the web} is certainly a different kind of complex
-multidimensional beast. ``The web'' is coined to be meant ``everything
-that runs in the browser''. While simplicity in the backend is mostly a
-matter of principles, any frontend developer is highly restricted on the
-entangled nature of the browser platform. Every call to an webpage
-results in a single \textbf{global DOM tree} managed by the browser
-engine. In the early days of the web a typical static page contained
-just a bunch of text nodes and some styling. Nowadays a typical webapp
-contains text nodes, functionality and styling from different resources
-and is expected to work highly dynamic and frictionless.
+The \textbf{state of the web} is certainly a different kind of
+complexity beast. ``The web'' is coined to ``everything that runs in the
+browser''. While simplicity in the backend is mostly a matter of
+principles, any frontend developer is highly restricted on the highly
+deterministic nature of the browser platform.
 
@@
 came to the conclusion, that only round about 5-6 \% of the transferred
-characters made for consumption.{[}1{]}
+characters made for human consumption.{[}1{]}
 
@@
 single point of failure one can argue the platform itself is the
-failure. In the recent years many frameworks, libraries and
-methodologies tried to simplify the way of building user interfaces by
-scoping assets and design rules into maintainable components. One of the
-most adopted technique was the \textbf{virtual DOM} approach essential
-ditching the old rule of separated HTML, CSS and JS in favor of
-\textbf{scoped JS components} (containing HTML and CSS). While
-frameworks are a valid approach to solve the problem they remain highly
-opinionated.
+failure. By design, every pageload results in a monolithic DOM tree
+managed by the browser engine. Whether rendering just a bunch of static
+text nodes or an ever changing webapp the underlying global nature of
+the DOM tree remains the same. Every additional service added to the
+webapp embodies another fold of complexity.
+
+In an non-deterministic runtime environment, encapsulation and
+modularization is a typical pattern to make complexity manageable and to
+accommodate future uncertainty.{[}2, p. 1{]} Today many websites tend
+towards becoming complex single-page applications (and therefore the
+average JS payload is soaring too). Consequently, the demands to the
+browser platform changed from a static page renderer to a
+\textbf{complexe UI machine}. Under the current situation only
+additional layers of abstraction can handle those complexity.
+
+In the recent years many \textbf{frameworks}, libraries and
+methodologies tried to tackle the global nature of the DOM by scoping
+assets and design rules into maintainable components. While the DOM
+can't be scoped, JS can. Therefore many of those frameworks, like
+ReactJS, AngularJS or VueJS just to name a few, ditched the old rule of
+separated HTML, CSS and JS in favor of an additional layer of abstracted
+JS components (containing content, markup and styling). Quiet often
+those frameworks mimic a MVC pattern on top of the browser engine which
+is a reasonable simple design pattern to build graphical user
+interfaces. This paper will apply the same pattern for building a
+browsernative microservice.While frameworks are a valid approach to
+build scalable web applications they remain highly opinionated, embody
+inherent complexity themselves and can change and break over time.
+Furthermore, they inflate the code with libraries. A standardized way
+for creating complex UI interfaces painlessly requires new build-in
+browser capabilities.
 
@@
 propose \emph{The Extensible Web Manifesto}.\footnote{\href{https://extensiblewebmanifesto.org/}{The
-  Extensible Web Manifesto}} The claim of the Manifesto is to enhance
-the current web with new low-level capacities. Those capacities should
-empower creators of the web to write more declarative code and therefore
-overcome known bottlenecks and artificial abstractions. Four years
-later, the new capabilities of JavaScript leapfrogged and many new
-low-level APIs brought to life. With the new APIs at hand a webdeveloper
-can create scoped and highly reusable microservices without additional
-libraries even directly in the browser console.
-
-\emph{Disclaimer:} Currently, many of the new standards proposed
-throughout this paper can only tried and tested inside Google Chrome or
-Opera. As the innovation cycle is pretty short it is expected to see
-those standards go live in all major browsers soon.In the meantime all
-of those functionalities can be safely \textbf{polyfilled}.
-
-\subsection{Microservices}\label{microservices}
-
-Microserve
-
-untangle the web
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-Betrachtet man andere populäre Webframeworks dieser Zeit, wie
-beispielsweise React oder Angular, lässt sich ziemlich schnell ein
-gemeinsames Designpattern ausmachen, wie Komponenten intern ihre
-Zuständen verwalten. Der Facebook Entwickler Dan Abramov hat dieses
-dichotome Pattern systematisch erfasst und in zwei Kategorien
-eingeteilt.
-
-Nach Abramov existieren zum einen Komponenten, die alleine für das
-\textbf{Darstellen von Information} zuständig sind. Diese Komponenten
-nehmen keinerlei Einfluss auf Informationen oder anderen Komponenten um
-sich herum. Sie sind im wahrsten Sinne passiv und fremdgesteuert über
-klar definierte Schnittstellen. Diese Komponenten sind häufig flexibel
-einsetzbar und hochgradig wiederwerwendbar. Abramov nennt diese
-Komponenten ``Presentational Components''.{[}2{]} Verortet man diese Art
-von Komponenten innerhalb des \emph{MVC Pattern}, sind die Komponenten
-reine \emph{View-Elemente}.
-
-Im Gegensatz dazu stehen Komponenten, die für die \textbf{Verarbeitung
-von Informationen} zuständig sind. Diese so genannten ``Container
-Components'' haben oft einen internen Zustand, den sie verändern können
-und an ihre Kindkomponenten weiterreichen können.{[}2{]} Im \emph{MVC
-Pattern} handelt es sich um die \emph{Controller}. In der funktionalen
-Programmiersprache Elm werden diese Elemente \emph{Updater} genannt, was
-ihre Funktion noch besser umschreibt.
-
-Ein üblicher eventgesteuerter Webservice setzt sich aus
-unterschiedlichsten Komponenten zusammen, die wiederum
-unterschiedlichste Eventlistener \& -emitter in sich subsummieren. Diese
-inhärente Komplexität verlangt geradezu nach einer klaren,
-deterministischen Struktur des Webservices, die das Zusammenspiel
-orchestriert. In der Analogie des Orchesters gesprochen, benötigt der
-Webservice (oder sogar die gesamte Webapplikation) einen Dirigenten, der
-für die Steuerung verantwortlich ist.
-
-\section{Microservices}\label{microservices-1}
-
-Opening up the case for \emph{Browsernative Microservices} brings up the
-question about the concept microservices in general. In fact the concept
-of microservices has many facets, stretching beyond disciplines and
-technical boundaries. It lacks a formal standardization but there are
+  Extensible Web Manifesto}} The claim of the manifesto is to enhance
+the current web platforms with new low-level capabilities. Those
+features should empower creators of the web to write more declarative
+code and therefore overcome known bottlenecks and artificial
+abstractions. Four years later, the enhancement of JavaScript
+leapfrogged and many new low-level APIs brought to life. With the new
+APIs at hand a webdeveloper can create scoped and highly reusable
+microservices without additional libraries even directly in the browser
+console.
+
+\emph{Disclaimer:} Currently, all new technologies proposed throughout
+this paper can be tried and tested in Google Chrome, Opera and the
+Safari Technology Preview. Usually the browser adoption time is short in
+time an therefore full support of all standards is expected soon.
+Meanwhile this standards can be used through \textbf{polyfills} on every
+important legacy browser.
+
+\section{Microservices}\label{microservices}
+
+Calling the case for \emph{Browsernative Microservices} brings up the
+question about the concept of microservices in general. In fact the
+concept of microservices has many facets, stretching beyond disciplines
+and technical boundaries. It lacks a formal standardization but subsumes
 certain ideas emergine from this pattern. As a primary source of truth
-this articles relies on the work of Sam Newman, who has written a
-comprehensive guide in \emph{Building Microservices}. The purpose of
-this section is to match those ideas against the manifestations of web
-components.
+this paper relies on the work of Sam Newman, who has written a
+comprehensive guide called \emph{Building Microservices} and the work of
+Fowler and Lewis. The purpose of this section is to match those ideas
+against the browser platform as targeted runtime.
 
@@
 together with other services seamlessly.{[}3, p. 2{]} or with the words
-of Fowler and Lewis: ``It is an approach to developing a single
-application as a suite of small services, each running in its own
-process and communicating with lightweight mechanisms,\ldots{}''{[}4{]}
-Microservices incorporate many ideas, like \emph{domain-driven design}
-where we try to represent the real world in our code.{[}3, p. 2{]} Or
-making use of \emph{continuous delivery} for pushing software rapidly
-through \emph{automated deployment} mechanisms in production.{[}4{]}
-And, last but not least, microservices utilizes the idea of small teams
-with a lot of product knowledge working mostly autonomous on their very
-own service with their very own set of tools and techniques.
-
-\subsection{Technical perspective}\label{technical-perspective}
+of Fowler and Lewis: ``\ldots{} the microservice architectural style is
+an approach to developing a single application as a suite of small
+services, each running in its own process and communicating with
+lightweight mechanisms, often an HTTP resource API''{[}4{]}
+Microservices incorporate many other ideas, like \emph{domain-driven
+design} where we try to represent the real world in our code.{[}3, p.
+2{]} Or making use of \emph{continuous delivery} for pushing software
+rapidly through \emph{automated deployment} mechanisms in
+production.{[}4{]} And, last but not least, microservices utilizes the
+idea of small teams with a lot of product knowledge working mostly
+autonomous on their very own service with their very own set of tools
+and techniques.
+
+\subsection{Componentization via
+Services}\label{componentization-via-services}
+
+``A~\textbf{component}~is a unit of software that is independently
+replaceable and upgradeable.''{[}4{]} Compared to libraries, a component
+service offers multiple advantages. A library is only loosely coupled to
+the implementation and therefore hard to track in functionality.
+Changing a library often requires an unforeseen amount of time fixing
+implementations. Therefore in the field we often see multiple versions
+of libraries deployed simultaneous (like with JQuery) which adds a huge
+amount of overhead on the page load. Eventually, dead code elimination
+won't work in highly coupled libraries. Components instead incorporate a
+much better evolution mechanism in the service contracts. Changing
+functionality won't break other components. A component can
+progressively enhanced which guarantees functionality throughout
+different versions.
+
+Compared to libraries componentized services offers a more explicit
+interface.{[}4{]} While the functionality of a library needs
+documentation, a component functionality is exposed via the components'
+signature and therefore highly accessible. HTML elements are explicitly
+controlled via their attributes and the new HTML service component are
+no exception to that.
+
+A microservice reinforces the \emph{Single Responsibility Principle}
+defined by Robert C. Martin: ``Gather together those things that change
+for the same reason and separate those things that change for different
+reasons.''{[}5{]}
+
+\subsection{Organized around Business
+Capabilities}\label{organized-around-business-capabilities}
+
+\begin{quote}
+``organizations which design systems \ldots{} are constrained to produce
+designs which are copies of the communication~structures~of these
+organizations''. {[}6{]}
+\end{quote}
+
+Microservice patterns are a product of real-world usage.{[}3, p. 1{]}
+Instead of splitting teams along the technology stack (UI Experts
+-\textgreater{} Middleware -\textgreater{} Database) a microservice
+approach model teams around \textbf{business capabilities}.{[}4{]}
+Consequently every team is capable of planning, designing, implementing,
+testing and maintaining their very own microservice. Across the stack
+every member gains high competence about the functionality of the
+service.
+
+By assigning service responsibility a team, the so called
+\textbf{Definition of Done} (DoD) shifts from ``accomplishing projects''
+to ``accomplishing products''. This new paradigm not only changes the
+administrative overhead like budgeting or resource allocation. It
+creates a kind of responsibility connection from the team to the
+service. Expectedly, those teams are more motivated within their very
+own service and exhibit a more sophisticated iteration time.{[}4{]}
+
+For many companies working in the spheres of the internet, the
+client-side is the most crucial area of business success. Todays
+client-side includes many devices of differing sizes and
+functionalities. As browser technologies, design guidelines and devices
+change in a fast paced manner it makes absolutely sense to shift
+responsibility towards the teams altogether. A \emph{browsernative
+microservice} needs commitment over its whole lifecycle.
+
+\subsection{Smart endpoints and dumb
+pipes}\label{smart-endpoints-and-dumb-pipes}
+
+Probably most microservices build for the browser are impure in
+functionality. They aim for side-effects like changing the UI given to
+input parameters or emitting browser events. Therefore a web developer
+must design communication channels for those microservices keeping in
+mind the nested nature of the DOM.
+
+\emph{Smart endpoints and dumb pipes} is coined to the approach of
+designing communication mostly decoupled and as cohesive as possible.
+This essentially removes most of the error-prone middleware altogether.
+In the analogy of the real world a message channel should look like
+sending a letter: Only two endpoints (sender and receiver) connoted as
+``smart'' and the letter itself looks mostly dumb (but unified) from the
+outside. Only the receiver is able to process the content message.
+
+Cited earlier in this paper microservices often rely on simple HTTP
+request-response with resource API's and lightweight messaging as
+communication protocols.{[}4{]} In the later parts of this paper the
+reader will see a generic \emph{browsernative microservice} split in the
+fashion of \textbf{MVC}. In the \textbf{VIEW} the microservice will rely
+on browser events and in-memory calls while the \textbf{CONTROLLER} have
+to deal with other resources over the fetch API.
+
+\subsection{Decentralized Governance}\label{decentralized-governance}
+
+\subsection{Decentralized Data
+Management}\label{decentralized-data-management}
+
+\subsection{Infrastructure Automation}\label{infrastructure-automation}
+
+\subsection{Design for failure}\label{design-for-failure}
+
+\subsection{Evolutionary Design}\label{evolutionary-design}
 
@@
 like APIs Databanks / Deployment
-
-Talking about microservices in a browsernative context isn't that far
-fetched as microservices themselves incorporate many ideas from the web.
-Exemplary, microservices often communicate via an HTTP request-response
-with resource API's and lightweight messaging.{[}4{]} Nevertheless,
-while (server-side) microservices offer wide ranges of technical
-possibilities, we must take into account that (client-side) browsers
-come with certain constraints and limitations.
-
-Fowler and Lewis issue a call for using services as components. A
-component~is regarded as a unit of software that is independently
-replaceable and upgradeable. The main advantage of a component in
-contrast of library is the possibility of an independent deployment. It
-aligns perfectly with the main goal of a microservice architecture to
-strip away most of the dependencies in favour of clean
-interfaces.{[}4{]}
-
-Components
-
-First of all, from a technical perspective, a microservice reinforces
-the \emph{Single Responsibility Principle} defined by Robert C. Martin:
-``Gather together those things that change for the same reason and
-separate those things that change for different reasons.''{[}5{]} An a
-way this principle tackles another often cited design principle of the
-\emph{seperation of concerns}. Web Components incorporate this principle
-in multiple ways while still remaining flexible.
 
@@
 The concept of microservices incooperates not only a technical
-perspective. Microservice patterns are a product of real-world
-usage.{[}3, p. 1{]} In a real world we typical have to deal with the so
-called \emph{Conway's Law}:
-
-\begin{quote}
-``organizations which design systems \ldots{} are constrained to produce
-designs which are copies of the communication~structures~of these
-organizations''. {[}6{]}
-\end{quote}
+perspective.
 
@@
 Like \emph{Custom Events}, \emph{Web Workers} had been around for a long
-time and therefore embrace full support among major browser vendors.
-They emerged at around 2009 when discussions about browser performance
-was in the early stages of development. Nevertheless, the addressed
-problem of \emph{Web Workers} is a fundamental language problem of JS
-itself.
+time and therefore embrace full support among major browsers. They
+emerged at around 2009 when discussions about browser performance was
+still in the early days. Nevertheless, the addressed problem of
+\emph{Web Workers} is a fundamental language problem of JS itself.
 
@@
 
-\hypertarget{ref-Abramov2015}{}
-{[}2{]} D. Abramov, ``Presentational and Container Components --
-Medium.'' 2015 {[}Online{]}. Available:
-\url{https://medium.com/@dan_abramov/smart-and-dumb-components-7ca2f9a7c7d0}.
-{[}Accessed: 01-Dec-2016{]}
+\hypertarget{ref-Baldwin2006}{}
+{[}2{]} C. Y. Baldwin and K. B. Clark, ``Modularity in the Design of
+Complex Engineering Systems,'' in \emph{Complex engineered systems:
+Science meets technology}, D. Braha, A. A. Minai, and Y. Bar-Yam, Eds.
+Berlin, Heidelberg: Springer Berlin Heidelberg, 2006, pp. 175--205
+{[}Online{]}. Available:
+\href{http://dx.doi.org/10.1007/3-540-32834-3\%7B/_\%7D9}{http://dx.doi.org/10.1007/3-540-32834-3\{\textbackslash{}\_\}9}
 

### 2017-02-07 build/paper.tex
pdf/tex build

@@
 of a humans life, simplicity aligns perception with our mental
-capacities. Understanding the word \textbf{simple} alone will outline
-the this paper on its own.
+capacities.
 
@@
 in a straightforward manner. It certainly has some overlapping's with
-\emph{easy}, but while \emph{easy} is more of a relative nature, simple
-can be laid out as a objective manner - a goal software development
-should pursuit and praise.
+easy, but while easy is more of a relative nature, simple can be laid
+out as a objective manner and therefore universally applicable.
 
@@
 direction. Layers of new abstractions wrestling with old legacy
-abstractions and mutating objects will get out of control. Subtle bugs
-start to creep in. Eventually the small piece of software ends up in a
-highly complected monolith which will determine future design decisions
-to a painful degree. Any company will be highly determined by the
-current state trying to ``keep the lights on''.
-
-On the other hand, any truly agile system architecture is laid out in a
-fine-grained manner. As Rich Hickey argues, design decisions should be
-made under the \textbf{impression of extending, substitution, moving,
-combining and repurposing}. The ability to reason about the program at
-any given time is crucial for future decisions and implementations.
-Recalling again the unidimensional nature of simplicity.
-
-The \emph{state of the web} is certainly a different kind of complex
-multidimensional beast. ``The web'' is coined to be meant ``everything
-that runs in the browser''. While simplicity in the backend is mostly a
-matter of principles, any frontend developer is highly restricted on the
-entangled nature of the browser platform.
+abstractions and mutation becomes untraceable. Subtle bugs start to
+creep in. Eventually the small piece of software may end up in a highly
+complected monolith which will determine future design decisions to a
+painful degree. Future strategies of the company/organization will be
+highly determined by the current state in the need of ``keeping the
+lights on''.
+
+On the other side of this dystopian example, the truly agile system
+architecture is laid out in a fine-grained manner. Software becomes
+pluggable, mutations traceable and modules interchangeable. As Rich
+Hickey argues, design decisions should be made under the
+\textbf{impression of extending, substitution, moving, combining and
+repurposing}. The ability to reason about the program at any given time
+is crucial for future decisions and implementations. Recalling again the
+unidimensional nature of simplicity.
+
+The \textbf{state of the web} is certainly a different kind of
+complexity beast. ``The web'' is coined to ``everything that runs in the
+browser''. While simplicity in the backend is mostly a matter of
+principles, any frontend developer is highly restricted on the highly
+deterministic nature of the browser platform.
 
@@
 single point of failure one can argue the platform itself is the
-failure. By design, every pageload results in a \textbf{global DOM tree}
-managed by the browser engine. Regardless of rendering just a bunch of
-static text nodes or an ever changing webapp the underlying global model
-remains the same. \textbf{Encapsulation and modularization} is a typical
-pattern in the computer industry to make complexity manageable and to
+failure. By design, every pageload results in a monolithic DOM tree
+managed by the browser engine. Whether rendering just a bunch of static
+text nodes or an ever changing webapp the underlying global nature of
+the DOM tree remains the same. Every additional service added to the
+webapp embodies another fold of complexity.
+
+In an non-deterministic runtime environment, encapsulation and
+modularization is a typical pattern to make complexity manageable and
 accommodate future uncertainty.{[}2, p. 1{]} Today many websites tend
@@
 browser platform changed from a static page renderer to a
-\textbf{complexe UI machine}. Without additional layers of abstraction
-modularization would be impossible to archive.
+\textbf{complexe UI machine}. Under the current situation only
+additional layers of abstraction can handle complexity.
 
 In the recent years many \textbf{frameworks}, libraries and
-methodologies tried to tackle the global nature of the DOM by scoping
-assets and design rules into maintainable components. While the DOM
-can't be scoped, JS can. Therefore many of those frameworks, like
-ReactJS, AngularJS or VueJS just to name a few, ditched the old rule of
-separated HTML, CSS and JS in favor of an additional layer of
-\textbf{abstracted JS components} (containing content, markup and
-styling). Quiet often those frameworks mimic a MVC pattern on top of the
-browser engine which is a reasonable simple design pattern to build
-graphical user interfaces. This paper will apply the same pattern for
-building a browsernative microservice.While frameworks are a valid
-approach to build scalable web applications they remain highly
-opinionated and can change and break over time. Furthermore, they
-inflate the code with libraries as well. A standardized way for creating
-complex UI interfaces painlessly requires new build-in browser
-capabilities.
+methodologies approached the global nature of the DOM by scoping assets
+and design rules into maintainable components. While the DOM can't be
+scoped, JS can. Therefore many frameworks, like ReactJS, AngularJS or
+VueJS just to name a few, ditched the old rule of separated HTML, CSS
+and JS in favor of an additional layer of abstracted JS components
+(containing content, markup and styling). Quiet often those frameworks
+mimic a MVC pattern on top of the browser engine which is a reasonable
+simple design pattern to build graphical user interfaces. While
+frameworks are a valid approach to build scalable web applications they
+remain highly opinionated, embody inherent complexity themselves and can
+change and break over time. Furthermore, code inflation in front-end is
+a crucial point for performance and third-party libraries are no
+exception on that. A standardized way for creating complex UI interfaces
+painlessly requires new build-in browser capabilities.
 
@@
 propose \emph{The Extensible Web Manifesto}.\footnote{\href{https://extensiblewebmanifesto.org/}{The
-  Extensible Web Manifesto}} The claim of the Manifesto is to enhance
+  Extensible Web Manifesto}} The claim of the manifesto is to enhance
 the current web platforms with new low-level capabilities. Those
@@
 abstractions. Four years later, the enhancement of JavaScript
-leapfrogged and many new low-level APIs brought to life. With the new
-APIs at hand a webdeveloper can create scoped and highly reusable
+leapfrogged and many new low-level APIs brought to life. With this new
+APIs at hand a vivid web developer can create scoped and highly reusable
 microservices without additional libraries even directly in the browser
@@
 
-\emph{Disclaimer:} Currently, all new technologies proposed throughout
-this paper can be tried and tested in Google Chrome, Opera and the
-Safari Technology Preview. Usually the browser adoption time is short in
-time an therefore full support of all standards is expected soon.
-Meanwhile this standards can be used through \textbf{polyfills} on every
-important legacy browser.
+\emph{Disclaimer:} This paper introduces many new browser build-ins with
+the focus on try and test. As the time of writing, examples can be tried
+frictionless in the consoles of the latest versions of \textbf{Google
+Chrome, Opera and Apple Safari}.\footnote{\href{http://jonrimmer.github.io/are-we-componentized-yet/}{Are
+  we componentized yet?}} On Mozilla Firefox technologies work behind a
+flag and Microsoft Edge implementation is unfortunately far behind. But
+Browser implementation changes quickly and soon technology adoption
+won't be an issue. Meanwhile new standards can be used through
+\textbf{polyfills} even on legacy browsers.
 
@@
 
-Calling the need for \emph{Browsernative Microservices} brings up the
+Calling the case for \emph{Browsernative Microservices} brings up the
 question about the concept of microservices in general. In fact the
@@
 this paper relies on the work of Sam Newman, who has written a
-comprehensive guide in \emph{Building Microservices} and the work of
-Fowler and Lewis. The purpose of this section is to match those ideas
-against the browser platform where different rules of development
-applies.
+comprehensive guide called \emph{Building Microservices} and the work of
+Fowler and Lewis. The purpose of this section is to match their ideas
+against the browser platform as targeted runtime.
 
@@
 together with other services seamlessly.{[}3, p. 2{]} or with the words
-of Fowler and Lewis: ``In short, the microservice architectural style is
+of Fowler and Lewis: ``\ldots{} the microservice architectural style is
 an approach to developing a single application as a suite of small
 services, each running in its own process and communicating with
-lightweight mechanisms, often an HTTP resource API''{[}4{]}
+lightweight mechanisms, often an HTTP resource API.''{[}4{]}
 Microservices incorporate many other ideas, like \emph{domain-driven
-design} where we try to represent the real world in our code.{[}3, p.
-2{]} Or making use of \emph{continuous delivery} for pushing software
-rapidly through \emph{automated deployment} mechanisms in
-production.{[}4{]} And, last but not least, microservices utilizes the
-idea of small teams with a lot of product knowledge working mostly
+design} where it pursues the incorporation of the real world in our
+code.{[}3, p. 2{]} Or making use of \emph{continuous delivery} for
+pushing software rapidly through \emph{automated deployment} mechanisms
+into production.{[}4{]} And, last but not least, microservices utilizes
+the idea of small teams with a lot of product knowledge working mostly
 autonomous on their very own service with their very own set of tools
@@
 ``A~\textbf{component}~is a unit of software that is independently
-replaceable and upgradeable.''{[}4{]} Compared to libraries a component
-service offers multiple advantages. A library is only loosely coupled to
-the implementation and therefore hard to track in functionality.
-Changing a library often requires an unforeseen amount of time fixing
-implementations. Therefore in the field we often see multiple versions
-of libraries deployed simultaneous (like with JQuery) which adds a huge
-amount of overhead on the page load. Eventually, dead code elimination
-won't work in highly coupled libraries. Components instead incorporate a
-much better evolution mechanism in the service contracts. Changing
-functionality won't break other components. A component can
-progressively enhanced which guarantees functionality throughout
-different versions.
+replaceable and upgradeable.'' {[}4{]} Components are the building
+blocks of the microservice but in fact they even share a lot of
+similarities just on a smaller scale. One of the similarity is the
+\emph{loose coupling principle}: changing and deploying one service
+shouldn't result in changing other parts of the system.{[}3, p. 30{]}.
+Same goes for the component which should know as little as it needs
+about the service to fulfil it's job.
+
+Any reader of the paper coming from the ReactJS / Redux world this
+concept of components might look familiar. Dan Abramov, the creator of
+Redux, once defined a simple dichotomous pattern for creating UI
+components. In his believe, there are \textbf{presentational components}
+only related with the concern about \emph{how things look}. This
+component literally doesn't know anything about the service in question
+which makes the component highly flexible and reusable. They are
+controlled solely from the outside, receiving data and dispatching
+unbiased events on user interaction.{[}5{]}
+
+On the other hand Dan Abramov outlines components which he calls
+\textbf{containers}. A container component is concerned with \emph{how
+things work}.{[}5{]} Containers acts as invisible wrappers around
+presentational components. Their job is to fetch data from child nodes,
+aggregating events, interacting with the model and push state back to
+the presentational components. In contrast to their presentational
+counterparts, containers can't life on its own without children.
+Conceptually, they show another important principle of components and
+microservices in general: \emph{high cohesion principle}.
+
+Whether designing a microservice or it's components we want related
+behavior sit together, and unrelated behavior to sit
+elsewhere.{\textbf{???}} Code quality can be measured just by counting
+the places changes in the code need to be made on changing behavior. In
+an arbitrary MVC system a \texttt{button} might be inserted in the VIEW,
+the CONTROLLER needs some adjustment and maybe the the MODEL, too. Three
+places for adjustment is a reasonable easy task for the brain,
+everything more violates the concept of simplicity.
+
+Using libraries in web development is common sense. But compared to
+libraries, a component service offers multiple advantages. A library is
+only loosely coupled to the implementation and therefore hard to track
+in functionality. Changing a library may result in an unforeseen amount
+of time fixing implementations. It is not unusual to see websites
+embodying different versions of the same library (like with JQuery).
+Another issue with libraries is dead code elimination which means the
+process of removing code that is never going to be executed. Newer build
+tools for the web, like Webpack 2 or Rollup offer this feature which
+relies heavily on the static structure on ES6 modules.{[}6{]} Contrary
+libraries for the browsers are traditionally ``shipped'' as
+\textbf{immediately-invoked functions}. Bootstrapping those closed
+functions is much harder to archive as even dead code is actually
+executed on load.
+
+Web components are necessary static ES6 modules and any capable bundler
+can choose which component is needed at runtime. Even more granular, a
+good web component is composed of static JS blocks which can be strapped
+away too.
+
+A web component is self-contained which means it embodys all needed
+functionality to get it's job done. Therefore it has a much better
+evolution mechanism in the service contracts. Changing functionality
+won't break other components. A component can progressively enhanced
+which guarantees functionality throughout different versions.
 
@@
 documentation, a component functionality is exposed via the components'
-signature and therefore highly accessible. HTML elements are usually
-controlled via their attributes. A generic HTML service therefore might
-look like this:
-\texttt{\textless{}my-button\ isPrimary\textgreater{}\textless{}/button\textgreater{}}.
+signature which comes in the fashion of an HTML element. For example, a
+button can consume attributes and becomes a primary buttons just by
+assignment.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{<my-botton}\OtherTok{ isPrimary}\KeywordTok{>}\NormalTok{Primary}\KeywordTok{</my-botton>}
+\end{Highlighting}
+\end{Shaded}
+
+A microservice reinforces the \emph{Single Responsibility Principle}
+defined by Robert C. Martin: ``Gather together those things that change
+for the same reason and separate those things that change for different
+reasons.''{[}7{]}
 
@@
 
-\subsection{Products not Projects}\label{products-not-projects}
+\begin{quote}
+``organizations which design systems \ldots{} are constrained to produce
+designs which are copies of the communication~structures~of these
+organizations''. {[}8{]}
+\end{quote}
+
+Emphasizing the human factor in microservices is a key feature.
+Microservices are a product of real-world usage.{[}3, p. 1{]} Instead of
+splitting team structures along the technology stack (UI Experts
+-\textgreater{} Middleware -\textgreater{} Database) a microservice
+approach model teams around \textbf{business capabilities}.{[}4{]}
+Consequently every team is capable of planning, designing, implementing,
+testing and maintaining their very own microservice. Along the
+technology stack every member gains high competence about the
+functionality of the service.
+
+Real-world domains tend to be complex and multifaceted. To unfold their
+complexity, domains can subdivided into bounded contexts.{[}3, p. 31{]}
+For example, dealing with customers is a business domain. Customers
+again have different contexts depending on their demands. One context
+can be sales, another context could be support. Every context makes
+different assumptions about the underlying model. Each bounded context
+draws an explicit interface where it decides what models to share with
+other contexts.{[}3, p. 30{]} Each context can derived into multiple
+microservices or, talking about \emph{browsernative microservices},
+interfaces to the customers.
+
+By assigning service responsibility to a team, the so called
+\textbf{Definition of Done} (DoD) shifts from ``accomplishing projects''
+to ``accomplishing products''. This new paradigm not only changes the
+administrative overhead like budgeting or resource allocation. It
+creates a kind of responsibility connection from the team to the
+service. Expectedly, those teams are more motivated within their very
+own service and exhibit a more sophisticated iteration time.{[}4{]}
+
+For many companies working in the spheres of the internet the
+client-side is highly important for their business. In fact, business
+goals and capabilities can be derived from the front-end needs. The
+state of the web is not only a story of numerous artifacts, it is also a
+story of an highly fragmented market along devices, operating systems,
+differing sizes and functionalities. Different devices again have
+different assumptions about the technology stack. Splitting teams along
+the stack results in an slow paced back and forth negotiation for every
+change to be made. As browser technologies, design guidelines and
+devices change in a fast paced manner it makes absolutely sense to shift
+responsibility towards the teams altogether. A \emph{browsernative
+microservice} embraces commitment over its whole lifecycle.
 
@@
 
-\subsection{Decentralized Governance}\label{decentralized-governance}
-
-\subsection{Decentralized Data
-Management}\label{decentralized-data-management}
+Microservices for the browsers aim for side-effects like changing the UI
+given to input parameters or emitting browser events. An isolated
+microservice won't make too much sense after all. To ensure
+communication between services, flexible yet powerful communication
+channels must be established.
+
+\emph{Smart endpoints and dumb pipes} is coined to the approach of
+designing communication mostly decoupled and as cohesive as
+possible.{[}4{]} In the analogy to the real world a message channel
+should look like sending a letter: Two smart endpoints (sender and
+receiver) and a mostly unified ``dumb'' letter format and channel. Each
+endpoint owns their specific domain logic. In the browsernative context
+we already emphasized the role of presentational components and
+containers. The ``smartness'' of presentational components is altering
+the UI and reacting on users input. The logic of a container follows the
+idea of an filter in a Unix sense - receiving a request, applying logic
+as appropriate and producing a response.{[}4{]}
+
+Cited earlier in this paper microservices often rely on simple HTTP
+request-response with resource API's and lightweight messaging as
+communication protocols.{[}4{]} Those technologies highly accessible and
+widely adopted and most probably usable for a browsernative service. In
+the implementation part of this paper a reader will explore the
+combination of REST protocols and a lightweight browser message bus in
+action.
+
+\subsection{Decentralization}\label{decentralization}
+
+Decentralization is a meta concept of microservices. As written earlier
+decentralization is part of the organizational structure by assigning
+service responsibility solely to the team. Or by gluing independent
+service bricks to a whole distributed system talking over the network
+protocols.
+
+This distributed nature allows teams to create their own technology
+stack, tools and services designed in the spirit of language- and
+platform independence - and share their knowledge with other
+parties.{[}4{]} In the recent years many big companies like Facebook,
+Google, Netflix and others followed that spirit and published their
+ideas and implementations open source. The previously mentioned ReactJS
+for example is a brainchild of Facebook. In fact, many tools and
+techniques are byproduct of vital interaction of concrete domain
+problems and their implementations.
+
+The spirit of freedom can't be applied to \emph{browsernative
+microservices} as the browser and its underlying DOM will be the
+limitation factor. Still there are some decentralization aspects around
+creating microservices for the browser worth to mention.
+
+JS is the widely accepted language of the web. Nevertheless, a
+microservice engineering team might choose another language for
+different reasons. Transpiling languages to JS as target language isn't
+exotic anymore. Languages like TypeScript, ClojureScript or PureScript
+compile to JS even exclusively. Once web components hit a critical mass
+there will be most likely some library support or foreign function
+interfaces towards ES6 modules (which are mandatory for the new
+specifications). With the rise of WebAssembly, a new low-level
+programming language for the browser, the determination on JS will
+hypothetical deteriorate and new quasi native languages for the web
+might gain traction.
+
+Another more real life decentralization aspect derives from the easiness
+of deployment browsernative microservices which makes them perfectly for
+open source. They are small, leightweight and can be tested live in the
+browser without any additional tooling.
+\href{https://webcomponents.org}{Webcomponents.org} for example is a
+registry for many ready-to-use components where even Google shares a lot
+of their material design elements.
+
+Last bot not least, decentralized governance in a web native context
+will be achieved just because of the encapsulation of the service.
+Contrary to the HTML, CSS and JS split model, the new autonomous service
+exhibits some sort of decentralization in creating the service.
 
@@
 
-The concept of microservices incooperates not only a technical
-perspective. Microservice patterns are a product of real-world
-usage.{[}3, p. 1{]} In a real world we typical have to deal with the so
-called \emph{Conway's Law}:
-
-\begin{quote}
-``organizations which design systems \ldots{} are constrained to produce
-designs which are copies of the communication~structures~of these
-organizations''. {[}6{]}
-\end{quote}
-
 Following this logic any company, whether it is web-related or not,
@@
 vast majority of companies developing for the web which are grouped
-around tasks.{[}7{]} A very common structure is formed by the technology
+around tasks.{[}9{]} A very common structure is formed by the technology
 stack (UX Designers, Frontend- \& Backend Developers) or by separating
@@
 essentially make it possible to create a scoped component but not caring
-too much on it's distribution{[}8{]}. Some people prefer the three specs
-{[}9{]}, but the majority advocating the four specs variant, which is
-listed on \sout{the official}
+too much on it's distribution{[}10{]}. Some people prefer the three
+specs {[}11{]}, but the majority advocating the four specs variant,
+which is listed on \sout{the official}
 \href{http://webcomponents.org}{webcomponents.org} website. For the
@@
 entire DOM API and means any properties/methods that you add to the
-class become part of the element's DOM interface.''{[}10{]} Like any
+class become part of the element's DOM interface.''{[}12{]} Like any
 other \emph{ES6 class}, \emph{custom element}s can be sub-classed
@@
   This block of code contains setup code, such as fetching resources or
-  rendering elements according to attributes.{[}10{]} For performance
+  rendering elements according to attributes.{[}12{]} For performance
   issues it's highly preferable to put much code in here.
@@
 Native DOM properties will reflect their values between HTML and JS
-automatically.{[}11, Para. 2.6.1{]}
+automatically.{[}13, Para. 2.6.1{]}
 
@@
 
-{[}{[} CSS Selectors {]}{]}
+\begin{itemize}
+\tightlist
+\item
+  ::slotted(selector)\\
+  Applies to distributed nodes and repaints them after distribution.
+  \texttt{Slotted} won't override outsides styles but can complement
+  them with unset style rules.
+\item
+  :host\\
+  The host property will add styles or change inherited ones inside
+  shadow DOM. Using \texttt{all:\ initial;} will ensure browser defaults
+  only.
+\item
+  :host(condition)\\
+  Like the previous one this node will style the shadow DOM but this
+  time based on attributes/conditions assigned to the hosting node.
+\item
+  :host-context(condition)\\
+  Like the previous one this node will style the shadow DOM but will
+  look after context set at the host node or even at the host ancestor.
+\end{itemize}
 
@@
 \textbf{context-aware custom elements}. A possible use case would be
-``theming'' a component (example taken from {[}12{]}):
+``theming'' a component (example taken from {[}14{]}):
 
@@
 universally. For a more in-depth guide, Google Engineer Eric Bidelman
-wroten a great primer on \emph{shadow DOM}{[}12{]}.
+wroten a great primer on \emph{shadow DOM}{[}14{]}.
 
@@
 amounts of API data and manipulating the DOM, runs on the same
-thread{[}13{]}. Putting a lot of work to the single main thread can slow
+thread{[}15{]}. Putting a lot of work to the single main thread can slow
 down the web service significantly. From time to time scripts can block
@@
 
+\hypertarget{ref-Abramov2015}{}
+{[}5{]} D. Abramov, ``Presentational and Container Components --
+Medium.'' 2015 {[}Online{]}. Available:
+\url{https://medium.com/@dan_abramov/smart-and-dumb-components-7ca2f9a7c7d0}.
+{[}Accessed: 01-Dec-2016{]}
+
+\hypertarget{ref-Rauschmayer2015}{}
+{[}6{]} D. A. Rauschmayer, ``Tree-shaking with webpack 2 and babel 6.''
+2015 {[}Online{]}. Available:
+\url{http://www.2ality.com/2015/12/webpack-tree-shaking.html}
+
 \hypertarget{ref-Martin}{}
-{[}5{]} R. C. Martin, ``The single responsibility principle.''
+{[}7{]} R. C. Martin, ``The single responsibility principle.''
 {[}Online{]}. Available:
@@
 \hypertarget{ref-Conway1968}{}
-{[}6{]} M. E. Conway, ``How do committees invent?'' 1968 {[}Online{]}.
+{[}8{]} M. E. Conway, ``How do committees invent?'' 1968 {[}Online{]}.
 Available: \url{http://www.melconway.com/Home/Committees_Paper.html}
@@
 \hypertarget{ref-Issa2016}{}
-{[}7{]} B. Issa, ``The way of the web.'' Polymer Summit 2016, Oct-2016
+{[}9{]} B. Issa, ``The way of the web.'' Polymer Summit 2016, Oct-2016
 {[}Online{]}. Available:
@@
 \hypertarget{ref-Buchner2016}{}
-{[}8{]} D. Buchner, ``Demythstifying web components,'' 2016
+{[}10{]} D. Buchner, ``Demythstifying web components,'' 2016
 {[}Online{]}. Available:
@@
 \hypertarget{ref-vanKesteren2014}{}
-{[}9{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
+{[}11{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
 {[}Online{]}. Available:
@@
 \hypertarget{ref-Bidelman2016}{}
-{[}10{]} E. Bidelman, ``Custom elements v1: reusable web components.''
+{[}12{]} E. Bidelman, ``Custom elements v1: reusable web components.''
 2016 {[}Online{]}. Available:
@@
 \hypertarget{ref-HTML}{}
-{[}11{]} \emph{HTML living standard --- last updated 11 january 2017}.
+{[}13{]} \emph{HTML living standard --- last updated 11 january 2017}.
 {[}Online{]}. Available: \url{https://html.spec.whatwg.org/multipage/}
@@
 \hypertarget{ref-Bidelman2016shadow}{}
-{[}12{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
+{[}14{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
 2016 {[}Online{]}. Available:
@@
 \hypertarget{ref-Bidelman2010}{}
-{[}13{]} E. Bidelman, ``The basics of web workers.'' 2010 {[}Online{]}.
+{[}15{]} E. Bidelman, ``The basics of web workers.'' 2010 {[}Online{]}.
 Available: \url{https://www.html5rocks.com/en/tutorials/workers/basics/}

### 2017-02-07 paper.tex
pdf/tex build

@@
 of a humans life, simplicity aligns perception with our mental
-capacities. This paper should be read under the impression of
-simplicity.
+capacities.
 
@@
 easy, but while easy is more of a relative nature, simple can be laid
-out as a objective manner - a goal software development should pursuit
-and praise.
+out as a objective manner and therefore universally applicable.
 
@@
 direction. Layers of new abstractions wrestling with old legacy
-abstractions and mutating objects will get out of control. Subtle bugs
-start to creep in. Eventually the small piece of software ends up in a
-highly complected monolith which will determine future design decisions
-to a painful degree. Any company will be highly determined by the
-current state trying to ``keep the lights on''.
-
-On the other hand, any truly agile system architecture is laid out in a
-fine-grained manner. As Rich Hickey argues, design decisions should be
-made under the \textbf{impression of extending, substitution, moving,
-combining and repurposing}. The ability to reason about the program at
-any given time is crucial for future decisions and implementations.
-Recalling again the unidimensional nature of simplicity.
+abstractions and mutation becomes untraceable. Subtle bugs start to
+creep in. Eventually the small piece of software may end up in a highly
+complected monolith which will determine future design decisions to a
+painful degree. Future strategies of the company/organization will be
+highly determined by the current state in the need of ``keeping the
+lights on''.
+
+On the other side of this dystopian example, the truly agile system
+architecture is laid out in a fine-grained manner. Software becomes
+pluggable, mutations traceable and modules interchangeable. As Rich
+Hickey argues, design decisions should be made under the
+\textbf{impression of extending, substitution, moving, combining and
+repurposing}. The ability to reason about the program at any given time
+is crucial for future decisions and implementations. Recalling again the
+unidimensional nature of simplicity.
 
@@
 In an non-deterministic runtime environment, encapsulation and
-modularization is a typical pattern to make complexity manageable and to
+modularization is a typical pattern to make complexity manageable and
 accommodate future uncertainty.{[}2, p. 1{]} Today many websites tend
@@
 \textbf{complexe UI machine}. Under the current situation only
-additional layers of abstraction can handle those complexity.
+additional layers of abstraction can handle complexity.
 
 In the recent years many \textbf{frameworks}, libraries and
-methodologies tried to tackle the global nature of the DOM by scoping
-assets and design rules into maintainable components. While the DOM
-can't be scoped, JS can. Therefore many of those frameworks, like
-ReactJS, AngularJS or VueJS just to name a few, ditched the old rule of
-separated HTML, CSS and JS in favor of an additional layer of abstracted
-JS components (containing content, markup and styling). Quiet often
-those frameworks mimic a MVC pattern on top of the browser engine which
-is a reasonable simple design pattern to build graphical user
-interfaces. This paper will apply the same pattern for building a
-browsernative microservice.While frameworks are a valid approach to
-build scalable web applications they remain highly opinionated, embody
-inherent complexity themselves and can change and break over time.
-Furthermore, they inflate the code with libraries. A standardized way
-for creating complex UI interfaces painlessly requires new build-in
-browser capabilities.
+methodologies approached the global nature of the DOM by scoping assets
+and design rules into maintainable components. While the DOM can't be
+scoped, JS can. Therefore many frameworks, like ReactJS, AngularJS or
+VueJS just to name a few, ditched the old rule of separated HTML, CSS
+and JS in favor of an additional layer of abstracted JS components
+(containing content, markup and styling). Quiet often those frameworks
+mimic a MVC pattern on top of the browser engine which is a reasonable
+simple design pattern to build graphical user interfaces. While
+frameworks are a valid approach to build scalable web applications they
+remain highly opinionated, embody inherent complexity themselves and can
+change and break over time. Furthermore, code inflation in front-end is
+a crucial point for performance and third-party libraries are no
+exception on that. A standardized way for creating complex UI interfaces
+painlessly requires new build-in browser capabilities.
 
@@
 comprehensive guide called \emph{Building Microservices} and the work of
-Fowler and Lewis. The purpose of this section is to match those ideas
+Fowler and Lewis. The purpose of this section is to match their ideas
 against the browser platform as targeted runtime.
@@
 services, each running in its own process and communicating with
-lightweight mechanisms, often an HTTP resource API''{[}4{]}
+lightweight mechanisms, often an HTTP resource API.''{[}4{]}
 Microservices incorporate many other ideas, like \emph{domain-driven
-design} where we try to represent the real world in our code.{[}3, p.
-2{]} Or making use of \emph{continuous delivery} for pushing software
-rapidly through \emph{automated deployment} mechanisms in
-production.{[}4{]} And, last but not least, microservices utilizes the
-idea of small teams with a lot of product knowledge working mostly
+design} where it pursues the incorporation of the real world in our
+code.{[}3, p. 2{]} Or making use of \emph{continuous delivery} for
+pushing software rapidly through \emph{automated deployment} mechanisms
+into production.{[}4{]} And, last but not least, microservices utilizes
+the idea of small teams with a lot of product knowledge working mostly
 autonomous on their very own service with their very own set of tools
@@
 ``A~\textbf{component}~is a unit of software that is independently
-replaceable and upgradeable.''{[}4{]} Compared to libraries, a component
-service offers multiple advantages. A library is only loosely coupled to
-the implementation and therefore hard to track in functionality.
-Changing a library often requires an unforeseen amount of time fixing
-implementations. Therefore in the field we often see multiple versions
-of libraries deployed simultaneous (like with JQuery) which adds a huge
-amount of overhead on the page load. Eventually, dead code elimination
-won't work in highly coupled libraries. Components instead incorporate a
-much better evolution mechanism in the service contracts. Changing
-functionality won't break other components. A component can
-progressively enhanced which guarantees functionality throughout
-different versions.
+replaceable and upgradeable.'' {[}4{]} Components are the building
+blocks of the microservice but in fact they even share a lot of
+similarities just on a smaller scale. One of the similarity is the
+\emph{loose coupling principle}: changing and deploying one service
+shouldn't result in changing other parts of the system.{[}3, p. 30{]}.
+Same goes for the component which should know as little as it needs
+about the service to fulfil it's job.
+
+Any reader of the paper coming from the ReactJS / Redux world this
+concept of components might look familiar. Dan Abramov, the creator of
+Redux, once defined a simple dichotomous pattern for creating UI
+components. In his believe, there are \textbf{presentational components}
+only related with the concern about \emph{how things look}. This
+component literally doesn't know anything about the service in question
+which makes the component highly flexible and reusable. They are
+controlled solely from the outside, receiving data and dispatching
+unbiased events on user interaction.{[}5{]}
+
+On the other hand Dan Abramov outlines components which he calls
+\textbf{containers}. A container component is concerned with \emph{how
+things work}.{[}5{]} Containers acts as invisible wrappers around
+presentational components. Their job is to fetch data from child nodes,
+aggregating events, interacting with the model and push state back to
+the presentational components. In contrast to their presentational
+counterparts, containers can't life on its own without children.
+Conceptually, they show another important principle of components and
+microservices in general: \emph{high cohesion principle}.
+
+Whether designing a microservice or it's components we want related
+behavior sit together, and unrelated behavior to sit elsewhere.{[}3, p.
+30{]} Code quality can be measured just by counting the places changes
+in the code need to be made on changing behavior. In an arbitrary MVC
+system a \texttt{button} might be inserted in the VIEW, the CONTROLLER
+needs some adjustment and maybe the the MODEL, too. Three places for
+adjustment is a reasonable easy task for the brain, everything more
+violates the concept of simplicity.
+
+Using libraries in web development is common sense. But compared to
+libraries, a component service offers multiple advantages. A library is
+only loosely coupled to the implementation and therefore hard to track
+in functionality. Changing a library may result in an unforeseen amount
+of time fixing implementations. It is not unusual to see websites
+embodying different versions of the same library (like with JQuery).
+Another issue with libraries is dead code elimination which means the
+process of removing code that is never going to be executed. Newer build
+tools for the web, like Webpack 2 or Rollup offer this feature which
+relies heavily on the static structure on ES6 modules.{[}6{]} Contrary
+libraries for the browsers are traditionally ``shipped'' as
+\textbf{immediately-invoked functions}. Bootstrapping those closed
+functions is much harder to archive as even dead code is actually
+executed on load.
+
+Web components are necessary static ES6 modules and any capable bundler
+can choose which component is needed at runtime. Even more granular, a
+good web component is composed of static JS blocks which can be strapped
+away too.
+
+A web component is self-contained which means it embodys all needed
+functionality to get it's job done. Therefore it has a much better
+evolution mechanism in the service contracts. Changing functionality
+won't break other components. A component can progressively enhanced
+which guarantees functionality throughout different versions.
 
@@
 documentation, a component functionality is exposed via the components'
-signature and therefore highly accessible. HTML elements are explicitly
-controlled via their attributes and the new HTML service component are
-no exception to that.
+signature which comes in the fashion of an HTML element. For example, a
+button can consume attributes and becomes a primary buttons just by
+assignment.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{<my-botton}\OtherTok{ isPrimary}\KeywordTok{>}\NormalTok{Primary}\KeywordTok{</my-botton>}
+\end{Highlighting}
+\end{Shaded}
 
@@
 for the same reason and separate those things that change for different
-reasons.''{[}5{]}
+reasons.''{[}7{]}
 
@@
 designs which are copies of the communication~structures~of these
-organizations''. {[}6{]}
+organizations''. {[}8{]}
 \end{quote}
 
-Microservice patterns are a product of real-world usage.{[}3, p. 1{]}
-Instead of splitting teams along the technology stack (UI Experts
+Emphasizing the human factor in microservices is a key feature.
+Microservices are a product of real-world usage.{[}3, p. 1{]} Instead of
+splitting team structures along the technology stack (UI Experts
 -\textgreater{} Middleware -\textgreater{} Database) a microservice
@@
 Consequently every team is capable of planning, designing, implementing,
-testing and maintaining their very own microservice. Across the stack
-every member gains high competence about the functionality of the
-service.
-
-By assigning service responsibility a team, the so called
+testing and maintaining their very own microservice. Along the
+technology stack every member gains high competence about the
+functionality of the service.
+
+Real-world domains tend to be complex and multifaceted. To unfold their
+complexity, domains can subdivided into bounded contexts.{[}3, p. 31{]}
+For example, dealing with customers is a business domain. Customers
+again have different contexts depending on their demands. One context
+can be sales, another context could be support. Every context makes
+different assumptions about the underlying model. Each bounded context
+draws an explicit interface where it decides what models to share with
+other contexts.{[}3, p. 30{]} Each context can derived into multiple
+microservices or, talking about \emph{browsernative microservices},
+interfaces to the customers.
+
+By assigning service responsibility to a team, the so called
 \textbf{Definition of Done} (DoD) shifts from ``accomplishing projects''
@@
 
-For many companies working in the spheres of the internet, the
-client-side is the most crucial area of business success. Todays
-client-side includes many devices of differing sizes and
-functionalities. As browser technologies, design guidelines and devices
-change in a fast paced manner it makes absolutely sense to shift
+For many companies working in the spheres of the internet the
+client-side is highly important for their business. In fact, business
+goals and capabilities can be derived from the front-end needs. The
+state of the web is not only a story of numerous artifacts, it is also a
+story of an highly fragmented market along devices, operating systems,
+differing sizes and functionalities. Different devices again have
+different assumptions about the technology stack. Splitting teams along
+the stack results in an slow paced back and forth negotiation for every
+change to be made. As browser technologies, design guidelines and
+devices change in a fast paced manner it makes absolutely sense to shift
 responsibility towards the teams altogether. A \emph{browsernative
-microservice} needs commitment over its whole lifecycle.
+microservice} embraces commitment over its whole lifecycle.
 
@@
 
-Probably most microservices build for the browser are impure in
-functionality. They aim for side-effects like changing the UI given to
-input parameters or emitting browser events. Therefore a web developer
-must design communication channels for those microservices keeping in
-mind the nested nature of the DOM.
+Microservices for the browsers aim for side-effects like changing the UI
+given to input parameters or emitting browser events. An isolated
+microservice won't make too much sense after all. To ensure
+communication between services, flexible yet powerful communication
+channels must be established.
 
 \emph{Smart endpoints and dumb pipes} is coined to the approach of
-designing communication mostly decoupled and as cohesive as possible.
-This essentially removes most of the error-prone middleware altogether.
-In the analogy of the real world a message channel should look like
-sending a letter: Only two endpoints (sender and receiver) connoted as
-``smart'' and the letter itself looks mostly dumb (but unified) from the
-outside. Only the receiver is able to process the content message.
+designing communication mostly decoupled and as cohesive as
+possible.{[}4{]} In the analogy to the real world a message channel
+should look like sending a letter: Two smart endpoints (sender and
+receiver) and a mostly unified ``dumb'' letter format and channel. Each
+endpoint owns their specific domain logic. In the browsernative context
+we already emphasized the role of presentational components and
+containers. The ``smartness'' of presentational components is altering
+the UI and reacting on users input. The logic of a container follows the
+idea of an filter in a Unix sense - receiving a request, applying logic
+as appropriate and producing a response.{[}4{]}
 
@@
 request-response with resource API's and lightweight messaging as
-communication protocols.{[}4{]} In the later parts of this paper the
-reader will see a generic \emph{browsernative microservice} split in the
-fashion of \textbf{MVC}. In the \textbf{VIEW} the microservice will rely
-on browser events and in-memory calls while the \textbf{CONTROLLER} have
-to deal with other resources over the fetch API.
-
-\subsection{Decentralized Governance}\label{decentralized-governance}
-
-\subsection{Decentralized Data
-Management}\label{decentralized-data-management}
+communication protocols.{[}4{]} Those technologies highly accessible and
+widely adopted and most probably usable for a browsernative service. In
+the implementation part of this paper a reader will explore the
+combination of REST protocols and a lightweight browser message bus in
+action.
+
+\subsection{Decentralization}\label{decentralization}
+
+Decentralization is a meta concept of microservices. As written earlier
+decentralization is part of the organizational structure by assigning
+service responsibility solely to the team. Or by gluing independent
+service bricks to a whole distributed system talking over the network
+protocols.
+
+This distributed nature allows teams to create their own technology
+stack, tools and services designed in the spirit of language- and
+platform independence - and share their knowledge with other
+parties.{[}4{]} In the recent years many big companies like Facebook,
+Google, Netflix and others followed that spirit and published their
+ideas and implementations open source. The previously mentioned ReactJS
+for example is a brainchild of Facebook. In fact, many tools and
+techniques are byproduct of vital interaction of concrete domain
+problems and their implementations.
+
+The spirit of freedom can't be applied to \emph{browsernative
+microservices} as the browser and its underlying DOM will be the
+limitation factor. Still there are some decentralization aspects around
+creating microservices for the browser worth to mention.
+
+JS is the widely accepted language of the web. Nevertheless, a
+microservice engineering team might choose another language for
+different reasons. Transpiling languages to JS as target language isn't
+exotic anymore. Languages like TypeScript, ClojureScript or PureScript
+compile to JS even exclusively. Once web components hit a critical mass
+there will be most likely some library support or foreign function
+interfaces towards ES6 modules (which are mandatory for the new
+specifications). With the rise of WebAssembly, a new low-level
+programming language for the browser, the determination on JS will
+hypothetical deteriorate and new quasi native languages for the web
+might gain traction.
+
+Another more real life decentralization aspect derives from the easiness
+of deployment browsernative microservices which makes them perfectly for
+open source. They are small, leightweight and can be tested live in the
+browser without any additional tooling.
+\href{https://webcomponents.org}{Webcomponents.org} for example is a
+registry for many ready-to-use components where even Google shares a lot
+of their material design elements.
+
+Last bot not least, decentralized governance in a web native context
+will be achieved just because of the encapsulation of the service.
+Contrary to the HTML, CSS and JS split model, the new autonomous service
+exhibits some sort of decentralization in creating the service.
 
@@
 
-The concept of microservices incooperates not only a technical
-perspective.
-
 Following this logic any company, whether it is web-related or not,

### 2017-02-07 build/paper.tex
pdf/tex build

@@
 Whether designing a microservice or it's components we want related
-behavior sit together, and unrelated behavior to sit
-elsewhere.{\textbf{???}} Code quality can be measured just by counting
-the places changes in the code need to be made on changing behavior. In
-an arbitrary MVC system a \texttt{button} might be inserted in the VIEW,
-the CONTROLLER needs some adjustment and maybe the the MODEL, too. Three
-places for adjustment is a reasonable easy task for the brain,
-everything more violates the concept of simplicity.
+behavior sit together, and unrelated behavior to sit elsewhere.{[}3, p.
+30{]} High cohesion can be perceived as the \emph{Single Responsibility
+Principle} defined by Robert C. Martin: ``Gather together those things
+that change for the same reason and separate those things that change
+for different reasons.''{[}6{]} In a very quick and dirty code quality
+analysis, the quality can be measured just by counting the places
+changes in the code occur. In an arbitrary MVC system a \texttt{button}
+might be inserted in the VIEW, the CONTROLLER needs some adjustment and
+maybe the the MODEL, too. Three places for adjustment is a reasonable
+easy task for the brain, everything more violates the concept of
+simplicity.
 
@@
 tools for the web, like Webpack 2 or Rollup offer this feature which
-relies heavily on the static structure on ES6 modules.{[}6{]} Contrary
+relies heavily on the static structure on ES6 modules.{[}7{]} Contrary
 libraries for the browsers are traditionally ``shipped'' as
@@
 \end{Shaded}
-
-A microservice reinforces the \emph{Single Responsibility Principle}
-defined by Robert C. Martin: ``Gather together those things that change
-for the same reason and separate those things that change for different
-reasons.''{[}7{]}
 
@@
 Another more real life decentralization aspect derives from the easiness
-of deployment browsernative microservices which makes them perfectly for
-open source. They are small, leightweight and can be tested live in the
-browser without any additional tooling.
+of deployment of \emph{browsernative microservices} which makes them a
+good fit for open source. They are small, leightweight and can be tested
+live in the browser without any additional tooling.
 \href{https://webcomponents.org}{Webcomponents.org} for example is a
-registry for many ready-to-use components where even Google shares a lot
-of their material design elements.
-
-Last bot not least, decentralized governance in a web native context
-will be achieved just because of the encapsulation of the service.
-Contrary to the HTML, CSS and JS split model, the new autonomous service
-exhibits some sort of decentralization in creating the service.
-
-\subsection{Infrastructure Automation}\label{infrastructure-automation}
-
-\subsection{Design for failure}\label{design-for-failure}
-
-\subsection{Evolutionary Design}\label{evolutionary-design}
-
-Shift of paradigms / BFF / Platform agnostic / changes in infrastructure
-like APIs Databanks / Deployment
-
-Most obvious ist the gathering of all related code under the umbrellar
-of a single HTML tag. Grouping together HTML, JS and CSS Code in a safe,
+registry for ready-to-use components of every scale where even Google
+shares a lot of their material design elements.
+
+Decentralization will be achieved just because of the encapsulation of
+the service. Grouping together HTML, JS and CSS Code in a safe,
 sandboxed environment exposes the possibility to build more cohesive and
@@
 sometimes biased) way to glue related parts together.
+
+\subsection{Infrastructure Automation}\label{infrastructure-automation}
+
+\subsection{Design for failure}\label{design-for-failure}
+
+Chapter about progressive enhancement
+
+\subsection{Evolutionary Design}\label{evolutionary-design}
+
+Shift of paradigms / BFF / Platform agnostic / changes in infrastructure
+like APIs Databanks / Deployment
+
+Most obvious ist the gathering of all related code under the umbrellar
+of a single HTML tag.
 
@@
 
+\section{Anatomy of an browsernative
+microservice}\label{anatomy-of-an-browsernative-microservice}
+
 dichotome Pattern
@@
 
-\section{Progressive Enhancement}\label{progressive-enhancement}
-
-Chapter about progressive enhancement
-
 \hypertarget{refs}{}
@@
 
+\hypertarget{ref-Martin}{}
+{[}6{]} R. C. Martin, ``The single responsibility principle.''
+{[}Online{]}. Available:
+\url{http://programmer.97things.oreilly.com/wiki/index.php/The_Single_Responsibility_Principle}
+
 \hypertarget{ref-Rauschmayer2015}{}
-{[}6{]} D. A. Rauschmayer, ``Tree-shaking with webpack 2 and babel 6.''
+{[}7{]} D. A. Rauschmayer, ``Tree-shaking with webpack 2 and babel 6.''
 2015 {[}Online{]}. Available:
 \url{http://www.2ality.com/2015/12/webpack-tree-shaking.html}
-
-\hypertarget{ref-Martin}{}
-{[}7{]} R. C. Martin, ``The single responsibility principle.''
-{[}Online{]}. Available:
-\url{http://programmer.97things.oreilly.com/wiki/index.php/The_Single_Responsibility_Principle}
 

### 2017-02-07 paper.tex
pdf/tex build

@@
 behavior sit together, and unrelated behavior to sit elsewhere.{[}3, p.
-30{]} Code quality can be measured just by counting the places changes
-in the code need to be made on changing behavior. In an arbitrary MVC
-system a \texttt{button} might be inserted in the VIEW, the CONTROLLER
-needs some adjustment and maybe the the MODEL, too. Three places for
-adjustment is a reasonable easy task for the brain, everything more
-violates the concept of simplicity.
+30{]} High cohesion can be perceived as the \emph{Single Responsibility
+Principle} defined by Robert C. Martin: ``Gather together those things
+that change for the same reason and separate those things that change
+for different reasons.''{[}6{]} In a very quick and dirty code quality
+analysis, the quality can be measured just by counting the places
+changes in the code occur. In an arbitrary MVC system a \texttt{button}
+might be inserted in the VIEW, the CONTROLLER needs some adjustment and
+maybe the the MODEL, too. Three places for adjustment is a reasonable
+easy task for the brain, everything more violates the concept of
+simplicity.
 
@@
 Another more real life decentralization aspect derives from the easiness
-of deployment browsernative microservices which makes them perfectly for
-open source. They are small, leightweight and can be tested live in the
-browser without any additional tooling.
-\href{https://webcomponents.org}{Webcomponents.org} for example is a
-registry for many ready-to-use components where even Google shares a lot
-of their material design elements.
-
-Last bot not least, decentralized governance in a web native context
-will be achieved just because of the encapsulation of the service.
-Contrary to the HTML, CSS and JS split model, the new autonomous service
-exhibits some sort of decentralization in creating the service.
+of deployment in a safe, sandboxed environment. No hidden overheads or
+required build tools. Thats why \emph{browsernative microservices} are
+well suited for sharing and open-sourcing.
+\href{https://webcomponents.org}{Webcomponents.org} is a registry for
+ready-to-use components of every scale where for example Google shares a
+lot of their material design elements.
 
@@
 
+Grouping together HTML, JS and CSS Code in a unified environment exposes
+the possibility to build rapidly and fast. In the typical global nature
+of web development those three pillars are separated. This circumstance
+left the developer switching back and forth between code bases
+developing a tricky (and sometimes biased) way to glue related parts
+together.
+
 \subsection{Design for failure}\label{design-for-failure}
+
+Chapter about progressive enhancement
 
@@
 Most obvious ist the gathering of all related code under the umbrellar
-of a single HTML tag. Grouping together HTML, JS and CSS Code in a safe,
-sandboxed environment exposes the possibility to build more cohesive and
-understandable services. In the typical global nature of web development
-those three pillars are separated. This circumstance left the developer
-switching back and forth between code bases developing a tricky (and
-sometimes biased) way to glue related parts together.
+of a single HTML tag.
 
@@
 the cascading nature of CSS, styles will still ``bleed in'' from
-ancestors to the \emph{light DOM} nodes. Therefore it's strongly
-recommended to begin every \emph{shadow DOM} with a \textbf{\emph{CSS
-reset:} \texttt{:host\ \{all:\ initial;\}}}.
+ancestors to the \emph{light DOM} nodes.
 
@@
 down the web service significantly. From time to time scripts can block
-or fail to whatever reason which leads to a frozen UI on the users side.
-A worker can overcome the bottleneck of the single-threaded nature with
-spawning a new \textbf{background thread}. Most of the browsers work can
-be leveraged to this new thread. Todays web architectures aims to
-leverage an increasing amout of proccessing to the client to avoid
-time-consuming roundtrips especially in mobiles networks.\footnote{Latency
-  numbers: https://gist.github.com/jboner/2841832} While many native
-APIs like \texttt{fetch} work seamlessly in the new thread, a worker has
-no access to the DOM at all.
+or fail for whatever reason which leads to a frozen UI on the users
+side. A worker can overcome the bottleneck of the single-threaded nature
+with spawning new \textbf{background threads}. Todays web architectures
+aims to leverage an increasing amout of processing to the client to
+avoid time-consuming roundtrips especially in mobiles
+networks.\footnote{Latency numbers:
+  https://gist.github.com/jboner/2841832} While many native APIs like
+\texttt{fetch} work seamlessly in the new thread, a worker has no access
+to the DOM at all.
 
@@
 
-Having no access to the DOM can be seen as hinderance,but i
-
-Macroperspektive / Composition
-
-\subsubsection{Assumptions about custom
-elements}\label{assumptions-about-custom-elements}
-
-Apart from the spec'd perspective there is mental model a webdeveloper
-might
-
-Creating and using webcomponents might require a new mental model how do
-design a
-
-containers
+\section{Anatomy of an browsernative
+microservice}\label{anatomy-of-an-browsernative-microservice}
+
+After grasping the principles and some technical background the paper
+should illustrate a practical example to \emph{browsernative
+microservices}. Googles library Polymer is a good start to learn about
+web components. They even offer a command line tool for creating a
+polymer skeleton. One of their most famous proof of concept is the
+so-called \href{https://shop.polymer-project.org/}{Polymer Shop} which
+is a fully-fledged online shop nested within a single custom element
+\texttt{\textless{}shop-app\textgreater{}}. It has some elements for
+routing, managing view, service worker caching, theming, etc. The whole
+shop runs as a single application only fetching and sending resources
+but never reloading altogether. Let's assume we work in a sales
+engineering team of the Polymer Shop and need to rebuild checkout
+process.
+
+A usual checkout has 3 to 4 steps:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Sign in or Sign up
+\item
+  Shipping details
+\item
+  Payment details
+\item
+  Review and place order
+\end{enumerate}
+
+Translated into a raw custom element HTML structure, the microservice
+might light like:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{<shop-checkout>}
+  \KeywordTok{<sign-in></sign-in>}
+  \KeywordTok{<shipping-details></shipping-details>}
+  \KeywordTok{<payment-details></payment-details>}
+  \KeywordTok{<place-order></place-order>}
+\KeywordTok{</shop-checkout>}
+\end{Highlighting}
+\end{Shaded}
+
+Pretty straight forward.
 

### 2017-02-09 paper.tex
pdf/tex build

@@
 
-On the other side of this dystopian example, the truly agile system
-architecture is laid out in a fine-grained manner. Software becomes
-pluggable, mutations traceable and modules interchangeable. As Rich
-Hickey argues, design decisions should be made under the
-\textbf{impression of extending, substitution, moving, combining and
-repurposing}. The ability to reason about the program at any given time
-is crucial for future decisions and implementations. Recalling again the
-unidimensional nature of simplicity.
+On the other side of this dystopian example, the truly modular system
+architecture shines in a approachable simplicity. The whole system is
+divided in pluggable parts, object mutations are traceable and modules
+interchangeable. As Rich Hickey argues, design decisions should be made
+under the \textbf{impression of extending, substitution, moving,
+combining and repurposing}. The ability to reason about the program at
+any given time is crucial for future decisions and implementations.
+Recalling again the unidimensional nature of simplicity.
 
@@
 character or around 125 pages of single-spaced text. Frederic Filloux
-calculated the ratio of real content on different newspaper websites and
-came to the conclusion, that only round about 5-6 \% of the transferred
-characters made for human consumption.{[}1{]}
+analyzed the payload on different newspaper websites and came to the
+conclusion, that only round about 5-6 \% of the transferred characters
+made for human consumption.{[}1{]}
 
@@
 text nodes or an ever changing webapp the underlying global nature of
-the DOM tree remains the same. Every additional service added to the
-webapp embodies another fold of complexity.
+the DOM tree remains the same. Every additional piece of software added
+to this global object embodies another fold of complexity.
 
@@
 leapfrogged and many new low-level APIs brought to life. With this new
-APIs at hand a vivid web developer can create scoped and highly reusable
-microservices without additional libraries even directly in the browser
-console.
+APIs at hand a vivid web developer can create robust websites without
+additional libraries.
 
 \emph{Disclaimer:} This paper introduces many new browser build-ins with
-the focus on try and test. As the time of writing, examples can be tried
-frictionless in the consoles of the latest versions of \textbf{Google
-Chrome, Opera and Apple Safari}.\footnote{\href{http://jonrimmer.github.io/are-we-componentized-yet/}{Are
+the focus on try and test. As the time of writing, many examples can be
+tried frictionless in the console of the latest versions of
+\textbf{Google Chrome, Opera and Apple Safari}.\footnote{\href{http://jonrimmer.github.io/are-we-componentized-yet/}{Are
   we componentized yet?}} On Mozilla Firefox technologies work behind a
@@
 
-Calling the case for \emph{Browsernative Microservices} brings up the
-question about the concept of microservices in general. In fact the
-concept of microservices has many facets, stretching beyond disciplines
-and technical boundaries. It lacks a formal standardization but subsumes
-certain ideas emergine from this pattern. As a primary source of truth
-this paper relies on the work of Sam Newman, who has written a
-comprehensive guide called \emph{Building Microservices} and the work of
-Fowler and Lewis. The purpose of this section is to match their ideas
-against the browser platform as targeted runtime.
-
-In a nutshell a microservice is a small, autonomous service that works
+In search of a better, simpler web architecture we might look on already
+established pattern that proofed to fulfill enterprise needs.
+Microservices are a good approach for tearing big monolithic systems
+into fine-grained simple services with explicit defined boundaries. In a
+nutshell a microservice is a small, autonomous service that works
 together with other services seamlessly.{[}3, p. 2{]} or with the words
@@
 autonomous on their very own service with their very own set of tools
-and techniques.
+and techniques. As a primary source of truth this paper relies on the
+work of Sam Newman, who has written a comprehensive guide called
+\emph{Building Microservices} and the work of Fowler and Lewis. The
+purpose of this section is to match their ideas against services on the
+browser platform.
 
@@
 replaceable and upgradeable.'' {[}4{]} Components are the building
-blocks of the microservice but in fact they even share a lot of
-similarities just on a smaller scale. One of the similarity is the
+blocks of microservices. While the service is a concrete piece of
+software a component is much more raw and single-minded following the
 \emph{loose coupling principle}: changing and deploying one service
 shouldn't result in changing other parts of the system.{[}3, p. 30{]}.
-Same goes for the component which should know as little as it needs
-about the service to fulfil it's job.
-
-Any reader of the paper coming from the ReactJS / Redux world this
-concept of components might look familiar. Dan Abramov, the creator of
-Redux, once defined a simple dichotomous pattern for creating UI
-components. In his believe, there are \textbf{presentational components}
-only related with the concern about \emph{how things look}. This
-component literally doesn't know anything about the service in question
-which makes the component highly flexible and reusable. They are
-controlled solely from the outside, receiving data and dispatching
-unbiased events on user interaction.{[}5{]}
+Small components and big microservices are similar on their principles.
+
+Any reader of the paper familiar with ReactJS / Redux the concept of
+components isn't anything new. Dan Abramov, the creator of Redux, once
+defined a simple dichotomous pattern for creating UI components. In his
+believe, there are \textbf{presentational components} only related with
+the concern about \emph{how things look}. This component literally
+doesn't know anything about the service in question which makes the
+component highly flexible and reusable. They are controlled solely from
+the outside, receiving data and dispatching unbiased events on user
+interaction.{[}5{]}
 
@@
 maybe the the MODEL, too. Three places for adjustment is a reasonable
-easy task for the brain, everything more violates the concept of
-simplicity.
+easy task for the brain.
 
@@
 libraries for the browsers are traditionally ``shipped'' as
-\textbf{immediately-invoked functions}. Bootstrapping those closed
-functions is much harder to archive as even dead code is actually
-executed on load.
-
-Web components are necessary static ES6 modules and any capable bundler
-can choose which component is needed at runtime. Even more granular, a
-good web component is composed of static JS blocks which can be strapped
-away too.
+\textbf{immediately-invoked functions}. Bootstrapping those ones is much
+harder to archive as dead code may be part of the function itself.
+
+Native web components are necessary static ES6 modules and any capable
+bundler can choose which component is needed at runtime. Even more
+granular, a good web component is composed of static JS blocks which can
+be strapped away too.
 
@@
 Real-world domains tend to be complex and multifaceted. To unfold their
-complexity, domains can subdivided into bounded contexts.{[}3, p. 31{]}
-For example, dealing with customers is a business domain. Customers
+complexity, domains can subdivided into \textbf{bounded contexts}.{[}3,
+p. 31{]} For example, customer service is a business domain. Customers
 again have different contexts depending on their demands. One context
@@
 receiver) and a mostly unified ``dumb'' letter format and channel. Each
-endpoint owns their specific domain logic. In the browsernative context
-we already emphasized the role of presentational components and
-containers. The ``smartness'' of presentational components is altering
-the UI and reacting on users input. The logic of a container follows the
-idea of an filter in a Unix sense - receiving a request, applying logic
-as appropriate and producing a response.{[}4{]}
+endpoint own its specific domain logic. In the browsernative context we
+already emphasized the role of presentational components and containers.
+The ``smartness'' of presentational components is altering the UI and
+reacting on users input. The logic of a container follows the idea of an
+filter in a Unix sense - receiving a request, applying logic as
+appropriate and producing a response.{[}4{]}
 
@@
 
-\subsection{Decentralization}\label{decentralization}
-
-Decentralization is a meta concept of microservices. As written earlier
-decentralization is part of the organizational structure by assigning
-service responsibility solely to the team. Or by gluing independent
-service bricks to a whole distributed system talking over the network
-protocols.
+By focusing the service into clear defined business boundaries, it is
+easier to define a smart API of the service which in our case will be a
+set of attributes for the HTML element bound to the service.
+
+\subsection{Decentralized Governance}\label{decentralized-governance}
+
+Microservices are separate entities and decentralization therefore is a
+meta concept of microservices. Distributed service responsibility owned
+by the team is one aspect. Gluing independent service bricks to a whole
+distributed system talking over the network protocols another one.
 
@@
 microservices} as the browser and its underlying DOM will be the
-limitation factor. Still there are some decentralization aspects around
-creating microservices for the browser worth to mention.
+limitation factor. Talking about the browser, a reader might be tempted
+to narrowly thinking about the VIEW layer only - which is not true
+anymore. In the recent years the browser engines grows to a kind of
+virtual machine: there are connectors to build-in databases,
+multithreading support, an ever-growing JS build-ins and even push
+notifications. So-called
+\emph{\href{https://developers.google.com/web/progressive-web-apps/}{Progressive
+Web Apps}}, a bunch of criteria for building good browserapps, can
+achieve a similar look and feel like native apps.
 
 JS is the widely accepted language of the web. Nevertheless, a
-microservice engineering team might choose another language for
-different reasons. Transpiling languages to JS as target language isn't
-exotic anymore. Languages like TypeScript, ClojureScript or PureScript
-compile to JS even exclusively. Once web components hit a critical mass
-there will be most likely some library support or foreign function
-interfaces towards ES6 modules (which are mandatory for the new
-specifications). With the rise of WebAssembly, a new low-level
-programming language for the browser, the determination on JS will
-hypothetical deteriorate and new quasi native languages for the web
-might gain traction.
+microservice engineering team might choose another language for various
+reasons. Transpiling languages to JS as target language isn't exotic
+anymore. Languages like TypeScript, ClojureScript or PureScript compile
+to JS even exclusively. Once web components hit a critical mass there
+will be most likely some library support or foreign function interfaces
+towards ES6 modules (which are mandatory for the new specifications).
+With the rise of WebAssembly, a new low-level programming language for
+the browser, the determination on JS will hypothetical deteriorate and
+new quasi native languages for the web might gain traction.
 
@@
 
+\subsection{Decentralized Data
+Management}\label{decentralized-data-management}
+
+Data Management in Microservice follows the same modular philosophy like
+the conceptual implementation. As mentioned earlier different bounded
+contexts make different assumptions of the underlying model. A
+\emph{browsernative microservices} takes this thought even further and
+expands it to the variety of devices. Decentralized decisions about
+conceptual models demand for decentralized data storage
+decisions.{[}4{]} Since network roundtrips are costly it is a good
+advice to only query as much data as needed and cache as much as
+possible.
+
+``Microservices prefer letting each service manage its own
+database.''{[}Fowler2014{]} Ben Issa, chief architect of ING Australia
+emphasizes this pragmatism on APIs in a conference talk. At ING the
+frontend demands drive the APIs, APIs are produces automatically and not
+even Issa knows how many APIs exists.{[}9{]} They are using a pattern
+called \textbf{backend for frontends} allowing the team working on a
+given UI also handle the server-side components.{[}3, p. 72{]} The
+perception on backend might change therefore from embodying logic to a
+data-silo only.
+
+To see this pattern in the field a reader might have a look at Facebooks
+\href{http://graphql.org/}{GraphQL} project. GraphQL is a query language
+for the frontend. The backend solely replies on the frontend needs.
+Another well documented \sout{even more powerful} approach in the field
+is Cognitects \href{http://www.datomic.com/}{Datomic}, where parts of
+the database will be reflected to the client. A so-called Transactor
+ensures ACID compliance.
+
+The upcoming microservice example assumes a generic build-in API next to
+the build-in frontend components. Instead of gluing frontend and backend
+together on runtime, the microservice is designed holistically
+containing both ends. To reduce network calls especially for mobile
+devices it is a good advice to cache data via a global Service Worker.
+Revamping offline capabilities even further data can be stored in a
+browser based database like PouchDB. For the sake of simplicity data
+management won't be conducted into depth throughout this paper.
+
 \subsection{Infrastructure Automation}\label{infrastructure-automation}
 
-Grouping together HTML, JS and CSS Code in a unified environment exposes
-the possibility to build rapidly and fast. In the typical global nature
-of web development those three pillars are separated. This circumstance
-left the developer switching back and forth between code bases
-developing a tricky (and sometimes biased) way to glue related parts
-together.
+In the global nature of web development the developer stage couldn't't
+completely decoupled from the production environment. This circumstance
+left developers switching back and forth between files developing tricky
+(and more often biased) ways to glue related parts together. With scoped
+components, code blocks can be developed more sane and conveyed into
+production without headache. Even more, in the scripted environment of
+the browser this can be implemented rapidly and continuous.
+
+Previously mentioned Ben Issa, described the ING workflow pretty simple.
+Every component deserves a own GIT repo containing
+
+\begin{itemize}
+\tightlist
+\item
+  Internationalization conformity (i18n)
+\item
+  Accessibility conformity (a11y)
+\item
+  Tests for the component
+\item
+  Demos of the component
+\item
+  Blueprints to mock the one to one APIs
+\item
+  Docs
+\end{itemize}
+
+Every check-in is handled as release candidate and can be independently
+deployed by a fully automated deployment machinery.{[}9{]} While this
+example is seems opinionated, it gives a sense of a mature component
+build for the web. Due to an exhaustive amount of testing and deployment
+tools for the browser an automated infrastructure shouldn't be a
+problem.
 
@@
 
-Chapter about progressive enhancement
+In theory a microservice is designed with a lot of emphasizes on
+real-time monitoring for both the architectural elements and business
+relevant metrics.{[}4{]} Due to the modular structure weak points can
+occur in the orchestration or the services. Testing and automation is a
+feasible task but failures occur within browsers anyway. For example,
+legacy browsers remain a problem for enhancing websites with new
+technologies.
+
+Regarding the evolution of the web, the ``next billion'' internet users
+most likely using Android, have decent specs mobile phones, use an
+evergreen browser but won't have a reliable internet connection.{[}10{]}
+While \emph{Progressive Enhancement} was once coined on the principle to
+build websites both for Browsers with JS and without, the new
+\emph{Progressive Enhancement} becomes more an ``offline first''
+principle. A \emph{browsernative microservice} therefore not only tries
+to cache data as much as possible, it should also bring in a lot of
+program logic as described in the previous chapters.
+
+Working in a JS heavy infrastructure demands for optimazation to avoid
+unexpected side-effects like the \emph{flash of unstyled content
+(FOUC)}. Googles Polymer propagates the \textbf{PRLP pattern}\footnote{\href{https://www.polymer-project.org/1.0/toolbox/server}{PRLP
+  pattern}}:
+
+\begin{itemize}
+\tightlist
+\item
+  Push critical resources for the initial route
+\item
+  Render initial route
+\item
+  Pre-cache remaining routes
+\item
+  Lazy-load and create remaining routes on demand
+\end{itemize}
+
+Following this pattern a critical resource can detect browser
+functionalities beforehand and switch to a \textbf{polyfill} instead of
+the latest browser optimized version. After the initial paint, critical
+resources like top-level microservices might be loaded asynchrony and
+registered.
 
 \subsection{Evolutionary Design}\label{evolutionary-design}
+
+microservices is a separate entity
+
+deployed as an isolated service
+
+monolithic horizontal scaling is not nessecarry
 
@@
 essentially make it possible to create a scoped component but not caring
-too much on it's distribution{[}10{]}. Some people prefer the three
-specs {[}11{]}, but the majority advocating the four specs variant,
+too much on it's distribution{[}11{]}. Some people prefer the three
+specs {[}12{]}, but the majority advocating the four specs variant,
 which is listed on \sout{the official}
@@
 entire DOM API and means any properties/methods that you add to the
-class become part of the element's DOM interface.''{[}12{]} Like any
+class become part of the element's DOM interface.''{[}13{]} Like any
 other \emph{ES6 class}, \emph{custom element}s can be sub-classed
@@
   This block of code contains setup code, such as fetching resources or
-  rendering elements according to attributes.{[}12{]} For performance
+  rendering elements according to attributes.{[}13{]} For performance
   issues it's highly preferable to put much code in here.
@@
 Native DOM properties will reflect their values between HTML and JS
-automatically.{[}13, Para. 2.6.1{]}
+automatically.{[}14, Para. 2.6.1{]}
 
@@
 \textbf{context-aware custom elements}. A possible use case would be
-``theming'' a component (example taken from {[}14{]}):
+``theming'' a component (example taken from {[}15{]}):
 
@@
 universally. For a more in-depth guide, Google Engineer Eric Bidelman
-wroten a great primer on \emph{shadow DOM}{[}14{]}.
+wroten a great primer on \emph{shadow DOM}{[}15{]}.
 
@@
 amounts of API data and manipulating the DOM, runs on the same
-thread{[}15{]}. Putting a lot of work to the single main thread can slow
+thread{[}16{]}. Putting a lot of work to the single main thread can slow
 down the web service significantly. From time to time scripts can block
@@
 
+http://www.pocketjavascript.com/blog/2015/11/23/introducing-pokedex-org
+
+bare minimum of centralized management of these services
+
+\section{Thinking further}\label{thinking-further}
+
+endgeräte werden besser
+
+multicore
+
+query languages from handheld
+
 dichotome Pattern
@@
 
+\hypertarget{ref-Lawson2016}{}
+{[}10{]} N. Lawson, ``Progressive enhancement isn't dead, but it smells
+funny.'' 2016 {[}Online{]}. Available:
+\url{https://nolanlawson.com/2016/10/13/progressive-enhancement-isnt-dead-but-it-smells-funny/}
+
 \hypertarget{ref-Buchner2016}{}
-{[}10{]} D. Buchner, ``Demythstifying web components,'' 2016
+{[}11{]} D. Buchner, ``Demythstifying web components,'' 2016
 {[}Online{]}. Available:
@@
 \hypertarget{ref-vanKesteren2014}{}
-{[}11{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
+{[}12{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
 {[}Online{]}. Available:
@@
 \hypertarget{ref-Bidelman2016}{}
-{[}12{]} E. Bidelman, ``Custom elements v1: reusable web components.''
+{[}13{]} E. Bidelman, ``Custom elements v1: reusable web components.''
 2016 {[}Online{]}. Available:
@@
 \hypertarget{ref-HTML}{}
-{[}13{]} \emph{HTML living standard --- last updated 11 january 2017}.
+{[}14{]} \emph{HTML living standard --- last updated 11 january 2017}.
 {[}Online{]}. Available: \url{https://html.spec.whatwg.org/multipage/}
@@
 \hypertarget{ref-Bidelman2016shadow}{}
-{[}14{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
+{[}15{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
 2016 {[}Online{]}. Available:
@@
 \hypertarget{ref-Bidelman2010}{}
-{[}15{]} E. Bidelman, ``The basics of web workers.'' 2010 {[}Online{]}.
+{[}16{]} E. Bidelman, ``The basics of web workers.'' 2010 {[}Online{]}.
 Available: \url{https://www.html5rocks.com/en/tutorials/workers/basics/}

### 2017-02-09 build/paper.tex
pdf/tex build

@@
 microservices} as the browser and its underlying DOM will be the
-limitation factor. Still there are some decentralization aspects around
-creating microservices for the browser worth to mention.
+limitation factor. Talking about the browser, a reader might be tempted
+to narrowly thinking about the VIEW layer only - which is not true
+anymore. In the recent years the browser engines grows to a kind of
+virtual machine: there are connectors to build-in databases,
+multithreading support, an ever-growing JS build-ins and even push
+notifications. So-called \emph{Progressive Web Apps\footnote{\href{https://developers.google.com/web/progressive-web-apps/}{Progressive
+  Web Apps}}}, a bunch of criteria for building good browserapps, can
+achieve a similar look and feel like native apps.
 
 JS is the widely accepted language of the web. Nevertheless, a
-microservice engineering team might choose another language for
-different reasons. Transpiling languages to JS as target language isn't
-exotic anymore. Languages like TypeScript, ClojureScript or PureScript
-compile to JS even exclusively. Once web components hit a critical mass
-there will be most likely some library support or foreign function
-interfaces towards ES6 modules (which are mandatory for the new
-specifications). With the rise of WebAssembly, a new low-level
-programming language for the browser, the determination on JS will
-hypothetical deteriorate and new quasi native languages for the web
-might gain traction.
+microservice engineering team might choose another language for various
+reasons. Transpiling languages to JS as target language isn't exotic
+anymore. Languages like TypeScript, ClojureScript or PureScript compile
+to JS even exclusively. Once web components hit a critical mass there
+will be most likely some library support or foreign function interfaces
+towards ES6 modules (which are mandatory for the new specifications).
+With the rise of WebAssembly, a new low-level programming language for
+the browser, the determination on JS will hypothetical deteriorate and
+new quasi native languages for the web might gain traction.
 
 Another more real life decentralization aspect derives from the easiness
-of deployment of \emph{browsernative microservices} which makes them a
-good fit for open source. They are small, leightweight and can be tested
-live in the browser without any additional tooling.
-\href{https://webcomponents.org}{Webcomponents.org} for example is a
-registry for ready-to-use components of every scale where even Google
-shares a lot of their material design elements.
-
-Decentralization will be achieved just because of the encapsulation of
-the service. Grouping together HTML, JS and CSS Code in a safe,
-sandboxed environment exposes the possibility to build more cohesive and
-understandable services. In the typical global nature of web development
-those three pillars are separated. This circumstance left the developer
-switching back and forth between code bases developing a tricky (and
-sometimes biased) way to glue related parts together.
+of deployment in a safe, sandboxed environment. No hidden overheads or
+required build tools. Thats why \emph{browsernative microservices} are
+well suited for sharing and open-sourcing. Webcomponents.org\footnote{\href{https://webcomponents.org}{Webcomponents.org}}
+is a registry for ready-to-use components of every scale where for
+example Google shares a lot of their material design elements.
+
+\subsection{Decentralized Data
+Management}\label{decentralized-data-management}
+
+Data Management in Microservice follows the same modular philosophy like
+the conceptual implementation. As mentioned earlier different bounded
+contexts make different assumptions of the underlying model. A
+\emph{browsernative microservices} takes this thought even further and
+expands it to the variety of devices. Decentralized decisions about
+conceptual models demand for decentralized data storage
+decisions.{[}4{]} Since network roundtrips are costly it is a good
+advice to only query as much data as needed and cache as much as
+possible.
+
+``Microservices prefer letting each service manage its own
+database.''{[}Fowler2014{]} Ben Issa, chief architect of ING Australia
+emphasizes this pragmatism on APIs in a conference talk. At ING the
+frontend demands drive the APIs, APIs are produces automatically and not
+even Issa knows how many APIs exists.{[}9{]} They are using a pattern
+called \textbf{backend for frontends} allowing the team working on a
+given UI also handle the server-side components.{[}3, p. 72{]} The
+perception on backend might change therefore from embodying logic to a
+data-silo only.
+
+To see this pattern in the field a reader might have a look at Facebooks
+GraphQL\footnote{\href{http://graphql.org/}{GraphQL}}. GraphQL is a
+query language for the frontend. The backend solely replies on the
+frontend needs. Another well documented \sout{even more powerful}
+approach in the field is Cognitects Datomic\footnote{\href{http://www.datomic.com/}{Datomic}},
+where parts of the database will be reflected to the client. A so-called
+Transactor ensures ACID compliance.
+
+The upcoming microservice example assumes a generic build-in API next to
+the build-in frontend components. Instead of gluing frontend and backend
+together on runtime, the microservice is designed holistically
+containing both ends. To reduce network calls especially for mobile
+devices it is a good advice to cache data via a global Service Worker.
+Revamping offline capabilities even further data can be stored in a
+browser based database like PouchDB. For the sake of simplicity data
+management won't be conducted into depth throughout this paper.
 
@@
 
+In the global nature of web development the developer stage couldn't't
+completely decoupled from the production environment. This circumstance
+left developers switching back and forth between files developing tricky
+(and more often biased) ways to glue related parts together. With scoped
+components, code blocks can be developed more sane and conveyed into
+production without headache. Even more, in the scripted environment of
+the browser this can be implemented rapidly and continuous.
+
+Previously mentioned Ben Issa, described the ING workflow pretty simple.
+Every component deserves a own GIT repo containing
+
+\begin{itemize}
+\tightlist
+\item
+  Internationalization conformity (i18n)
+\item
+  Accessibility conformity (a11y)
+\item
+  Tests for the component
+\item
+  Demos of the component
+\item
+  Blueprints to mock the one to one APIs
+\item
+  Docs
+\end{itemize}
+
+Every check-in is handled as release candidate and can be independently
+deployed by a fully automated deployment machinery.{[}9{]} While this
+example is seems opinionated, it gives a sense of a mature component
+build for the web. Due to an exhaustive amount of testing and deployment
+tools for the browser an automated infrastructure shouldn't be a
+problem.
+
 \subsection{Design for failure}\label{design-for-failure}
 
-Chapter about progressive enhancement
+In theory a microservice is designed with a lot of emphasizes on
+real-time monitoring for both the architectural elements and business
+relevant metrics.{[}4{]} Due to the modular structure weak points can
+occur in the orchestration or the services. Testing and automation is a
+feasible task but failures occur within browsers anyway. For example,
+legacy browsers remain a problem for enhancing websites with new
+technologies.
+
+Regarding the evolution of the web, the ``next billion'' internet users
+most likely using Android, have decent specs mobile phones, use an
+evergreen browser but won't have a reliable internet connection.{[}10{]}
+While \emph{Progressive Enhancement} was once coined on the principle to
+build websites both for Browsers with JS and without, the new
+\emph{Progressive Enhancement} becomes more an ``offline first''
+principle. A \emph{browsernative microservice} therefore not only tries
+to cache data as much as possible, it should also bring in a lot of
+program logic as described in the previous chapters.
+
+Working in a JS heavy infrastructure demands for optimazation to avoid
+unexpected side-effects like the \emph{flash of unstyled content
+(FOUC)}. Googles Polymer propagates the a general-purpose \textbf{PRLP
+pattern}\footnote{\href{https://www.polymer-project.org/1.0/toolbox/server}{PRLP
+  pattern}}:
+
+\begin{itemize}
+\tightlist
+\item
+  Push critical resources for the initial route
+\item
+  Render initial route
+\item
+  Pre-cache remaining routes
+\item
+  Lazy-load and create remaining routes on demand
+\end{itemize}
+
+Following this pattern a critical resource can detect browser
+functionalities beforehand and switch to a \textbf{polyfill} instead of
+the latest browser optimized version. After the initial paint, critical
+resources like top-level microservices or other app logic can be loaded
+and registered.
 
@@
 
-Shift of paradigms / BFF / Platform agnostic / changes in infrastructure
-like APIs Databanks / Deployment
-
-Most obvious ist the gathering of all related code under the umbrellar
-of a single HTML tag.
-
-Secondly, the sub-standard \emph{custom elements} introduces so called
-lifecycle methods and a getter/setter interface exposing the
-functionality to the developer. Event handling, for example, can be
-registered in place which is much more declarative than assigning event
-listeners from the outside. Of course, this events can be pushed down to
-nested tags, allowing an increasingly granular system design. This
-approach will be explained further in the upcoming sections.
-
-Following this logic any company, whether it is web-related or not,
-should be devided in units grouped around a destinct business service to
-optimise the workflow. Fowler and Lewis outlines this approach as an
-``alignment of business capabilities''{[}4{]} While this kind of
-structure may be true for companies like Google or Amazon, there is a
-vast majority of companies developing for the web which are grouped
-around tasks.{[}9{]} A very common structure is formed by the technology
-stack (UX Designers, Frontend- \& Backend Developers) or by separating
-teams along the product lifecycle (development, testing, deployment).
-
-Advocators from the microservice approach propose a different model.
-best described by . Web components are one (but important) way to tie up
-those diciplines as one component can host a single independent business
-service. Combined with a flexible backend service these components can
-be huge gain over the cumbersome functional organizational approach.
-
-The ideas transcending from the microservice approach offers plenty of
-choices and decisions how to proceed with designing a program or to
-structure a process.
+Using modular web architecture isn't a perceptually new approach.
+
+Angular, React
+
+Interopt
+
+Freedom of choice
 
@@
 amounts of API data and manipulating the DOM, runs on the same
-thread{[}15{]}. Putting a lot of work to the single main thread can slow
+thread{[}16{]}. Putting a lot of work to the single main thread can slow
 down the web service significantly. From time to time scripts can block
-or fail to whatever reason which leads to a frozen UI on the users side.
-A worker can overcome the bottleneck of the single-threaded nature with
-spawning a new \textbf{background thread}. Most of the browsers work can
-be leveraged to this new thread. Todays web architectures aims to
-leverage an increasing amout of proccessing to the client to avoid
-time-consuming roundtrips especially in mobiles networks.\footnote{Latency
-  numbers: https://gist.github.com/jboner/2841832} While many native
-APIs like \texttt{fetch} work seamlessly in the new thread, a worker has
-no access to the DOM at all.
+or fail for whatever reason which leads to a frozen UI on the users
+side. A worker can overcome the bottleneck of the single-threaded nature
+with spawning new \textbf{background threads}. Todays web architectures
+aims to leverage an increasing amout of processing to the client to
+avoid time-consuming roundtrips especially in mobiles
+networks.\footnote{Latency numbers:
+  https://gist.github.com/jboner/2841832} While many native APIs like
+\texttt{fetch} work seamlessly in the new thread, a worker has no access
+to the DOM at all.
 
@@
 
-Having no access to the DOM can be seen as hinderance,but i
-
-Macroperspektive / Composition
-
-\subsubsection{Assumptions about custom
-elements}\label{assumptions-about-custom-elements}
-
-Apart from the spec'd perspective there is mental model a webdeveloper
-might
-
-Creating and using webcomponents might require a new mental model how do
-design a
-
-containers
-
 \section{Anatomy of an browsernative
 microservice}\label{anatomy-of-an-browsernative-microservice}
+
+After grasping the principles and some technical background the paper
+should illustrate a practical example to \emph{browsernative
+microservices}. Googles library Polymer is a good start to learn about
+web components. They even offer a command line tool for creating a
+polymer skeleton. One of their most famous proof of concept is the
+so-called \href{https://shop.polymer-project.org/}{Polymer Shop} which
+is a fully-fledged online shop nested within a single custom element
+\texttt{\textless{}shop-app\textgreater{}}. It has some elements for
+routing, managing view, service worker caching, theming, etc. The whole
+shop runs as a single application only fetching and sending resources
+but never reloading altogether. Let's assume we work in a sales
+engineering team of the Polymer Shop and need to rebuild checkout
+process.
+
+A usual checkout has 3 to 4 steps:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Sign in or Sign up
+\item
+  Shipping details
+\item
+  Payment details
+\item
+  Review and place order
+\end{enumerate}
+
+Translated into a raw custom element HTML structure, the microservice
+might light like:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{<shop-checkout>}
+  \KeywordTok{<sign-in></sign-in>}
+  \KeywordTok{<shipping-details></shipping-details>}
+  \KeywordTok{<payment-details></payment-details>}
+  \KeywordTok{<place-order></place-order>}
+\KeywordTok{</shop-checkout>}
+\end{Highlighting}
+\end{Shaded}
+
+Pretty straight forward.
+
+http://www.pocketjavascript.com/blog/2015/11/23/introducing-pokedex-org
+
+bare minimum of centralized management of these services
+
+\section{Thinking further}\label{thinking-further}
+
+endgeräte werden besser
+
+multicore
+
+query languages from handheld
+
+Most obvious ist the gathering of all related code under the umbrellar
+of a single HTML tag.
+
+Secondly, the sub-standard \emph{custom elements} introduces so called
+lifecycle methods and a getter/setter interface exposing the
+functionality to the developer. Event handling, for example, can be
+registered in place which is much more declarative than assigning event
+listeners from the outside. Of course, this events can be pushed down to
+nested tags, allowing an increasingly granular system design. This
+approach will be explained further in the upcoming sections.
+
+Following this logic any company, whether it is web-related or not,
+should be devided in units grouped around a destinct business service to
+optimise the workflow. Fowler and Lewis outlines this approach as an
+``alignment of business capabilities''{[}4{]} While this kind of
+structure may be true for companies like Google or Amazon, there is a
+vast majority of companies developing for the web which are grouped
+around tasks.{[}9{]} A very common structure is formed by the technology
+stack (UX Designers, Frontend- \& Backend Developers) or by separating
+teams along the product lifecycle (development, testing, deployment).
+
+Advocators from the microservice approach propose a different model.
+best described by . Web components are one (but important) way to tie up
+those diciplines as one component can host a single independent business
+service. Combined with a flexible backend service these components can
+be huge gain over the cumbersome functional organizational approach.
 

### 2017-02-13 paper.tex
pdf/tex build

@@
 ``juggle'' a few things at a time. Rich Hickey, the inventor of the
-programming language \emph{Clojure} gave an inspirational keynote on the
-topic of \textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
+programming language Clojure gave an inspirational keynote on the topic
+of \textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
   Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} In every sphere
@@
 multithreading support, an ever-growing JS build-ins and even push
-notifications. So-called
-\emph{\href{https://developers.google.com/web/progressive-web-apps/}{Progressive
-Web Apps}}, a bunch of criteria for building good browserapps, can
+notifications. So-called \emph{Progressive Web Apps\footnote{\href{https://developers.google.com/web/progressive-web-apps/}{Progressive
+  Web Apps}}}, a bunch of criteria for building good browserapps, can
 achieve a similar look and feel like native apps.
@@
 required build tools. Thats why \emph{browsernative microservices} are
-well suited for sharing and open-sourcing.
-\href{https://webcomponents.org}{Webcomponents.org} is a registry for
-ready-to-use components of every scale where for example Google shares a
-lot of their material design elements.
+well suited for sharing and open-sourcing. Webcomponents.org\footnote{\href{https://webcomponents.org}{Webcomponents.org}}
+is a registry for ready-to-use components of every scale where for
+example Google shares a lot of their material design elements.
 
@@
 contexts make different assumptions of the underlying model. A
-\emph{browsernative microservices} takes this thought even further and
-expands it to the variety of devices. Decentralized decisions about
-conceptual models demand for decentralized data storage
-decisions.{[}4{]} Since network roundtrips are costly it is a good
-advice to only query as much data as needed and cache as much as
-possible.
+\emph{browsernative microservices} takes this idea even further and
+expands it to the fragmented world of electronic devices. Decentralized
+decisions about conceptual models demand for decentralized data storage
+decisions.{[}4{]} Todays web architectures aims to leverage an
+increasing amout of processing to the client to avoid time-consuming
+roundtrips especially in mobiles networks.\footnote{Latency numbers:
+  https://gist.github.com/jboner/2841832} Since network roundtrips are
+costly it is a good advice to only query as much data as needed and
+cache as much as possible.
 
@@
 emphasizes this pragmatism on APIs in a conference talk. At ING the
-frontend demands drive the APIs, APIs are produces automatically and not
-even Issa knows how many APIs exists.{[}9{]} They are using a pattern
-called \textbf{backend for frontends} allowing the team working on a
-given UI also handle the server-side components.{[}3, p. 72{]} The
+frontend demands drive the backend APIs, APIs are produces automatically
+and not even Issa knows how many APIs exists.{[}9{]} They are using a
+pattern called \textbf{backend for frontends} allowing the team working
+on a given UI also handle the server-side components.{[}3, p. 72{]} The
 perception on backend might change therefore from embodying logic to a
@@
 To see this pattern in the field a reader might have a look at Facebooks
-\href{http://graphql.org/}{GraphQL} project. GraphQL is a query language
-for the frontend. The backend solely replies on the frontend needs.
-Another well documented \sout{even more powerful} approach in the field
-is Cognitects \href{http://www.datomic.com/}{Datomic}, where parts of
-the database will be reflected to the client. A so-called Transactor
-ensures ACID compliance.
-
-The upcoming microservice example assumes a generic build-in API next to
-the build-in frontend components. Instead of gluing frontend and backend
-together on runtime, the microservice is designed holistically
-containing both ends. To reduce network calls especially for mobile
-devices it is a good advice to cache data via a global Service Worker.
-Revamping offline capabilities even further data can be stored in a
-browser based database like PouchDB. For the sake of simplicity data
-management won't be conducted into depth throughout this paper.
+GraphQL\footnote{\href{http://graphql.org/}{GraphQL}}. GraphQL is a
+query language for the frontend. The backend solely replies on the
+frontend needs. Another well documented \sout{even more powerful}
+approach in the field is Cognitects Datomic\footnote{\href{http://www.datomic.com/}{Datomic}},
+where parts of the database will be reflected to the client. A so-called
+Transactor ensures ACID compliance.
+
+The simplified microservice example later in this paper assumes a
+generic build-in API next to the build-in frontend components. Instead
+of gluing frontend and backend together on runtime, the microservice is
+designed holistically containing both front- and backends. To reduce
+network calls especially for mobile devices it is a good advice to cache
+data via a global Service Worker. Revamping offline capabilities even
+further data can be stored in a browser based database like PouchDB. For
+the sake of simplicity data management won't be explored into depth
+throughout this paper.
 
@@
 
-In the global nature of web development the developer stage couldn't't
+In the global nature of web development the developer stage couldn't
 completely decoupled from the production environment. This circumstance
@@
 Previously mentioned Ben Issa, described the ING workflow pretty simple.
-Every component deserves a own GIT repo containing
+Every component deserves a own \textbf{git repo} containing
 
@@
 relevant metrics.{[}4{]} Due to the modular structure weak points can
-occur in the orchestration or the services. Testing and automation is a
-feasible task but failures occur within browsers anyway. For example,
-legacy browsers remain a problem for enhancing websites with new
-technologies.
+occur in the orchestration of the services. Testing and automation is a
+feasible task but failures may occur in another end users setting
+undetected. Legacy browsers for example remain a problem for enhancing
+websites with new technologies.
 
@@
 
-Working in a JS heavy infrastructure demands for optimazation to avoid
+Working in a JS heavy infrastructure demands for optimization to avoid
 unexpected side-effects like the \emph{flash of unstyled content
-(FOUC)}. Googles Polymer propagates the \textbf{PRLP pattern}\footnote{\href{https://www.polymer-project.org/1.0/toolbox/server}{PRLP
+(FOUC)}. Googles Polymer propagates the a general-purpose \textbf{PRLP
+pattern}\footnote{\href{https://www.polymer-project.org/1.0/toolbox/server}{PRLP
   pattern}}:
@@
 the latest browser optimized version. After the initial paint, critical
-resources like top-level microservices might be loaded asynchrony and
-registered.
+resources like top-level microservices or other app logic can be loaded
+and registered.
 
@@
 
-microservices is a separate entity
-
-deployed as an isolated service
-
-monolithic horizontal scaling is not nessecarry
-
-Shift of paradigms / BFF / Platform agnostic / changes in infrastructure
-like APIs Databanks / Deployment
-
-Most obvious ist the gathering of all related code under the umbrellar
-of a single HTML tag.
-
-Secondly, the sub-standard \emph{custom elements} introduces so called
-lifecycle methods and a getter/setter interface exposing the
-functionality to the developer. Event handling, for example, can be
-registered in place which is much more declarative than assigning event
-listeners from the outside. Of course, this events can be pushed down to
-nested tags, allowing an increasingly granular system design. This
-approach will be explained further in the upcoming sections.
-
-Following this logic any company, whether it is web-related or not,
-should be devided in units grouped around a destinct business service to
-optimise the workflow. Fowler and Lewis outlines this approach as an
-``alignment of business capabilities''{[}4{]} While this kind of
-structure may be true for companies like Google or Amazon, there is a
-vast majority of companies developing for the web which are grouped
-around tasks.{[}9{]} A very common structure is formed by the technology
-stack (UX Designers, Frontend- \& Backend Developers) or by separating
-teams along the product lifecycle (development, testing, deployment).
-
-Advocators from the microservice approach propose a different model.
-best described by . Web components are one (but important) way to tie up
-those diciplines as one component can host a single independent business
-service. Combined with a flexible backend service these components can
-be huge gain over the cumbersome functional organizational approach.
-
-The ideas transcending from the microservice approach offers plenty of
-choices and decisions how to proceed with designing a program or to
-structure a process.
+Microservices tend to become smaller over time. An evolutionary design
+approach puts emphasizes on decomposition and scrapping the service.
+``The key property of a component is the notion of independent
+replacement and upgradeability.''{[}4{]} Therefore we can safely change
+and chop services. Lazy parts of the system which won't change to often
+should be separated from parts undergoing a lot of churn.{[}4{]} Parts
+that needs coupled changes could should be moved together or should be
+even merged.
+
+This flexible approach fits good in the world of browser based
+development. In the last decade we have seen a lot of of changes in the
+way we develop for the web. New technologies like Angular and React are
+handy to use but add some future uncertainty.
+
+In the next section of this paper a lot of standardized technologies
+around the concept of \textbf{Web Components} for the browsers will be
+described. \emph{Browsernative Microservices} should be perceived as
+complementary technology. Contrary to Angular, React and other
+frameworks they have a strong interop with existing systems and can be
+used with them together.
+
+Andrew Rota, developer at Wayfair, came up with the idea to compose
+small web components around a managing React system.{[}11{]} As web
+components are native elements there is no difference to use a native
+\texttt{button} or a native \texttt{custom-button}. With this pattern a
+web developer can make use of the encapsulated advantages of web
+components while still making use of the declarative event management
+from various frameworks. Whatever new framework will be on the rise
+within the next years, this approach allows rapid decomposition and
+reassembling.
 
@@
 essentially make it possible to create a scoped component but not caring
-too much on it's distribution{[}11{]}. Some people prefer the three
-specs {[}12{]}, but the majority advocating the four specs variant,
+too much on it's distribution{[}12{]}. Some people prefer the three
+specs {[}13{]}, but the majority advocating the four specs variant,
 which is listed on \sout{the official}
@@
 entire DOM API and means any properties/methods that you add to the
-class become part of the element's DOM interface.''{[}13{]} Like any
+class become part of the element's DOM interface.''{[}14{]} Like any
 other \emph{ES6 class}, \emph{custom element}s can be sub-classed
@@
   This block of code contains setup code, such as fetching resources or
-  rendering elements according to attributes.{[}13{]} For performance
+  rendering elements according to attributes.{[}14{]} For performance
   issues it's highly preferable to put much code in here.
@@
 attributes to HTML with \texttt{this.setAttributes(...)} respectively.
-On \textbf{insertion time} html attributes might raise their hand with
+On insertion time HTML attributes might raise their hand with
 \texttt{this.hasAttributes(...)} and \texttt{this.getAttributes(...)}.
 Native DOM properties will reflect their values between HTML and JS
-automatically.{[}14, Para. 2.6.1{]}
+automatically.{[}15, Para. 2.6.1{]}
 
@@
 \textbf{mental model} behind \emph{web compontents}. A custom element is
-similar to a named function where attributes treated as \textbf{input
-variables}. In the hierarchical nature of DOM, input can occur either
-top-down via assignments and bottom-up via captured events. The same
-goes true when talking about output. Even though it seems obvious, it
-might be helpful to keep this point in mind.
+similar to a named function where attributes treated as input variables.
+In the hierarchical nature of DOM, input can occur either top-down via
+assignments and bottom-up via captured events. The same goes true when
+talking about output. Even though it seems obvious, it might be helpful
+to keep this point in mind.
 
@@
 One aspect didn't mentioned yet is the possibility of creating
-sub-classes of \textbf{build-in elements} by extending the native
-Interfaces like the \texttt{HTMLButtonElement} interface. While this
-functionality is perfectly spec'd it is strongly rejected by some
-browser vendors.\footnote{https://github.com/w3c/webcomponents/issues/509}
-Most likely the spec will change in future in one or other way on this
-issue and therefore \textbf{customized build-in elements} left out of
-this paper intentionally.
+sub-classes of build-in elements by extending the native Interfaces like
+the \texttt{HTMLButtonElement} interface. While this functionality is
+perfectly spec'd it is strongly rejected by some browser
+vendors.\footnote{https://github.com/w3c/webcomponents/issues/509} Most
+likely the spec will change in future in one or other way on this issue
+and therefore customized build-in elements left out of this paper
+intentionally.
 
@@
 \textbf{context-aware custom elements}. A possible use case would be
-``theming'' a component (example taken from {[}15{]}):
+``theming'' a component (example taken from {[}16{]}):
 
@@
 universally. For a more in-depth guide, Google Engineer Eric Bidelman
-wroten a great primer on \emph{shadow DOM}{[}15{]}.
+wroten a great primer on \emph{shadow DOM}{[}16{]}.
 
@@
 amounts of API data and manipulating the DOM, runs on the same
-thread{[}16{]}. Putting a lot of work to the single main thread can slow
+thread{[}17{]}. Putting a lot of work to the single main thread can slow
 down the web service significantly. From time to time scripts can block
-or fail for whatever reason which leads to a frozen UI on the users
-side. A worker can overcome the bottleneck of the single-threaded nature
-with spawning new \textbf{background threads}. Todays web architectures
-aims to leverage an increasing amout of processing to the client to
-avoid time-consuming roundtrips especially in mobiles
-networks.\footnote{Latency numbers:
-  https://gist.github.com/jboner/2841832} While many native APIs like
-\texttt{fetch} work seamlessly in the new thread, a worker has no access
-to the DOM at all.
+or fail for whatever reason which leads to a frozen or crashed UI. A
+worker can overcome the bottleneck of the single-threaded nature with
+spawning new \textbf{background threads} which allows the UI to stay
+responsive even when computation-heavy tasks should be carried out.
+Effectively, a worker thread makes use of the multi core CPUs most
+devices have nowadays. To grasp the full potential of workers, a reader
+might dive deeper into the Angular 2 architecture, where most of the
+application layer is abstracted from the main rendering thread into
+worker threads.\footnote{\href{https://docs.google.com/document/d/1M9FmT05Q6qpsjgvH1XvCm840yn2eWEg0PMskSQz7k4E}{Angular
+  2 Rendering Architecture}}
 
@@
 
-After grasping the principles and some technical background the paper
-should illustrate a practical example to \emph{browsernative
-microservices}. Googles library Polymer is a good start to learn about
-web components. They even offer a command line tool for creating a
-polymer skeleton. One of their most famous proof of concept is the
-so-called \href{https://shop.polymer-project.org/}{Polymer Shop} which
-is a fully-fledged online shop nested within a single custom element
+After getting some confidence in microservice principles and some
+technical background the paper should illustrate a practical example to
+\emph{browsernative microservices}. Googles library Polymer is a good
+place for learning about web components in depth. They even offer a
+command line tool for creating a polymer skeleton. One of their most
+famous proof of concept is the so-called
+\href{https://shop.polymer-project.org/}{Polymer Shop} which is a
+fully-fledged online shop nested within a single custom root element
 \texttt{\textless{}shop-app\textgreater{}}. It has some elements for
@@
 but never reloading altogether. Let's assume we work in a sales
-engineering team of the Polymer Shop and need to rebuild checkout
-process.
+engineering team of the Polymer Shop and need to rebuild the checkout
+microservice.
 
@@
 
-Translated into a raw custom element HTML structure, the microservice
-might light like:
+Translated into a raw custom element HTML structure, the top-level
+microservice might look like:
 
@@
 
-Pretty straight forward.
-
-http://www.pocketjavascript.com/blog/2015/11/23/introducing-pokedex-org
-
-bare minimum of centralized management of these services
+\subsection{Concepts}\label{concepts}
+
+The core concept of the newly created \texttt{shop-checkout} is a clean
+MVC pattern where the model is a worker managed by the root element. The
+role of the root element offers a bare minimum of centralized
+orchestration as a kind of message broker. Messages from child nodes
+will just forwarded to the worker and answers passed back to the child
+nodes. Therefore, the basic microservice communication looks like this
+
+\begin{verbatim}
+VIEW root           |               MODEL worker
+                    |
+Event  --------Event msg--------->    Msg
+    |               |               Handler
+    |               |                   |
+Idle                |               Effekt
+    |               |                   |
+  Msg               |                   |
+Handler <-----Action msg----------  Action                                      
+                    |                   
+                    |
+\end{verbatim}
+
+Effects are yielded by the asynchronous operation of messages and create
+effects returned to sender. The VIEW side of the root element forwards
+all the messages and won't be bothered about the content.
 
@@
 query languages from handheld
+
+Most obvious ist the gathering of all related code under the umbrellar
+of a single HTML tag.
+
+Secondly, the sub-standard \emph{custom elements} introduces so called
+lifecycle methods and a getter/setter interface exposing the
+functionality to the developer. Event handling, for example, can be
+registered in place which is much more declarative than assigning event
+listeners from the outside. Of course, this events can be pushed down to
+nested tags, allowing an increasingly granular system design. This
+approach will be explained further in the upcoming sections.
+
+Following this logic any company, whether it is web-related or not,
+should be devided in units grouped around a destinct business service to
+optimise the workflow. Fowler and Lewis outlines this approach as an
+``alignment of business capabilities''{[}4{]} While this kind of
+structure may be true for companies like Google or Amazon, there is a
+vast majority of companies developing for the web which are grouped
+around tasks.{[}9{]} A very common structure is formed by the technology
+stack (UX Designers, Frontend- \& Backend Developers) or by separating
+teams along the product lifecycle (development, testing, deployment).
+
+Advocators from the microservice approach propose a different model.
+best described by . Web components are one (but important) way to tie up
+those diciplines as one component can host a single independent business
+service. Combined with a flexible backend service these components can
+be huge gain over the cumbersome functional organizational approach.
 
@@
 
+\hypertarget{ref-Rota2015}{}
+{[}11{]} A. Rota, ``React.js conf 2015 - the complementarity of react
+and web components.'' 2015 {[}Online{]}. Available:
+\url{https://www.youtube.com/watch?t=124\&v=g0TD0efcwVg}
+
 \hypertarget{ref-Buchner2016}{}
-{[}11{]} D. Buchner, ``Demythstifying web components,'' 2016
+{[}12{]} D. Buchner, ``Demythstifying web components,'' 2016
 {[}Online{]}. Available:
@@
 \hypertarget{ref-vanKesteren2014}{}
-{[}12{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
+{[}13{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
 {[}Online{]}. Available:
@@
 \hypertarget{ref-Bidelman2016}{}
-{[}13{]} E. Bidelman, ``Custom elements v1: reusable web components.''
+{[}14{]} E. Bidelman, ``Custom elements v1: reusable web components.''
 2016 {[}Online{]}. Available:
@@
 \hypertarget{ref-HTML}{}
-{[}14{]} \emph{HTML living standard --- last updated 11 january 2017}.
+{[}15{]} \emph{HTML living standard --- last updated 11 january 2017}.
 {[}Online{]}. Available: \url{https://html.spec.whatwg.org/multipage/}
@@
 \hypertarget{ref-Bidelman2016shadow}{}
-{[}15{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
+{[}16{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
 2016 {[}Online{]}. Available:
@@
 \hypertarget{ref-Bidelman2010}{}
-{[}16{]} E. Bidelman, ``The basics of web workers.'' 2010 {[}Online{]}.
+{[}17{]} E. Bidelman, ``The basics of web workers.'' 2010 {[}Online{]}.
 Available: \url{https://www.html5rocks.com/en/tutorials/workers/basics/}

### 2017-02-13 build/paper.tex
pdf/tex build

@@
 
-The upcoming microservice example assumes a generic build-in API next to
-the build-in frontend components. Instead of gluing frontend and backend
-together on runtime, the microservice is designed holistically
-containing both ends. To reduce network calls especially for mobile
-devices it is a good advice to cache data via a global Service Worker.
-Revamping offline capabilities even further data can be stored in a
-browser based database like PouchDB. For the sake of simplicity data
-management won't be conducted into depth throughout this paper.
+The simplified microservice example later in this paper assumes a
+generic build-in API next to the build-in frontend components. Instead
+of gluing frontend and backend together on runtime, the microservice is
+designed holistically containing both front- and backends. To reduce
+network calls especially for mobile devices it is a good advice to cache
+data via a global Service Worker. Revamping offline capabilities even
+further data can be stored in a browser based database like PouchDB. For
+the sake of simplicity data management won't be explored into depth
+throughout this paper.
 
@@
 
-Working in a JS heavy infrastructure demands for optimazation to avoid
+Working in a JS heavy infrastructure demands for optimization to avoid
 unexpected side-effects like the \emph{flash of unstyled content
@@
 
-Using modular web architecture isn't a perceptually new approach.
-
-Angular, React
-
-Interopt
-
-Freedom of choice
+Microservices tend to become smaller over time. An evolutionary design
+approach puts emphasizes on decomposition and scrapping the service.
+``The key property of a component is the notion of independent
+replacement and upgradeability.''{[}4{]} Therefore we can safely change
+and chop services. Lazy parts of the system which won't change to often
+should be separated from parts undergoing a lot of churn.{[}4{]} Parts
+that needs coupled changes could should be moved together or should be
+even merged.
+
+This flexible approach fits good in the world of browser based
+development. In the last decade we have seen a lot of of changes in the
+way we develop for the web. New technologies like Angular and React are
+handy to use but add some future uncertainty.
+
+In the next section of this paper a lot of standardized technologies
+around the concept of \textbf{Web Components} for the browsers will be
+described. \emph{Browsernative Microservices} should be perceived as
+complementary technology. Contrary to Angular, React and other
+frameworks they have a strong interop with existing systems and can be
+used with them together.
+
+Andrew Rota, developer at Wayfair, came up with the idea to compose
+small web components around a managing React system.{[}11{]} As web
+components are native elements there is no difference to use a native
+\texttt{button} or a native \texttt{custom-button}. With this pattern a
+web developer can make use of the encapsulated advantages of web
+components while still making use of the declarative event management
+from various frameworks. Whatever new framework will be on the rise
+within the next years, this approach allows rapid decomposition and
+reassembling.
 
@@
     \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
-    \KeywordTok{const} \NormalTok{template }\OperatorTok{=}
+    \KeywordTok{const} \NormalTok{helloTemplate }\OperatorTok{=}
      \VariableTok{document}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'#hello'}\NormalTok{)}\OperatorTok{;}
     \KeywordTok{this}\NormalTok{.}\AttributeTok{shadowRoot}
-     \NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{template}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
+     \NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{helloTemplate}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
    \OperatorTok{\}}
@@
 
-After grasping the principles and some technical background the paper
-should illustrate a practical example to \emph{browsernative
-microservices}. Googles library Polymer is a good start to learn about
-web components. They even offer a command line tool for creating a
-polymer skeleton. One of their most famous proof of concept is the
-so-called \href{https://shop.polymer-project.org/}{Polymer Shop} which
-is a fully-fledged online shop nested within a single custom element
-\texttt{\textless{}shop-app\textgreater{}}. It has some elements for
-routing, managing view, service worker caching, theming, etc. The whole
-shop runs as a single application only fetching and sending resources
-but never reloading altogether. Let's assume we work in a sales
-engineering team of the Polymer Shop and need to rebuild checkout
-process.
+After getting some confidence in microservice principles and some
+technical background the paper should illustrate a practical example to
+\emph{browsernative microservices}. Googles library Polymer is a good
+place for learning about web components in depth and make use of their
+simple command line tools. One of their most famous proof of concept is
+the so-called \href{https://shop.polymer-project.org/}{Polymer Shop}
+which is a fully-fledged online shop nested within a single custom root
+element \texttt{\textless{}shop-app\textgreater{}}. It has some elements
+for routing, managing view, service worker caching, theming, etc. The
+whole shop runs as a single application only fetching and sending
+resources but never reloading altogether. Let's assume we work in a
+sales engineering team of the Polymer Shop and need to rebuild the
+checkout microservice.
 
@@
 
-Translated into a raw custom element HTML structure, the microservice
-might light like:
+Translated into a raw custom element HTML structure, the top-level
+microservice might look like the following snippet:
 
@@
 
-Pretty straight forward.
-
-http://www.pocketjavascript.com/blog/2015/11/23/introducing-pokedex-org
-
-bare minimum of centralized management of these services
+\subsection{Service Root}\label{service-root}
+
+The core concept of the newly created \texttt{shop-checkout} is a clean
+MVC pattern where the model is a worker managed by the root element. The
+role of the root element offers a bare minimum of centralized
+orchestration as a \textbf{kind of message broker}. Messages from child
+nodes will just forwarded to the worker and answers passed back to the
+child nodes. Therefore, the basic microservice communication looks like
+this
+
+\begin{verbatim}
+VIEW root           |               MODEL worker
+                    |
+Event  --------Event msg--------->    Msg
+    |               |               Handler
+    |               |                   |
+Idle                |               Effekt
+    |               |                   |
+  Msg               |                   |
+Handler <-----Action msg----------  Action                                      
+                    |                   
+                    |
+\end{verbatim}
+
+Effects are yielded by the asynchronous operation of messages and create
+actions returned to sender. Effects can be created by external messages,
+like subscription to an WebSocket, too. Effects may be created by
+fetching additional resources from the server. The VIEW side of the root
+element forwards all the messages and won't be bothered about the
+content.
 

### 2017-02-16 paper.tex
pdf/tex build

@@
-\documentclass[]{assets/latex/ieee}
+\documentclass[]{article}
 \usepackage{lmodern}
@@
 
-\section{Simplicity and the state of the
-web}\label{simplicity-and-the-state-of-the-web}
+{
+\setcounter{tocdepth}{3}
+\tableofcontents
+}
+\section{Simplicity and the web}\label{simplicity-and-the-web}
 
@@
 Software development is undoubtedly rich in complexity and full of
-subtle pitfalls for the human brain. In a typical scenario, a piece of
-software evolves over time in one or another \sout{opinionated}
-direction. Layers of new abstractions wrestling with old legacy
-abstractions and mutation becomes untraceable. Subtle bugs start to
-creep in. Eventually the small piece of software may end up in a highly
-complected monolith which will determine future design decisions to a
-painful degree. Future strategies of the company/organization will be
-highly determined by the current state in the need of ``keeping the
-lights on''.
-
-On the other side of this dystopian example, the truly modular system
-architecture shines in a approachable simplicity. The whole system is
-divided in pluggable parts, object mutations are traceable and modules
-interchangeable. As Rich Hickey argues, design decisions should be made
-under the \textbf{impression of extending, substitution, moving,
-combining and repurposing}. The ability to reason about the program at
-any given time is crucial for future decisions and implementations.
-Recalling again the unidimensional nature of simplicity.
-
-The \textbf{state of the web} is certainly a different kind of
-complexity beast. ``The web'' is coined to ``everything that runs in the
-browser''. While simplicity in the backend is mostly a matter of
-principles, any frontend developer is highly restricted on the highly
-deterministic nature of the browser platform.
+subtle pitfalls. In a typical scenario, a piece of software evolves over
+time in one or another opinionated direction. Layers of new abstractions
+wrestling with old legacy abstractions and mutation becomes untraceable.
+Subtle bugs start to creep in. Eventually the small piece of software
+may end up in a highly complected monolith which will determine future
+design decisions to a painful degree. Future strategies of the
+company/organization will be highly determined by the current state in
+the need of ``keeping the lights on''.
+
+On the other side of this dystopian scenario, the truly modular system
+architecture abandons many of those inconsistencies. The whole system is
+divided in pluggable parts, object mutation is either traceable or
+avoided altogether in favor of immutable data structures. As Rich Hickey
+argues, design decisions should be made under the \textbf{impression of
+extending, substitution, moving, combining and repurposing}. The ability
+to reason about the program at any given time is crucial for future
+decisions and implementations. Recalling again the unidimensional nature
+of simplicity.
+
+Simplicity in ``the web'', read as a loose generalization of
+``everything that runs in the browser'', is certainly a story full of
+misconceptions and will be explained further along. While simplicity in
+the backend is mostly a matter of principles and patterns, any
+browser-based frontend is restricted on the highly deterministic nature
+of the browser platform.
 
@@
 text nodes or an ever changing webapp the underlying global nature of
-the DOM tree remains the same. Every additional piece of software added
-to this global object embodies another fold of complexity.
+the DOM tree remains the same. Every additional piece of code added to
+the webpage will invisibly add another fold of complexity to this global
+object.
 
@@
 modularization is a typical pattern to make complexity manageable and
-accommodate future uncertainty.{[}2, p. 1{]} Today many websites tend
-towards becoming complex single-page applications (and therefore the
-average JS payload is soaring too). Consequently, the demands to the
-browser platform changed from a static page renderer to a
-\textbf{complexe UI machine}. Under the current situation only
-additional layers of abstraction can handle complexity.
+accommodate future uncertainty.{[}2, p. 1{]} Since years the average JS
+payload is steadily rising which can be interpreted as a trend towards
+more dynamic websites. The demands to the browser platform changed from
+a static page renderer to a \textbf{dynamic UI machine} without changing
+the underlying architecture significantly. Under the current situation
+only additional layers of abstraction can wrestle complexity.
 
@@
 and design rules into maintainable components. While the DOM can't be
-scoped, JS can. Therefore many frameworks, like ReactJS, AngularJS or
-VueJS just to name a few, ditched the old rule of separated HTML, CSS
-and JS in favor of an additional layer of abstracted JS components
-(containing content, markup and styling). Quiet often those frameworks
-mimic a MVC pattern on top of the browser engine which is a reasonable
-simple design pattern to build graphical user interfaces. While
-frameworks are a valid approach to build scalable web applications they
-remain highly opinionated, embody inherent complexity themselves and can
-change and break over time. Furthermore, code inflation in front-end is
-a crucial point for performance and third-party libraries are no
-exception on that. A standardized way for creating complex UI interfaces
-painlessly requires new build-in browser capabilities.
-
-In 2013 thinkers, creators and browser vendors joined together to
-propose \emph{The Extensible Web Manifesto}.\footnote{\href{https://extensiblewebmanifesto.org/}{The
-  Extensible Web Manifesto}} The claim of the manifesto is to enhance
+scoped, JS can. Many frameworks, like ReactJS, AngularJS or VueJS just
+to name a few, ditched the old rule of separated HTML, CSS and JS in
+favor of an additional layer of abstracted JS components (containing
+content, markup and styling). Quiet often those frameworks mimic a MVC
+pattern on top of the browser engine which is a reasonable simple design
+pattern to build graphical user interfaces. While frameworks are a valid
+approach to build scalable web applications they remain highly
+opinionated, embody inherent complexity themselves and can change and
+break over time. Another downside is code inflation which is a crucial
+point for performance and third-party libraries are no exception on
+that. A standardized way for creating complex UI interfaces painlessly
+requires new build-in browser capabilities.
+
+In the year 2013 thinkers, creators and browser vendors joined together
+to propose \emph{The Extensible Web Manifesto}.\footnote{\href{https://extensiblewebmanifesto.org/}{The
+  Extensible Web Manifesto}} The claim of the manifesto was to enhance
 the current web platforms with new low-level capabilities. Those
@@
 leapfrogged and many new low-level APIs brought to life. With this new
-APIs at hand a vivid web developer can create robust websites without
-additional libraries.
+APIs at hand a vivid web developer can create robust websites with less
+code and less additional libraries. This paper is an approach to unfold
+these \textbf{browsernative} technologies to create overall simple and
+resilient \textbf{microservices} for the browser
 
@@
 services, each running in its own process and communicating with
-lightweight mechanisms, often an HTTP resource API.''{[}4{]}
-Microservices incorporate many other ideas, like \emph{domain-driven
-design} where it pursues the incorporation of the real world in our
-code.{[}3, p. 2{]} Or making use of \emph{continuous delivery} for
-pushing software rapidly through \emph{automated deployment} mechanisms
-into production.{[}4{]} And, last but not least, microservices utilizes
-the idea of small teams with a lot of product knowledge working mostly
-autonomous on their very own service with their very own set of tools
-and techniques. As a primary source of truth this paper relies on the
-work of Sam Newman, who has written a comprehensive guide called
-\emph{Building Microservices} and the work of Fowler and Lewis. The
-purpose of this section is to match their ideas against services on the
-browser platform.
+lightweight mechanisms, often an HTTP resource API.''{[}4{]} This
+definition aligns well with the idea of \textbf{Single-Page
+applications} (SPAs) in the browser composed of own services or
+sometimes other third-party services.
+
+Further along, microservices incorporate many other ideas, like
+\emph{domain-driven design} where it pursues the incorporation of real
+world structure in the code.{[}3, p. 2{]} Or making use of
+\emph{continuous delivery} for pushing software rapidly through
+\emph{automated deployment} mechanisms into production.{[}4{]} And, last
+but not least, microservices utilizes the idea of small teams with a lot
+of product knowledge working mostly autonomous on their very own service
+with their very own set of tools and techniques. As a primary source of
+truth this paper relies on the work of Sam Newman, who has written a
+comprehensive guide called \emph{Building Microservices} and the work of
+Fowler and Lewis. The purpose of this section is to gain confidence
+about the microservices architecture in the context of the browser
+platform.
 
@@
 
-Any reader of the paper familiar with ReactJS / Redux the concept of
+Any reader of the paper familiar with ReactJS / Redux, the concept of
 components isn't anything new. Dan Abramov, the creator of Redux, once
@@
 
-On the other hand Dan Abramov outlines components which he calls
+On the other hand Abramov described components he named
 \textbf{containers}. A container component is concerned with \emph{how
@@
 behavior sit together, and unrelated behavior to sit elsewhere.{[}3, p.
-30{]} High cohesion can be perceived as the \emph{Single Responsibility
-Principle} defined by Robert C. Martin: ``Gather together those things
+30{]} High cohesion can be expressed in a dynamic way as the
+\emph{Single Responsibility Principle}: ``Gather together those things
 that change for the same reason and separate those things that change
@@
 analysis, the quality can be measured just by counting the places
-changes in the code occur. In an arbitrary MVC system a \texttt{button}
-might be inserted in the VIEW, the CONTROLLER needs some adjustment and
-maybe the the MODEL, too. Three places for adjustment is a reasonable
-easy task for the brain.
-
-Using libraries in web development is common sense. But compared to
+changes in the code occur. In an arbitrary MVC system when a
+\texttt{button} is dropped into the VIEW part, the CONTROLLER needs some
+adjustment for sure and maybe the MODEL, too. Three places for
+adjustment is a reasonable easy task for the brain. In the field of
+browser based development those principle is often violated in the
+separated entities HTML, JS and CSS.
+
+Using libraries in web development is a common sense. But compared to
 libraries, a component service offers multiple advantages. A library is
@@
 tools for the web, like Webpack 2 or Rollup offer this feature which
-relies heavily on the static structure on ES6 modules.{[}7{]} Contrary
-libraries for the browsers are traditionally ``shipped'' as
-\textbf{immediately-invoked functions}. Bootstrapping those ones is much
+relies heavily on the static structure on ES6 modules.{[}7{]} Libraries
+for the browsers are traditionally ``shipped'' as non static
+\textbf{immediately-invoked functions}. Bootstrapping those is much
 harder to archive as dead code may be part of the function itself.
@@
 bundler can choose which component is needed at runtime. Even more
-granular, a good web component is composed of static JS blocks which can
-be strapped away too.
-
-A web component is self-contained which means it embodys all needed
-functionality to get it's job done. Therefore it has a much better
-evolution mechanism in the service contracts. Changing functionality
-won't break other components. A component can progressively enhanced
-which guarantees functionality throughout different versions.
+granular, a good web component is composed of static JS dependencies
+which can be deflated or strapped away too.
+
+A \emph{component for the web}, or web component, is
+\textbf{self-contained} which means it embodys all needed functionality
+to get it's job done. Therefore it has a much better evolution mechanism
+in the service contracts. Changing functionality won't break other
+components. A component can progressively enhanced which guarantees
+functionality throughout different versions.
 
@@
 other contexts.{[}3, p. 30{]} Each context can derived into multiple
-microservices or, talking about \emph{browsernative microservices},
+microservices or, talking about \emph{Browsernative Microservices},
 interfaces to the customers.
@@
 receiver) and a mostly unified ``dumb'' letter format and channel. Each
-endpoint own its specific domain logic. In the browsernative context we
+endpoint owns it specific domain logic. In the browsernative context we
 already emphasized the role of presentational components and containers.
@@
 the implementation part of this paper a reader will explore the
-combination of REST protocols and a lightweight browser message bus in
-action.
+combination of RESTish protocols and a lightweight browser message bus
+in action.
 
@@
 
-Microservices are separate entities and decentralization therefore is a
+Microservices are separate entities and decentralization is therefore a
 meta concept of microservices. Distributed service responsibility owned
@@
 stack, tools and services designed in the spirit of language- and
-platform independence - and share their knowledge with other
+platform independence and share their knowledge with other
 parties.{[}4{]} In the recent years many big companies like Facebook,
@@
 
-The spirit of freedom can't be applied to \emph{browsernative
-microservices} as the browser and its underlying DOM will be the
-limitation factor. Talking about the browser, a reader might be tempted
-to narrowly thinking about the VIEW layer only - which is not true
-anymore. In the recent years the browser engines grows to a kind of
-virtual machine: there are connectors to build-in databases,
-multithreading support, an ever-growing JS build-ins and even push
-notifications. So-called \emph{Progressive Web Apps\footnote{\href{https://developers.google.com/web/progressive-web-apps/}{Progressive
+The spirit of freedom can't be applied universally to
+\emph{Browsernative Microservices} as the browser and its underlying DOM
+will be the limitation factor to a certain degree. Talking about the
+browser, a reader might be tempted to narrowly thinking of the obvious
+VIEW layer only - which is not true anymore. In the recent years the
+browser engines grows to a kind of virtual machine: there are connectors
+to build-in databases, multithreading support, an ever-growing JS
+build-ins and even speech synthesis. So-called \emph{Progressive Web
+Apps\footnote{\href{https://developers.google.com/web/progressive-web-apps}{Progressive
   Web Apps}}}, a bunch of criteria for building good browserapps, can
-achieve a similar look and feel like native apps.
+achieve a similar look and feel like native apps. Services like
+NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}},
+effectively compiling ``the web'' to native code, lower the boundary
+between native and browser code even further.
 
@@
 to JS even exclusively. Once web components hit a critical mass there
-will be most likely some library support or foreign function interfaces
+will be most likely some library support or foreign function interface
 towards ES6 modules (which are mandatory for the new specifications).
@@
 Another more real life decentralization aspect derives from the easiness
-of deployment in a safe, sandboxed environment. No hidden overheads or
-required build tools. Thats why \emph{browsernative microservices} are
-well suited for sharing and open-sourcing. Webcomponents.org\footnote{\href{https://webcomponents.org}{Webcomponents.org}}
-is a registry for ready-to-use components of every scale where for
-example Google shares a lot of their material design elements.
+of deployment in a safe, sandboxed environment. Web components virtually
+ship no overhead or require obscure build tools. This makes them ideal
+candidates for sharing and open-sourcing. Webcomponents.org\footnote{\href{https://webcomponents.org}{Webcomponents.org}}
+is a registry for ready-to-use components of every scale and purpose
+where for even Google shares a lot of their material design elements.
 
@@
 Data Management in Microservice follows the same modular philosophy like
-the conceptual implementation. As mentioned earlier different bounded
-contexts make different assumptions of the underlying model. A
+the service implementation. As mentioned earlier different bounded
+contexts make different assumptions of the underlying models. A
 \emph{browsernative microservices} takes this idea even further and
@@
 emphasizes this pragmatism on APIs in a conference talk. At ING the
-frontend demands drive the backend APIs, APIs are produces automatically
-and not even Issa knows how many APIs exists.{[}9{]} They are using a
-pattern called \textbf{backend for frontends} allowing the team working
-on a given UI also handle the server-side components.{[}3, p. 72{]} The
-perception on backend might change therefore from embodying logic to a
-data-silo only.
+frontend demands tailor the backend APIs, APIs may be produced
+automatically and not even Issa knows how many APIs exists.{[}9{]} They
+are using a pattern called \textbf{backend for frontends} empowering the
+team working to craft their UI and backend in a one-to-one
+relationship.{[}3, p. 72{]}
 
@@
 query language for the frontend. The backend solely replies on the
-frontend needs. Another well documented \sout{even more powerful}
-approach in the field is Cognitects Datomic\footnote{\href{http://www.datomic.com/}{Datomic}},
+frontend needs. Another well documented example in the field is
+Cognitects Datomic\footnote{\href{http://www.datomic.com/}{Datomic}},
 where parts of the database will be reflected to the client. A so-called
@@
 The simplified microservice example later in this paper assumes a
-generic build-in API next to the build-in frontend components. Instead
-of gluing frontend and backend together on runtime, the microservice is
-designed holistically containing both front- and backends. To reduce
-network calls especially for mobile devices it is a good advice to cache
-data via a global Service Worker. Revamping offline capabilities even
-further data can be stored in a browser based database like PouchDB. For
-the sake of simplicity data management won't be explored into depth
-throughout this paper.
+generic build-in API accompanied by build-in frontend components.
+Instead of gluing frontend and backend together on runtime, the
+microservice is designed holistically containing both front- and
+backends. To reduce network calls especially for mobile devices it is a
+good advice to cache data (f.e. in a global Service Worker). Revamping
+offline capabilities even further data can be stored in a browser based
+database like PouchDB. For the sake of simplicity data management won't
+be explored into depth throughout this paper.
 
@@
 
-In the global nature of web development the developer stage couldn't
+In the global nature of web development the development couldn't
 completely decoupled from the production environment. This circumstance
 left developers switching back and forth between files developing tricky
-(and more often biased) ways to glue related parts together. With scoped
-components, code blocks can be developed more sane and conveyed into
-production without headache. Even more, in the scripted environment of
-the browser this can be implemented rapidly and continuous.
-
-Previously mentioned Ben Issa, described the ING workflow pretty simple.
+opinionated (and more often biased) ways to glue related parts together.
+With scoped components, code blocks can be developed more sane and
+conveyed into production without headache. Even more, in the scripted
+environment of the browser this can be implemented rapidly and
+continuous.
+
+Previously mentioned Ben Issa, described the ING standard workflow.
 Every component deserves a own \textbf{git repo} containing
@@
 Every check-in is handled as release candidate and can be independently
-deployed by a fully automated deployment machinery.{[}9{]} While this
-example is seems opinionated, it gives a sense of a mature component
-build for the web. Due to an exhaustive amount of testing and deployment
-tools for the browser an automated infrastructure shouldn't be a
-problem.
+deployed by a fully automated deployment machinery.{[}9{]} Even though
+this example is an opinionated perception it gives a sense of a mature
+component build for the web. Due to an exhaustive amount of testing and
+deployment tools for the browser an automated infrastructure shouldn't
+be a problem.
 
@@
 While \emph{Progressive Enhancement} was once coined on the principle to
-build websites both for Browsers with JS and without, the new
-\emph{Progressive Enhancement} becomes more an ``offline first''
-principle. A \emph{browsernative microservice} therefore not only tries
-to cache data as much as possible, it should also bring in a lot of
-program logic as described in the previous chapters.
+build websites both for Browsers with JS and HTML only, the new
+\emph{Progressive Enhancement} tends towards an ``\textbf{offline
+first}'' build principle. A \emph{Browsernative Microservice} therefore
+not only tries to cache data as much as possible, it should also bring
+in a lot of program logic as described in the previous chapters.
 
@@
 replacement and upgradeability.''{[}4{]} Therefore we can safely change
-and chop services. Lazy parts of the system which won't change to often
+and chop services. Lazy parts of the system which won't change often
 should be separated from parts undergoing a lot of churn.{[}4{]} Parts
@@
 development. In the last decade we have seen a lot of of changes in the
-way we develop for the web. New technologies like Angular and React are
-handy to use but add some future uncertainty.
-
-In the next section of this paper a lot of standardized technologies
-around the concept of \textbf{Web Components} for the browsers will be
-described. \emph{Browsernative Microservices} should be perceived as
-complementary technology. Contrary to Angular, React and other
-frameworks they have a strong interop with existing systems and can be
-used with them together.
+way we develop for the web. New approaches like the virtual DOM approach
+found their way into mainstream web development followed by frameworks
+and libraries.
+
+\emph{Browsernative Microservices} should be perceived as complementary
+technology. Contrary to Angular, React and other frameworks they have a
+strong interop with existing systems and can be used with them together.
 
@@
 within the next years, this approach allows rapid decomposition and
-reassembling.
+reassembling towards a new system.
 
@@
 
-\emph{Custom elements} are the fundamental building blocks for web
-components. Essentially, they provide a way to create custom HTML tags
-enriched with behavior, design and functionality. An obligatory
-\textbf{HelloWorld} will help to grasp the spec:
+Custom elements are the fundamental building blocks for web components
+introducing the \emph{Single Responsibility Principle} to the browser.
+In essence, they provide a way to create \textbf{custom HTML tags}
+subsuming behavior, design and functionality. An obligatory
+\textbf{HelloWorld} will give a flavor about the spec:
 
@@
   \KeywordTok{this}\NormalTok{.}\AttributeTok{onclick} \OperatorTok{=} \NormalTok{e }\OperatorTok{=>} \AttributeTok{alert}\NormalTok{(}\StringTok{"hello"}\NormalTok{)}\OperatorTok{;}
-  \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(...)}\OperatorTok{;}
  \OperatorTok{\}}
@@
 
-Most obvious, this spec relies on the new \emph{ES6 Class Syntax} in
-favor of the original prototype-based inheritance model.
-``Extending~\texttt{HTMLElement}~ensures the custom element inherits the
-entire DOM API and means any properties/methods that you add to the
-class become part of the element's DOM interface.''{[}14{]} Like any
-other \emph{ES6 class}, \emph{custom element}s can be sub-classed
-further on with the typical \texttt{extends} inheritance.
+This example should be self-explanatory. Notably, custom elements come
+in the fashion of \emph{ES6 Classes}\footnote{\href{https://developer.mozilla.org/en/docs/Web/JavaScript/Reference/Classes}{JavaScript
+  Classes}} in favor of the normal JavaScript prototype-based
+inheritance model. This class must inherit the base \texttt{HTMLElement}
+interface which~``ensures the newly created element inherits the entire
+DOM API and any properties/methods that you add to the class become part
+of the element's DOM interface.''{[}14{]} Like any other \emph{ES6
+class}, the new element can be specialized further on using the typical
+\texttt{extends} inheritance.
 
@@
 \texttt{this} which points to the element itself. Instead of querying
-and assigning behavior AFTER creation of the node in question
-\texttt{this} functionality allows a \textbf{declarative programming
-style}. Assigning functionality happens right in place BEFORE creation
-or insertion of the DOM. The so called \emph{fat-arrow}
-(\texttt{=\textgreater{}}) is just a new feature of ES6 and nothing more
-than an anonymous \texttt{function()}.
-
-After definition, the element needs to be registered in the new global
-build-in \texttt{customElements} with an tag name like
-\texttt{\textless{}hello-world\textgreater{}}. Mind the dash inside the
-tag name to conform the spec. Finally, the new element can go live
-inside the HTML Document \texttt{index.html}.
+and assigning behavior AFTER creation of the node, custom elements ship
+their functionality PRIOR initialization of the element. The so called
+\emph{fat-arrow} (\texttt{=\textgreater{}}) is just a new ES6 syntax
+feature for an anonymous function declaration.
+
+After declaration, the new HTML element needs to be registered in the
+global build-in \texttt{customElements} object with an tag name like
+\texttt{\textless{}hello-world\textgreater{}} acting as key to the
+element. Mind the dash inside the tag name to conform the spec. Finally,
+the new element can go live inside the HTML Document
+\texttt{index.html}.
 
@@
   method is called straight after the \texttt{constructor()} on insert.
-  This block of code contains setup code, such as fetching resources or
-  rendering elements according to attributes.{[}14{]} For performance
-  issues it's highly preferable to put much code in here.
+  Typically, this method contains setup code, such as fetching resources
+  or rendering elements according to attributes.{[}14{]} For a fast
+  initial render of the page, it is highly preferable to put many
+  proceedings in favor of the constructor.
 \item
@@
 As previously mentioned, the custom elements must \texttt{extend} the
-\texttt{HTMLElement}. Consequently, the new element inherits properties
-and methods from it (and it's parent \texttt{Element}) and things like
-\texttt{id,\ class,\ addEventListner,\ ...} run out-of-the-box.
-Additionally, it is possible to define custom attributes using the
-\emph{custom elements'} \textbf{getter / setter interface} to steer the
-behavior of the element.
+\texttt{HTMLElement} interface. Therefore, the new element inherits base
+properties and methods commonly used in all HTML elements like
+\texttt{id,\ class,\ addEventListner} . Additionally, it is possible to
+define custom attributes using the \emph{custom elements'}
+\textbf{getter / setter interface} to steer the behavior of the element.
 
@@
 
-While getters and setters work great in the JS world they fail crossing
-the boundaries to the corresponding HTML node. Declaring
+Native DOM properties, like \texttt{id} or \texttt{onclick}, reflect
+their values between HTML and JS.{[}15, Para. 2.6.1{]} For example
+declaring the HTML
+\texttt{\textless{}hello-world\ id="hello"\textgreater{}} like this
+equals to assign the ID in JS like \texttt{Node.id\ =\ "hello"}. This
+behavior won't work out-of-the-box with methods or properties defined
+setters. For example declaring
 \texttt{\textless{}hello-world\ sayhello="mars"\textgreater{}\textless{}/hello-world\textgreater{}}
-would't work in the previous setup. A common workaround is archived by
-using the previous mentioned \texttt{attributeChangedCallback} lifecycle
-method to \textbf{reflect changing HTML attributes to JS} and/or map JS
-attributes to HTML with \texttt{this.setAttributes(...)} respectively.
-On insertion time HTML attributes might raise their hand with
-\texttt{this.hasAttributes(...)} and \texttt{this.getAttributes(...)}.
-Native DOM properties will reflect their values between HTML and JS
-automatically.{[}15, Para. 2.6.1{]}
+would't call the \texttt{sayhello} function in the previous setup.
+
+A common workaround to bind HTML and JS behavior together is archived by
+using the aforementioned lifecycle method
+\texttt{attributeChangedCallback} to \textbf{bind changing HTML
+attributes to JS properties} and to map JS attributes to HTML with
+\texttt{this.setAttributes(...)} respectively. On insertion time HTML
+attributes might trigger custom JS methods retrieving HTML attributes
+using \texttt{this.getAttributes(...)} method.
 
@@
 
-The second most important concept of \emph{web components} rewards to
-the \emph{shadow DOM} spec. In terms of complexity this spec outpacing
-all others by far. A \emph{shadow DOM} is basically an isolated DOM tree
-living inside an another (hosting) DOM tree. The spec refers the hosting
-tree as \emph{light DOM tree} and the attached DOM as \emph{shadow DOM
-tree}. Conceptually, the \emph{shadow DOM} issues a single important
-topic in software development: \textbf{Encapsulation}. While the first
-spec \emph{custom elements} provides a sufficient way to encapsulate JS
+A \emph{shadow DOM} is basically an isolated DOM tree living inside an
+another (hosting) DOM tree. The spec refers the hosting tree as
+\emph{light DOM tree} and the attached DOM as \emph{shadow DOM tree}.
+Conceptually, the \emph{shadow DOM} issues a single important topic in
+software development: \textbf{Encapsulation}. While the first spec
+\emph{custom elements} provides a sufficient way to encapsulate JS
 behavior, \emph{shadow DOM} coined strongly to in the direction of style
@@
 \KeywordTok{</hello-world>}
+
 \CommentTok{<!-- COMPONENT STARTS HERE -->}
@@
     \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
-    \KeywordTok{const} \NormalTok{template }\OperatorTok{=}
-     \VariableTok{document}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'#hello'}\NormalTok{)}\OperatorTok{;}
-    \KeywordTok{this}\NormalTok{.}\AttributeTok{shadowRoot}
-     \NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{template}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
+    \KeywordTok{const} \NormalTok{helloTemplate }\OperatorTok{=} \VariableTok{document}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'#hello'}\NormalTok{)}\OperatorTok{;}
+    \KeywordTok{this}\NormalTok{.}\VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{helloTemplate}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
    \OperatorTok{\}}
@@
 event, strongly recommended. Once the imported HTML document comes into
-scope, activation follows a very similar pattern like the previously
-mentioned \emph{HTML templates}:
+scope, activation follows a very similar pattern like the aforementioned
+\emph{HTML templates}:
 
@@
 \texttt{\textless{}hello-world\textgreater{}\textless{}/hello-world\textgreater{}}.
-Preliminary, the component needs proper configuration as the last
-\texttt{HelloWorld} example wouldn't work like it is currently. The next
-section will elaborate the right configuration and composition of a
-component to work out-of-the-box.
+Preliminary, the component needs proper configuration. The next section
+will elaborate the right configuration and composition of a component to
+work out-of-the-box.
 
@@
 Despite the discrepancies among browser vendors \emph{HTML Imports}
-should still be part of the paper and future \emph{web components} as no
-other native browser technology can bundle up CSS, JS and HTML that
-efficient. As of today, January 2017, only Googles Chrome and related
-Opera browser supporting the full \emph{web components} spec and,
-despite from \emph{HTML Imports}, all other browser vendors most likely
-will catch up with \emph{Custom Elements} and \emph{Shadow DOM} within
-this year. In the meantime, the full \emph{web components} stack can be
-\textbf{polyfilled} and used across all browsers.
+should still be part of the paper as no other native browser technology
+can bundle up CSS, JS and HTML that efficient.
 
@@
 Naming events after the emitting tag makes the API almost
-self-explanatory. Fortunately the \texttt{detail} property can transmit
-even objects witch allows the developer to craft almost all
-functionality inside one event. The further sections will elaborate a
+self-explanatory. The \texttt{detail} property can be loaded with
+primitives as well as objects. The further sections will elaborate a
 feasible architecture for building a scaling microservice architecture.
@@
 Elements} section, the pattern of \textbf{extending native elements}
-should be somewhat dismissed as it may be implemented outside of Chrome.
-Nevertheless, it is possible to create own kind of quasi native buttons
-when chaining a \emph{CustomEvent} directly after the native click
-event.
+should be somewhat dismissed as it may never implemented outside of
+Chrome. Nevertheless, it is possible to create own kind of quasi native
+buttons when chaining a \emph{CustomEvent} directly after the native
+click event.
 
@@
 (native) child elements. In this case, the \emph{Custom Element} catches
-events from children, buffers or rebuild them and eventuall fires an
+events from children, buffers or rebuild them and eventually fires an
 event towards the document root.
@@
 Like \emph{Custom Events}, \emph{Web Workers} had been around for a long
-time and therefore embrace full support among major browsers. They
-emerged at around 2009 when discussions about browser performance was
-still in the early days. Nevertheless, the addressed problem of
-\emph{Web Workers} is a fundamental language problem of JS itself.
+time and therefore enjoy full support among major browsers. They emerged
+at around 2009 when discussions about browser performance was still in
+the early days. Nevertheless, the addressed problem of \emph{Web
+Workers} is a fundamental language problem of JS itself.
 
@@
 responsive even when computation-heavy tasks should be carried out.
-Effectively, a worker thread makes use of the multi core CPUs most
-devices have nowadays. To grasp the full potential of workers, a reader
-might dive deeper into the Angular 2 architecture, where most of the
-application layer is abstracted from the main rendering thread into
-worker threads.\footnote{\href{https://docs.google.com/document/d/1M9FmT05Q6qpsjgvH1XvCm840yn2eWEg0PMskSQz7k4E}{Angular
+Furthermore, a worker thread adds a performance advantage embracing the
+multi core CPU architecture most devices running on today. To grasp the
+full potential of workers, a reader might dive deeper into the Angular 2
+architecture, where most of the application layer is abstracted from the
+main rendering thread into worker threads.\footnote{\href{https://docs.google.com/document/d/1M9FmT05Q6qpsjgvH1XvCm840yn2eWEg0PMskSQz7k4E}{Angular
   2 Rendering Architecture}}
@@
 
-\section{Anatomy of an browsernative
-microservice}\label{anatomy-of-an-browsernative-microservice}
-
-After getting some confidence in microservice principles and some
-technical background the paper should illustrate a practical example to
+\section{Building an browsernative
+microservice}\label{building-an-browsernative-microservice}
+
+After getting confidence in microservice principles and technical
+background the paper should bring them both together to form a
 \emph{browsernative microservices}. Googles library Polymer is a good
-place for learning about web components in depth. They even offer a
-command line tool for creating a polymer skeleton. One of their most
-famous proof of concept is the so-called
-\href{https://shop.polymer-project.org/}{Polymer Shop} which is a
-fully-fledged online shop nested within a single custom root element
-\texttt{\textless{}shop-app\textgreater{}}. It has some elements for
-routing, managing view, service worker caching, theming, etc. The whole
-shop runs as a single application only fetching and sending resources
-but never reloading altogether. Let's assume we work in a sales
-engineering team of the Polymer Shop and need to rebuild the checkout
-microservice.
+place for learning about web components in depth and make use of their
+simple command line tools. One of their most famous proof of concept is
+the so-called \href{https://shop.polymer-project.org/}{Polymer Shop}
+which is a fully-fledged online shop nested within a single root element
+\texttt{\textless{}shop-app\textgreater{}}. This app made of several
+main views and many more invisible custom elements for routing, service
+worker caching, theming, etc. The whole shop runs as a single
+application only fetching and sending resources and switching views.
+Let's assume we work in a sales engineering team of the Polymer Shop and
+need to rebuild the checkout microservice. Needless to say that many
+ideas mentioned here are opinions of the author and by no means
+standardized patterns.
+
+The current checkout can be found at
+https://shop.polymer-project.org/checkout. Currently the checkout is a
+single, 671 lines of code long Polymer component incorporation all
+required fields for sign in, shipping, billing and summarizing the
+order. In the spirit of microservices we will split up the component to
+their smaller components. The shopping cart data is pulled out of a
+local storage JSON entity set up previously by another custom-element.
+Another in-memory opportunity is storing the data in the build-in
+database IndexedDB.
+
+\subsection{Checkout microservice}\label{checkout-microservice}
 
@@
 
-Translated into a raw custom element HTML structure, the top-level
-microservice might look like:
+Translated into a raw \textbf{Custom Element} HTML structure, the
+top-level microservice might look like the following snippet:
 
@@
 
-\subsection{Concepts}\label{concepts}
-
-The core concept of the newly created \texttt{shop-checkout} is a clean
-MVC pattern where the model is a worker managed by the root element. The
-role of the root element offers a bare minimum of centralized
-orchestration as a kind of message broker. Messages from child nodes
-will just forwarded to the worker and answers passed back to the child
-nodes. Therefore, the basic microservice communication looks like this
+Messages from child nodes will just forwarded to the worker and answers
+passed back to the child nodes. Therefore, the basic microservice
+communication looks like this
 
@@
 Effects are yielded by the asynchronous operation of messages and create
-effects returned to sender. The VIEW side of the root element forwards
-all the messages and won't be bothered about the content.
+actions returned to sender. Effects can be created by external messages,
+like subscription to an WebSocket, too. Effects may be created by
+fetching additional resources from the server. The VIEW side of the root
+element forwards all the messages and won't be bothered about the
+content.
 

### 2017-02-17 build/paper.tex
pdf/tex build

@@
 nutshell a microservice is a small, autonomous service that works
-together with other services seamlessly.{[}3, p. 2{]} or with the words
+together with other services seamlessly.{[}3, p. 2{]} Or with the words
 of Fowler and Lewis: ``\ldots{} the microservice architectural style is
@@
 services, each running in its own process and communicating with
-lightweight mechanisms, often an HTTP resource API.''{[}4{]}
-Microservices incorporate many other ideas, like \emph{domain-driven
-design} where it pursues the incorporation of the real world in our
-code.{[}3, p. 2{]} Or making use of \emph{continuous delivery} for
-pushing software rapidly through \emph{automated deployment} mechanisms
-into production.{[}4{]} And, last but not least, microservices utilizes
-the idea of small teams with a lot of product knowledge working mostly
-autonomous on their very own service with their very own set of tools
-and techniques. As a primary source of truth this paper relies on the
-work of Sam Newman, who has written a comprehensive guide called
-\emph{Building Microservices} and the work of Fowler and Lewis. The
-purpose of this section is to match their ideas against services on the
-browser platform.
+lightweight mechanisms, often an HTTP resource API.''{[}4{]} Just yet at
+this point a reader might spot some similarities with microservices and
+the browser-based development: Both wrestling the problem of monolithic
+architecture and both using lightweight communication mechanisms. In
+fact, many big companies of ``the web'' like Amazon or Netflix
+successfully transformed their monolithic system into a service based
+system which gives a glimpse of the power behind microservices.{[}4{]}
+
+Microservices incorporate many ideas from developing scalable software,
+like \emph{domain-driven design} where it pursues the incorporation of
+real world structure in the code.{[}3, p. 2{]} Or making use of
+\emph{continuous delivery} for pushing software rapidly through
+\emph{automated deployment} mechanisms into production.{[}4{]}
+Furthermore, microservices transcendent the technical perspective and
+reaches into the team organization. As a primary source of truth this
+paper relies on the work of Sam Newman, who has written a comprehensive
+guide called \emph{Building Microservices} and the work of Fowler and
+Lewis. The purpose of this section is to gain confidence about the
+microservices architecture in the context of the browser platform.
 
@@
 replaceable and upgradeable.'' {[}4{]} Components are the building
-blocks of microservices. While the service is a concrete piece of
-software a component is much more raw and single-minded following the
-\emph{loose coupling principle}: changing and deploying one service
-shouldn't result in changing other parts of the system.{[}3, p. 30{]}.
-Small components and big microservices are similar on their principles.
-
-Any reader of the paper familiar with ReactJS / Redux the concept of
-components isn't anything new. Dan Abramov, the creator of Redux, once
-defined a simple dichotomous pattern for creating UI components. In his
-believe, there are \textbf{presentational components} only related with
-the concern about \emph{how things look}. This component literally
-doesn't know anything about the service in question which makes the
-component highly flexible and reusable. They are controlled solely from
-the outside, receiving data and dispatching unbiased events on user
-interaction.{[}5{]}
-
-On the other hand Dan Abramov outlines components which he calls
-\textbf{containers}. A container component is concerned with \emph{how
-things work}.{[}5{]} Containers acts as invisible wrappers around
-presentational components. Their job is to fetch data from child nodes,
-aggregating events, interacting with the model and push state back to
-the presentational components. In contrast to their presentational
-counterparts, containers can't life on its own without children.
-Conceptually, they show another important principle of components and
-microservices in general: \emph{high cohesion principle}.
-
-Whether designing a microservice or it's components we want related
-behavior sit together, and unrelated behavior to sit elsewhere.{[}3, p.
-30{]} High cohesion can be perceived as the \emph{Single Responsibility
-Principle} defined by Robert C. Martin: ``Gather together those things
-that change for the same reason and separate those things that change
-for different reasons.''{[}6{]} In a very quick and dirty code quality
-analysis, the quality can be measured just by counting the places
-changes in the code occur. In an arbitrary MVC system a \texttt{button}
-might be inserted in the VIEW, the CONTROLLER needs some adjustment and
-maybe the the MODEL, too. Three places for adjustment is a reasonable
-easy task for the brain.
-
-Using libraries in web development is common sense. But compared to
+blocks of microservices. And microservices are the building blocks of
+applications. Essentially the difference between microservices and
+components is just the level of abstraction. Whether a concrete
+microservice or a much more generic component, both following pretty
+much the same principles. Therefore this paper referring to both parts
+when talking about \textbf{services}.
+
+The first principle of services is the \textbf{loose coupling
+principle}: changing and deploying one service shouldn't result in
+changing other parts of the system.{[}3, p. 30{]}. Mutations or shadowed
+variables, which is happening a lot in CSS, making it hard to keep
+changes ought to only affect one place in the application. A
+\emph{browsernative microservice} therefore pushing encapsulation and
+avoiding variable mutations outside its scope as much as possible.
+Practically, CSS will be scoped and JS fosters immutable JS entities and
+avoids variables leaking into the global namespace preferring the
+\texttt{let} or \texttt{const} declaration in favor of the old
+\texttt{var} declaration.
+
+The second principle of services is the \textbf{high cohesion
+principle}: Whether designing a microservice or it's components we want
+related behavior sit together, and unrelated behavior to sit
+elsewhere.{[}3, p. 30{]} High cohesion can be expressed in a dynamic way
+as the \emph{Single Responsibility Principle}: ``Gather together those
+things that change for the same reason and separate those things that
+change for different reasons.''{[}5{]} In a very quick and dirty code
+quality analysis, the quality can be measured just by counting the
+places changes in the code occur in order to implement a functionality.
+In an arbitrary MVC system when a \texttt{button} is dropped into the
+VIEW part, the CONTROLLER needs some adjustment and maybe the MODEL,
+too. Three places for adjustment is a reasonable easy task for the
+brain. In the field of browser based development those principle is
+often violated in the separated entities HTML, JS and CSS. A
+\emph{browsernative microservice} aims for combining all of the
+resources cohesively in a single place and file.
+
+A \emph{component for the web}, or web component, is
+\textbf{self-contained} which means it embodys all needed functionality
+to get it's job done. Therefore it has a much better evolution mechanism
+in the service contracts. Changing functionality won't break other
+components. A component can progressively enhanced which guarantees
+functionality throughout different versions.
+
+Using libraries in web development is a common sense. But compared to
 libraries, a component service offers multiple advantages. A library is
@@
 Another issue with libraries is dead code elimination which means the
-process of removing code that is never going to be executed. Newer build
+process of removing code that is never going to be executed. New build
 tools for the web, like Webpack 2 or Rollup offer this feature which
-relies heavily on the static structure on ES6 modules.{[}7{]} Contrary
-libraries for the browsers are traditionally ``shipped'' as
-\textbf{immediately-invoked functions}. Bootstrapping those ones is much
-harder to archive as dead code may be part of the function itself.
-
-Native web components are necessary static ES6 modules and any capable
-bundler can choose which component is needed at runtime. Even more
-granular, a good web component is composed of static JS blocks which can
-be strapped away too.
-
-A web component is self-contained which means it embodys all needed
-functionality to get it's job done. Therefore it has a much better
-evolution mechanism in the service contracts. Changing functionality
-won't break other components. A component can progressively enhanced
-which guarantees functionality throughout different versions.
-
-Compared to libraries componentized services offers a more explicit
+relies heavily on the static structure on ES6 modules.{[}6{]} Libraries
+for the browsers are traditionally ``shipped'' as non static
+\textbf{immediately-invoked functions}. Bootstrapping those is much
+harder to archive as dead code may be part of the function itself. A
+\emph{webnative microservices} should therefore embrace the high
+cohesion principle and abandon most of the libraries in favor of static
+ES6 module to lower the amount of dead code and to increase performance
+especially at the first render.
+
+Another advantage components have over libraries is the more explicit
 interface.{[}4{]} While the functionality of a library needs
 documentation, a component functionality is exposed via the components'
-signature which comes in the fashion of an HTML element. For example, a
-button can consume attributes and becomes a primary buttons just by
-assignment.
+signature which comes in the fashion of an HTML element in
+\emph{browsernative microservices}. For example, a button component can
+consume attributes and becomes a primary buttons just by assignment
+
+.
 
@@
 designs which are copies of the communication~structures~of these
-organizations''. {[}8{]}
+organizations''. {[}7{]}
 \end{quote}
@@
 emphasizes this pragmatism on APIs in a conference talk. At ING the
-frontend demands drive the backend APIs, APIs are produces automatically
-and not even Issa knows how many APIs exists.{[}9{]} They are using a
-pattern called \textbf{backend for frontends} allowing the team working
-on a given UI also handle the server-side components.{[}3, p. 72{]} The
-perception on backend might change therefore from embodying logic to a
-data-silo only.
+frontend demands tailor the backend APIs, APIs may be produced
+automatically and not even Issa knows how many APIs exists.{[}8{]} They
+are using a pattern called \textbf{backend for frontends} empowering the
+team working to craft their UI and backend in a one-to-one
+relationship.{[}3, p. 72{]}
 
@@
 Every check-in is handled as release candidate and can be independently
-deployed by a fully automated deployment machinery.{[}9{]} While this
-example is seems opinionated, it gives a sense of a mature component
-build for the web. Due to an exhaustive amount of testing and deployment
-tools for the browser an automated infrastructure shouldn't be a
-problem.
+deployed by a fully automated deployment machinery.{[}8{]} Even though
+this example is an opinionated perception it gives a sense of a mature
+component build for the web. Due to an exhaustive amount of testing and
+deployment tools for the browser an automated infrastructure shouldn't
+be a problem.
 
@@
 most likely using Android, have decent specs mobile phones, use an
-evergreen browser but won't have a reliable internet connection.{[}10{]}
+evergreen browser but won't have a reliable internet connection.{[}9{]}
 While \emph{Progressive Enhancement} was once coined on the principle to
-build websites both for Browsers with JS and without, the new
-\emph{Progressive Enhancement} becomes more an ``offline first''
-principle. A \emph{browsernative microservice} therefore not only tries
-to cache data as much as possible, it should also bring in a lot of
-program logic as described in the previous chapters.
+build websites both for Browsers with JS and HTML only, the new
+\emph{Progressive Enhancement} tends towards an ``\textbf{offline
+first}'' build principle. A \emph{Browsernative Microservice} therefore
+not only tries to cache data as much as possible, it should also bring
+in a lot of program logic as described in the previous chapters.
 
@@
 development. In the last decade we have seen a lot of of changes in the
-way we develop for the web. New technologies like Angular and React are
-handy to use but add some future uncertainty.
-
-In the next section of this paper a lot of standardized technologies
-around the concept of \textbf{Web Components} for the browsers will be
-described. \emph{Browsernative Microservices} should be perceived as
-complementary technology. Contrary to Angular, React and other
-frameworks they have a strong interop with existing systems and can be
-used with them together.
+way we develop for the web. New approaches like the virtual DOM approach
+found their way into mainstream web development followed by frameworks
+and libraries.
+
+\emph{Browsernative Microservices} should be perceived as complementary
+technology. Contrary to Angular, React and other frameworks they have a
+strong interop with existing systems and can be used with them together.
 
 Andrew Rota, developer at Wayfair, came up with the idea to compose
-small web components around a managing React system.{[}11{]} As web
+small web components around a managing React system.{[}10{]} As web
 components are native elements there is no difference to use a native
@@
 essentially make it possible to create a scoped component but not caring
-too much on it's distribution{[}12{]}. Some people prefer the three
-specs {[}13{]}, but the majority advocating the four specs variant,
+too much on it's distribution{[}11{]}. Some people prefer the three
+specs {[}12{]}, but the majority advocating the four specs variant,
 which is listed on \sout{the official}
@@
 
-Most obvious, this spec relies on the new \emph{ES6 Class Syntax} in
-favor of the original prototype-based inheritance model.
-``Extending~\texttt{HTMLElement}~ensures the custom element inherits the
-entire DOM API and means any properties/methods that you add to the
-class become part of the element's DOM interface.''{[}14{]} Like any
-other \emph{ES6 class}, \emph{custom element}s can be sub-classed
-further on with the typical \texttt{extends} inheritance.
+This example should be self-explanatory. Notably, custom elements come
+in the fashion of \emph{ES6 Classes}\footnote{\href{https://developer.mozilla.org/en/docs/Web/JavaScript/Reference/Classes}{JavaScript
+  Classes}} in favor of the normal JavaScript prototype-based
+inheritance model. This class must inherit the base \texttt{HTMLElement}
+interface which~``ensures the newly created element inherits the entire
+DOM API and any properties/methods that you add to the class become part
+of the element's DOM interface.''{[}13{]} Like any other \emph{ES6
+class}, the new element can be specialized further on using the typical
+\texttt{extends} inheritance.
 
@@
   method is called straight after the \texttt{constructor()} on insert.
-  This block of code contains setup code, such as fetching resources or
-  rendering elements according to attributes.{[}14{]} For performance
-  issues it's highly preferable to put much code in here.
+  Typically, this method contains setup code, such as fetching resources
+  or rendering elements according to attributes.{[}13{]} For a fast
+  initial render of the page, it is highly preferable to put many
+  proceedings in favor of the constructor.
 \item
@@
 
-While getters and setters work great in the JS world they fail crossing
-the boundaries to the corresponding HTML node. Declaring
+Native DOM properties, like \texttt{id} or \texttt{onclick}, reflect
+their values between HTML and JS.{[}14, Para. 2.6.1{]} For example
+declaring the HTML
+\texttt{\textless{}hello-world\ id="hello"\textgreater{}} like this
+equals to assign the ID in JS like \texttt{Node.id\ =\ "hello"}. This
+behavior won't work out-of-the-box with methods or properties defined
+setters. For example declaring
 \texttt{\textless{}hello-world\ sayhello="mars"\textgreater{}\textless{}/hello-world\textgreater{}}
-would't work in the previous setup. A common workaround is archived by
-using the previous mentioned \texttt{attributeChangedCallback} lifecycle
-method to \textbf{reflect changing HTML attributes to JS} and/or map JS
-attributes to HTML with \texttt{this.setAttributes(...)} respectively.
-On insertion time HTML attributes might raise their hand with
-\texttt{this.hasAttributes(...)} and \texttt{this.getAttributes(...)}.
-Native DOM properties will reflect their values between HTML and JS
-automatically.{[}15, Para. 2.6.1{]}
+would't call the \texttt{sayhello} function in the previous setup.
+
+A common workaround to bind HTML and JS behavior together is archived by
+using the aforementioned lifecycle method
+\texttt{attributeChangedCallback} to \textbf{bind changing HTML
+attributes to JS properties} and to map JS attributes to HTML with
+\texttt{this.setAttributes(...)} respectively. On insertion time HTML
+attributes might trigger custom JS methods retrieving HTML attributes
+using \texttt{this.getAttributes(...)} method.
 
@@
 \textbf{context-aware custom elements}. A possible use case would be
-``theming'' a component (example taken from {[}16{]}):
+``theming'' a component (example taken from {[}15{]}):
 
@@
 universally. For a more in-depth guide, Google Engineer Eric Bidelman
-wroten a great primer on \emph{shadow DOM}{[}16{]}.
+wroten a great primer on \emph{shadow DOM}{[}15{]}.
 
@@
     \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
-    \KeywordTok{const} \NormalTok{helloTemplate }\OperatorTok{=}
-     \VariableTok{document}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'#hello'}\NormalTok{)}\OperatorTok{;}
-    \KeywordTok{this}\NormalTok{.}\AttributeTok{shadowRoot}
-     \NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{helloTemplate}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
+    \KeywordTok{const} \NormalTok{helloTemplate }\OperatorTok{=} \VariableTok{document}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'#hello'}\NormalTok{)}\OperatorTok{;}
+    \KeywordTok{this}\NormalTok{.}\VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{helloTemplate}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
    \OperatorTok{\}}
@@
 amounts of API data and manipulating the DOM, runs on the same
-thread{[}17{]}. Putting a lot of work to the single main thread can slow
+thread{[}16{]}. Putting a lot of work to the single main thread can slow
 down the web service significantly. From time to time scripts can block
@@
 
-\section{Anatomy of an browsernative
-microservice}\label{anatomy-of-an-browsernative-microservice}
-
-After getting some confidence in microservice principles and some
-technical background the paper should illustrate a practical example to
-\emph{browsernative microservices}. Googles library Polymer is a good
-place for learning about web components in depth and make use of their
-simple command line tools. One of their most famous proof of concept is
-the so-called \href{https://shop.polymer-project.org/}{Polymer Shop}
-which is a fully-fledged online shop nested within a single custom root
-element \texttt{\textless{}shop-app\textgreater{}}. It has some elements
-for routing, managing view, service worker caching, theming, etc. The
-whole shop runs as a single application only fetching and sending
-resources but never reloading altogether. Let's assume we work in a
-sales engineering team of the Polymer Shop and need to rebuild the
-checkout microservice.
-
-A usual checkout has 3 to 4 steps:
+\section{Building an browsernative
+microservice}\label{building-an-browsernative-microservice}
+
+After getting confidence in microservice principles and technical
+background the paper should bring them both together to form a
+\emph{browsernative microservices}. Needless to say the described
+example is overall simplified to only illustrate the connection between
+browsernative technologies and microservice patterns. Googles library
+Polymer is a good place for learning about web components in depth and
+make use of their simple command line tools. One of their most famous
+proof of concept is the so-called
+\href{https://shop.polymer-project.org/}{Polymer Shop} which is a
+fully-fledged online shop nested within a single root element
+\texttt{\textless{}shop-app\textgreater{}}. This app made of several
+main views and many more invisible custom elements for routing, service
+worker caching, theming, etc. The whole shop runs as a single
+application only fetching and sending resources and switching views.
+Let's assume we work in a sales engineering team of the Polymer Shop and
+need to rebuild the checkout microservice.
+
+The current checkout can be found at
+https://shop.polymer-project.org/checkout. Currently the checkout is a
+single, 671 lines of code long Polymer component incorporation all
+required fields for sign in, shipping, billing and summarizing the
+order. In the spirit of microservices we will split up the component to
+their smaller components. The shopping cart data is pulled out of a
+local storage JSON entity set up previously by another custom-element.
+Another in-memory opportunity is storing the data in the build-in
+database IndexedDB.
+
+\subsection{Checkout microservice}\label{checkout-microservice}
+
+Like a usual checkout this example has 3 to 4 steps:
 
@@
 
-Translated into a raw custom element HTML structure, the top-level
-microservice might look like the following snippet:
+Translated into a raw \textbf{Custom Element} HTML structure, the
+top-level microservice might look like the following snippet:
 
@@
 \begin{Highlighting}[]
+\NormalTok{>shop-checkout.html}
 \KeywordTok{<shop-checkout>}
@@
 
-\subsection{Service Root}\label{service-root}
-
-The core concept of the newly created \texttt{shop-checkout} is a clean
-MVC pattern where the model is a worker managed by the root element. The
-role of the root element offers a bare minimum of centralized
-orchestration as a \textbf{kind of message broker}. Messages from child
-nodes will just forwarded to the worker and answers passed back to the
-child nodes. Therefore, the basic microservice communication looks like
-this
+Yet already we see the simplicity around web components as they persue a
+clear structure. Each of the child components should act independently
+over other child nodes utilizing the loose coupling principle. Each
+child ships all the HTML, CSS and JS code needed to fulfil its work
+following the high cohesion principle. Each component may contain
+different views to accommodate different bounded contexts resulting from
+different devices. And last but not least, all of them communicate over
+an unobstrusive message bus via the service root component
+\texttt{shop-checkout}.
+
+Before diving deeper into implementation, its worth to clarify the
+\textbf{architectural pattern} behind components. Any reader of the
+paper came across ReactJS / Redux, the concept of components may look
+familiar. Dan Abramov, the creator of Redux, once defined a simple
+dichotomous pattern for creating UI components. Firstly, he came up with
+the idea of \textbf{presentational components} only related with the
+concern about \emph{how things look}. This component literally doesn't
+know anything about the service in question which makes the component
+highly flexible and reusable. They are controlled solely from the
+outside, receiving data and dispatching unbiased events on user
+interaction.{[}17{]} Most probably every presentational component
+embodies more HTML/CSS markup and less JS code. It should encapsulate
+its styles from bleeding out and protect its styles being overwritten.
+Furthermore, it may contain several templates to change it's look on
+different demands.
+
+Secondly, Abramov described components he refers as \textbf{containers}.
+A container component is concerned with \emph{how things work}.{[}17{]}
+Containers acts as invisible wrappers around presentational components
+acting more in the sense of UNIX filters. Their job is to fetch data
+from child nodes, aggregating events, interacting with the model and
+push state back to the presentational components. Consequently, they
+might contain more JS and less, if any, HTML markup. We probably don't
+need to utilize ShadowDOM as no styles are involved.
+
+Last but not least, there are \textbf{build-in components}, which is
+every traditional HTMLElement. They are solely controlled and styled
+from the outside and must therfore wrapped in presentationals and/or
+containers.
+
+Lets start the service description top-down with the \textbf{root
+container}.
+
+\subsubsection{Root container}\label{root-container}
+
+Messages from child nodes will just forwarded to the worker and answers
+passed back to the child nodes. Therefore, the basic microservice
+communication looks like this
 
@@
 vast majority of companies developing for the web which are grouped
-around tasks.{[}9{]} A very common structure is formed by the technology
+around tasks.{[}8{]} A very common structure is formed by the technology
 stack (UX Designers, Frontend- \& Backend Developers) or by separating
@@
 
+\hypertarget{ref-Martin}{}
+{[}5{]} R. C. Martin, ``The single responsibility principle.''
+{[}Online{]}. Available:
+\url{http://programmer.97things.oreilly.com/wiki/index.php/The_Single_Responsibility_Principle}
+
+\hypertarget{ref-Rauschmayer2015}{}
+{[}6{]} D. A. Rauschmayer, ``Tree-shaking with webpack 2 and babel 6.''
+2015 {[}Online{]}. Available:
+\url{http://www.2ality.com/2015/12/webpack-tree-shaking.html}
+
+\hypertarget{ref-Conway1968}{}
+{[}7{]} M. E. Conway, ``How do committees invent?'' 1968 {[}Online{]}.
+Available: \url{http://www.melconway.com/Home/Committees_Paper.html}
+
+\hypertarget{ref-Issa2016}{}
+{[}8{]} B. Issa, ``The way of the web.'' Polymer Summit 2016, Oct-2016
+{[}Online{]}. Available:
+\url{https://www.youtube.com/watch?v=8ZTFEhPBJEE}
+
+\hypertarget{ref-Lawson2016}{}
+{[}9{]} N. Lawson, ``Progressive enhancement isn't dead, but it smells
+funny.'' 2016 {[}Online{]}. Available:
+\url{https://nolanlawson.com/2016/10/13/progressive-enhancement-isnt-dead-but-it-smells-funny/}
+
+\hypertarget{ref-Rota2015}{}
+{[}10{]} A. Rota, ``React.js conf 2015 - the complementarity of react
+and web components.'' 2015 {[}Online{]}. Available:
+\url{https://www.youtube.com/watch?t=124\&v=g0TD0efcwVg}
+
+\hypertarget{ref-Buchner2016}{}
+{[}11{]} D. Buchner, ``Demythstifying web components,'' 2016
+{[}Online{]}. Available:
+\url{http://www.backalleycoder.com/2016/08/26/demythstifying-web-components/}
+
+\hypertarget{ref-vanKesteren2014}{}
+{[}12{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
+{[}Online{]}. Available:
+\url{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}
+
+\hypertarget{ref-Bidelman2016}{}
+{[}13{]} E. Bidelman, ``Custom elements v1: reusable web components.''
+2016 {[}Online{]}. Available:
+\url{https://developers.google.com/web/fundamentals/primers/customelements/}.
+{[}Accessed: 01-Dec-2016{]}
+
+\hypertarget{ref-HTML}{}
+{[}14{]} \emph{HTML living standard --- last updated 11 january 2017}.
+{[}Online{]}. Available: \url{https://html.spec.whatwg.org/multipage/}
+
+\hypertarget{ref-Bidelman2016shadow}{}
+{[}15{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
+2016 {[}Online{]}. Available:
+\url{https://developers.google.com/web/fundamentals/getting-started/primers/shadowdom}
+
+\hypertarget{ref-Bidelman2010}{}
+{[}16{]} E. Bidelman, ``The basics of web workers.'' 2010 {[}Online{]}.
+Available: \url{https://www.html5rocks.com/en/tutorials/workers/basics/}
+
 \hypertarget{ref-Abramov2015}{}
-{[}5{]} D. Abramov, ``Presentational and Container Components --
+{[}17{]} D. Abramov, ``Presentational and Container Components --
 Medium.'' 2015 {[}Online{]}. Available:
@@
 
-\hypertarget{ref-Martin}{}
-{[}6{]} R. C. Martin, ``The single responsibility principle.''
-{[}Online{]}. Available:
-\url{http://programmer.97things.oreilly.com/wiki/index.php/The_Single_Responsibility_Principle}
-
-\hypertarget{ref-Rauschmayer2015}{}
-{[}7{]} D. A. Rauschmayer, ``Tree-shaking with webpack 2 and babel 6.''
-2015 {[}Online{]}. Available:
-\url{http://www.2ality.com/2015/12/webpack-tree-shaking.html}
-
-\hypertarget{ref-Conway1968}{}
-{[}8{]} M. E. Conway, ``How do committees invent?'' 1968 {[}Online{]}.
-Available: \url{http://www.melconway.com/Home/Committees_Paper.html}
-
-\hypertarget{ref-Issa2016}{}
-{[}9{]} B. Issa, ``The way of the web.'' Polymer Summit 2016, Oct-2016
-{[}Online{]}. Available:
-\url{https://www.youtube.com/watch?v=8ZTFEhPBJEE}
-
-\hypertarget{ref-Lawson2016}{}
-{[}10{]} N. Lawson, ``Progressive enhancement isn't dead, but it smells
-funny.'' 2016 {[}Online{]}. Available:
-\url{https://nolanlawson.com/2016/10/13/progressive-enhancement-isnt-dead-but-it-smells-funny/}
-
-\hypertarget{ref-Rota2015}{}
-{[}11{]} A. Rota, ``React.js conf 2015 - the complementarity of react
-and web components.'' 2015 {[}Online{]}. Available:
-\url{https://www.youtube.com/watch?t=124\&v=g0TD0efcwVg}
-
-\hypertarget{ref-Buchner2016}{}
-{[}12{]} D. Buchner, ``Demythstifying web components,'' 2016
-{[}Online{]}. Available:
-\url{http://www.backalleycoder.com/2016/08/26/demythstifying-web-components/}
-
-\hypertarget{ref-vanKesteren2014}{}
-{[}13{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
-{[}Online{]}. Available:
-\url{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}
-
-\hypertarget{ref-Bidelman2016}{}
-{[}14{]} E. Bidelman, ``Custom elements v1: reusable web components.''
-2016 {[}Online{]}. Available:
-\url{https://developers.google.com/web/fundamentals/primers/customelements/}.
-{[}Accessed: 01-Dec-2016{]}
-
-\hypertarget{ref-HTML}{}
-{[}15{]} \emph{HTML living standard --- last updated 11 january 2017}.
-{[}Online{]}. Available: \url{https://html.spec.whatwg.org/multipage/}
-
-\hypertarget{ref-Bidelman2016shadow}{}
-{[}16{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
-2016 {[}Online{]}. Available:
-\url{https://developers.google.com/web/fundamentals/getting-started/primers/shadowdom}
-
-\hypertarget{ref-Bidelman2010}{}
-{[}17{]} E. Bidelman, ``The basics of web workers.'' 2010 {[}Online{]}.
-Available: \url{https://www.html5rocks.com/en/tutorials/workers/basics/}
-
 \end{document}

### 2017-02-20 paper.tex
pdf/tex build

@@
 services, each running in its own process and communicating with
-lightweight mechanisms, often an HTTP resource API.''{[}4{]} This
-definition aligns well with the idea of \textbf{Single-Page
-applications} (SPAs) in the browser composed of own services or
-sometimes other third-party services.
-
-Further along, microservices incorporate many other ideas, like
-\emph{domain-driven design} where it pursues the incorporation of real
-world structure in the code.{[}3, p. 2{]} Or making use of
+lightweight mechanisms, often an HTTP resource API.''{[}4{]} Just yet at
+this point a reader might spot some similarities with microservices and
+the browser-based development: Both wrestling the problem of monolithic
+architecture and both using lightweight communication mechanisms. In
+fact, many big companies of ``the web'' like Amazon or Netflix
+successfully transformed their monolithic system into a service based
+system which gives a glimpse of the power behind microservices.{[}4{]}
+
+Microservices incorporate many ideas from developing scalable software,
+like \emph{domain-driven design} where it pursues the incorporation of
+real world structure in the code.{[}3, p. 2{]} Or making use of
 \emph{continuous delivery} for pushing software rapidly through
-\emph{automated deployment} mechanisms into production.{[}4{]} And, last
-but not least, microservices utilizes the idea of small teams with a lot
-of product knowledge working mostly autonomous on their very own service
-with their very own set of tools and techniques. As a primary source of
-truth this paper relies on the work of Sam Newman, who has written a
-comprehensive guide called \emph{Building Microservices} and the work of
-Fowler and Lewis. The purpose of this section is to gain confidence
-about the microservices architecture in the context of the browser
-platform.
+\emph{automated deployment} mechanisms into production.{[}4{]}
+Furthermore, microservices transcendent the technical perspective and
+reaches into the team organization. As a primary source of truth this
+paper relies on the work of Sam Newman, who has written a comprehensive
+guide called \emph{Building Microservices} and the work of Fowler and
+Lewis. The purpose of this section is to gain confidence about the
+microservices architecture in the context of the browser platform.
 
@@
 replaceable and upgradeable.'' {[}4{]} Components are the building
-blocks of microservices. While the service is a concrete piece of
-software a component is much more raw and single-minded following the
-\emph{loose coupling principle}: changing and deploying one service
-shouldn't result in changing other parts of the system.{[}3, p. 30{]}.
-Small components and big microservices are similar on their principles.
-
-Any reader of the paper familiar with ReactJS / Redux, the concept of
-components isn't anything new. Dan Abramov, the creator of Redux, once
-defined a simple dichotomous pattern for creating UI components. In his
-believe, there are \textbf{presentational components} only related with
-the concern about \emph{how things look}. This component literally
-doesn't know anything about the service in question which makes the
-component highly flexible and reusable. They are controlled solely from
-the outside, receiving data and dispatching unbiased events on user
-interaction.{[}5{]}
-
-On the other hand Abramov described components he named
-\textbf{containers}. A container component is concerned with \emph{how
-things work}.{[}5{]} Containers acts as invisible wrappers around
-presentational components. Their job is to fetch data from child nodes,
-aggregating events, interacting with the model and push state back to
-the presentational components. In contrast to their presentational
-counterparts, containers can't life on its own without children.
-Conceptually, they show another important principle of components and
-microservices in general: \emph{high cohesion principle}.
-
-Whether designing a microservice or it's components we want related
-behavior sit together, and unrelated behavior to sit elsewhere.{[}3, p.
-30{]} High cohesion can be expressed in a dynamic way as the
-\emph{Single Responsibility Principle}: ``Gather together those things
-that change for the same reason and separate those things that change
-for different reasons.''{[}6{]} In a very quick and dirty code quality
-analysis, the quality can be measured just by counting the places
-changes in the code occur. In an arbitrary MVC system when a
-\texttt{button} is dropped into the VIEW part, the CONTROLLER needs some
-adjustment for sure and maybe the MODEL, too. Three places for
-adjustment is a reasonable easy task for the brain. In the field of
-browser based development those principle is often violated in the
-separated entities HTML, JS and CSS.
+blocks of microservices. And microservices are the building blocks of
+applications. Essentially the difference between microservices and
+components is just the level of abstraction. Whether a concrete
+microservice or a much more generic component, both following pretty
+much the same principles. Therefore this paper referring to both parts
+when talking about \textbf{services}.
+
+The first principle of services is the \textbf{loose coupling
+principle}: changing and deploying one service shouldn't result in
+changing other parts of the system.{[}3, p. 30{]}. Mutations or shadowed
+variables, which is happening a lot in CSS, making it hard to keep
+changes ought to only affect one place in the application. A
+\emph{browsernative microservice} therefore pushing encapsulation and
+avoiding variable mutations outside its scope as much as possible.
+Practically, CSS will be scoped and JS fosters immutable JS entities and
+avoids variables leaking into the global namespace preferring the
+\texttt{let} or \texttt{const} declaration in favor of the old
+\texttt{var} declaration.
+
+The second principle of services is the \textbf{high cohesion
+principle}: Whether designing a microservice or it's components we want
+related behavior sit together, and unrelated behavior to sit
+elsewhere.{[}3, p. 30{]} High cohesion can be expressed in a dynamic way
+as the \emph{Single Responsibility Principle}: ``Gather together those
+things that change for the same reason and separate those things that
+change for different reasons.''{[}5{]} In a very quick and dirty code
+quality analysis, the quality can be measured just by counting the
+places changes in the code occur in order to implement a functionality.
+In an arbitrary MVC system when a \texttt{button} is dropped into the
+VIEW part, the CONTROLLER needs some adjustment and maybe the MODEL,
+too. Three places for adjustment is a reasonable easy task for the
+brain. In the field of browser based development those principle is
+often violated in the separated entities HTML, JS and CSS. A
+\emph{browsernative microservice} aims for combining all of the
+resources cohesively in a single place and file.
+
+A \emph{component for the web}, or web component, is
+\textbf{self-contained} which means it embodys all needed functionality
+to get it's job done. Therefore it has a much better evolution mechanism
+in the service contracts. Changing functionality won't break other
+components. A component can progressively enhanced which guarantees
+functionality throughout different versions.
 
@@
 Another issue with libraries is dead code elimination which means the
-process of removing code that is never going to be executed. Newer build
+process of removing code that is never going to be executed. New build
 tools for the web, like Webpack 2 or Rollup offer this feature which
-relies heavily on the static structure on ES6 modules.{[}7{]} Libraries
+relies heavily on the static structure on ES6 modules.{[}6{]} Libraries
 for the browsers are traditionally ``shipped'' as non static
 \textbf{immediately-invoked functions}. Bootstrapping those is much
-harder to archive as dead code may be part of the function itself.
-
-Native web components are necessary static ES6 modules and any capable
-bundler can choose which component is needed at runtime. Even more
-granular, a good web component is composed of static JS dependencies
-which can be deflated or strapped away too.
-
-A \emph{component for the web}, or web component, is
-\textbf{self-contained} which means it embodys all needed functionality
-to get it's job done. Therefore it has a much better evolution mechanism
-in the service contracts. Changing functionality won't break other
-components. A component can progressively enhanced which guarantees
-functionality throughout different versions.
-
-Compared to libraries componentized services offers a more explicit
+harder to archive as dead code may be part of the function itself. A
+\emph{webnative microservices} should therefore embrace the high
+cohesion principle and abandon most of the libraries in favor of static
+ES6 module to lower the amount of dead code and to increase performance
+especially at the first render.
+
+Another advantage components have over libraries is the more explicit
 interface.{[}4{]} While the functionality of a library needs
 documentation, a component functionality is exposed via the components'
-signature which comes in the fashion of an HTML element. For example, a
-button can consume attributes and becomes a primary buttons just by
-assignment.
+signature which comes in the fashion of an HTML element in
+\emph{browsernative microservices}. For example, a button component can
+consume attributes and becomes a primary buttons just by assignment
+
+.
 
@@
 frontend demands tailor the backend APIs, APIs may be produced
-automatically and not even Issa knows how many APIs exists.{[}9{]} They
+automatically and not even Issa knows how many APIs exists.{[}8{]} They
 are using a pattern called \textbf{backend for frontends} empowering the
@@
 Every check-in is handled as release candidate and can be independently
-deployed by a fully automated deployment machinery.{[}9{]} Even though
+deployed by a fully automated deployment machinery.{[}8{]} Even though
 this example is an opinionated perception it gives a sense of a mature
@@
 most likely using Android, have decent specs mobile phones, use an
-evergreen browser but won't have a reliable internet connection.{[}10{]}
+evergreen browser but won't have a reliable internet connection.{[}9{]}
 While \emph{Progressive Enhancement} was once coined on the principle to
@@
 Andrew Rota, developer at Wayfair, came up with the idea to compose
-small web components around a managing React system.{[}11{]} As web
+small web components around a managing React system.{[}10{]} As web
 components are native elements there is no difference to use a native
@@
 DOM API and any properties/methods that you add to the class become part
-of the element's DOM interface.''{[}14{]} Like any other \emph{ES6
+of the element's DOM interface.''{[}13{]} Like any other \emph{ES6
 class}, the new element can be specialized further on using the typical
@@
   Typically, this method contains setup code, such as fetching resources
-  or rendering elements according to attributes.{[}14{]} For a fast
+  or rendering elements according to attributes.{[}13{]} For a fast
   initial render of the page, it is highly preferable to put many
@@
 Native DOM properties, like \texttt{id} or \texttt{onclick}, reflect
-their values between HTML and JS.{[}15, Para. 2.6.1{]} For example
+their values between HTML and JS.{[}14, Para. 2.6.1{]} For example
 declaring the HTML
@@
     \DataTypeTok{bubbles}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,} \CommentTok{//important!}
-    \DataTypeTok{detail}\OperatorTok{:} \StringTok{'Contains string or object'}
+    \DataTypeTok{detail}\OperatorTok{:} \StringTok{'Contains scalar or object'}
    \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
@@
 background the paper should bring them both together to form a
-\emph{browsernative microservices}. Googles library Polymer is a good
-place for learning about web components in depth and make use of their
-simple command line tools. One of their most famous proof of concept is
-the so-called \href{https://shop.polymer-project.org/}{Polymer Shop}
-which is a fully-fledged online shop nested within a single root element
+\emph{browsernative microservices}. Needless to say the described
+example is overall simplified to only illustrate the connection between
+browsernative technologies and microservice patterns. Googles library
+Polymer is a good place for learning about web components in depth and
+make use of their simple command line tools. One of their most famous
+proof of concept is the so-called
+\href{https://shop.polymer-project.org/}{Polymer Shop} which is a
+fully-fledged online shop nested within a single root element
 \texttt{\textless{}shop-app\textgreater{}}. This app made of several
@@
 Let's assume we work in a sales engineering team of the Polymer Shop and
-need to rebuild the checkout microservice. Needless to say that many
-ideas mentioned here are opinions of the author and by no means
-standardized patterns.
+need to rebuild the checkout microservice.
 
@@
 required fields for sign in, shipping, billing and summarizing the
-order. In the spirit of microservices we will split up the component to
-their smaller components. The shopping cart data is pulled out of a
+order. In the spirit of microservices we will split up the microservice
+into fine grained components. The shopping cart data is pulled out of a
 local storage JSON entity set up previously by another custom-element.
@@
 
-A usual checkout has 3 to 4 steps:
+Like a usual checkout this example has 3 to 4 steps:
 
@@
 
-Messages from child nodes will just forwarded to the worker and answers
-passed back to the child nodes. Therefore, the basic microservice
-communication looks like this
+Yet already we see the simplicity arouse from web components as they
+persue a clear structure. Each of the child components should act
+independently over other child nodes utilizing the loose coupling
+principle. Each child ships all the HTML, CSS and JS code needed to
+fulfil its work following the high cohesion principle. Each component
+may contain different views to accommodate different bounded contexts
+resulting from different devices. And last but not least, all of them
+communicate over an unobstrusive message bus via the service root
+component \texttt{shop-checkout}.
+
+Before diving deeper into implementation, its worth to clarify the
+\textbf{architectural pattern} behind components. Any reader of the
+paper came across ReactJS / Redux, the concept of components may look
+familiar. Dan Abramov, the creator of Redux, once defined a simple
+dichotomous pattern for creating UI components. Firstly, he came up with
+the idea of \textbf{presentational components} only related with the
+concern about \emph{how things look}. This component literally doesn't
+know anything about the service in question which makes the component
+highly flexible and reusable. They are controlled solely from the
+outside, receiving data and dispatching unbiased events on user
+interaction.{[}17{]} Most probably every presentational component
+embodies more HTML/CSS markup and less JS code. It should encapsulate
+its styles from bleeding out and protect its styles being overwritten.
+Furthermore, it may contain several templates to change it's look on
+different demands.
+
+Secondly, Abramov described components he refers as \textbf{containers}.
+A container component is concerned with \emph{how things work}.{[}17{]}
+Containers acts as invisible wrappers around presentational components
+acting more in the sense of UNIX filters. Their job is to fetch data
+from child nodes, aggregating events, interacting with the model and
+push state back to the presentational components. Consequently, they
+might contain more JS and less, if any, HTML markup. We probably don't
+need to utilize ShadowDOM as no styles are involved.
+
+Last but not least, there are \textbf{build-in components}, which is
+every build-in HTMLElement like the HTMLButtonElement. They are solely
+controllable and styleable from the outside and must therefore wrapped
+in presentational components and/or containers.
+
+Lets start the service description top-down beginning with the
+\textbf{root container} managing the overall service.
+
+\subsubsection{Root container}\label{root-container}
+
+The root container \texttt{shop-checkout} is basically just an
+encapsulation layer in terms of service functionalities. Encapsulation
+of CSS won't be necessary at this point as no styling is involved. On
+initialization the \texttt{shop-checkout} might perform following tasks:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Importing child components with \textbf{HTML Imports}
+\item
+  Initialize the service in a logical top to bottom order like in the
+  snippet above
+\item
+  Initializing a \textbf{model thread} with
+  \texttt{new\ Worker(\textquotesingle{}checkout-model.js\textquotesingle{})};
+\item
+  Setting up a \texttt{postMessage} stub to dispatch messages to the
+  model
+\item
+  Setting up an event listener for messages from the model
+\item
+  Setting up an event listener for \textbf{custom events} from the child
+  nodes
+\end{enumerate}
+
+On runtime the root container acts mostly like a \textbf{message
+dispatcher} implementing the microservice idea of smart endpoints and
+dumb pipes.
+
+Every custom child node is eligible aggregate subordinate events from
+their children, buffer them and interact with the model trough via a
+\textbf{dumb message system}. To talk to the model any child node can
+dispatch a \textbf{unified event containing a unified message object}
+like the following example illustrates:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{// Within every child node class}
+\NormalTok{set }\AttributeTok{dispatch}\NormalTok{(msg) }\OperatorTok{\{}
+  \KeywordTok{this}\NormalTok{.}\AttributeTok{dispatchEvent}\NormalTok{(}
+    \KeywordTok{new} \AttributeTok{CustomEvent}\NormalTok{(}\StringTok{'checkout'}\OperatorTok{,} \OperatorTok{\{}
+      \DataTypeTok{bubbles}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,}
+      \DataTypeTok{detail}\OperatorTok{:} \NormalTok{msg}
+    \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+  \NormalTok{)}\OperatorTok{;}
+\OperatorTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+Calling \texttt{this.dispatch(msg)} within a child node will trigger an
+event bubbling upstream. The service root will implement a simple event
+listener for \texttt{checkout} events. In a second step the service root
+add information about the original target (like
+\emph{tag-name\#id.className}) to the message so the message can be
+processed in the model thread and eventually pushed back correctly to
+the sender.
+
+Summarizing the unidirectional model-view communication from the service
+root perspective looks like following sketch:
 
 \begin{verbatim}
-VIEW root           |               MODEL worker
+VIEW thread         |           MODEL thread
                     |
-Event  --------Event msg--------->    Msg
-    |               |               Handler
-    |               |                   |
-Idle                |               Effekt
-    |               |                   |
-  Msg               |                   |
-Handler <-----Action msg----------  Action                                      
+Checkout ------Event Msg--------->    Msg
+  Event             |                Handler
+                    |                   |
+                    |                Effect
+                    |                   |
+  Msg               |                Postmessage
+Handler <-----Action Msg----------    Action
                     |                   
@@
 actions returned to sender. Effects can be created by external messages,
-like subscription to an WebSocket, too. Effects may be created by
-fetching additional resources from the server. The VIEW side of the root
-element forwards all the messages and won't be bothered about the
-content.
+like subscription to an WebSocket, too. Effects may be created with
+additional resources from the server or syncing with local storage like
+in the polymer-shop.
+
+While the msg handlers are basic ``dumb'' switches the
+\textbf{smartness} solely arouse from intelligent controllers processing
+the message. Every component must implement a \textbf{unified socket} in
+it's service contracts like a property on the custom element:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{// Within every child node class}
+\NormalTok{set }\AttributeTok{receive}\NormalTok{(msg) }\OperatorTok{\{}
+  \CommentTok{// do something smart}
+\OperatorTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+This communication flow is designed to \textbf{fail graceful} as simply
+nothing happens if one or another switch is not defined. Furthermore it
+incorporates lots of possibilities for \textbf{evolutionary design} as
+child nodes can be loosely dropped in or teared apart. Middleware
+container wrappers can be pulled in between child nodes and service root
+to do some extra work and and even the message object can be enhanced
+freely as long as it won't break with the message conventions. Changing
+or enhancing any functionality requires only to touch the controllers in
+the child nodes in question and the endpoint section at the model.
+Furthermore, \textbf{infrastructure automation} might be achieved by
+dynamically evaluating the mounted child nodes and registering the
+``backend'' model switches on build time. Due to it's standardized
+message system the components are testable within standardized tests.
+
+To work in a larger scale, we might define a base component which has
+the default message and socket properties in its guts and extend every
+new element from this base class.
+
+Another yet more powerful approach can be achieved by calculating the
+new shape of the DOM in the worker directly and just patch the DOM in
+the view layer.\footnote{A good example of this approach can be found at
+  \href{http://www.pocketjavascript.com/blog/2015/11/23/introducing-pokedex-org}{Introducing
+  Pokedex}} This approach is certainly more complicated but worth to
+evaluate further more.
+
+\subsubsection{Presentational
+components}\label{presentational-components}
 
@@
 
-endgeräte werden besser
-
-multicore
-
-query languages from handheld
+cohesive HTML JS CSS
 

### 2017-02-21 build/paper.tex
pdf/tex build

@@
 components is just the level of abstraction. Whether a concrete
-microservice or a much more generic component, both following pretty
-much the same principles. Therefore this paper referring to both parts
-when talking about \textbf{services}.
+microservice or a much more generic component, both share a similar set
+of principles. Therefore this paper referring to both parts when talking
+about \textbf{services}.
 
@@
 Practically, CSS will be scoped and JS fosters immutable JS entities and
-avoids variables leaking into the global namespace preferring the
-\texttt{let} or \texttt{const} declaration in favor of the old
-\texttt{var} declaration.
+avoids variables leaking into the global namespace.
 
@@
 places changes in the code occur in order to implement a functionality.
-In an arbitrary MVC system when a \texttt{button} is dropped into the
-VIEW part, the CONTROLLER needs some adjustment and maybe the MODEL,
-too. Three places for adjustment is a reasonable easy task for the
-brain. In the field of browser based development those principle is
-often violated in the separated entities HTML, JS and CSS. A
-\emph{browsernative microservice} aims for combining all of the
-resources cohesively in a single place and file.
-
-A \emph{component for the web}, or web component, is
-\textbf{self-contained} which means it embodys all needed functionality
-to get it's job done. Therefore it has a much better evolution mechanism
-in the service contracts. Changing functionality won't break other
-components. A component can progressively enhanced which guarantees
-functionality throughout different versions.
-
-Using libraries in web development is a common sense. But compared to
-libraries, a component service offers multiple advantages. A library is
-only loosely coupled to the implementation and therefore hard to track
-in functionality. Changing a library may result in an unforeseen amount
-of time fixing implementations. It is not unusual to see websites
-embodying different versions of the same library (like with JQuery).
-Another issue with libraries is dead code elimination which means the
-process of removing code that is never going to be executed. New build
-tools for the web, like Webpack 2 or Rollup offer this feature which
-relies heavily on the static structure on ES6 modules.{[}6{]} Libraries
-for the browsers are traditionally ``shipped'' as non static
-\textbf{immediately-invoked functions}. Bootstrapping those is much
-harder to archive as dead code may be part of the function itself. A
-\emph{webnative microservices} should therefore embrace the high
-cohesion principle and abandon most of the libraries in favor of static
-ES6 module to lower the amount of dead code and to increase performance
-especially at the first render.
-
-Another advantage components have over libraries is the more explicit
-interface.{[}4{]} While the functionality of a library needs
-documentation, a component functionality is exposed via the components'
-signature which comes in the fashion of an HTML element in
-\emph{browsernative microservices}. For example, a button component can
-consume attributes and becomes a primary buttons just by assignment
-
-.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{<my-botton}\OtherTok{ isPrimary}\KeywordTok{>}\NormalTok{Primary}\KeywordTok{</my-botton>}
-\end{Highlighting}
-\end{Shaded}
+An arbitrary threefold MVC system should require a maximum of three
+changes to implement or change functionalities. The problem in browser
+based development is not only the global paradigm which makes changes
+deliberately unpredictable. The high cohesion principle is violated by
+the traditional separation along the technological entities
+\textbf{HTML, JS and CSS}. Understanding the relation of HTML markup and
+another CSS file adds incidental complexity. Different approaches
+emerged over the recent years to join the forces. Most notable the
+dedicated frontend language \emph{elm}\footnote{\href{http://elm-lang.org/}{elm
+  lang}} where model, view and controller sit in one place. A
+browsernative service aims for a combinement of the web native trinity
+HTML, JS and CSS.
+
+Traditional web development relies on \textbf{libraries} to enhance the
+service capabilities of the web platform. Compared to libraries, a
+component service offers multiple advantages for building, deployment
+and shipping. As expressed earlier a \emph{components for the web}, or
+web component, is \textbf{self-contained} which means it embodies all
+needed functionality to get it's job done. Therefore it has a much
+better evolution mechanism in the service contracts. Changing
+functionality won't break other services. A component can progressively
+enhanced which guarantees functionality throughout different versions. A
+library is only loosely coupled to the implementation and therefore hard
+to track in functionality. Changing a library may result in an
+unforeseen amount of time fixing implementations. It is not unusual to
+see websites embodying different versions of the same library to
+guarantee functionality but lowers page performance significantly.
+
+Another issue where web components stand out is related to performance
+and especially the critical first page rendering. Libraries for the
+browsers are traditionally ``shipped'' as non static immediately-invoked
+functions run immediately on page load. Following the \emph{Google RAIL}
+model, a user-centric performance measurement, a page load ought be less
+than 1 second to catch up user's attention.{[}6{]} There are many ways
+for optimizing the critical first render but as a rule of thumb a
+build-in web component might be always superior to libraries in terms of
+first rendering. The fine-grained lifecycle methods which will be
+described in the technical section of this paper give the developer far
+reaching optimization opportunities.
+
+The last argument in favor of components over libraries is the more
+explicit interface.{[}4{]} While the functionality of a library needs
+documentation to be accessible a component functionality is exposed via
+the components' signature. The markup language HTML is by design
+equipped with an expressive syntax and steering a web component using
+attributes and values should be most straightforward even to
+unexperienced web developers.
 
@@
 technology stack every member gains high competence about the
-functionality of the service.
+functionality of the service architecture which can be critical for
+evolving the service over its whole lifecycle.
 
 Real-world domains tend to be complex and multifaceted. To unfold their
-complexity, domains can subdivided into \textbf{bounded contexts}.{[}3,
-p. 31{]} For example, customer service is a business domain. Customers
-again have different contexts depending on their demands. One context
-can be sales, another context could be support. Every context makes
-different assumptions about the underlying model. Each bounded context
-draws an explicit interface where it decides what models to share with
-other contexts.{[}3, p. 30{]} Each context can derived into multiple
-microservices or, talking about \emph{Browsernative Microservices},
-interfaces to the customers.
+complexity, domains can subdivided into \textbf{bounded context}.{[}3,
+p. 31{]} For example, customer service is a business domain but with
+varying bounded context. One context can be sales, another context could
+be support. Every context makes different assumptions about the
+underlying model and draws an explicit interface where it decides what
+to share with other contexts.{[}3, p. 30{]} Evaluating each bounded
+context within each business domain will eventually shape the data
+structures and services in question. This methodology can be iterated
+over and over again. A sales service for example might be evaluated to
+different sales context resulting in differing interfaces on devices
+shaped by browsernative microservices.
 
@@
 
-For many companies working in the spheres of the internet the
-client-side is highly important for their business. In fact, business
-goals and capabilities can be derived from the front-end needs. The
-state of the web is not only a story of numerous artifacts, it is also a
-story of an highly fragmented market along devices, operating systems,
-differing sizes and functionalities. Different devices again have
-different assumptions about the technology stack. Splitting teams along
-the stack results in an slow paced back and forth negotiation for every
-change to be made. As browser technologies, design guidelines and
-devices change in a fast paced manner it makes absolutely sense to shift
-responsibility towards the teams altogether. A \emph{browsernative
-microservice} embraces commitment over its whole lifecycle.
+For many companies working in the spheres of the internet the client
+side is highly important for their business. In fact, business goals and
+capabilities can be derived from the front-end needs. The state of the
+web is not only a story of numerous artifacts, it is also a story of an
+highly fragmented market along devices, operating systems, differing
+sizes and functionalities. Different devices again have different
+assumptions about the technology stack. Splitting teams along the stack
+results in an slow paced back and forth negotiation for every change to
+be made. As browser technologies, design guidelines and devices change
+frequently it makes absolutely sense to shift responsibility towards the
+teams altogether. By focusing the service into clear defined business
+boundaries, it is easier to define a smart API of the service.
 
@@
 
-Microservices for the browsers aim for side-effects like changing the UI
-given to input parameters or emitting browser events. An isolated
-microservice won't make too much sense after all. To ensure
-communication between services, flexible yet powerful communication
-channels must be established.
-
-\emph{Smart endpoints and dumb pipes} is coined to the approach of
-designing communication mostly decoupled and as cohesive as
-possible.{[}4{]} In the analogy to the real world a message channel
-should look like sending a letter: Two smart endpoints (sender and
-receiver) and a mostly unified ``dumb'' letter format and channel. Each
-endpoint owns it specific domain logic. In the browsernative context we
-already emphasized the role of presentational components and containers.
-The ``smartness'' of presentational components is altering the UI and
-reacting on users input. The logic of a container follows the idea of an
-filter in a Unix sense - receiving a request, applying logic as
-appropriate and producing a response.{[}4{]}
-
-Cited earlier in this paper microservices often rely on simple HTTP
-request-response with resource API's and lightweight messaging as
-communication protocols.{[}4{]} Those technologies highly accessible and
-widely adopted and most probably usable for a browsernative service. In
-the implementation part of this paper a reader will explore the
-combination of RESTish protocols and a lightweight browser message bus
-in action.
-
-By focusing the service into clear defined business boundaries, it is
-easier to define a smart API of the service which in our case will be a
-set of attributes for the HTML element bound to the service.
+To ensure the microservice functionality among teams and different
+services requires thoughtful decentralization. Emphasizing once more the
+real-world capabilities of microservices a message channel architecture
+can be derived from patterns known from traditional postal services. A
+physical letter has only two smart endpoints entitled to read and
+process the message while packaging likewise the connection itself is
+maximized towards unification. \emph{Smart endpoints and dumb pipes} is
+coined to the approach of designing communication mostly decoupled and
+as cohesive as possible.{[}4{]} Applying this rules to the web platform
+can result into building unified JSON message objects passed along
+``dumb'' middleware components. A typical browser event comes close to
+this definition and suits arguably well for in-memory communication
+between web components and microservices.
+
+Microservices heavily rely on simple HTTP request-response with resource
+APIs and lightweight messaging.{[}4{]} REST APIs are technology agnostic
+which aligns well with the idea of decentralization.{[}3, p. 247{]} The
+advantage of this overall simple communication model is the suitability
+for both frontend-backend likewise backend-backend communication. A
+service therefore can evolve from an heavy backend with a lot of network
+roundtrips to a leaner backend seamlessly. The browser build-in fetch
+API which is essentially a HTTP request can be heavily incorporated into
+a browsernative microservice to ensure communication between services.
 
@@
 
-Microservices are separate entities and decentralization is therefore a
-meta concept of microservices. Distributed service responsibility owned
-by the team is one aspect. Gluing independent service bricks to a whole
-distributed system talking over the network protocols another one.
+Microservices are separate entities and decentralization is important to
+ensure autonomy. Distributed service responsibility owned by the team
+promoting shared governance and technology agnostic communication
+channels work towards a mostly decoupled service architecture.
 
@@
 ideas and implementations open source. The previously mentioned ReactJS
-for example is a brainchild of Facebook. In fact, many tools and
-techniques are byproduct of vital interaction of concrete domain
-problems and their implementations.
+for example is a product of Facebooks need to ensure a consistent
+frontend experience. In fact, many tools and techniques are byproduct of
+vital interaction of concrete domain problems and their implementations.
 
 The spirit of freedom can't be applied universally to
-\emph{Browsernative Microservices} as the browser and its underlying DOM
+\emph{browsernative microservices} as the browser and its underlying DOM
 will be the limitation factor to a certain degree. Talking about the
@@
 ship no overhead or require obscure build tools. This makes them ideal
-candidates for sharing and open-sourcing. Webcomponents.org\footnote{\href{https://webcomponents.org}{Webcomponents.org}}
+candidates for sharing and open source publishing.
+Webcomponents.org\footnote{\href{https://webcomponents.org}{Webcomponents.org}}
 is a registry for ready-to-use components of every scale and purpose
@@
 
-Data Management in Microservice follows the same modular philosophy like
-the service implementation. As mentioned earlier different bounded
+Data Management in a microservice follows the same modular philosophy
+like the service implementation. As mentioned earlier different bounded
 contexts make different assumptions of the underlying models. A
@@
 costly it is a good advice to only query as much data as needed and
-cache as much as possible.
+cache as much as possible. The build-in LocalStorage or its successor
+IndexedDB are mature persistence technologies and libraries like
+PouchDB\footnote{\href{https://pouchdb.com/}{PouchDB}} even offer
+backend syncing out of the box.
 
@@
 generic build-in API accompanied by build-in frontend components.
-Instead of gluing frontend and backend together on runtime, the
+Instead of gluing frontend and backend together on runtime the
 microservice is designed holistically containing both front- and
-backends. To reduce network calls especially for mobile devices it is a
-good advice to cache data (f.e. in a global Service Worker). Revamping
-offline capabilities even further data can be stored in a browser based
-database like PouchDB. For the sake of simplicity data management won't
-be explored into depth throughout this paper.
+backends. For the sake of simplicity data management won't be explored
+into depth throughout this paper.
 
@@
 component build for the web. Due to an exhaustive amount of testing and
-deployment tools for the browser an automated infrastructure shouldn't
-be a problem.
+deployment tools for JS an automated infrastructure shouldn't be a
+problem.
 
@@
 
-\emph{Browsernative Microservices} should be perceived as complementary
+\emph{Browsernative microservices} should be perceived as complementary
 technology. Contrary to Angular, React and other frameworks they have a
@@
 specs {[}12{]}, but the majority advocating the four specs variant,
-which is listed on \sout{the official}
+which is listed on the quasi-official
 \href{http://webcomponents.org}{webcomponents.org} website. For the
@@
 required fields for sign in, shipping, billing and summarizing the
-order. In the spirit of microservices we will split up the component to
-their smaller components. The shopping cart data is pulled out of a
+order. In the spirit of microservices we will split up the microservice
+into fine grained components. The shopping cart data is pulled out of a
 local storage JSON entity set up previously by another custom-element.
-Another in-memory opportunity is storing the data in the build-in
-database IndexedDB.
 
@@
 
-Yet already we see the simplicity around web components as they persue a
-clear structure. Each of the child components should act independently
-over other child nodes utilizing the loose coupling principle. Each
-child ships all the HTML, CSS and JS code needed to fulfil its work
-following the high cohesion principle. Each component may contain
-different views to accommodate different bounded contexts resulting from
-different devices. And last but not least, all of them communicate over
-an unobstrusive message bus via the service root component
-\texttt{shop-checkout}.
+Yet already we see the simplicity arouse from web components as they
+persue a clear structure. Each of the child components should act
+independently over other child nodes utilizing the loose coupling
+principle. Each child ships all the HTML, CSS and JS code needed to
+fulfil its work following the high cohesion principle. Each component
+may contain different views to accommodate different bounded contexts
+resulting from different devices. And last but not least, all of them
+communicate over an unobstrusive message bus via the service root
+component \texttt{shop-checkout}.
 
@@
 Last but not least, there are \textbf{build-in components}, which is
-every traditional HTMLElement. They are solely controlled and styled
-from the outside and must therfore wrapped in presentationals and/or
-containers.
-
-Lets start the service description top-down with the \textbf{root
-container}.
+every build-in HTMLElement like the HTMLButtonElement. They are solely
+controllable and styleable from the outside and must therefore wrapped
+in presentational components and/or containers.
+
+Lets start the service description top-down beginning with the
+\textbf{root container} managing the overall service.
 
@@
 
-Messages from child nodes will just forwarded to the worker and answers
-passed back to the child nodes. Therefore, the basic microservice
-communication looks like this
+The root container \texttt{shop-checkout} is basically just an
+encapsulation layer in terms of service functionalities. Encapsulation
+of CSS won't be necessary at this point as no styling is involved. On
+initialization the \texttt{shop-checkout} might perform following tasks:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Importing child components with \textbf{HTML Imports}
+\item
+  Initialize the service in a logical top to bottom order like in the
+  snippet above
+\item
+  Initializing a \textbf{model thread} with
+  \texttt{new\ Worker(\textquotesingle{}checkout-model.js\textquotesingle{})};
+\item
+  Setting up a \texttt{postMessage} stub to dispatch messages to the
+  model
+\item
+  Setting up an event listener for messages from the model
+\item
+  Setting up an event listener for \textbf{custom events} from the child
+  nodes
+\end{enumerate}
+
+On runtime the root container acts mostly like a \textbf{message
+dispatcher} implementing the microservice idea of smart endpoints and
+dumb pipes.
+
+Every custom child node is eligible aggregate subordinate events from
+their children, buffer them and interact with the model trough via a
+\textbf{dumb message system}. To talk to the model any child node can
+dispatch a \textbf{unified event containing a unified message object}
+like the following example illustrates:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{// Within every child node class}
+\NormalTok{set }\AttributeTok{dispatch}\NormalTok{(msg) }\OperatorTok{\{}
+  \KeywordTok{this}\NormalTok{.}\AttributeTok{dispatchEvent}\NormalTok{(}
+    \KeywordTok{new} \AttributeTok{CustomEvent}\NormalTok{(}\StringTok{'checkout'}\OperatorTok{,} \OperatorTok{\{}
+      \DataTypeTok{bubbles}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,}
+      \DataTypeTok{detail}\OperatorTok{:} \NormalTok{msg}
+    \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+  \NormalTok{)}\OperatorTok{;}
+\OperatorTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+Calling \texttt{this.dispatch(msg)} within a child node will trigger an
+event bubbling upstream. The service root will implement a simple event
+listener for \texttt{checkout} events. In a second step the service root
+add information about the original target (like
+\emph{tag-name\#id.className}) to the message so the message can be
+processed in the model thread and eventually pushed back correctly to
+the sender.
+
+Summarizing the unidirectional model-view communication from the service
+root perspective looks like following sketch:
 
 \begin{verbatim}
-VIEW root           |               MODEL worker
+VIEW thread         |           MODEL thread
                     |
-Event  --------Event msg--------->    Msg
-    |               |               Handler
-    |               |                   |
-Idle                |               Effekt
-    |               |                   |
-  Msg               |                   |
-Handler <-----Action msg----------  Action                                      
+Checkout ------Event Msg--------->    Msg
+  Event             |                Handler
+                    |                   |
+                    |                Effect
+                    |                   |
+  Msg               |                Postmessage
+Handler <-----Action Msg----------    Action
                     |                   
@@
 
-\hypertarget{ref-Rauschmayer2015}{}
-{[}6{]} D. A. Rauschmayer, ``Tree-shaking with webpack 2 and babel 6.''
-2015 {[}Online{]}. Available:
-\url{http://www.2ality.com/2015/12/webpack-tree-shaking.html}
+\hypertarget{ref-Kearney2016}{}
+{[}6{]} M. Kearney, ``Measure performance with the rail model.'' 2016
+{[}Online{]}. Available:
+\url{https://developers.google.com/web/fundamentals/performance/rail}
 

### 2017-02-22 build/paper.tex
pdf/tex build

@@
 On the other side of this dystopian scenario, the truly modular system
-architecture abandons many of those inconsistencies. The whole system is
-divided in pluggable parts, object mutation is either traceable or
-avoided altogether in favor of immutable data structures. As Rich Hickey
-argues, design decisions should be made under the \textbf{impression of
-extending, substitution, moving, combining and repurposing}. The ability
-to reason about the program at any given time is crucial for future
-decisions and implementations. Recalling again the unidimensional nature
-of simplicity.
+architecture abandons many of those potential inconsistencies. The whole
+system is divided in pluggable parts, object mutation is either
+traceable or avoided altogether in favor of immutable data structures.
+As Rich Hickey argues, design decisions should be made under the
+\textbf{impression of extending, substitution, moving, combining and
+repurposing}. The ability to reason about the program at any given time
+is crucial for future decisions and implementations. Recalling again the
+unidimensional nature of simplicity.
 
@@
 ``everything that runs in the browser'', is certainly a story full of
-misconceptions and will be explained further along. While simplicity in
-the backend is mostly a matter of principles and patterns, any
-browser-based frontend is restricted on the highly deterministic nature
-of the browser platform.
+misconceptions. While simplicity in the backend is mostly a matter of
+principles and patterns, any browser-based frontend is restricted on the
+highly deterministic nature of the browser platform.
 
@@
 currently around 2.5 MB.\footnote{\href{http://httparchive.org/trends.php}{HTTPArchive
-  Trends}} Subtracting images, fonts or other content the size of HTML,
-CSS and JS sums up to a total average of 550 kb. One character weights
-around 1 byte which means an average webpage is delivering 550.000
-character or around 125 pages of single-spaced text. Frederic Filloux
-analyzed the payload on different newspaper websites and came to the
-conclusion, that only round about 5-6 \% of the transferred characters
-made for human consumption.{[}1{]}
+  Trends}} Leaving images, fonts or other content aside the size of
+HTML, CSS and JS sums up to a total average of 550 kb. One character
+weights around 1 byte which means an average webpage is delivering
+550.000 character or around 125 pages of single-spaced text. Frederic
+Filloux analyzed the payload on different newspaper websites and came to
+the conclusion, that only round about 5-6 \% of the transferred
+characters made for human consumption.{[}1{]}
 
@@
 
-\emph{Disclaimer:} This paper introduces many new browser build-ins with
-the focus on try and test. As the time of writing, many examples can be
-tried frictionless in the console of the latest versions of
-\textbf{Google Chrome, Opera and Apple Safari}.\footnote{\href{http://jonrimmer.github.io/are-we-componentized-yet/}{Are
-  we componentized yet?}} On Mozilla Firefox technologies work behind a
-flag and Microsoft Edge implementation is unfortunately far behind. But
-Browser implementation changes quickly and soon technology adoption
-won't be an issue. Meanwhile new standards can be used through
-\textbf{polyfills} even on legacy browsers.
-
 \section{Microservices}\label{microservices}
 
-In search of a better, simpler web architecture we might look on already
-established pattern that proofed to fulfill enterprise needs.
+In search of a better, simpler web architecture one might look on
+already established pattern that proofed to fulfill enterprise needs.
 Microservices are a good approach for tearing big monolithic systems
@@
 services, each running in its own process and communicating with
-lightweight mechanisms, often an HTTP resource API.''{[}4{]} Just yet at
-this point a reader might spot some similarities with microservices and
-the browser-based development: Both wrestling the problem of monolithic
+lightweight mechanisms, often an HTTP resource API.''{[}4{]} Yet at this
+point a reader might spot some similarities with microservices and the
+browser-based development: Both wrestling the problem of monolithic
 architecture and both using lightweight communication mechanisms. In
@@
 successfully transformed their monolithic system into a service based
-system which gives a glimpse of the power behind microservices.{[}4{]}
-
-Microservices incorporate many ideas from developing scalable software,
-like \emph{domain-driven design} where it pursues the incorporation of
-real world structure in the code.{[}3, p. 2{]} Or making use of
-\emph{continuous delivery} for pushing software rapidly through
-\emph{automated deployment} mechanisms into production.{[}4{]}
+system which gives a taste of the power behind microservices.{[}4{]}
+
+Microservices incorporate a wide array of ideas from developing scalable
+software, like \emph{domain-driven design} where it pursues the
+incorporation of real world structure in the code.{[}3, p. 2{]} Or
+making use of \emph{continuous delivery} for pushing software rapidly
+through \emph{automated deployment} mechanisms into production.{[}4{]}
 Furthermore, microservices transcendent the technical perspective and
@@
 testing and maintaining their very own microservice. Along the
-technology stack every member gains high competence about the
-functionality of the service architecture which can be critical for
-evolving the service over its whole lifecycle.
+technology stack every member gains high competence about the the
+service architecture which can be critical for evolving the service over
+its whole lifecycle.
 
@@
 context within each business domain will eventually shape the data
-structures and services in question. This methodology can be iterated
-over and over again. A sales service for example might be evaluated to
-different sales context resulting in differing interfaces on devices
-shaped by browsernative microservices.
-
-By assigning service responsibility to a team, the so called
+persistency model likewise the interface to the service. This
+methodology can be iterated over and over again. A sales service for
+example might be evaluated to different sales context resulting in
+differing interfaces on differing devices shaped by browsernative
+microservices. By focusing the service into clear defined business
+boundaries, it is easier to define a smart API of the service.
+
+Assigning service responsibility to a team, the so called
 \textbf{Definition of Done} (DoD) shifts from ``accomplishing projects''
@@
 administrative overhead like budgeting or resource allocation. It
-creates a kind of responsibility connection from the team to the
-service. Expectedly, those teams are more motivated within their very
-own service and exhibit a more sophisticated iteration time.{[}4{]}
+creates a kind of responsibility connection from the team to the service
+which can be best described as \textbf{shared governance} model. ``Each
+team collectively share responsibility for evolving the technical vision
+of the system.''{[}3, p. 247{]} Expectedly, those teams are more
+motivated within their very own service and exhibit a more sophisticated
+iteration time.{[}4{]}
 
@@
 side is highly important for their business. In fact, business goals and
-capabilities can be derived from the front-end needs. The state of the
-web is not only a story of numerous artifacts, it is also a story of an
-highly fragmented market along devices, operating systems, differing
-sizes and functionalities. Different devices again have different
-assumptions about the technology stack. Splitting teams along the stack
-results in an slow paced back and forth negotiation for every change to
-be made. As browser technologies, design guidelines and devices change
-frequently it makes absolutely sense to shift responsibility towards the
-teams altogether. By focusing the service into clear defined business
-boundaries, it is easier to define a smart API of the service.
+capabilities can be derived from frontend needs. The state of the web is
+not only a story of numerous artifacts, it is also a story of an highly
+fragmented market along devices, operating systems, differing sizes and
+functionalities. Different devices again have different assumptions
+about the technology stack. Splitting teams along the stack results in
+an slow paced back and forth negotiation for every change to be made
+make iteration time expensive. As browser technologies, design
+guidelines and devices change frequently it makes absolutely sense to
+shift responsibility towards the teams altogether.
 
@@
 Microservices heavily rely on simple HTTP request-response with resource
-APIs and lightweight messaging.{[}4{]} REST APIs are technology agnostic
-which aligns well with the idea of decentralization.{[}3, p. 247{]} The
-advantage of this overall simple communication model is the suitability
-for both frontend-backend likewise backend-backend communication. A
-service therefore can evolve from an heavy backend with a lot of network
-roundtrips to a leaner backend seamlessly. The browser build-in fetch
-API which is essentially a HTTP request can be heavily incorporated into
-a browsernative microservice to ensure communication between services.
+APIs and lightweight messaging.{[}4{]} Newman puts his stand on
+technologic-agnostic REST APIs to free data persistence from
+implementation constraints.{[}3, p. 247{]} The advantage of this overall
+simple communication model is the suitability for both frontend-backend
+likewise backend-backend communication. A service therefore can evolve
+from an heavy backend with a lot of network roundtrips to a leaner
+backend seamlessly. The browser build-in \textbf{fetch API} which is
+essentially a HTTP request can be heavily incorporated into a
+browsernative microservice to ensure communication to services in the
+backend.
 
@@
 Microservices are separate entities and decentralization is important to
-ensure autonomy. Distributed service responsibility owned by the team
-promoting shared governance and technology agnostic communication
-channels work towards a mostly decoupled service architecture.
-
-This distributed nature allows teams to create their own technology
+ensure autonomy. This paper already described fragmented services
+bounded to singular business context choreographed by simple
+communication protocols developed and evolved by autonomous teams.
+
+This distributed nature empowers teams to create their own technology
 stack, tools and services designed in the spirit of language- and
@@
 VIEW layer only - which is not true anymore. In the recent years the
-browser engines grows to a kind of virtual machine: there are connectors
-to build-in databases, multithreading support, an ever-growing JS
-build-ins and even speech synthesis. So-called \emph{Progressive Web
-Apps\footnote{\href{https://developers.google.com/web/progressive-web-apps}{Progressive
-  Web Apps}}}, a bunch of criteria for building good browserapps, can
-achieve a similar look and feel like native apps. Services like
-NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}},
-effectively compiling ``the web'' to native code, lower the boundary
-between native and browser code even further.
+major browser engines grow to to a fully-fledged app deployment platform
+offering connectors to build-in databases, multithreading support and
+ever-growing JS build-ins like speech synthesis or push notifications.
+So-called \emph{Progressive Web Apps\footnote{\href{https://developers.google.com/web/progressive-web-apps}{Progressive
+  Web Apps}}}, a bunch of criteria for building good browser apps, can
+achieve a similar look and feel like native apps. And last but not least
+services like NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}}
+effectively compiling ``the web'' to native machine code lowering the
+boundary between native and browser code even further.
 
@@
 microservice engineering team might choose another language for various
-reasons. Transpiling languages to JS as target language isn't exotic
-anymore. Languages like TypeScript, ClojureScript or PureScript compile
-to JS even exclusively. Once web components hit a critical mass there
-will be most likely some library support or foreign function interface
-towards ES6 modules (which are mandatory for the new specifications).
-With the rise of WebAssembly, a new low-level programming language for
-the browser, the determination on JS will hypothetical deteriorate and
-new quasi native languages for the web might gain traction.
+reasons. Transpiling languages to JS as target language is a stable
+solution nowadays. Languages like TypeScript, ClojureScript or
+PureScript compile to JS even exclusively. Once web components hit a
+critical mass there will be most likely some library support or foreign
+function interface towards ES6 modules (which are mandatory for the new
+specifications). With the rise of WebAssembly, a new low-level
+programming language for the browser, the determination on JS will
+hypothetical deteriorate and new quasi native languages for the web
+might gain traction.
 
@@
 of deployment in a safe, sandboxed environment. Web components virtually
-ship no overhead or require obscure build tools. This makes them ideal
-candidates for sharing and open source publishing.
-Webcomponents.org\footnote{\href{https://webcomponents.org}{Webcomponents.org}}
+ship no overhead or require dedicated build tools. This makes them ideal
+candidates for sharing and open source publishing similar to the largest
+JS package registry NPM. In the spirit of NPM web components can be
+perceived as frontend packages with an HTML interface instead of a JS
+signature. Webcomponents.org\footnote{\href{https://webcomponents.org}{Webcomponents.org}}
 is a registry for ready-to-use components of every scale and purpose
@@
 PouchDB\footnote{\href{https://pouchdb.com/}{PouchDB}} even offer
-backend syncing out of the box.
+backend adapters for syncing out of the box.
 
@@
 
+Microservices tend to increase complexity as this model adds a sheer
+number of moving parts to the system whereas requires proper
+orchestration.{[}3, p. 246{]} Arguably every more sophisticated web
+developer already came across build tools like Webpack or infrastructure
+automation tools like Gulp. Testing and deploying web components
+shouldn't be an obstacle in development.
+
 In the global nature of web development the development couldn't
@@
 opinionated (and more often biased) ways to glue related parts together.
-With scoped components, code blocks can be developed more sane and
-conveyed into production without headache. Even more, in the scripted
-environment of the browser this can be implemented rapidly and
-continuous.
-
-Previously mentioned Ben Issa, described the ING standard workflow.
-Every component deserves a own \textbf{git repo} containing
+Bret Victor, UI designer at Apple defined the importance of an
+\emph{immediate feedback principle} for developing user
+interfaces.\footnote{\href{https://vimeo.com/36579366}{Bret Victor -
+  Inventing on Principle}} In his talk he emphasizes the importance of
+an immediate connection between the creator of a product and product
+itself. Any change must results in an immediate visible feedback. Web
+components catch up with this principle as they allow isolated
+development within a single file containing all bits and pieces of the
+web component. Every major browsers devtools offer a direct file
+manipulation functionality so development can take place directly in
+place.
+
+When it comes to standardized deployment guidelines previously mentioned
+Ben Issa, described the ING standard workflow. Every component deserves
+a own \textbf{git repo} containing
 
@@
 
-Every check-in is handled as release candidate and can be independently
-deployed by a fully automated deployment machinery.{[}8{]} Even though
-this example is an opinionated perception it gives a sense of a mature
-component build for the web. Due to an exhaustive amount of testing and
-deployment tools for JS an automated infrastructure shouldn't be a
-problem.
+This example should made clear that synthetic tests of the component are
+part of the service right from the start. Every check-in is handled as
+release candidate and can be independently tested and deployed by a
+fully automated machinery.{[}8{]} Even though this example is an
+opinionated perception it gives a sense of a mature component build for
+the web. Due to an exhaustive amount of testing and deployment tools for
+JS an automated infrastructure shouldn't be a problem.
 
@@
 pieces.
+
+\emph{Disclaimer:} This paper introduces many new browser build-ins with
+the focus on try and test. As the time of writing, many examples can be
+tried frictionless in the console of the latest versions of
+\textbf{Google Chrome, Opera and Apple Safari}.\footnote{\href{http://jonrimmer.github.io/are-we-componentized-yet/}{Are
+  we componentized yet?}} On Mozilla Firefox technologies work behind a
+flag and Microsoft Edge implementation is unfortunately far behind. But
+Browser implementation changes quickly and soon technology adoption
+won't be an issue. Meanwhile new standards can be used through
+\textbf{polyfills} even on legacy browsers.
 
@@
 After getting confidence in microservice principles and technical
-background the paper should bring them both together to form a
+background the paper should briefly join them to form a
 \emph{browsernative microservices}. Needless to say the described
@@
 \texttt{\textless{}shop-app\textgreater{}}. This app made of several
-main views and many more invisible custom elements for routing, service
+main views and many more invisible wrapper elements for routing, service
 worker caching, theming, etc. The whole shop runs as a single
-application only fetching and sending resources and switching views.
+application fetching and updating remote resources and switching views.
 Let's assume we work in a sales engineering team of the Polymer Shop and
@@
 https://shop.polymer-project.org/checkout. Currently the checkout is a
-single, 671 lines of code long Polymer component incorporation all
-required fields for sign in, shipping, billing and summarizing the
-order. In the spirit of microservices we will split up the microservice
-into fine grained components. The shopping cart data is pulled out of a
-local storage JSON entity set up previously by another custom-element.
+single, 671 lines of code long Polymer component including all required
+fields for sign in, shipping, billing and summarizing the order. In the
+spirit of microservices we will split up the microservice into fine
+grained components. The shopping cart data is pulled out of a local
+storage JSON entity set up previously by another custom-element.
 
@@
 
-Like a usual checkout this example has 3 to 4 steps:
+By breaking down the service the team defined the business boundaries
+within the checkout process and ended up with following granular service
+blocks:
 
@@
 communicate over an unobstrusive message bus via the service root
-component \texttt{shop-checkout}.
-
-Before diving deeper into implementation, its worth to clarify the
+component \texttt{\textless{}shop-checkout\textgreater{}}.
+
+Before diving deeper into implementation, its worth to clarify an
 \textbf{architectural pattern} behind components. Any reader of the
@@
 Containers acts as invisible wrappers around presentational components
-acting more in the sense of UNIX filters. Their job is to fetch data
-from child nodes, aggregating events, interacting with the model and
-push state back to the presentational components. Consequently, they
-might contain more JS and less, if any, HTML markup. We probably don't
-need to utilize ShadowDOM as no styles are involved.
-
-Last but not least, there are \textbf{build-in components}, which is
-every build-in HTMLElement like the HTMLButtonElement. They are solely
-controllable and styleable from the outside and must therefore wrapped
-in presentational components and/or containers.
+acting in the sense of UNIX filters. Their job is to fetch data from
+child nodes, aggregating events, interacting with the model and push
+state back to the presentational components. Consequently they might
+contain more JS and less if any HTML markup. We probably don't need to
+utilize ShadowDOM as no styles are involved.
+
+Last but not least, we can expand the model for illustrational purposes
+to \textbf{native components} which is every build-in HTMLElement like
+the HTMLButtonElement. They are solely controllable and styleable from
+the outside and must therefore wrapped in presentational components
+and/or containers.
 
@@
 
-The root container \texttt{shop-checkout} is basically just an
-encapsulation layer in terms of service functionalities. Encapsulation
-of CSS won't be necessary at this point as no styling is involved. On
-initialization the \texttt{shop-checkout} might perform following tasks:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Importing child components with \textbf{HTML Imports}
-\item
-  Initialize the service in a logical top to bottom order like in the
-  snippet above
-\item
-  Initializing a \textbf{model thread} with
-  \texttt{new\ Worker(\textquotesingle{}checkout-model.js\textquotesingle{})};
-\item
-  Setting up a \texttt{postMessage} stub to dispatch messages to the
-  model
-\item
-  Setting up an event listener for messages from the model
-\item
-  Setting up an event listener for \textbf{custom events} from the child
-  nodes
-\end{enumerate}
-
-On runtime the root container acts mostly like a \textbf{message
-dispatcher} implementing the microservice idea of smart endpoints and
-dumb pipes.
-
-Every custom child node is eligible aggregate subordinate events from
-their children, buffer them and interact with the model trough via a
-\textbf{dumb message system}. To talk to the model any child node can
-dispatch a \textbf{unified event containing a unified message object}
-like the following example illustrates:
+The root container \texttt{\textless{}shop-checkout\textgreater{}} is
+basically just an encapsulation layer in terms of service
+functionalities. Encapsulation of CSS won't be necessary at this point
+as no styling is involved. The simplified root container might look like
+the following code snippet
 
@@
 \begin{Highlighting}[]
-\CommentTok{// Within every child node class}
-\NormalTok{set }\AttributeTok{dispatch}\NormalTok{(msg) }\OperatorTok{\{}
+\NormalTok{> shop-checkout.html}
+\CommentTok{<!-- IMPORTS -->}
+\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"sign-in.html"}\OtherTok{ async}\KeywordTok{>}
+\NormalTok{...}
+\CommentTok{<!-- VIEW -->}
+\KeywordTok{<shop-checkout>}
+  \KeywordTok{<sign-in></sign-in>}
+  \NormalTok{...}
+\KeywordTok{</shop-checkout>}
+
+\CommentTok{<!-- CONTROLLER -->}
+\KeywordTok{<script>}
+\KeywordTok{class} \NormalTok{ShopCheckout }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
+    \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
+    \KeywordTok{this}\NormalTok{.}\AttributeTok{MODEL} \OperatorTok{=} \KeywordTok{new} \AttributeTok{Worker}\NormalTok{(}\StringTok{'checkout-model.js'}\NormalTok{)}\OperatorTok{;}
+  \OperatorTok{\}}
+
+  \AttributeTok{connectedCallback}\NormalTok{() }\OperatorTok{\{}
+    \CommentTok{// listen to msg from MODEL}
+    \KeywordTok{this}\NormalTok{.}\VariableTok{MODEL}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'message'}\OperatorTok{,} \NormalTok{[Msg Handler])}
+    \CommentTok{// listen to child nodes}
+    \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'checkout'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>} \OperatorTok{\{}
+      \CommentTok{// delegate event to handler}
+      \KeywordTok{this}\NormalTok{.}\AttributeTok{_sendToMODEL} \OperatorTok{=} \NormalTok{e}\OperatorTok{;} 
+    \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+  \OperatorTok{\}}
+
+  \NormalTok{set }\AttributeTok{_sendToMODEL}\NormalTok{(e) }\OperatorTok{\{}
+    \KeywordTok{const} \NormalTok{id }\OperatorTok{=} \VariableTok{e}\NormalTok{.}\VariableTok{target}\NormalTok{.}\AttributeTok{id}\OperatorTok{,} 
+          \NormalTok{name }\OperatorTok{=} \VariableTok{e}\NormalTok{.}\VariableTok{target}\NormalTok{.}\AttributeTok{localName}\OperatorTok{,}
+          \NormalTok{load }\OperatorTok{=} \VariableTok{e}\NormalTok{.}\AttributeTok{detail}\OperatorTok{,}
+          \NormalTok{letter }\OperatorTok{=} \VariableTok{Object}\NormalTok{.}\AttributeTok{assign}\NormalTok{(}\OperatorTok{\{\},} \OperatorTok{\{} \NormalTok{id}\OperatorTok{,} \NormalTok{name}\OperatorTok{,} \NormalTok{load }\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+    \CommentTok{// send to MODEL}
+    \KeywordTok{this}\NormalTok{.}\VariableTok{MODEL}\NormalTok{.}\AttributeTok{postMessage}\NormalTok{(letter)}\OperatorTok{;} 
+  \OperatorTok{\}}
+  \NormalTok{...}
+\OperatorTok{\}}    
+\VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'shop-checkout'}\OperatorTok{,} \NormalTok{shopCheckout)}\OperatorTok{;}
+\OperatorTok{<}\SpecialStringTok{/script>}
+\end{Highlighting}
+\end{Shaded}
+
+The purpose of this simplified code snippet is to outline the MVC
+threefold in the service root. On runtime the root container acts mostly
+like a \textbf{message dispatcher} implementing the microservice
+principle of smart endpoints and dumb pipes.
+
+In order to interact with the MODEL, every child node must implement the
+\textbf{dedicated \texttt{checkout}} custom event (which will be
+explained further on in the next section). Summarizing the
+unidirectional communication from the service root perspective looks
+like following plot:
+
+\begin{verbatim}
+VIEW thread                 |               MODEL thread
+                            |
++----------+                |               +----------+
+| Checkout | ----------Event Msg--------->  |  Msg     |
+|  Event   |                |               | Handler  |
++----------+                |               +----------+
+                            |                   |
+                            |                   |
+                            |               +----------+
+                            |               | Generate |
+                            |               |  Effect  |
+                            |               +----------+
+                            |                   |
+                            |                   |
++----------+                |               +----------+
+|  Msg     |                |               | Postmsg  |
+| Handler  | <--------Action Msg----------  |  Action  |
++----------+                |               +----------+    
+                            |
+\end{verbatim}
+
+Effects are yielded by the asynchronous operation of messages and create
+actions returned to sender by the message handlers. Effects can be
+created by external messages, like subscription to an WebSocket, too.
+Effects may be created with additional resources from the server or
+syncing with local storage like in the polymer-shop.
+
+While the msg handlers are basic ``dumb'' switch statements the
+\textbf{smartness} solely arouse from intelligent controllers processing
+the message. This communication model offers lots of possibilities for
+\textbf{evolutionary design} as child nodes can be loosely dropped in
+the service root and communication might \textbf{fail graceful} as
+simply nothing happens if one or another switch is not defined.
+
+\subsubsection{Container components}\label{container-components}
+
+The second layer of the checkout microservice like
+\texttt{\textless{}sign-in\textgreater{}} or
+\texttt{\textless{}shipping-details\textgreater{}} still contain a lot
+of logic and no or less styling. Their job is to control their
+underlying presentational leaf components, aggregating events and to
+define a \textbf{set of actions towards the model} . Every container
+mounted directly under the service root may have a dedicated area inside
+the worker thread where it's fulfil his duties. For example, a typical
+\texttt{\textless{}sign-in\textgreater{}} contains merely two fields
+username and password and a submit button. The component listens for the
+submit action, aggregating the credentials and might add some semantics
+to it like \emph{action: `SIGNIN SUBMITTED'}. Fields and action message
+will be dispatched towards the model for further processing like
+initiating a authorization process.
+
+Every container component is eligible to aggregate subordinate events
+from their children, buffer them and interact with the model trough via
+a \textbf{unified message system}. A base class, from which
+\texttt{\textless{}sign-in\textgreater{}} or
+\texttt{\textless{}shipping-details\textgreater{}} can be extended can
+look like this:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{class} \NormalTok{CheckoutContainerBase }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+  \KeywordTok{static} \AttributeTok{getActions}\NormalTok{() }\OperatorTok{\{}
+    \ControlFlowTok{return} \NormalTok{[...]}
+  \OperatorTok{\}}
+  \NormalTok{set }\AttributeTok{_dispatch}\NormalTok{(msg) }\OperatorTok{\{}
     \KeywordTok{this}\NormalTok{.}\AttributeTok{dispatchEvent}\NormalTok{(}
@@
   \OperatorTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-Calling \texttt{this.dispatch(msg)} within a child node will trigger an
-event bubbling upstream. The service root will implement a simple event
-listener for \texttt{checkout} events. In a second step the service root
-add information about the original target (like
-\emph{tag-name\#id.className}) to the message so the message can be
-processed in the model thread and eventually pushed back correctly to
-the sender.
-
-Summarizing the unidirectional model-view communication from the service
-root perspective looks like following sketch:
-
-\begin{verbatim}
-VIEW thread         |           MODEL thread
-                    |
-Checkout ------Event Msg--------->    Msg
-  Event             |                Handler
-                    |                   |
-                    |                Effect
-                    |                   |
-  Msg               |                Postmessage
-Handler <-----Action Msg----------    Action
-                    |                   
-                    |
-\end{verbatim}
-
-Effects are yielded by the asynchronous operation of messages and create
-actions returned to sender. Effects can be created by external messages,
-like subscription to an WebSocket, too. Effects may be created with
-additional resources from the server or syncing with local storage like
-in the polymer-shop.
-
-While the msg handlers are basic ``dumb'' switches the
-\textbf{smartness} solely arouse from intelligent controllers processing
-the message. Every component must implement a \textbf{unified socket} in
-it's service contracts like a property on the custom element:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{// Within every child node class}
-\NormalTok{set }\AttributeTok{receive}\NormalTok{(msg) }\OperatorTok{\{}
-  \CommentTok{// do something smart}
+  \NormalTok{set }\AttributeTok{_receive}\NormalTok{(msg) }\OperatorTok{\{}
+  \CommentTok{// switch to methods}
+    \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\KeywordTok{this}\NormalTok{.}\AttributeTok{localName} \OperatorTok{+} \StringTok{"received msg"}\NormalTok{)}\OperatorTok{;}
+    \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(msg)}\OperatorTok{;}
+  \OperatorTok{\}}
 \OperatorTok{\}}
@@
 
-This communication flow is designed to \textbf{fail graceful} as simply
-nothing happens if one or another switch is not defined. Furthermore it
-incorporates lots of possibilities for \textbf{evolutionary design} as
-child nodes can be loosely dropped in or teared apart. Middleware
-container wrappers can be pulled in between child nodes and service root
-to do some extra work and and even the message object can be enhanced
-freely as long as it won't break with the message conventions. Changing
-or enhancing any functionality requires only to touch the controllers in
-the child nodes in question and the endpoint section at the model.
-Furthermore, \textbf{infrastructure automation} might be achieved by
-dynamically evaluating the mounted child nodes and registering the
-``backend'' model switches on build time. Due to it's standardized
-message system the components are testable within standardized tests.
-
-To work in a larger scale, we might define a base component which has
-the default message and socket properties in its guts and extend every
-new element from this base class.
-
-Another yet more powerful approach can be achieved by calculating the
-new shape of the DOM in the worker directly and just patch the DOM in
-the view layer.\footnote{A good example of this approach can be found at
-  \href{http://www.pocketjavascript.com/blog/2015/11/23/introducing-pokedex-org}{Introducing
-  Pokedex}} This approach is certainly more complicated but worth to
-evaluate further more.
+Extending the \texttt{CheckoutContainerBase} will equipped every
+container node with the unified message interface. Calling
+\texttt{this.\_dispatch(msg)} within a within the container will trigger
+an event bubbling upstream. The service root will implement a simple
+event listener for \texttt{checkout} events. After receiving an answer
+from the MODEL a service root can \texttt{querySelector} the container
+node in question and push forward the answer over the \texttt{\_receive}
+property.
+
+There might be even more middleware containers pulled in between
+presentational leafs and service root component to do some extra work
+without caring about bubbling events altogether. Changing or enhancing
+any functionality requires only to touch the controllers in the
+containers in question and the endpoint section at the model.
+\textbf{Infrastructure automation} might be achieved by dynamically
+evaluating the mounted containers beneath the service root and modeling
+the ``backend'' model accordingly. Due to it's standardized message
+system the containers are testable within standardized tests.
 
@@
 
+In comparison to the former presentational component is build mostly
+
 \section{Thinking further}\label{thinking-further}
@@
 
-Most obvious ist the gathering of all related code under the umbrellar
-of a single HTML tag.
-
-Secondly, the sub-standard \emph{custom elements} introduces so called
-lifecycle methods and a getter/setter interface exposing the
-functionality to the developer. Event handling, for example, can be
-registered in place which is much more declarative than assigning event
-listeners from the outside. Of course, this events can be pushed down to
-nested tags, allowing an increasingly granular system design. This
-approach will be explained further in the upcoming sections.
-
-Following this logic any company, whether it is web-related or not,
-should be devided in units grouped around a destinct business service to
-optimise the workflow. Fowler and Lewis outlines this approach as an
-``alignment of business capabilities''{[}4{]} While this kind of
-structure may be true for companies like Google or Amazon, there is a
-vast majority of companies developing for the web which are grouped
-around tasks.{[}8{]} A very common structure is formed by the technology
-stack (UX Designers, Frontend- \& Backend Developers) or by separating
-teams along the product lifecycle (development, testing, deployment).
-
-Advocators from the microservice approach propose a different model.
-best described by . Web components are one (but important) way to tie up
-those diciplines as one component can host a single independent business
-service. Combined with a flexible backend service these components can
-be huge gain over the cumbersome functional organizational approach.
-
-dichotome Pattern
-
-Ein üblicher eventgesteuerter Webservice setzt sich aus
-unterschiedlichsten Komponenten zusammen, die wiederum
-unterschiedlichste Eventlistener \& -emitter in sich subsummieren. Diese
-inhärente Komplexität verlangt geradezu nach einer klaren,
-deterministischen Struktur des Webservices, die das Zusammenspiel
-orchestriert. In der Analogie des Orchesters gesprochen, benötigt der
-Webservice (oder sogar die gesamte Webapplikation) einen Dirigenten, der
-für die Steuerung verantwortlich ist.
-
-http://alistair.cockburn.us/Hexagonal+architecture
-
-MVC Pattern
-
-Pure frontend vs heavy backend
+virtual DOM
 

### 2017-02-23 paper.tex
pdf/tex build

@@
 pattern to build graphical user interfaces. While frameworks are a valid
-approach to build scalable web applications they remain highly
+approach for building scalable web applications they remain highly
 opinionated, embody inherent complexity themselves and can change and
 break over time. Another downside is code inflation which is a crucial
-point for performance and third-party libraries are no exception on
-that. A standardized way for creating complex UI interfaces painlessly
-requires new build-in browser capabilities.
+point for performance. All of those bottlenecks in the web demand for a
+new standardized way for creating and evolving complex web services.
 
@@
 to propose \emph{The Extensible Web Manifesto}.\footnote{\href{https://extensiblewebmanifesto.org/}{The
-  Extensible Web Manifesto}} The claim of the manifesto was to enhance
-the current web platforms with new low-level capabilities. Those
-features should empower creators of the web to write more declarative
-code and therefore overcome known bottlenecks and artificial
-abstractions. Four years later, the enhancement of JavaScript
-leapfrogged and many new low-level APIs brought to life. With this new
-APIs at hand a vivid web developer can create robust websites with less
-code and less additional libraries. This paper is an approach to unfold
-these \textbf{browsernative} technologies to create overall simple and
-resilient \textbf{microservices} for the browser
+  Extensible Web Manifesto}} The manifestos claim was to enhance the
+current web platforms with new low-level capabilities. Those features
+should empower creators of the web to write more declarative code and
+abandon known problems and artificial abstractions. Four years later,
+the enhancement of JavaScript leapfrogged and many new low-level APIs
+brought to life. With this new APIs at hand a vivid web developer can
+create robust websites with less code and less additional libraries.
+This paper is an approach to unfold these \textbf{browsernative}
+technologies to create overall simple and resilient
+\textbf{microservices} for the browser
+
+\section{Microservices}\label{microservices}
+
+In search of a better, simpler web architecture one might look on
+already established pattern that proofed to fulfill enterprise needs.
+Microservices are a good approach for tearing big monolithic systems
+into fine-grained simple services with explicit defined boundaries. In a
+nutshell a microservice is a small, autonomous service that works
+together with other services seamlessly.{[}3, p. 2{]} Or with the words
+of Fowler and Lewis: ``\ldots{} the microservice architectural style is
+an approach to developing a single application as a suite of small
+services, each running in its own process and communicating with
+lightweight mechanisms, often an HTTP resource API.''{[}4{]} Yet at this
+point a reader might spot some similarities with microservices and the
+browser-based development: Both wrestling the problem of monolithic
+architecture and both using lightweight communication mechanisms. In
+fact, many big companies of ``the web'' like Amazon or Netflix
+successfully transformed their monolithic system into a service based
+system which gives a taste of the power behind microservices.{[}4{]}
+
+Microservices incorporate a wide array of ideas from developing scalable
+software, like \emph{domain-driven design} where it pursues the
+incorporation of real world structure in the code.{[}3, p. 2{]} Or
+making use of \emph{continuous delivery} for pushing software rapidly
+through \emph{automated deployment} mechanisms into production.{[}4{]}
+Furthermore, microservices transcendent the technical perspective and
+reaches into the team organization. As a primary source of truth this
+paper relies on the work of Sam Newman, who has written a comprehensive
+guide called \emph{Building Microservices} and the work of Fowler and
+Lewis. The purpose of this section is to gain confidence about the
+microservices architecture in the context of the browser platform.
+
+\subsection{Componentization via
+Services}\label{componentization-via-services}
+
+``A~\textbf{component}~is a unit of software that is independently
+replaceable and upgradeable.'' {[}4{]} Components are the building
+blocks of microservices. And microservices are the building blocks of
+applications. Essentially the difference between microservices and
+components is just the level of abstraction. Whether a concrete
+microservice or a much more generic component, both share a similar set
+of principles. Therefore this paper referring to both parts when talking
+about \textbf{services}.
+
+The first principle of services is the \textbf{loose coupling
+principle}: changing and deploying one service shouldn't result in
+changing other parts of the system.{[}3, p. 30{]}. Mutations or shadowed
+variables, which is happening a lot in CSS, making it hard to keep
+changes ought to only affect one place in the application. A
+\emph{browsernative microservice} therefore pushing encapsulation and
+avoiding variable mutations outside its scope as much as possible.
+Practically, CSS will be scoped and JS fosters immutable JS entities and
+avoids variables leaking into the global namespace.
+
+The second principle of services is the \textbf{high cohesion
+principle}: Whether designing a microservice or it's components we want
+related behavior sit together, and unrelated behavior to sit
+elsewhere.{[}3, p. 30{]} High cohesion can be expressed in a dynamic way
+as the \emph{Single Responsibility Principle}: ``Gather together those
+things that change for the same reason and separate those things that
+change for different reasons.''{[}5{]} In a very quick and dirty code
+quality analysis, the quality can be measured just by counting the
+places changes in the code occur in order to implement a functionality.
+An arbitrary threefold MVC system should require a maximum of three
+changes to implement or change functionalities. The problem in browser
+based development is not only the global paradigm which makes changes
+deliberately unpredictable. The high cohesion principle is violated by
+the traditional separation along the siloed entities \textbf{HTML, JS
+and CSS}. Understanding the relation of HTML markup and another CSS file
+adds incidental complexity. Different approaches emerged over the recent
+years to join the forces. Most notable the dedicated frontend language
+\emph{elm}\footnote{\href{http://elm-lang.org/}{elm lang}} where MODEL,
+VIEW and CONTROLLER sit in one place. A browsernative service aims for a
+combinement of the web native trinity HTML, JS and CSS.
+
+Traditional web development relies on \textbf{libraries} to enhance the
+service capabilities of the web platform. Compared to libraries, a
+component service offers multiple advantages for building, deployment
+and shipping. As expressed earlier a \emph{components for the web}, or
+web component, is \textbf{self-contained} which means it embodies all
+needed functionality to get it's job done. Therefore it has a much
+better evolution mechanism in the service contracts. Changing
+functionality won't break other services. A component can progressively
+enhanced which guarantees functionality throughout different versions. A
+library is only loosely coupled to the implementation and therefore hard
+to track in functionality. Changing a library may result in an
+unforeseen amount of time fixing implementations. It is not unusual to
+see websites embodying different versions of the same library to
+guarantee functionality but lowers page performance significantly.
+
+Another issue where web components stand out is related to performance
+and especially the critical first page rendering. Libraries for the
+browsers are traditionally ``shipped'' as non static immediately-invoked
+functions run immediately on page load. Following the \emph{Google RAIL}
+model, a user-centric performance measurement, a page load ought be less
+than 1 second to catch up user's attention.{[}6{]} There are many ways
+for optimizing the critical first render but as a rule of thumb a
+build-in web component might be always superior to libraries in terms of
+first rendering. The fine-grained lifecycle methods which will be
+described in the technical section of this paper give the developer far
+reaching optimization opportunities.
+
+The last argument in favor of components over libraries is the more
+explicit interface.{[}4{]} While the functionality of a library needs
+documentation to be accessible a component functionality is exposed via
+the components' signature. The markup language HTML is by design
+equipped with an expressive syntax and steering a web component using
+attributes and values should be most straightforward even to
+unexperienced web developers.
+
+\subsection{Organized around Business
+Capabilities}\label{organized-around-business-capabilities}
+
+\begin{quote}
+``organizations which design systems \ldots{} are constrained to produce
+designs which are copies of the communication~structures~of these
+organizations''. {[}7{]}
+\end{quote}
+
+Emphasizing the human factor in microservices is a key feature.
+Microservices are a product of real-world usage.{[}3, p. 1{]} Instead of
+splitting team structures along the technology stack (UI Experts
+-\textgreater{} Middleware -\textgreater{} Database) a microservice
+approach model teams around \textbf{business capabilities}.{[}4{]}
+Consequently every team is capable of planning, designing, implementing,
+testing and maintaining their very own microservice. Along the
+technology stack every member gains high competence about the the
+service architecture which can be critical for evolving the service over
+its whole lifecycle.
+
+Real-world domains tend to be complex and multifaceted. To unfold their
+complexity, domains can subdivided into \textbf{bounded context}.{[}3,
+p. 31{]} For example, customer service is a business domain but with
+varying bounded context. One context can be sales, another context could
+be support. Every context makes different assumptions about the
+underlying model and draws an explicit interface where it decides what
+to share with other contexts.{[}3, p. 30{]} Evaluating each bounded
+context within each business domain will eventually shape the data
+persistency model likewise the interface to the service. This
+methodology can be iterated over and over again. A sales service for
+example might be evaluated to different sales context resulting in
+differing interfaces on differing devices shaped by browsernative
+microservices. By focusing the service into clear defined business
+boundaries, it is easier to define a smart API of the service.
+
+Assigning service responsibility to a team, the so called
+\textbf{Definition of Done} (DoD) shifts from ``accomplishing projects''
+to ``accomplishing products''. This new paradigm not only changes the
+administrative overhead like budgeting or resource allocation. It
+creates a kind of responsibility connection from the team to the service
+which can be best described as \textbf{shared governance} model. ``Each
+team collectively share responsibility for evolving the technical vision
+of the system.''{[}3, p. 247{]} Expectedly, those teams are more
+motivated within their very own service and exhibit a more sophisticated
+iteration time.{[}4{]}
+
+For many companies working in the spheres of the internet the client
+side is highly important for their business. In fact, business goals and
+capabilities can be derived from frontend needs. The state of the web is
+not only a story of numerous artifacts, it is also a story of an highly
+fragmented market along devices, operating systems, differing sizes and
+functionalities. Different devices again have different assumptions
+about the technology stack. Splitting teams along the stack results in
+an slow paced back and forth negotiation for every change to be made
+make iteration time expensive. As browser technologies, design
+guidelines and devices change frequently it makes absolutely sense to
+shift responsibility towards the teams altogether.
+
+\subsection{Smart endpoints and dumb
+pipes}\label{smart-endpoints-and-dumb-pipes}
+
+To ensure the microservice functionality among teams and different
+services requires thoughtful decentralization. Emphasizing once more the
+real-world capabilities of microservices a message channel architecture
+can be derived from patterns known from traditional postal services. A
+physical letter has only two smart endpoints entitled to read and
+process the message while packaging likewise the connection itself is
+maximized towards unification. \emph{Smart endpoints and dumb pipes} is
+coined to the approach of designing communication mostly decoupled and
+as cohesive as possible.{[}4{]} Applying this rules to the web platform
+can result into building unified JSON message objects passed along
+``dumb'' middleware components. A typical browser event comes close to
+this definition and suits arguably well for in-memory communication
+between web components and microservices.
+
+Microservices heavily rely on simple HTTP request-response with resource
+APIs and lightweight messaging.{[}4{]} Newman puts his recommendation on
+technologic-agnostic REST APIs to free data persistence from
+implementation constraints.{[}3, p. 247{]} The advantage of this overall
+simple communication model is the suitability for both frontend-backend
+likewise backend-backend communication. A service therefore can evolve
+from an heavy backend with a lot of network roundtrips to a leaner
+backend seamlessly. The browser build-in \textbf{fetch API} which is
+essentially a HTTP request can be heavily incorporated into a
+browsernative microservice to ensure communication to services in the
+backend.
+
+\subsection{Decentralized Governance}\label{decentralized-governance}
+
+Microservices are separate entities and decentralization is important to
+ensure autonomy. This paper already described fragmented services
+bounded to singular business context choreographed by simple
+communication protocols developed and evolved by autonomous teams.
+
+This distributed nature empowers teams to create their own technology
+stack, tools and services designed in the spirit of language- and
+platform independence and share their knowledge with other
+parties.{[}4{]} In the recent years many big companies like Facebook,
+Google, Netflix and others followed that spirit and published their
+ideas and implementations open source. The previously mentioned ReactJS
+for example is a product of Facebooks need to ensure a consistent
+frontend experience. In fact, many tools and techniques are byproduct of
+vital interaction of concrete domain problems and their implementations.
+
+The spirit of freedom can't be applied universally to
+\emph{browsernative microservices} as the browser and its underlying DOM
+will be the limitation factor to a certain degree. Talking about the
+browser, a reader might be tempted to narrowly thinking of the obvious
+VIEW layer only - which is not true anymore. In the recent years the
+major browser engines grow to to a fully-fledged app deployment platform
+offering connectors to build-in databases, multithreading support and
+ever-growing JS build-ins like speech synthesis or push notifications.
+So-called \emph{Progressive Web Apps\footnote{\href{https://developers.google.com/web/progressive-web-apps}{Progressive
+  Web Apps}}}, a bunch of criteria for building good browser apps, can
+achieve a similar look and feel like native apps. And last but not least
+services like NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}}
+effectively compiling ``the web'' to native machine code lowering the
+boundary between native and browser code even further.
+
+JS is the widely accepted language of the web. Nevertheless, a
+microservice engineering team might choose another language for various
+reasons. Transpiling languages to JS as target language is a stable
+solution nowadays. Languages like TypeScript, ClojureScript or
+PureScript compile to JS even exclusively. Once web components hit a
+critical mass there will be most likely some library support or foreign
+function interface towards ES6 modules (which are mandatory for the new
+specifications). With the rise of WebAssembly, a new low-level
+programming language for the browser, the determination on JS will
+hypothetical deteriorate and new quasi native languages for the web
+might gain traction.
+
+Another more real life decentralization aspect derives from the easiness
+of deployment in a safe, sandboxed environment. Web components virtually
+ship no overhead or require dedicated build tools. This makes them ideal
+candidates for sharing and open source publishing similar to the largest
+JS package registry NPM. In the spirit of NPM web components can be
+perceived as frontend packages with an HTML interface instead of a JS
+signature. Webcomponents.org\footnote{\href{https://webcomponents.org}{Webcomponents.org}}
+is a registry for ready-to-use components of every scale and purpose
+where for even Google shares a lot of their material design elements.
+
+\subsection{Decentralized Data
+Management}\label{decentralized-data-management}
+
+Data Management in a microservice follows the same modular philosophy
+like the service implementation. As mentioned earlier different bounded
+contexts make different assumptions of the underlying models. A
+\emph{browsernative microservices} takes this idea even further and
+expands it to the fragmented world of electronic devices. Decentralized
+decisions about conceptual models demand for decentralized data storage
+decisions.{[}4{]} Todays web architectures aims to leverage an
+increasing amout of processing to the client to avoid time-consuming
+roundtrips especially in mobiles networks.\footnote{Latency numbers:
+  https://gist.github.com/jboner/2841832} Since network roundtrips are
+costly it is a good advice to only query as much data as needed and
+cache as much as possible. The build-in LocalStorage or its successor
+IndexedDB are mature persistence technologies and libraries like
+PouchDB\footnote{\href{https://pouchdb.com/}{PouchDB}} even offer
+backend adapters for syncing out of the box.
+
+``Microservices prefer letting each service manage its own
+database.''{[}Fowler2014{]} Ben Issa, chief architect of ING Australia
+emphasizes this pragmatism on APIs in a conference talk. At ING the
+frontend demands tailor the backend APIs, APIs may be produced
+automatically and not even Issa knows how many APIs exists.{[}8{]} They
+are using a pattern called \textbf{backend for frontends} empowering the
+team working to craft their UI and backend in a one-to-one
+relationship.{[}3, p. 72{]}
+
+To see this pattern in the field a reader might have a look at Facebooks
+GraphQL\footnote{\href{http://graphql.org/}{GraphQL}}. GraphQL is a
+query language for the frontend. The backend solely replies on the
+frontend needs. Another well documented example in the field is
+Cognitects Datomic\footnote{\href{http://www.datomic.com/}{Datomic}},
+where parts of the database will be reflected to the client. A so-called
+Transactor ensures ACID compliance.
+
+The simplified microservice example later in this paper assumes a
+generic build-in API accompanied by build-in frontend components.
+Instead of gluing frontend and backend together on runtime the
+microservice is designed holistically containing both front- and
+backends. For the sake of simplicity data management won't be explored
+into depth throughout this paper.
+
+\subsection{Infrastructure Automation}\label{infrastructure-automation}
+
+Microservices tend to increase complexity as this model adds a sheer
+number of moving parts to the system whereas requires proper
+orchestration.{[}3, p. 246{]} Arguably every more sophisticated web
+developer already came across build tools like Webpack or infrastructure
+automation tools like Gulp. Testing and deploying web components
+shouldn't be an obstacle in development.
+
+In the global nature of web development the development couldn't
+completely decoupled from the production environment. This circumstance
+left developers switching back and forth between files developing tricky
+opinionated (and more often biased) ways to glue related parts together.
+Bret Victor, UI designer at Apple defined the importance of an
+\emph{immediate feedback principle} for developing user
+interfaces.\footnote{\href{https://vimeo.com/36579366}{Bret Victor -
+  Inventing on Principle}} In his talk he emphasizes the importance of
+an immediate connection between the creator of a product and product
+itself. Any change must results in an immediate visible feedback. Web
+components catch up with this principle as they allow isolated
+development within a single file containing all bits and pieces of the
+web component. Every major browser devtool offers a direct file
+manipulation functionality so development can be even in place.
+
+When it comes to standardized deployment guidelines previously mentioned
+Ben Issa, described the ING standard workflow. Every component deserves
+a own \textbf{git repo} containing
+
+\begin{itemize}
+\tightlist
+\item
+  Internationalization conformity (i18n)
+\item
+  Accessibility conformity (a11y)
+\item
+  Tests for the component
+\item
+  Demos of the component
+\item
+  Blueprints to mock the one to one APIs
+\item
+  Docs
+\end{itemize}
+
+Even though this example is an opinionated perception it gives a sense
+of a mature component build for the web. This example should made clear
+that all parts of the component put together in one place. Every
+check-in is handled as release candidate and can be independently tested
+and deployed by a fully automated machinery.{[}8{]} Due to an exhaustive
+amount of testing and deployment tools for JS an automated
+infrastructure shouldn't be a problem.
+
+\subsection{Design for failure}\label{design-for-failure}
+
+In theory a microservice is designed with focus on monitoring for both
+the architectural elements and business relevant metrics.{[}4{]} Due to
+the modular structure weak points can occur in the orchestration of the
+services. A microservice should track down every communication flow and
+provide defaults and meaningful error messages where communication
+stuck. Testing every single component with predefined synthetic events
+ensures functionality. Nevertheless, browser support may vary and legacy
+browsers remain a general problem for enhancing websites with new
+technologies and therefore demand further configuration.
+
+Combinment of the resources in the browser always demanded for
+optimization to avoid unexpected side-effects like the \emph{flash of
+unstyled content (FOUC)}. Googles Polymer propagates the a
+general-purpose pattern called \textbf{PRLP}\footnote{\href{https://www.polymer-project.org/1.0/toolbox/server}{PRLP
+  pattern}}:
+
+\begin{itemize}
+\tightlist
+\item
+  Push critical resources for the initial route
+\item
+  Render initial route
+\item
+  Pre-cache remaining routes
+\item
+  Lazy-load and create remaining routes on demand
+\end{itemize}
+
+Following this pattern a critical resource can evaluate browser maturity
+beforehand and switch to a \textbf{polyfill} or another fallback
+solution instead of the latest browser optimized version. After the
+initial paint, critical resources like top-level microservices or other
+app logic can be loaded and registered.
+
+Regarding the evolution of the web, the ``next billion'' internet users
+most likely using Android, have decent specs mobile phones, use an
+evergreen browser but won't have a reliable internet connection.{[}9{]}
+While \emph{Progressive Enhancement} was once coined on the principle to
+build websites both for Browsers with JS and HTML only, the new
+\emph{Progressive Enhancement} tends towards an \textbf{offline first}
+principle avoiding network connectivity failures. A \emph{browsernative
+microservice} therefore not only tries to cache data as much as
+possible, it should also bring in a lot of program logic as described in
+the previous chapters.
+
+\subsection{Evolutionary Design}\label{evolutionary-design}
+
+Microservices tend to become smaller over time. An evolutionary design
+approach puts emphasizes on decomposition and scrapping the service.
+``The key property of a component is the notion of independent
+replacement and upgradeability.''{[}4{]} Therefore we can safely change
+and chop services. Lazy components of the system which won't change
+often should be separated from parts undergoing a lot of churn.{[}4{]}
+And services which change for the same reason might be moved together or
+even could be merged.
+
+Being flexible when developing for the web is a selling point as
+innovation cycles for browser development is fast paced and technologies
+can change quickly. Frontend related hardware, software and
+methodologies innovate rapidly over time.
+
+\emph{Browsernative microservices} should be perceived as complementary
+technology in contrast to full-service frameworks like Angular. Being
+native technology pursues a strong interopt approach with existing
+systems. Andrew Rota for example came up with the pattern to use small,
+encapsulated and stateless web components as leaves in the tree of React
+components instead of native HTML elements.{[}10{]} Even React code can
+profit from the expressiveness of custom components. As web components
+are native after registration there is no difference in using a native
+\texttt{button} over a \texttt{meaningful-button}. Most likely there
+will be always some cutting edge framework promising advantages over
+native code. Whatever new framework will be on the rise within the next
+years native components can eliminate future uncertainty and allowing
+rapid reassembling towards new architectures.
+
+\section{W3C specifications}\label{w3c-specifications}
+
+For building a native microservice running on the ``bare-metal'' browser
+engine requires a bunch of new specifications and assumptions. Most
+importantly the quasi specification \textbf{Web Components} is needed.
+\emph{Web Components} is not a real standard. It's an amalgam of APIs
+from multiple w3c specs which can be used independently, too. A
+webdeveloper may choose one spec and embrace the freedom in architecture
+which can be combined with other frameworks/libraries.
+
+Depending on the context, some people argue for only two specs which
+essentially make it possible to create a scoped component but not caring
+too much on it's distribution{[}11{]}. Some people prefer the three
+specs {[}12{]}, but the majority advocating the four specs variant,
+which is listed on the quasi-official
+\href{http://webcomponents.org}{webcomponents.org} website. For the
+purpose of this article, all four specs will be discussed briefly to
+provide a rough understanding. It is not meant to cover all bits and
+pieces.
 
@@
 
-\section{Microservices}\label{microservices}
-
-In search of a better, simpler web architecture we might look on already
-established pattern that proofed to fulfill enterprise needs.
-Microservices are a good approach for tearing big monolithic systems
-into fine-grained simple services with explicit defined boundaries. In a
-nutshell a microservice is a small, autonomous service that works
-together with other services seamlessly.{[}3, p. 2{]} Or with the words
-of Fowler and Lewis: ``\ldots{} the microservice architectural style is
-an approach to developing a single application as a suite of small
-services, each running in its own process and communicating with
-lightweight mechanisms, often an HTTP resource API.''{[}4{]} Just yet at
-this point a reader might spot some similarities with microservices and
-the browser-based development: Both wrestling the problem of monolithic
-architecture and both using lightweight communication mechanisms. In
-fact, many big companies of ``the web'' like Amazon or Netflix
-successfully transformed their monolithic system into a service based
-system which gives a glimpse of the power behind microservices.{[}4{]}
-
-Microservices incorporate many ideas from developing scalable software,
-like \emph{domain-driven design} where it pursues the incorporation of
-real world structure in the code.{[}3, p. 2{]} Or making use of
-\emph{continuous delivery} for pushing software rapidly through
-\emph{automated deployment} mechanisms into production.{[}4{]}
-Furthermore, microservices transcendent the technical perspective and
-reaches into the team organization. As a primary source of truth this
-paper relies on the work of Sam Newman, who has written a comprehensive
-guide called \emph{Building Microservices} and the work of Fowler and
-Lewis. The purpose of this section is to gain confidence about the
-microservices architecture in the context of the browser platform.
-
-\subsection{Componentization via
-Services}\label{componentization-via-services}
-
-``A~\textbf{component}~is a unit of software that is independently
-replaceable and upgradeable.'' {[}4{]} Components are the building
-blocks of microservices. And microservices are the building blocks of
-applications. Essentially the difference between microservices and
-components is just the level of abstraction. Whether a concrete
-microservice or a much more generic component, both following pretty
-much the same principles. Therefore this paper referring to both parts
-when talking about \textbf{services}.
-
-The first principle of services is the \textbf{loose coupling
-principle}: changing and deploying one service shouldn't result in
-changing other parts of the system.{[}3, p. 30{]}. Mutations or shadowed
-variables, which is happening a lot in CSS, making it hard to keep
-changes ought to only affect one place in the application. A
-\emph{browsernative microservice} therefore pushing encapsulation and
-avoiding variable mutations outside its scope as much as possible.
-Practically, CSS will be scoped and JS fosters immutable JS entities and
-avoids variables leaking into the global namespace preferring the
-\texttt{let} or \texttt{const} declaration in favor of the old
-\texttt{var} declaration.
-
-The second principle of services is the \textbf{high cohesion
-principle}: Whether designing a microservice or it's components we want
-related behavior sit together, and unrelated behavior to sit
-elsewhere.{[}3, p. 30{]} High cohesion can be expressed in a dynamic way
-as the \emph{Single Responsibility Principle}: ``Gather together those
-things that change for the same reason and separate those things that
-change for different reasons.''{[}5{]} In a very quick and dirty code
-quality analysis, the quality can be measured just by counting the
-places changes in the code occur in order to implement a functionality.
-In an arbitrary MVC system when a \texttt{button} is dropped into the
-VIEW part, the CONTROLLER needs some adjustment and maybe the MODEL,
-too. Three places for adjustment is a reasonable easy task for the
-brain. In the field of browser based development those principle is
-often violated in the separated entities HTML, JS and CSS. A
-\emph{browsernative microservice} aims for combining all of the
-resources cohesively in a single place and file.
-
-A \emph{component for the web}, or web component, is
-\textbf{self-contained} which means it embodys all needed functionality
-to get it's job done. Therefore it has a much better evolution mechanism
-in the service contracts. Changing functionality won't break other
-components. A component can progressively enhanced which guarantees
-functionality throughout different versions.
-
-Using libraries in web development is a common sense. But compared to
-libraries, a component service offers multiple advantages. A library is
-only loosely coupled to the implementation and therefore hard to track
-in functionality. Changing a library may result in an unforeseen amount
-of time fixing implementations. It is not unusual to see websites
-embodying different versions of the same library (like with JQuery).
-Another issue with libraries is dead code elimination which means the
-process of removing code that is never going to be executed. New build
-tools for the web, like Webpack 2 or Rollup offer this feature which
-relies heavily on the static structure on ES6 modules.{[}6{]} Libraries
-for the browsers are traditionally ``shipped'' as non static
-\textbf{immediately-invoked functions}. Bootstrapping those is much
-harder to archive as dead code may be part of the function itself. A
-\emph{webnative microservices} should therefore embrace the high
-cohesion principle and abandon most of the libraries in favor of static
-ES6 module to lower the amount of dead code and to increase performance
-especially at the first render.
-
-Another advantage components have over libraries is the more explicit
-interface.{[}4{]} While the functionality of a library needs
-documentation, a component functionality is exposed via the components'
-signature which comes in the fashion of an HTML element in
-\emph{browsernative microservices}. For example, a button component can
-consume attributes and becomes a primary buttons just by assignment
-
-.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{<my-botton}\OtherTok{ isPrimary}\KeywordTok{>}\NormalTok{Primary}\KeywordTok{</my-botton>}
-\end{Highlighting}
-\end{Shaded}
-
-\subsection{Organized around Business
-Capabilities}\label{organized-around-business-capabilities}
-
-\begin{quote}
-``organizations which design systems \ldots{} are constrained to produce
-designs which are copies of the communication~structures~of these
-organizations''. {[}7{]}
-\end{quote}
-
-Emphasizing the human factor in microservices is a key feature.
-Microservices are a product of real-world usage.{[}3, p. 1{]} Instead of
-splitting team structures along the technology stack (UI Experts
--\textgreater{} Middleware -\textgreater{} Database) a microservice
-approach model teams around \textbf{business capabilities}.{[}4{]}
-Consequently every team is capable of planning, designing, implementing,
-testing and maintaining their very own microservice. Along the
-technology stack every member gains high competence about the
-functionality of the service.
-
-Real-world domains tend to be complex and multifaceted. To unfold their
-complexity, domains can subdivided into \textbf{bounded contexts}.{[}3,
-p. 31{]} For example, customer service is a business domain. Customers
-again have different contexts depending on their demands. One context
-can be sales, another context could be support. Every context makes
-different assumptions about the underlying model. Each bounded context
-draws an explicit interface where it decides what models to share with
-other contexts.{[}3, p. 30{]} Each context can derived into multiple
-microservices or, talking about \emph{Browsernative Microservices},
-interfaces to the customers.
-
-By assigning service responsibility to a team, the so called
-\textbf{Definition of Done} (DoD) shifts from ``accomplishing projects''
-to ``accomplishing products''. This new paradigm not only changes the
-administrative overhead like budgeting or resource allocation. It
-creates a kind of responsibility connection from the team to the
-service. Expectedly, those teams are more motivated within their very
-own service and exhibit a more sophisticated iteration time.{[}4{]}
-
-For many companies working in the spheres of the internet the
-client-side is highly important for their business. In fact, business
-goals and capabilities can be derived from the front-end needs. The
-state of the web is not only a story of numerous artifacts, it is also a
-story of an highly fragmented market along devices, operating systems,
-differing sizes and functionalities. Different devices again have
-different assumptions about the technology stack. Splitting teams along
-the stack results in an slow paced back and forth negotiation for every
-change to be made. As browser technologies, design guidelines and
-devices change in a fast paced manner it makes absolutely sense to shift
-responsibility towards the teams altogether. A \emph{browsernative
-microservice} embraces commitment over its whole lifecycle.
-
-\subsection{Smart endpoints and dumb
-pipes}\label{smart-endpoints-and-dumb-pipes}
-
-Microservices for the browsers aim for side-effects like changing the UI
-given to input parameters or emitting browser events. An isolated
-microservice won't make too much sense after all. To ensure
-communication between services, flexible yet powerful communication
-channels must be established.
-
-\emph{Smart endpoints and dumb pipes} is coined to the approach of
-designing communication mostly decoupled and as cohesive as
-possible.{[}4{]} In the analogy to the real world a message channel
-should look like sending a letter: Two smart endpoints (sender and
-receiver) and a mostly unified ``dumb'' letter format and channel. Each
-endpoint owns it specific domain logic. In the browsernative context we
-already emphasized the role of presentational components and containers.
-The ``smartness'' of presentational components is altering the UI and
-reacting on users input. The logic of a container follows the idea of an
-filter in a Unix sense - receiving a request, applying logic as
-appropriate and producing a response.{[}4{]}
-
-Cited earlier in this paper microservices often rely on simple HTTP
-request-response with resource API's and lightweight messaging as
-communication protocols.{[}4{]} Those technologies highly accessible and
-widely adopted and most probably usable for a browsernative service. In
-the implementation part of this paper a reader will explore the
-combination of RESTish protocols and a lightweight browser message bus
-in action.
-
-By focusing the service into clear defined business boundaries, it is
-easier to define a smart API of the service which in our case will be a
-set of attributes for the HTML element bound to the service.
-
-\subsection{Decentralized Governance}\label{decentralized-governance}
-
-Microservices are separate entities and decentralization is therefore a
-meta concept of microservices. Distributed service responsibility owned
-by the team is one aspect. Gluing independent service bricks to a whole
-distributed system talking over the network protocols another one.
-
-This distributed nature allows teams to create their own technology
-stack, tools and services designed in the spirit of language- and
-platform independence and share their knowledge with other
-parties.{[}4{]} In the recent years many big companies like Facebook,
-Google, Netflix and others followed that spirit and published their
-ideas and implementations open source. The previously mentioned ReactJS
-for example is a brainchild of Facebook. In fact, many tools and
-techniques are byproduct of vital interaction of concrete domain
-problems and their implementations.
-
-The spirit of freedom can't be applied universally to
-\emph{Browsernative Microservices} as the browser and its underlying DOM
-will be the limitation factor to a certain degree. Talking about the
-browser, a reader might be tempted to narrowly thinking of the obvious
-VIEW layer only - which is not true anymore. In the recent years the
-browser engines grows to a kind of virtual machine: there are connectors
-to build-in databases, multithreading support, an ever-growing JS
-build-ins and even speech synthesis. So-called \emph{Progressive Web
-Apps\footnote{\href{https://developers.google.com/web/progressive-web-apps}{Progressive
-  Web Apps}}}, a bunch of criteria for building good browserapps, can
-achieve a similar look and feel like native apps. Services like
-NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}},
-effectively compiling ``the web'' to native code, lower the boundary
-between native and browser code even further.
-
-JS is the widely accepted language of the web. Nevertheless, a
-microservice engineering team might choose another language for various
-reasons. Transpiling languages to JS as target language isn't exotic
-anymore. Languages like TypeScript, ClojureScript or PureScript compile
-to JS even exclusively. Once web components hit a critical mass there
-will be most likely some library support or foreign function interface
-towards ES6 modules (which are mandatory for the new specifications).
-With the rise of WebAssembly, a new low-level programming language for
-the browser, the determination on JS will hypothetical deteriorate and
-new quasi native languages for the web might gain traction.
-
-Another more real life decentralization aspect derives from the easiness
-of deployment in a safe, sandboxed environment. Web components virtually
-ship no overhead or require obscure build tools. This makes them ideal
-candidates for sharing and open-sourcing. Webcomponents.org\footnote{\href{https://webcomponents.org}{Webcomponents.org}}
-is a registry for ready-to-use components of every scale and purpose
-where for even Google shares a lot of their material design elements.
-
-\subsection{Decentralized Data
-Management}\label{decentralized-data-management}
-
-Data Management in Microservice follows the same modular philosophy like
-the service implementation. As mentioned earlier different bounded
-contexts make different assumptions of the underlying models. A
-\emph{browsernative microservices} takes this idea even further and
-expands it to the fragmented world of electronic devices. Decentralized
-decisions about conceptual models demand for decentralized data storage
-decisions.{[}4{]} Todays web architectures aims to leverage an
-increasing amout of processing to the client to avoid time-consuming
-roundtrips especially in mobiles networks.\footnote{Latency numbers:
-  https://gist.github.com/jboner/2841832} Since network roundtrips are
-costly it is a good advice to only query as much data as needed and
-cache as much as possible.
-
-``Microservices prefer letting each service manage its own
-database.''{[}Fowler2014{]} Ben Issa, chief architect of ING Australia
-emphasizes this pragmatism on APIs in a conference talk. At ING the
-frontend demands tailor the backend APIs, APIs may be produced
-automatically and not even Issa knows how many APIs exists.{[}8{]} They
-are using a pattern called \textbf{backend for frontends} empowering the
-team working to craft their UI and backend in a one-to-one
-relationship.{[}3, p. 72{]}
-
-To see this pattern in the field a reader might have a look at Facebooks
-GraphQL\footnote{\href{http://graphql.org/}{GraphQL}}. GraphQL is a
-query language for the frontend. The backend solely replies on the
-frontend needs. Another well documented example in the field is
-Cognitects Datomic\footnote{\href{http://www.datomic.com/}{Datomic}},
-where parts of the database will be reflected to the client. A so-called
-Transactor ensures ACID compliance.
-
-The simplified microservice example later in this paper assumes a
-generic build-in API accompanied by build-in frontend components.
-Instead of gluing frontend and backend together on runtime, the
-microservice is designed holistically containing both front- and
-backends. To reduce network calls especially for mobile devices it is a
-good advice to cache data (f.e. in a global Service Worker). Revamping
-offline capabilities even further data can be stored in a browser based
-database like PouchDB. For the sake of simplicity data management won't
-be explored into depth throughout this paper.
-
-\subsection{Infrastructure Automation}\label{infrastructure-automation}
-
-In the global nature of web development the development couldn't
-completely decoupled from the production environment. This circumstance
-left developers switching back and forth between files developing tricky
-opinionated (and more often biased) ways to glue related parts together.
-With scoped components, code blocks can be developed more sane and
-conveyed into production without headache. Even more, in the scripted
-environment of the browser this can be implemented rapidly and
-continuous.
-
-Previously mentioned Ben Issa, described the ING standard workflow.
-Every component deserves a own \textbf{git repo} containing
-
-\begin{itemize}
-\tightlist
-\item
-  Internationalization conformity (i18n)
-\item
-  Accessibility conformity (a11y)
-\item
-  Tests for the component
-\item
-  Demos of the component
-\item
-  Blueprints to mock the one to one APIs
-\item
-  Docs
-\end{itemize}
-
-Every check-in is handled as release candidate and can be independently
-deployed by a fully automated deployment machinery.{[}8{]} Even though
-this example is an opinionated perception it gives a sense of a mature
-component build for the web. Due to an exhaustive amount of testing and
-deployment tools for the browser an automated infrastructure shouldn't
-be a problem.
-
-\subsection{Design for failure}\label{design-for-failure}
-
-In theory a microservice is designed with a lot of emphasizes on
-real-time monitoring for both the architectural elements and business
-relevant metrics.{[}4{]} Due to the modular structure weak points can
-occur in the orchestration of the services. Testing and automation is a
-feasible task but failures may occur in another end users setting
-undetected. Legacy browsers for example remain a problem for enhancing
-websites with new technologies.
-
-Regarding the evolution of the web, the ``next billion'' internet users
-most likely using Android, have decent specs mobile phones, use an
-evergreen browser but won't have a reliable internet connection.{[}9{]}
-While \emph{Progressive Enhancement} was once coined on the principle to
-build websites both for Browsers with JS and HTML only, the new
-\emph{Progressive Enhancement} tends towards an ``\textbf{offline
-first}'' build principle. A \emph{Browsernative Microservice} therefore
-not only tries to cache data as much as possible, it should also bring
-in a lot of program logic as described in the previous chapters.
-
-Working in a JS heavy infrastructure demands for optimization to avoid
-unexpected side-effects like the \emph{flash of unstyled content
-(FOUC)}. Googles Polymer propagates the a general-purpose \textbf{PRLP
-pattern}\footnote{\href{https://www.polymer-project.org/1.0/toolbox/server}{PRLP
-  pattern}}:
-
-\begin{itemize}
-\tightlist
-\item
-  Push critical resources for the initial route
-\item
-  Render initial route
-\item
-  Pre-cache remaining routes
-\item
-  Lazy-load and create remaining routes on demand
-\end{itemize}
-
-Following this pattern a critical resource can detect browser
-functionalities beforehand and switch to a \textbf{polyfill} instead of
-the latest browser optimized version. After the initial paint, critical
-resources like top-level microservices or other app logic can be loaded
-and registered.
-
-\subsection{Evolutionary Design}\label{evolutionary-design}
-
-Microservices tend to become smaller over time. An evolutionary design
-approach puts emphasizes on decomposition and scrapping the service.
-``The key property of a component is the notion of independent
-replacement and upgradeability.''{[}4{]} Therefore we can safely change
-and chop services. Lazy parts of the system which won't change often
-should be separated from parts undergoing a lot of churn.{[}4{]} Parts
-that needs coupled changes could should be moved together or should be
-even merged.
-
-This flexible approach fits good in the world of browser based
-development. In the last decade we have seen a lot of of changes in the
-way we develop for the web. New approaches like the virtual DOM approach
-found their way into mainstream web development followed by frameworks
-and libraries.
-
-\emph{Browsernative Microservices} should be perceived as complementary
-technology. Contrary to Angular, React and other frameworks they have a
-strong interop with existing systems and can be used with them together.
-
-Andrew Rota, developer at Wayfair, came up with the idea to compose
-small web components around a managing React system.{[}10{]} As web
-components are native elements there is no difference to use a native
-\texttt{button} or a native \texttt{custom-button}. With this pattern a
-web developer can make use of the encapsulated advantages of web
-components while still making use of the declarative event management
-from various frameworks. Whatever new framework will be on the rise
-within the next years, this approach allows rapid decomposition and
-reassembling towards a new system.
-
-\section{W3C specifications}\label{w3c-specifications}
-
-For building a native microservice running on the ``bare-metal'' browser
-engine requires a bunch of new specifications and assumptions. Most
-importantly the quasi specification \textbf{Web Components} is needed.
-\emph{Web Components} is not a real standard. It's an amalgam of APIs
-from multiple w3c specs which can be used independently, too. A
-webdeveloper may choose one spec and embrace the freedom in architecture
-which can be combined with other frameworks/libraries.
-
-Depending on the context, some people argue for only two specs which
-essentially make it possible to create a scoped component but not caring
-too much on it's distribution{[}11{]}. Some people prefer the three
-specs {[}12{]}, but the majority advocating the four specs variant,
-which is listed on \sout{the official}
-\href{http://webcomponents.org}{webcomponents.org} website. For the
-purpose of this article, all four specs will be discussed briefly to
-provide a rough understanding. It is not meant to cover all bits and
-pieces.
-
 \subsection{\texorpdfstring{Custom Elements
@@
  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
-  \KeywordTok{super}\NormalTok{()}\OperatorTok{;} \CommentTok{// mandatory!}
+  \KeywordTok{super}\NormalTok{()}\OperatorTok{;} \CommentTok{// mandatory in constructor}
   \KeywordTok{this}\NormalTok{.}\AttributeTok{onclick} \OperatorTok{=} \NormalTok{e }\OperatorTok{=>} \AttributeTok{alert}\NormalTok{(}\StringTok{"hello"}\NormalTok{)}\OperatorTok{;}
@@
 \KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
- \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}\NormalTok{...}\OperatorTok{\}}
  \NormalTok{set }\AttributeTok{sayhello}\NormalTok{(val) }\OperatorTok{\{}
@@
  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
-  \NormalTok{...}
+   \CommentTok{// this.onclick...}
    \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
@@
 https://shop.polymer-project.org/checkout. Currently the checkout is a
-single, 671 lines of code long Polymer component incorporation all
-required fields for sign in, shipping, billing and summarizing the
-order. In the spirit of microservices we will split up the microservice
-into fine grained components. The shopping cart data is pulled out of a
-local storage JSON entity set up previously by another custom-element.
-Another in-memory opportunity is storing the data in the build-in
-database IndexedDB.
+single, 671 lines of code long Polymer component including all required
+fields for sign in, shipping, billing and summarizing the order. In the
+spirit of microservices we will split up the microservice into fine
+grained components. The shopping cart data is pulled out of a local
+storage JSON entity set up previously by another custom-element.
 
@@
 \begin{Highlighting}[]
-\CommentTok{// Within every child node class}
-\NormalTok{set }\AttributeTok{dispatch}\NormalTok{(msg) }\OperatorTok{\{}
+\NormalTok{> shop-checkout.html}
+\CommentTok{<!-- IMPORTS -->}
+\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"sign-in.html"}\OtherTok{ async}\KeywordTok{>}
+\NormalTok{...}
+\CommentTok{<!-- VIEW -->}
+\KeywordTok{<shop-checkout>}
+  \KeywordTok{<sign-in></sign-in>}
+  \NormalTok{...}
+\KeywordTok{</shop-checkout>}
+
+\CommentTok{<!-- CONTROLLER -->}
+\KeywordTok{<script>}
+\KeywordTok{class} \NormalTok{ShopCheckout }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
+    \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
+    \KeywordTok{this}\NormalTok{.}\AttributeTok{MODEL} \OperatorTok{=} \KeywordTok{new} \AttributeTok{Worker}\NormalTok{(}\StringTok{'checkout-model.js'}\NormalTok{)}\OperatorTok{;}
+  \OperatorTok{\}}
+
+  \AttributeTok{connectedCallback}\NormalTok{() }\OperatorTok{\{}
+    \CommentTok{// listen to msg from MODEL}
+    \KeywordTok{this}\NormalTok{.}\VariableTok{MODEL}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'message'}\OperatorTok{,} \NormalTok{[Msg Handler])}
+    \CommentTok{// listen to child nodes}
+    \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'checkout'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>} \OperatorTok{\{}
+      \CommentTok{// delegate event to handler}
+      \KeywordTok{this}\NormalTok{.}\AttributeTok{_sendToMODEL} \OperatorTok{=} \NormalTok{e}\OperatorTok{;} 
+    \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+  \OperatorTok{\}}
+
+  \NormalTok{set }\AttributeTok{_sendToMODEL}\NormalTok{(e) }\OperatorTok{\{}
+    \KeywordTok{const} \NormalTok{id }\OperatorTok{=} \VariableTok{e}\NormalTok{.}\VariableTok{target}\NormalTok{.}\AttributeTok{id}\OperatorTok{,} 
+          \NormalTok{name }\OperatorTok{=} \VariableTok{e}\NormalTok{.}\VariableTok{target}\NormalTok{.}\AttributeTok{localName}\OperatorTok{,}
+          \NormalTok{load }\OperatorTok{=} \VariableTok{e}\NormalTok{.}\AttributeTok{detail}\OperatorTok{,}
+          \NormalTok{letter }\OperatorTok{=} \VariableTok{Object}\NormalTok{.}\AttributeTok{assign}\NormalTok{(}\OperatorTok{\{\},} \OperatorTok{\{} \NormalTok{id}\OperatorTok{,} \NormalTok{name}\OperatorTok{,} \NormalTok{load }\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+    \CommentTok{// send to MODEL}
+    \KeywordTok{this}\NormalTok{.}\VariableTok{MODEL}\NormalTok{.}\AttributeTok{postMessage}\NormalTok{(letter)}\OperatorTok{;} 
+  \OperatorTok{\}}
+  \NormalTok{...}
+\OperatorTok{\}}    
+\VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'shop-checkout'}\OperatorTok{,} \NormalTok{shopCheckout)}\OperatorTok{;}
+\OperatorTok{<}\SpecialStringTok{/script>}
+\end{Highlighting}
+\end{Shaded}
+
+The purpose of this simplified code snippet is to outline the MVC
+threefold in the service root. On runtime the root container acts mostly
+like a \textbf{message dispatcher} implementing the microservice
+principle of smart endpoints and dumb pipes.
+
+In order to interact with the MODEL, every child node must implement the
+\textbf{dedicated \texttt{checkout}} custom event (which will be
+explained further on in the next section). Summarizing the
+unidirectional communication from the service root perspective looks
+like following plot:
+
+\begin{verbatim}
+VIEW thread                 |               MODEL thread
+                            |
++----------+                |               +----------+
+| Checkout | ----------Event Msg--------->  |  Msg     |
+|  Event   |                |               | Handler  |
++----------+                |               +----------+
+                            |                   |
+                            |                   |
+                            |               +----------+
+                            |               | Generate |
+                            |               |  Effect  |
+                            |               +----------+
+                            |                   |
+                            |                   |
++----------+                |               +----------+
+|  Msg     |                |               | Postmsg  |
+| Handler  | <--------Action Msg----------  |  Action  |
++----------+                |               +----------+    
+                            |
+\end{verbatim}
+
+Effects are yielded by the asynchronous operation of messages and create
+actions returned to sender by the message handlers. Effects can be
+created by external messages, like subscription to an WebSocket, too.
+Effects may be created with additional resources from the server or
+syncing with local storage like in the polymer-shop.
+
+While the msg handlers are basic ``dumb'' switch statements the
+\textbf{smartness} solely arouse from intelligent controllers processing
+the message. This communication model offers lots of possibilities for
+\textbf{evolutionary design} as child nodes can be loosely dropped in
+the service root and communication might \textbf{fail graceful} as
+simply nothing happens if one or another switch is not defined.
+
+\subsubsection{Container components}\label{container-components}
+
+The second layer of the checkout microservice like
+\texttt{\textless{}sign-in\textgreater{}} or
+\texttt{\textless{}shipping-details\textgreater{}} still contain a lot
+of logic and no or less styling. Their job is to control their
+underlying presentational leaf components, aggregating events and to
+define a \textbf{set of actions towards the model} . Every container
+mounted directly under the service root may have a dedicated area inside
+the worker thread where it's fulfil his duties. For example, a typical
+\texttt{\textless{}sign-in\textgreater{}} contains merely two fields
+username and password and a submit button. The component listens for the
+submit action, aggregating the credentials and might add some semantics
+to it like \emph{action: `SIGNIN SUBMITTED'}. Fields and action message
+will be dispatched towards the model for further processing like
+initiating a authorization process.
+
+Every container component is eligible to aggregate subordinate events
+from their children, buffer them and interact with the model trough via
+a \textbf{unified message system}. A base class, from which
+\texttt{\textless{}sign-in\textgreater{}} or
+\texttt{\textless{}shipping-details\textgreater{}} can be extended can
+look like this:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{class} \NormalTok{CheckoutContainerBase }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+  \CommentTok{// static methods for testing...}
+  \NormalTok{set }\AttributeTok{_dispatch}\NormalTok{(msg) }\OperatorTok{\{}
     \KeywordTok{this}\NormalTok{.}\AttributeTok{dispatchEvent}\NormalTok{(}
@@
 
-Most obvious ist the gathering of all related code under the umbrellar
-of a single HTML tag.
-
-Secondly, the sub-standard \emph{custom elements} introduces so called
-lifecycle methods and a getter/setter interface exposing the
-functionality to the developer. Event handling, for example, can be
-registered in place which is much more declarative than assigning event
-listeners from the outside. Of course, this events can be pushed down to
-nested tags, allowing an increasingly granular system design. This
-approach will be explained further in the upcoming sections.
-
-Following this logic any company, whether it is web-related or not,
-should be devided in units grouped around a destinct business service to
-optimise the workflow. Fowler and Lewis outlines this approach as an
-``alignment of business capabilities''{[}4{]} While this kind of
-structure may be true for companies like Google or Amazon, there is a
-vast majority of companies developing for the web which are grouped
-around tasks.{[}8{]} A very common structure is formed by the technology
-stack (UX Designers, Frontend- \& Backend Developers) or by separating
-teams along the product lifecycle (development, testing, deployment).
-
-Advocators from the microservice approach propose a different model.
-best described by . Web components are one (but important) way to tie up
-those diciplines as one component can host a single independent business
-service. Combined with a flexible backend service these components can
-be huge gain over the cumbersome functional organizational approach.
-
-dichotome Pattern
-
-Ein üblicher eventgesteuerter Webservice setzt sich aus
-unterschiedlichsten Komponenten zusammen, die wiederum
-unterschiedlichste Eventlistener \& -emitter in sich subsummieren. Diese
-inhärente Komplexität verlangt geradezu nach einer klaren,
-deterministischen Struktur des Webservices, die das Zusammenspiel
-orchestriert. In der Analogie des Orchesters gesprochen, benötigt der
-Webservice (oder sogar die gesamte Webapplikation) einen Dirigenten, der
-für die Steuerung verantwortlich ist.
-
-http://alistair.cockburn.us/Hexagonal+architecture
-
-MVC Pattern
-
-Pure frontend vs heavy backend
+virtual DOM
+
+declarative event management
 
@@
 {[}3{]} S. Newman, \emph{Building microservices}. O'Reilly Media, Inc.,
-2016 {[}Online{]}. Available:
-\url{http://www.ebook.de/de/product/22539693/sam_newmann_building_microservices.html}
+2016.
 

### 2017-02-24 build/paper.tex
pdf/tex build

@@
 \newcommand{\NormalTok}[1]{{#1}}
+\usepackage{longtable,booktabs}
+% Fix footnotes in tables (requires footnote package)
+\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
 \usepackage[normalem]{ulem}
@@
 }
-\section{Simplicity and the web}\label{simplicity-and-the-web}
+\section{1. Simplicity and the web}\label{simplicity-and-the-web}
 
@@
 
-Derived from the ancient Latin word \textbf{simplex}, simple can be
-understood as ``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
+Derived from the Latin word \textbf{simplex}, simple can be understood
+as ``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
   Dictionary}} which points directly to the unidimensional aspect. While
@@
 to propose \emph{The Extensible Web Manifesto}.\footnote{\href{https://extensiblewebmanifesto.org/}{The
-  Extensible Web Manifesto}} The claim of the manifesto was to enhance
-the current web platforms with new low-level capabilities. Those
-features should empower creators of the web to write more declarative
-code and therefore overcome known bottlenecks and artificial
-abstractions. Four years later, the enhancement of JavaScript
-leapfrogged and many new low-level APIs brought to life. With this new
-APIs at hand a vivid web developer can create robust websites with less
-code and less additional libraries. This paper is an approach to unfold
-these \textbf{browsernative} technologies to create overall simple and
-resilient \textbf{microservices} for the browser
-
-\section{Microservices}\label{microservices}
+  Extensible Web Manifesto}} The manifestos claim was to enhance the
+current web platforms with new low-level capabilities. Those features
+should empower creators of the web to write more declarative code and
+abandon known problems and artificial abstractions. Four years later,
+the enhancement of JavaScript leapfrogged and many new low-level APIs
+brought to life. With this new APIs at hand a vivid web developer can
+create robust websites with less code and less additional libraries.
+This paper is an approach to unfold these \textbf{browsernative}
+technologies to create overall simple and resilient
+\textbf{microservices} for the browser
+
+\section{2. Microservices}\label{microservices}
 
@@
 Microservices incorporate a wide array of ideas from developing scalable
-software, like \emph{domain-driven design} where it pursues the
+software, like \textbf{domain-driven design} where it pursues the
 incorporation of real world structure in the code.{[}3, p. 2{]} Or
-making use of \emph{continuous delivery} for pushing software rapidly
-through \emph{automated deployment} mechanisms into production.{[}4{]}
+making use of \textbf{continuous delivery} for pushing software rapidly
+through \textbf{automated deployment} mechanisms into production.{[}4{]}
 Furthermore, microservices transcendent the technical perspective and
@@
 paper relies on the work of Sam Newman, who has written a comprehensive
-guide called \emph{Building Microservices} and the work of Fowler and
-Lewis. The purpose of this section is to gain confidence about the
-microservices architecture in the context of the browser platform.
-
-\subsection{Componentization via
+guide towards microservices{[}3{]} and the work of Fowler and
+Lewis{[}4{]}. The purpose of this section is to gain confidence about
+the microservices architecture in the context of the browser platform.
+
+\subsection{2.1. Componentization via
 Services}\label{componentization-via-services}
@@
 related behavior sit together, and unrelated behavior to sit
-elsewhere.{[}3, p. 30{]} High cohesion can be expressed in a dynamic way
-as the \emph{Single Responsibility Principle}: ``Gather together those
-things that change for the same reason and separate those things that
-change for different reasons.''{[}5{]} In a very quick and dirty code
-quality analysis, the quality can be measured just by counting the
+elsewhere.{[}3, p. 30{]} High cohesion can be enhanced towards the more
+dynamic \textbf{Single Responsibility Principle}: ``Gather together
+those things that change for the same reason and separate those things
+that change for different reasons.''{[}5{]} In a very quick and dirty
+code quality analysis, the quality can be measured just by counting the
 places changes in the code occur in order to implement a functionality.
@@
 deliberately unpredictable. The high cohesion principle is violated by
-the traditional separation along the technological entities
-\textbf{HTML, JS and CSS}. Understanding the relation of HTML markup and
-another CSS file adds incidental complexity. Different approaches
-emerged over the recent years to join the forces. Most notable the
-dedicated frontend language \emph{elm}\footnote{\href{http://elm-lang.org/}{elm
-  lang}} where model, view and controller sit in one place. A
-browsernative service aims for a combinement of the web native trinity
-HTML, JS and CSS.
+the traditional separation along the siloed entities \textbf{HTML, JS
+and CSS}. Understanding the relation of HTML markup and another CSS file
+adds incidental complexity. Different approaches emerged over the recent
+years to join the forces. Most notable the dedicated frontend language
+\emph{elm}\footnote{\href{http://elm-lang.org/}{elm lang}} where MODEL,
+VIEW and CONTROLLER sit in one place. A browsernative service aims for a
+combinement of the web native trinity HTML, JS and CSS.
 
@@
 
-\subsection{Organized around Business
+\subsection{2.2. Organized around Business
 Capabilities}\label{organized-around-business-capabilities}
@@
 
-\subsection{Smart endpoints and dumb
+\subsection{2.3. Smart endpoints and dumb
 pipes}\label{smart-endpoints-and-dumb-pipes}
@@
 Microservices heavily rely on simple HTTP request-response with resource
-APIs and lightweight messaging.{[}4{]} Newman puts his stand on
+APIs and lightweight messaging.{[}4{]} Newman puts his recommendation on
 technologic-agnostic REST APIs to free data persistence from
@@
 
-\subsection{Decentralized Governance}\label{decentralized-governance}
+\subsection{2.4. Decentralized
+Governance}\label{decentralized-governance}
 
@@
 ever-growing JS build-ins like speech synthesis or push notifications.
-So-called \emph{Progressive Web Apps\footnote{\href{https://developers.google.com/web/progressive-web-apps}{Progressive
-  Web Apps}}}, a bunch of criteria for building good browser apps, can
+So-called \textbf{Progressive Web Apps}\footnote{\href{https://developers.google.com/web/progressive-web-apps}{Progressive
+  Web Apps}}, a bunch of criteria for building good browser apps, can
 achieve a similar look and feel like native apps. And last but not least
@@
 
-\subsection{Decentralized Data
+\subsection{2.5. Decentralized Data
 Management}\label{decentralized-data-management}
@@
 
-\subsection{Infrastructure Automation}\label{infrastructure-automation}
+\subsection{2.6. Infrastructure
+Automation}\label{infrastructure-automation}
 
@@
 Bret Victor, UI designer at Apple defined the importance of an
-\emph{immediate feedback principle} for developing user
+\textbf{immediate feedback principle} for developing user
 interfaces.\footnote{\href{https://vimeo.com/36579366}{Bret Victor -
@@
 development within a single file containing all bits and pieces of the
-web component. Every major browsers devtools offer a direct file
-manipulation functionality so development can take place directly in
-place.
+web component. Every major browser devtool offers a direct file
+manipulation functionality so development can be even in place.
 
 When it comes to standardized deployment guidelines previously mentioned
-Ben Issa, described the ING standard workflow. Every component deserves
-a own \textbf{git repo} containing
+Ben Issa, described the ING standard workflow. Every component comes in
+its own \textbf{git repo} containing:
 
@@
 
-This example should made clear that synthetic tests of the component are
-part of the service right from the start. Every check-in is handled as
-release candidate and can be independently tested and deployed by a
-fully automated machinery.{[}8{]} Even though this example is an
-opinionated perception it gives a sense of a mature component build for
-the web. Due to an exhaustive amount of testing and deployment tools for
-JS an automated infrastructure shouldn't be a problem.
-
-\subsection{Design for failure}\label{design-for-failure}
-
-In theory a microservice is designed with a lot of emphasizes on
-real-time monitoring for both the architectural elements and business
-relevant metrics.{[}4{]} Due to the modular structure weak points can
-occur in the orchestration of the services. Testing and automation is a
-feasible task but failures may occur in another end users setting
-undetected. Legacy browsers for example remain a problem for enhancing
-websites with new technologies.
+Even though this example is an opinionated perception it gives a sense
+of a mature component build for the web. This example should made clear
+that all parts of the component put together in one place. Every
+check-in is handled as release candidate and can be independently tested
+and deployed by a fully automated machinery.{[}8{]} Due to an exhaustive
+amount of testing and deployment tools for JS an automated
+infrastructure shouldn't be a problem.
+
+\subsection{2.7. Design for failure}\label{design-for-failure}
+
+In theory a microservice is designed with focus on monitoring for both
+the architectural elements and business relevant metrics.{[}4{]} Due to
+the modular structure weak points can occur in the orchestration of the
+services. A microservice should track down every communication flow and
+provide defaults and meaningful error messages where communication
+stuck. Testing every single component with predefined synthetic events
+ensures functionality. Nevertheless, browser support may vary and legacy
+browsers remain a general problem for enhancing websites with new
+technologies and therefore demand further configuration.
+
+Combinment of the resources in the browser always demanded for
+optimization to avoid unexpected side-effects like \emph{flash of
+unstyled content}. Googles Polymer propagates the a general-purpose
+pattern called \textbf{PRLP}\footnote{\href{https://www.polymer-project.org/1.0/toolbox/server}{PRLP
+  pattern}}:
+
+\begin{itemize}
+\tightlist
+\item
+  Push critical resources for the initial route
+\item
+  Render initial route
+\item
+  Pre-cache remaining routes
+\item
+  Lazy-load and create remaining routes on demand
+\end{itemize}
+
+Following this pattern a critical resource can evaluate browser maturity
+beforehand and switch to a \textbf{polyfill} or another fallback
+solution instead of the latest browser optimized version. After the
+initial paint, critical resources like top-level microservices or other
+app logic can be loaded and registered.
 
@@
 evergreen browser but won't have a reliable internet connection.{[}9{]}
-While \emph{Progressive Enhancement} was once coined on the principle to
-build websites both for Browsers with JS and HTML only, the new
-\emph{Progressive Enhancement} tends towards an ``\textbf{offline
-first}'' build principle. A \emph{Browsernative Microservice} therefore
-not only tries to cache data as much as possible, it should also bring
-in a lot of program logic as described in the previous chapters.
-
-Working in a JS heavy infrastructure demands for optimization to avoid
-unexpected side-effects like the \emph{flash of unstyled content
-(FOUC)}. Googles Polymer propagates the a general-purpose \textbf{PRLP
-pattern}\footnote{\href{https://www.polymer-project.org/1.0/toolbox/server}{PRLP
-  pattern}}:
-
-\begin{itemize}
-\tightlist
-\item
-  Push critical resources for the initial route
-\item
-  Render initial route
-\item
-  Pre-cache remaining routes
-\item
-  Lazy-load and create remaining routes on demand
-\end{itemize}
-
-Following this pattern a critical resource can detect browser
-functionalities beforehand and switch to a \textbf{polyfill} instead of
-the latest browser optimized version. After the initial paint, critical
-resources like top-level microservices or other app logic can be loaded
-and registered.
-
-\subsection{Evolutionary Design}\label{evolutionary-design}
+While \textbf{Progressive Enhancement} was once related to build
+websites both for browsers with and without JS support the demands
+changed tends towards an \textbf{offline first} principle avoiding
+network connectivity failures.{[}9{]} A \emph{browsernative
+microservice} therefore not only tries to cache data as much as
+possible, it should also bring in a lot of program logic as described in
+the previous chapters.
+
+\subsection{2.8. Evolutionary Design}\label{evolutionary-design}
 
@@
 replacement and upgradeability.''{[}4{]} Therefore we can safely change
-and chop services. Lazy parts of the system which won't change often
-should be separated from parts undergoing a lot of churn.{[}4{]} Parts
-that needs coupled changes could should be moved together or should be
-even merged.
-
-This flexible approach fits good in the world of browser based
-development. In the last decade we have seen a lot of of changes in the
-way we develop for the web. New approaches like the virtual DOM approach
-found their way into mainstream web development followed by frameworks
-and libraries.
+and chop services. Lazy components of the system which won't change
+often should be separated from parts undergoing a lot of churn.{[}4{]}
+And services which change for the same reason might be moved together or
+even could be merged.
+
+Pursuing flexibility for web development is a selling point as
+innovation cycles for browser development is fast paced and technologies
+can change quickly. Frontend related hardware, software and
+methodologies innovate rapidly over time.
 
 \emph{Browsernative microservices} should be perceived as complementary
-technology. Contrary to Angular, React and other frameworks they have a
-strong interop with existing systems and can be used with them together.
-
-Andrew Rota, developer at Wayfair, came up with the idea to compose
-small web components around a managing React system.{[}10{]} As web
-components are native elements there is no difference to use a native
-\texttt{button} or a native \texttt{custom-button}. With this pattern a
-web developer can make use of the encapsulated advantages of web
-components while still making use of the declarative event management
-from various frameworks. Whatever new framework will be on the rise
-within the next years, this approach allows rapid decomposition and
-reassembling towards a new system.
-
-\section{W3C specifications}\label{w3c-specifications}
+technology in contrast to full-service frameworks like Angular. Being
+native technology pursues a strong interopt approach with existing
+systems. Andrew Rota for example came up with a pattern using small,
+encapsulated and stateless web components as leaves in the tree of React
+components instead of native HTML elements.{[}10{]} Even React can
+eventually profit from the expressiveness of custom components using a
+custom \texttt{\textless{}meaningful-button\textgreater{}} over a native
+\texttt{\textless{}button\textgreater{}}. Most likely there will be
+always some cutting edge framework promising advantages over native
+code. Whatever new framework will be on the rise within the next years
+native components can eliminate future uncertainty and allowing rapid
+reassembling towards new architectures.
+
+\section{3. W3C specifications}\label{w3c-specifications}
 
@@
 importantly the quasi specification \textbf{Web Components} is needed.
-\emph{Web Components} is not a real standard. It's an amalgam of APIs
-from multiple w3c specs which can be used independently, too. A
-webdeveloper may choose one spec and embrace the freedom in architecture
-which can be combined with other frameworks/libraries.
+Web Components is not a real standard. It's an amalgam of APIs from
+multiple w3c specs which can be used independently. A web developer may
+choose one spec and embrace the freedom in architecture which can be
+combined with other frameworks/libraries.
 
@@
 \emph{Disclaimer:} This paper introduces many new browser build-ins with
-the focus on try and test. As the time of writing, many examples can be
+the focus on testability. As the time of writing, many examples can be
 tried frictionless in the console of the latest versions of
@@
 
-\subsection{\texorpdfstring{Custom Elements
-\href{http://w3c.github.io/webcomponents/spec/custom/}{(w3c)}}{Custom Elements (w3c)}}\label{custom-elements-w3c}
-
-Custom elements are the fundamental building blocks for web components
-introducing the \emph{Single Responsibility Principle} to the browser.
-In essence, they provide a way to create \textbf{custom HTML tags}
+\subsection{\texorpdfstring{3.1. Custom Elements
+\href{http://w3c.github.io/webcomponents/spec/custom/}{(w3c)}}{3.1. Custom Elements (w3c)}}\label{custom-elements-w3c}
+
+\emph{Custom elements} are the fundamental building blocks for web
+components introducing the \textbf{Single Responsibility Principle} to
+the browser. In short, they provide a way to create custom HTML tags
 subsuming behavior, design and functionality. An obligatory
@@
 
-This example should be self-explanatory. Notably, custom elements come
-in the fashion of \emph{ES6 Classes}\footnote{\href{https://developer.mozilla.org/en/docs/Web/JavaScript/Reference/Classes}{JavaScript
+This example should be almost self-explanatory in functionality.
+\emph{Custom elements} come in the fashion of \textbf{ES6
+Classes}\footnote{\href{https://developer.mozilla.org/en/docs/Web/JavaScript/Reference/Classes}{JavaScript
   Classes}} in favor of the normal JavaScript prototype-based
-inheritance model. This class must inherit the base \texttt{HTMLElement}
-interface which~``ensures the newly created element inherits the entire
-DOM API and any properties/methods that you add to the class become part
-of the element's DOM interface.''{[}13{]} Like any other \emph{ES6
-class}, the new element can be specialized further on using the typical
-\texttt{extends} inheritance.
-
-The beauty of \emph{custom elements} comes with the keyword
-\texttt{this} which points to the element itself. Instead of querying
-and assigning behavior AFTER creation of the node, custom elements ship
-their functionality PRIOR initialization of the element. The so called
-\emph{fat-arrow} (\texttt{=\textgreater{}}) is just a new ES6 syntax
-feature for an anonymous function declaration.
-
-After declaration, the new HTML element needs to be registered in the
-global build-in \texttt{customElements} object with an tag name like
-\texttt{\textless{}hello-world\textgreater{}} acting as key to the
-element. Mind the dash inside the tag name to conform the spec. Finally,
-the new element can go live inside the HTML Document
-\texttt{index.html}.
-
-\subsubsection{Lifecycle methods}\label{lifecycle-methods}
-
-In addition to the \texttt{constructor()}, the spec defines so called
-\emph{lifecycle callbacks} for controlling the \textbf{behaviour in the
-DOM}. Many popular frameworks like ReactJS or AngularJS rely on similar
-approaches:
+inheritance model which was part of an older specification. Every valid
+element must extend the base \texttt{HTMLElement} interface
+which~``ensures the newly created element inherits the entire DOM API
+and any properties/methods that you add to the class become part of the
+element's DOM interface.''{[}13{]} Like any other ES6 class any
+\emph{custom element} can be specialized further using the typical
+inheritance model allowing higher levels of abstraction.
+
+The beauty of \emph{custom elements} comes with the \textbf{bounded
+\texttt{this} keyword} which points to the element itself. Instead of
+querying and assigning behavior after creation of the node, custom
+elements ship their functionality on initialization of the element. The
+so called \emph{fat-arrow} (\texttt{=\textgreater{}}) is just a new ES6
+syntax feature for an anonymous function declaration.
+
+After declaration the new HTML element needs to be registered in the
+global build-in \texttt{customElements} object with an dedicated tag
+name acting as key to the element. Mind the dash inside the tag name to
+conform the spec. Finally, the new element can be mounted inside the
+HTML document.
+
+\subsubsection{3.1.1. Lifecycle methods}\label{lifecycle-methods}
+
+In addition to the constructor which runs procedures on initialization,
+the spec defines \textbf{lifecycle callbacks} for controlling elements
+behavior towards DOM interaction. Many popular frameworks like React or
+Angular rely on similar approaches:
 
@@
 \item
-  \texttt{connectedCallback()}\\
-  Called upon the time of \emph{connecting or upgrading the node} which
-  means the moment the node is rendered inside the DOM. Typically this
-  method is called straight after the \texttt{constructor()} on insert.
-  Typically, this method contains setup code, such as fetching resources
-  or rendering elements according to attributes.{[}13{]} For a fast
-  initial render of the page, it is highly preferable to put many
-  proceedings in favor of the constructor.
-\item
-  \texttt{disconnectedCallback()}\\
-  Called upon the time of \emph{node removal}. Cleanup code like
-  removing eventListeners or disconnecting websockets can be put here.
-\item
-  \texttt{attributeChangedCallback(attrName,\ oldVal,\ newVal)}\\
-  This method provides an \emph{Onchange handler} that runs for certain
-  attributes called with three values as defined in the signature. It is
-  meant to control an elements' transition from on \texttt{oldVal} to a
-  \texttt{newVal}. Due to performance issues, this callback is only
-  triggered for attributes registered in an \emph{observedAttributes}
-  array.
-\item
-  \texttt{adoptedCallback()}\\
-  Called when moving the node \emph{between documents}.
+  connectedCallback()\\
+  Called upon the time of \textbf{connecting or upgrading the node}
+  which means the moment the node is rendered inside the DOM. Typically
+  this method is called straight after the constructor if the node is
+  inserted directly. For a faster initial render of the page it is
+  highly preferable to put many proceedings in this method. Usually this
+  method contains setup code such as fetching resources or rendering
+  elements according to attributes.{[}13{]}
+\item
+  disconnectedCallback()\\
+  Called upon the time of \textbf{node removal}. Cleanup code like
+  removing event listeners or disconnecting web sockets can be put here.
+\item
+  attributeChangedCallback(attrName, oldVal, newVal)\\
+  This method provides an \textbf{onchange handler} for certain elements
+  attributes. This method is used to guide elements' transition from an
+  old value to a new state. Due to performance issues this callback is
+  only triggered for attributes registered in a dedicated array shipped
+  with the element.
+\item
+  adoptedCallback()\\
+  Called when moving the node \textbf{between documents}. This method
+  comes handy when using HTML imports described later.
 \end{itemize}
 
-\subsubsection{Custom attributes}\label{custom-attributes}
-
-As previously mentioned, the custom elements must \texttt{extend} the
-\texttt{HTMLElement} interface. Therefore, the new element inherits base
-properties and methods commonly used in all HTML elements like
-\texttt{id,\ class,\ addEventListner} . Additionally, it is possible to
-define custom attributes using the \emph{custom elements'}
-\textbf{getter / setter interface} to steer the behavior of the element.
+\subsubsection{3.1.2. Custom attributes}\label{custom-attributes}
+
+As mentioned earlier any custom elements must extend the
+\texttt{HTMLElement} interface ensuring base properties and methods used
+in throughout all HTML elements like \textbf{id, class, addEventListner
+\ldots{}}. Additionally, it is possible to define custom attributes
+using the \emph{custom elements'} \textbf{getter / setter interface} to
+steer the behavior of the element.
 
@@
 \KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
- \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}\NormalTok{...}\OperatorTok{\}}
   \NormalTok{set }\AttributeTok{sayhello}\NormalTok{(val) }\OperatorTok{\{}
@@
   \OperatorTok{\}}
-\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+\OperatorTok{\}}
 \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}\OperatorTok{;}
-\CommentTok{// Instantiation}
+\CommentTok{// Instantiation via JS instead of HTML}
 \KeywordTok{var} \NormalTok{el }\OperatorTok{=} \KeywordTok{new} \AttributeTok{HelloWorld}\NormalTok{()}\OperatorTok{;}
@@
 
-Native DOM properties, like \texttt{id} or \texttt{onclick}, reflect
-their values between HTML and JS.{[}14, Para. 2.6.1{]} For example
-declaring the HTML
-\texttt{\textless{}hello-world\ id="hello"\textgreater{}} like this
-equals to assign the ID in JS like \texttt{Node.id\ =\ "hello"}. This
-behavior won't work out-of-the-box with methods or properties defined
-setters. For example declaring
-\texttt{\textless{}hello-world\ sayhello="mars"\textgreater{}\textless{}/hello-world\textgreater{}}
-would't call the \texttt{sayhello} function in the previous setup.
-
-A common workaround to bind HTML and JS behavior together is archived by
-using the aforementioned lifecycle method
-\texttt{attributeChangedCallback} to \textbf{bind changing HTML
-attributes to JS properties} and to map JS attributes to HTML with
-\texttt{this.setAttributes(...)} respectively. On insertion time HTML
-attributes might trigger custom JS methods retrieving HTML attributes
-using \texttt{this.getAttributes(...)} method.
-
-Concluding this section, a reader might already discover the
-\textbf{mental model} behind \emph{web compontents}. A custom element is
-similar to a named function where attributes treated as input variables.
-In the hierarchical nature of DOM, input can occur either top-down via
-assignments and bottom-up via captured events. The same goes true when
-talking about output. Even though it seems obvious, it might be helpful
-to keep this point in mind.
-
-\subsubsection{Customized build-in
+Native DOM properties always reflect their values between HTML and
+JS.{[}14, Para. 2.6.1{]} Declaring
+\texttt{\textless{}hello-world\ id="hello"\textgreater{}} equals to the
+JS declaration \texttt{new\ HelloWorld().id\ =\ "hello"}.
+
+This behavior won't work out-of-the-box with methods defined with
+setters as they are strictly JS only . Mounting
+\texttt{\textless{}hello-world\ sayhello="mars"\textgreater{}} would't
+call the \texttt{sayhello} method in the previous setup. Value
+reflection can be implemented inside the \emph{custom elements} using
+the native methods \texttt{getAttributes} and \texttt{setAttributes}.
+Using them exhaustively throughout lifecycles methods the new components
+can configured to read and listen to HTML attributes accordingly.
+
+Designing a \emph{custom element} this way creates HTML elements with
+named attribute interfaces reaching deep into JS functionality. With
+this mental model in mind a web developer can create highly dynamic web
+components.
+
+\subsubsection{3.1.3. Customized build-in
 elements}\label{customized-build-in-elements}
 
-One aspect didn't mentioned yet is the possibility of creating
-sub-classes of build-in elements by extending the native Interfaces like
-the \texttt{HTMLButtonElement} interface. While this functionality is
-perfectly spec'd it is strongly rejected by some browser
-vendors.\footnote{https://github.com/w3c/webcomponents/issues/509} Most
-likely the spec will change in future in one or other way on this issue
-and therefore customized build-in elements left out of this paper
+One aspect didn't mentioned yet is the possibility of extending other
+build-in elements by extending other interfaces instead of the
+\texttt{HTMLElement} interface. While this functionality is perfectly
+spec'd it is strongly rejected by some browser vendors.\footnote{https://github.com/w3c/webcomponents/issues/509}
+Most likely the spec will change in future in one or other way on this
+issue and therefore customized build-in elements left out of this paper
 intentionally.
 
-\subsection{\texorpdfstring{Shadow DOM
-\href{http://w3c.github.io/webcomponents/spec/shadow/}{(w3c)}}{Shadow DOM (w3c)}}\label{shadow-dom-w3c}
-
-A \emph{shadow DOM} is basically an isolated DOM tree living inside an
-another (hosting) DOM tree. The spec refers the hosting tree as
-\emph{light DOM tree} and the attached DOM as \emph{shadow DOM tree}.
-Conceptually, the \emph{shadow DOM} issues a single important topic in
-software development: \textbf{Encapsulation}. While the first spec
-\emph{custom elements} provides a sufficient way to encapsulate JS
-behavior, \emph{shadow DOM} coined strongly to in the direction of style
-encapsulation.
-
-With an ever increasing complexity of an single-page application, the
+\subsection{\texorpdfstring{3.2. Shadow DOM
+\href{http://w3c.github.io/webcomponents/spec/shadow/}{(w3c)}}{3.2. Shadow DOM (w3c)}}\label{shadow-dom-w3c}
+
+A \emph{shadow DOM} is just an isolated DOM tree living inside an
+another DOM tree. The spec refers the hosting tree as \textbf{light DOM
+tree} and the attached DOM as \textbf{shadow DOM tree}. Conceptually
+\emph{shadow DOM} issues a single important issue for building scalable
+software which is namely \textbf{encapsulation}. While custom elements
+provide a good way to encapsulate JS behavior \emph{shadow DOM} tends
+strongly to the direction of style and event encapsulation.
+
+With an ever increasing complexity of single-page applications the
 global nature of the DOM creates a daunting situation for code
 organization and leads over times to highly fragmented bits of CSS and
-obscure CSS selectors or html wrappers. Of course, this situation lowers
-code clarity and reusability dramatically. The only solution which won't
-break with the existing global paradigm effectively is to allow separate
+obscured CSS selectors. Of course this situation lowers code clarity and
+reusability dramatically. The only solution which won't break with the
+existing global paradigm of the DOM effectively is to allow separate
 pieces of encapsulated code sit on top of the global DOM - introducing
-the shadowed DOM approach!
+the shadowed DOM approach.
 
@@
  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
-  \NormalTok{...}
    \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
@@
 The new global method \texttt{attachShadow} adds a new document root to
-the \texttt{HelloWorld} which has the same properties as a normal DOM.
-Therefore, invoking \texttt{innerHTML} method would fill the new
-document (fragment) with some arbitrary content. Note that
-\texttt{shadowRoot} is marked as \textbf{open} which ensures that some
-events can bubble out and outside JS can reach in the new root. Nested
-children nodes and other content in the light DOM are ``shadowed'' by
-the new root and must be invited in by so called \texttt{slots}.
-
-\subsubsection{Slots}\label{slots}
+the \texttt{HelloWorld} which has the same properties as a normal, light
+document object. Note that \texttt{shadowRoot} is marked as
+\textbf{open} which ensures that some events can bubble out and outside
+JS can reach in the new root.
+
+Filling the \emph{shadow DOM} with an \texttt{innerHTML} string is
+rather impractical. To fill a \emph{shadow DOM} with life usually it
+invites light DOM child nodes nested under the hosting node using a
+technique called \texttt{slots}.
+
+\subsubsection{3.2.1. Slots}\label{slots}
 
 Contradicting to the simplified \texttt{HelloWorld} example, a
-\emph{shadow DOM} shouldn't contain any \sout{valuable} content. While
-technical possible any change of an element would require deeply nested
-calls from the \emph{light DOM} to the \emph{shadow DOM} to update the
-element in place. That's why \emph{shadow DOM} should be perceived more
-as \textbf{static HTML template} and provide therefore a kind of
-internal frame for the render engine. \texttt{Slots} are placeholders
-for \emph{light DOM} nodes used to mark the endpoints in question.
+\emph{shadow DOM} shouldn't contain dynamic content. Changing or
+interacting with the paragraph node from the example would require
+nested JS calls querying the hosting node, entering the \emph{shadow
+DOM} and applying a function. Imported JS behavior from third-party
+libraries in the \emph{light DOM} can't be used inside the \emph{shadow
+DOM}, too.
+
+That's why the shadowed documents should be more perceived as
+\textbf{static documents} filled and managed solely by the render
+engine. \texttt{Slots} are target areas for \emph{light DOM} nodes used
+to mark the endpoints in question.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> index.html}
+\KeywordTok{<HelloWorld-with-ShadowDOM>}
+  \CommentTok{<!--}
+\CommentTok{    All child nodes will be moved inside the}
+\CommentTok{    shadowRoot if shadowRoot.innerHTML = '<slot></slot>'}
+\CommentTok{    -->}
+  \KeywordTok{<p>}\NormalTok{hello I will be scoped}\KeywordTok{</p>}
+\KeywordTok{</HelloWorld-with-ShadowDOM>}
+\end{Highlighting}
+\end{Shaded}
 
@@
 \emph{shadow DOM}. Their just rendered in place. It's an subtle but
-important difference towards handling a node. JS behaviour and CSS
-styles applied in the \emph{light DOM} will still be valid in the
-\emph{shadow DOM}. The render engine literally taking the nodes and
-putting them inside the \texttt{slot}. This procedure is commonly
-referred as \textbf{flattening} of the DOM trees.
-
-\paragraph{Named slots}\label{named-slots}
-
-A named slot is the preferable way for clear code organization. Taking
-for example
-\texttt{\textless{}slot\ name="hello"\textgreater{}Drop\ me\ a\ "hello"\ node\textless{}/slot\textgreater{}}
-targets all direct \emph{light DOM} child nodes of the hosting node
-matching the slot name like
-\texttt{\textless{}div\ slot="hello"\textgreater{}\textless{}/div\textgreater{}}.
+important difference towards handling a node. All JS behavior and CSS
+styles applied in the \emph{light DOM} will be valid in the \emph{shadow
+DOM}. The render engine literally taking the nodes and drop them inside
+the \texttt{slot} tag. This procedure is commonly referred as
+\textbf{flattening} of the DOM trees.
+
+It's possible to add semantics to the \emph{shadow DOM} in naming the
+slots which frees the \emph{light DOM} from the responsiblity to deliver
+nodes in a correct top-to-bottom order. Combining \emph{shadow DOM} with
+HTML templates provides the web developer with a flexible \textbf{HTML
+template engine}.
+
+\begin{longtable}[]{@{}ll@{}}
+\toprule
+FROM: light DOM & TO: shadow DOM\tabularnewline
+\midrule
+\endhead
+\texttt{\textless{}p\ slot="hello"\textgreater{}Named\textless{}/p\textgreater{}}
+&
+\texttt{\textless{}slot\ name="hello"\textgreater{}hello\ only\textless{}/slot\textgreater{}}\tabularnewline
+\texttt{\textless{}p\textgreater{}Unnamed\textless{}/p\textgreater{}} &
+\texttt{\textless{}slot\textgreater{}Unnamed\ nodes\textless{}/slot\textgreater{}}\tabularnewline
+\bottomrule
+\end{longtable}
+
 Writing a little documentation inside the
 \texttt{\textless{}slot\textgreater{}} tag is considered as a good
-practice as it will be rendered only if no matching \emph{light DOM}
-node is available. This functionality makes a \emph{custom element}
-pretty much self-explanatory.
-
-\paragraph{Unnamed slots}\label{unnamed-slots}
-
-Inside a so-called \emph{default slot} which looks like
-\texttt{\textless{}slot\textgreater{}Unnamed\ content\ goes\ here\textless{}/slot\textgreater{}},
-the render engine expands all direct \emph{light DOM} children without a
-\texttt{slot} attribute. In case of multiple default slots, the first
-slot takes it all.
-
-\subsubsection{Styling}\label{styling}
+practice as it provides the developer with visual clues what nodes must
+be delivered. This functionality makes a \emph{shadow DOM} pretty much
+self-explanatory. Inside a default slot tag the render engine expands
+all \emph{light DOM} children without a named \texttt{slot} attribution.
+
+\subsubsection{3.2.2. Styling}\label{styling}
 
@@
 \emph{shadow DOM} are not affected by any styling from outside. Nodes
-which are declared outside and distributed via \texttt{slots} will be
-styled in the \emph{light DOM} and can be additionally painted in the
-\emph{shadow DOM} through the new CSS-Selector \texttt{::slotted()}.
+which are distributed will be styled in the \emph{light DOM} and can be
+additionally painted in the \emph{shadow DOM}.
 
@@
 minimize the global stylings to some base styling for uniformity of the
-web site while leaving the specific stylings to the component. Due to
-the cascading nature of CSS, styles will still ``bleed in'' from
-ancestors to the \emph{light DOM} nodes.
-
-Regarding the importance style encapsulation, a couple of new CSS rules
-emerged that are exclusively targeting the \emph{shadow DOM}. The table
-below outlines styling possibilities for the use INSIDE the \emph{shadow
-DOM}:
+web site while leaving the specific stylings to the component. It is
+possible to \textbf{reset all styles} inside the \emph{light DOM} before
+distributing nodes using the \textbf{\texttt{all:\ initial}} reset. To
+ensure a consistent look between different shadow roots this technique
+should be used carefully.
+
+Regarding the importance style encapsulation, a couple of \textbf{new
+CSS rules} emerged that are exclusively targeting the \emph{shadow DOM}.
+The table below outlines styling possibilities for the use INSIDE the
+\emph{shadow DOM}:
 
@@
   Applies to distributed nodes and repaints them after distribution.
-  \texttt{Slotted} won't override outsides styles but can complement
-  them with unset style rules.
+  Slotted won't override outsides styles but can complement them
+  previously unset style rules.
 \item
@@
   The host property will add styles or change inherited ones inside
-  shadow DOM. Using \texttt{all:\ initial;} will ensure browser defaults
-  only.
+  shadow DOM. Aforementioned style resets can be placed here.
 \item
   :host(condition)\\
-  Like the previous one this node will style the shadow DOM but this
-  time based on attributes/conditions assigned to the hosting node.
+  Like the previous rule this selector will style the shadow DOM but
+  this time based on attributes/conditions assigned to the hosting node.
 \item
   :host-context(condition)\\
-  Like the previous one this node will style the shadow DOM but will
-  look after context set at the host node or even at the host ancestor.
+  Like the previous rule this selector will style the shadow DOM but
+  will look after context set at the host node or even at the host
+  ancestor.
 \end{itemize}
 
-Using the \emph{functional selector} of \texttt{:host()} or even the
-only-functional \texttt{:host-contest()} allows the creation of
-\textbf{context-aware custom elements}. A possible use case would be
-``theming'' a component (example taken from {[}15{]}):
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{> index.html}
-\KeywordTok{<body}\OtherTok{ class=}\StringTok{"darktheme"}\KeywordTok{>}
-  \KeywordTok{<fancy-tabs>}
-    \NormalTok{...}
-  \KeywordTok{</fancy-tabs>}
-\KeywordTok{</body>}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{> fancy-tabs shadowRoot}
-\NormalTok{<style>}
-\DecValTok{:}\NormalTok{host-context(}\FloatTok{.darktheme}\NormalTok{) }\KeywordTok{\{}
-\ErrorTok{ } \KeywordTok{color:} \DataTypeTok{white}\KeywordTok{;}
-\ErrorTok{ } \KeywordTok{background:} \DataTypeTok{black}\KeywordTok{;}
-\KeywordTok{\}}
-\NormalTok{</style>}
-\end{Highlighting}
-\end{Shaded}
-
-\subsubsection{JS Behavior}\label{js-behavior}
+Using the \textbf{functional selectors} \texttt{host()} or
+\texttt{host-context()} allows the creation of context-aware custom
+elements. A possible usecase would be ``theming'' a component.
+
+\subsubsection{3.2.3. Behavior}\label{behavior}
 
@@
 (the \emph{light DOM}) and not the part of the \emph{shadow DOM}. On the
-other hand there are numerous scenarios where JS is just concerned with
-\textbf{styling or animation of an element}.In this case it might be
-more straightforward to apply JS inside the \emph{shadow DOM} to avoid
-mixing with business logic handlers.
-
-Drilling down to a \emph{light DOM} node from an \emph{shadow DOM}
-context is not possible with querying the node directly with
-\texttt{.querySelector()} or \texttt{.getElementById()} as the node is
-not part of the context. To get a distributed node in question it needs
-the way over the slot node and call \texttt{slot.assinedNodes()} to
-receive an array of distributed node(s) which can be accessed and
-manipulated like any other node. Calling \texttt{.assignedNodes()} on an
-empty \texttt{slot} returns an empty array.
-
-Wrapping up this section, \emph{shadow DOM} provides a non-hacky way to
-create uniform looking \emph{custom elements} and even enhance styling
-possibilities without adding much overhead. Still, for smaller
-components with only one or two child nodes, just a little styling
-and/or no structured redistribution a \emph{shadow DOM} might be to hard
-to reason about. Eventually it all depends on the question of ``how hard
-is it to implement it without shadow DOM'' - which can't be answered
-universally. For a more in-depth guide, Google Engineer Eric Bidelman
-wroten a great primer on \emph{shadow DOM}{[}15{]}.
-
-So far, there is still a missing link between \emph{light DOM} and
-\emph{shadow DOM}. The observant reader may have already noticed the
-weak point in the \texttt{HelloWorld} example: how to ``vitalize'' the
-\emph{shadow DOM}. Recapturing the last \texttt{HelloWorld} example a
-string of markup was assigned to the \texttt{shadowRoot.innerHTML}
-property. While it works perfectly fine in this simple case, a string of
+other hand there are numerous scenarios where JS is used for styling or
+animation of an element. In this case it might be more straightforward
+to \textbf{apply JS inside the \emph{shadow DOM}} to avoid mixing with
+logic handlers with styling.
+
+It seems that a \emph{light DOM} node is in the \emph{shadow DOM}
+context after distribution. Visually this may be true but logically the
+node stays in the normal document. To reach a distributed node from the
+\emph{shadow DOM} to apply some JS behavior for styling it needs the
+extra way over the slot node. Calling \texttt{assinedNodes()} on the
+hosting slot element returns a linkage to the distributed node which can
+be accessed and manipulated like in the \emph{light DOM} context.
+
+Wrapping up this section \emph{shadow DOM} provides a non-hacky way to
+create uniform looking custom elements and even enhance styling
+possibilities without adding overhead. For small components with just a
+little styling a \emph{shadow DOM} might be over engineered. Eventually
+it all depends on the question of ``how hard is it to implement it
+without shadow DOM'' - which can't be answered universally. For a more
+in-depth guide, Google Engineer Eric Bidelman wroten a great primer on
+\emph{shadow DOM}{[}15{]}.
+
+There is still a missing link between \emph{light DOM} and \emph{shadow
+DOM}. The observant reader may have already noticed the weak point in
+the HelloWorld example: how to ``vitalize'' the \emph{shadow DOM} with
+slot or other element structuring instead of \texttt{innerHTML} strings.
+While strings works perfectly fine in this simple case a string of
 markup is rather cumbersome and error-prone and doesn't scale well. When
@@
 life hard for developers to work with it because it requires manual
-indentation and is out of syntax highlighting. That's the time templates
-come into play.
-
-\subsection{\texorpdfstring{HTML Templates
-\href{https://www.w3.org/TR/html5/scripting-1.html\#the-template-element}{(w3c)}}{HTML Templates (w3c)}}\label{html-templates-w3c}
+indentation of code blocks and is typically out of syntax highlighting.
+That's the time HTML templates come into play.
+
+\subsection{\texorpdfstring{3.3. HTML Templates
+\href{https://www.w3.org/TR/html5/scripting-1.html\#the-template-element}{(w3c)}}{3.3. HTML Templates (w3c)}}\label{html-templates-w3c}
 
@@
 \KeywordTok{class} \NormalTok{CheckoutContainerBase }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
-  \KeywordTok{static} \AttributeTok{getActions}\NormalTok{() }\OperatorTok{\{}
-    \ControlFlowTok{return} \NormalTok{[...]}
-  \OperatorTok{\}}
+  \CommentTok{// static methods for testing...}
   \NormalTok{set }\AttributeTok{_dispatch}\NormalTok{(msg) }\OperatorTok{\{}
@@
 
+declarative event management
+
 \hypertarget{refs}{}

### 2017-02-27 build/paper.tex
pdf/tex build

@@
 \IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
-\usepackage[normalem]{ulem}
-% avoid problems with \sout in headers with hyperref:
-\pdfstringdefDisableCommands{\renewcommand{\sout}{}}
 \IfFileExists{parskip.sty}{%
@@
 
-Computers can scale, humans can't. Ever since a program or complex
-system made by humans has been constrained by humans mental
-capabilities. Like in the analogy of juggling balls, our brain can just
-``juggle'' a few things at a time. Rich Hickey, the inventor of the
-programming language Clojure gave an inspirational keynote on the topic
-of \textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
+Computers can scale, humans can't. Ever since a complex system made by
+humans has been constrained by humans mental capabilities. Like in the
+analogy of juggling balls our brain can just ``juggle'' a few things at
+a time. Rich Hickey, the inventor of the programming language Clojure
+gave an inspirational keynote on the topic of
+\textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
   Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} In every sphere
-of a humans life, simplicity aligns perception with our mental
+of a humans life simplicity aligns perception with our mental
 capacities.
@@
 as ``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
-  Dictionary}} which points directly to the unidimensional aspect. While
+  Dictionary}} which points towards an unidimensional nature. While
 complexity describes the multilayered und entangled nature of
@@
 in a straightforward manner. It certainly has some overlapping's with
-easy, but while easy is more of a relative nature, simple can be laid
+easy, but while easy is more of a relative spirit, simple can be laid
 out as a objective manner and therefore universally applicable.
@@
 Software development is undoubtedly rich in complexity and full of
-subtle pitfalls. In a typical scenario, a piece of software evolves over
-time in one or another opinionated direction. Layers of new abstractions
-wrestling with old legacy abstractions and mutation becomes untraceable.
-Subtle bugs start to creep in. Eventually the small piece of software
-may end up in a highly complected monolith which will determine future
-design decisions to a painful degree. Future strategies of the
-company/organization will be highly determined by the current state in
-the need of ``keeping the lights on''.
-
-On the other side of this dystopian scenario, the truly modular system
+subtle pitfalls. In a typical scenario a growing software project
+evolves in one or another opinionated direction over time. Layers of
+abstractions wrestling with aged legacy code requiring layers of
+middleware. Mutating assets create subtle bugs and so forth. Eventually
+the small piece of software may end up in a highly complected monolith
+which will determine future design decisions to a painful degree.
+Opaqueness of the system will slow down innovation to a minimum in the
+need of ``keeping the lights on''.
+
+On the other side of this dystopian scenario a truly modular system
 architecture abandons many of those potential inconsistencies. The whole
@@
 Simplicity in ``the web'', read as a loose generalization of
-``everything that runs in the browser'', is certainly a story full of
+``everything that runs in the browser'' is certainly a story full of
 misconceptions. While simplicity in the backend is mostly a matter of
@@
 favor of an additional layer of abstracted JS components (containing
-content, markup and styling). Quiet often those frameworks mimic a MVC
-pattern on top of the browser engine which is a reasonable simple design
-pattern to build graphical user interfaces. While frameworks are a valid
-approach for building scalable web applications they remain highly
-opinionated, embody inherent complexity themselves and can change and
-break over time. Another downside is code inflation which is a crucial
-point for performance. All of those bottlenecks in the web demand for a
-new standardized way for creating and evolving complex web services.
+structure, behavior and sometimes styling). Quiet often those frameworks
+mimic a MVC pattern on top of the browser engine which is a reasonable
+simple design pattern to build graphical user interfaces. While
+frameworks are a valid approach for building scalable web applications
+they remain highly opinionated, embody inherent complexity themselves
+and can change and break over time. Another downside is code inflation
+which is a crucial point for performance. All of those bottlenecks in
+the web demand for a new standardized way for creating and evolving
+complex web services.
 
@@
 function interface towards ES6 modules (which are mandatory for the new
-specifications). With the rise of WebAssembly, a new low-level
-programming language for the browser, the determination on JS will
-hypothetical deteriorate and new quasi native languages for the web
-might gain traction.
+specifications).
 
@@
 \subsection{\texorpdfstring{3.1. Custom Elements
-\href{http://w3c.github.io/webcomponents/spec/custom/}{(w3c)}}{3.1. Custom Elements (w3c)}}\label{custom-elements-w3c}
+\href{https://html.spec.whatwg.org/multipage/scripting.html\#custom-elements}{(whatwg)}}{3.1. Custom Elements (whatwg)}}\label{custom-elements-whatwg}
 
@@
 \begin{Highlighting}[]
-\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
+\OperatorTok{>} \VariableTok{HelloWorld}\NormalTok{.}\AttributeTok{js}
 \KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
@@
 using the \emph{custom elements'} \textbf{getter / setter interface} to
-steer the behavior of the element.
+steer the behavior of the element. Note that the keywords \textbf{get}
+and \textbf{set} as well as the previously used \textbf{constructor} are
+optional!
 
@@
     \KeywordTok{this}\NormalTok{.}\AttributeTok{_hello} \OperatorTok{=} \NormalTok{val}\OperatorTok{;}
-    \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\KeywordTok{this}\NormalTok{.}\AttributeTok{_hello}\NormalTok{)}\OperatorTok{;}
   \OperatorTok{\}}
@@
 \KeywordTok{var} \NormalTok{el }\OperatorTok{=} \KeywordTok{new} \AttributeTok{HelloWorld}\NormalTok{()}\OperatorTok{;}
-\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello} \OperatorTok{=} \StringTok{"earth"}\OperatorTok{;}
-\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello}\OperatorTok{;}\CommentTok{//"earth"}
+\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello} \OperatorTok{=} \StringTok{"earth"}\OperatorTok{;} \CommentTok{// "Call" the setter}
+\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello}\OperatorTok{;}\CommentTok{// Yields "earth"}
 \end{Highlighting}
@@
 the \texttt{HelloWorld} which has the same properties as a normal, light
-document object. Note that \texttt{shadowRoot} is marked as
+document object. Note that the \texttt{shadowRoot} object is marked as
 \textbf{open} which ensures that some events can bubble out and outside
@@
 markup is rather cumbersome and error-prone and doesn't scale well. When
-putting quotes inside another quotes things break quickly. It makes the
-life hard for developers to work with it because it requires manual
-indentation of code blocks and is typically out of syntax highlighting.
-That's the time HTML templates come into play.
+putting quotes inside other quotes things break quickly. Strings make
+development harder because code editor features like indentation or
+syntax highlighting won't be supported. The HTML templates is set to
+fill this gap.
 
 \subsection{\texorpdfstring{3.3. HTML Templates
-\href{https://www.w3.org/TR/html5/scripting-1.html\#the-template-element}{(w3c)}}{3.3. HTML Templates (w3c)}}\label{html-templates-w3c}
-
-Among all other new standards \emph{HTML templates} are the most mature
-and adopted standard in the browser environment. All major browsers,
-except from Internet Explorer, support this standard.
-
-One core concept in templates is efficiency: Whatever dropped inside a
-\texttt{template} tag \sout{bucket} will be parsed on runtime - but not
-constructed into the \emph{content tree}. It remains plain HTML Markup
-sitting somewhere in the document until the time of activation.
-
-Activation typically takes four steps:
+\href{https://html.spec.whatwg.org/multipage/scripting.html\#the-template-element}{(whatwg)}}{3.3. HTML Templates (whatwg)}}\label{html-templates-whatwg}
+
+Among all other new specifications \emph{HTML templates} are the most
+mature and adopted standard in the browser environment. All major
+browsers support it since years.
+
+One core concept in templates is browser performance. Elements inside a
+\texttt{template} tag will be parsed on runtime - but not constructed
+and rendered into the content tree. They're remaining plain HTML markup
+sitting in the document until the time of activation.
+
+Activation usually takes four steps:
 
@@
   const content = node.content;\\
-  -\textgreater{} Returns a \emph{DocumentFragment} object. Handling is
-  straight forward content.querySelector(`img').src = `logo.png';
-\item
-  \textbf{Optional: Cloning the \emph{DocumentFragment} for multiple
-  use}\\
+  Returns a DocumentFragment object.
+\item
+  \textbf{Optional: Cloning the fragment for multiple use}\\
   const clone = content.cloneNode(``deep'');
@@
 
-As easy and minimal \emph{HTML templates} are, they're missing out a
-crucial feature other template implementations usually have. As
-templates are basically just dump containers for HTML Markup, there is
-no way to define some logic as \textbf{placeholders} where content
-should appear. Of course, with heavy use of JS things could be modeled
-this way. The idiomatic way tends more towards a \emph{Shadow DOM \&
-HTML templates} symbiosis.
+As easy and minimal \emph{HTML templates} are they're missing out a
+crucial feature other template implementations used to have. As
+templates are basically just containers for HTML markup there is no
+idiomatic way to define \textbf{placeholders} for dynamic content.
+Templates could be mock up this way using JS for altering the content
+but the much cleaner way leverages the previously described slot
+technique from shadow DOM.
 
@@
 \begin{Highlighting}[]
-\NormalTok{> index.html}
+\NormalTok{> hello-world-component.html}
 \KeywordTok{<hello-world>}
- \KeywordTok{<p}\OtherTok{ id=}\StringTok{"sendto"}\OtherTok{ slot=}\StringTok{"placeholder"}\KeywordTok{>}
-  \NormalTok{Hello World Web Component  }
- \KeywordTok{</p>}
-\KeywordTok{</hello-world>}
-
-\CommentTok{<!-- COMPONENT STARTS HERE -->}
   \KeywordTok{<template}\OtherTok{ id=}\StringTok{"hello"}\KeywordTok{>}
- \CommentTok{<!-- STYLES -->}
+  \CommentTok{<!-- styling -->}
     \KeywordTok{<style>}
-  \FloatTok{#stylewrapper} \KeywordTok{\{}
+      \FloatTok{#hellowrap} \KeywordTok{\{}
         \KeywordTok{font-weight:} \DataTypeTok{bold}\KeywordTok{;}
@@
     \KeywordTok{</style>}
- \CommentTok{<!-- CONTENT -->}
- \KeywordTok{<div}\OtherTok{ id=}\StringTok{"stylewrapper"}\KeywordTok{>}
+
+    \CommentTok{<!-- structure -->}
+    \KeywordTok{<div}\OtherTok{ id=}\StringTok{"hellowrap"}\KeywordTok{>}
       \KeywordTok{<slot}\OtherTok{ name=}\StringTok{"placeholder"}\KeywordTok{>}
@@
 
+  \CommentTok{<!-- light DOM / typical inserted at index.html -->}
+  \KeywordTok{<p}\OtherTok{ slot=}\StringTok{"placeholder"}\KeywordTok{>}
+    \NormalTok{Hello World Web Component}
+  \KeywordTok{</p>}
+\KeywordTok{</hello-world>}
+
 \KeywordTok{<script>}
- \CommentTok{// Switched to anonymous class notation}
- \CommentTok{// for keeping associated code together.}
+  \KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+    \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
+      \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
+      \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+      \KeywordTok{const} \NormalTok{temp }\OperatorTok{=} \KeywordTok{this}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'template#hello'}\NormalTok{)}\OperatorTok{;}
+      \KeywordTok{this}\NormalTok{.}\VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{temp}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
+    \OperatorTok{\}}
+  \OperatorTok{\}}
+  \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}\OperatorTok{;}
+\OperatorTok{<}\SpecialStringTok{/script>}
+\end{Highlighting}
+\end{Shaded}
+
+The updated \texttt{HelloWorld} component looks already pretty mature.
+It combines all the previous mentioned standards into one HTML file.
+Custom elements serves the logic, shadow DOM scopes the styles and
+\emph{HTML Templates} efficiently glues light DOM and shadow DOM
+together. This separation of concerns comes with a surplus in
+flexibility. In a real world scenario \texttt{HelloWorld} could
+reference multiple \emph{HTML Templates} and switch them around without
+any fuss. Even further a developer might split up templates into named
+\textbf{STYLE} templates and \textbf{CONTENT} templates to increase
+reusability even further.
+
+The last standard in the row of four is not concerned with the internals
+implementation of a web component. HTML Imports serves the need for an
+efficient distribution mechanism of components and other HTML resources.
+
+\subsection{\texorpdfstring{3.4. HTML Imports
+\href{https://www.w3.org/TR/html-imports/}{(w3c)}}{3.4. HTML Imports (w3c)}}\label{html-imports-w3c}
+
+Importing the \texttt{HelloWorld} component is a one-liner:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> index.html}
+\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"Hello.html"}\OtherTok{ async}\KeywordTok{>}
+
+\KeywordTok{<hello-world>}
+  \KeywordTok{<p}\OtherTok{ slot=}\StringTok{"placeholder"}\KeywordTok{>}
+    \NormalTok{Hello World Web Component}
+  \KeywordTok{</p>}
+\KeywordTok{</hello-world>}
+\end{Highlighting}
+\end{Shaded}
+
+The \texttt{async} flag is optional but recommended like any other
+fetching event. Once the imported HTML document comes into scope
+activation follows a very similar process compared to the aforementioned
+HTML templates:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  \textbf{Querying the link node}
+\item
+  \textbf{Parsing the content and preparing the render}\\
+  const content = linknode.import; -\textgreater{} Unlike the \emph{HTML
+  template} a complete document object is constructed.
+\item
+  \textbf{Optional: Cloning some nodes for multiple use}
+\item
+  \textbf{Appending the clone/original to destination}
+\end{enumerate}
+
+Again this is the imperative way to handle a generic \emph{HTML Import}.
+In the declarative world of web components a component is
+\textbf{parsed, auto-activated and anchored} solely by its' tag name
+\texttt{\textless{}hello-world\textgreater{}\textless{}/hello-world\textgreater{}}
+on purpose. The very own lifecycle method \textbf{adoptedCallback()} in
+custom elements shows the strong interconnection between those
+standards.
+
+\emph{HTML imports} can import everything which is wrappable with HTML
+markup. Stylesheets, scripts, documents, media files and even further
+imports statements can form a semantic \emph{HTML import} statement. The
+far reaching possibilities of a single standard has its downsides.
+According to Mozilla \emph{HTML imports} are not fully compatible with
+the dependency model of the new ES6 modules.\footnote{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}
+Under the current situation Mozilla and Apple and Microsoft won't
+implement \emph{HTML Imports} soon if any. Currently only Googles blink
+web engine supports \emph{HTML Imports} as they are the driving force
+behind the web components specs in general. Despite the discrepancies
+among browser vendors \emph{HTML Imports} are part of this paper as no
+other native standard can ship bundled HTML, CSS and JS which is a core
+concept behind web components.
+
+\subsection{\texorpdfstring{3.5. Custom Events
+\href{https://dom.spec.whatwg.org/\#interface-customevent}{(whatwg)}}{3.5. Custom Events (whatwg)}}\label{custom-events-whatwg}
+
+Events are first-class citizens in the browser. After all they provide a
+neat communication channel for dynamic interactions. \emph{Custom
+Events} are part of the DOM since years but with the rise of web
+components they will most likely become an indispensable building block
+of web components.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> hello-world-component.html}
+
+\KeywordTok{<hello-world>}
+ \KeywordTok{<button>}\NormalTok{Launch CustomEvent}\KeywordTok{</button>}
+\KeywordTok{</hello-world>}
+
+\KeywordTok{<script>}
  \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} 
@@
     \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
-    \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
-    \KeywordTok{const} \NormalTok{helloTemplate }\OperatorTok{=} \VariableTok{document}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'#hello'}\NormalTok{)}\OperatorTok{;}
-    \KeywordTok{this}\NormalTok{.}\VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{helloTemplate}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
+
+   \CommentTok{// Craft a helloevent}
+    \KeywordTok{const} \NormalTok{helloevent }\OperatorTok{=} \KeywordTok{new} \AttributeTok{CustomEvent}\NormalTok{(}\StringTok{'helloevent'}\OperatorTok{,} \OperatorTok{\{}
+    \DataTypeTok{bubbles}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,}
+    \DataTypeTok{detail}\OperatorTok{:} \StringTok{'Contains scalar or object'}
+   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+
+   \CommentTok{// Launch helloevent if child button clicks}
+   \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'click'}\OperatorTok{,} \NormalTok{click }\OperatorTok{=>} \OperatorTok{\{}
+    \KeywordTok{this}\NormalTok{.}\AttributeTok{dispatchEvent}\NormalTok{(helloevent)}
+    \VariableTok{click}\NormalTok{.}\AttributeTok{stopPropagation}\NormalTok{()}\OperatorTok{;}
+   \OperatorTok{\}}\NormalTok{)}
   \OperatorTok{\}}
@@
 
-The updated \texttt{HelloWorld} component looks already pretty mature.
-It combines all the previous mentioned standards into one blob of HTML.
-\emph{Custom Elements} serves the logic, \emph{Shadow DOM} scopes the
-styles and \emph{HTML Templates} efficiently glues DOM and \emph{Shadow
-DOM} together. This separation of concerns comes with a huge gain in
-flexibility. In a real world scenario \texttt{HelloWorld} would
-contain/reference multiple \emph{HTML Templates} and could switch them
-around without any fuss. A developer might to split up templates into
-\textbf{STYLE} templates and \textbf{CONTENT} templates to increase
-reusability even further.
-
-The last standard in the row of four is not concerned with the internals
-of a \emph{web component}. \emph{HTML Imports} serves the need for an
-efficient distribution mechanism of components.
-
-\subsection{\texorpdfstring{HTML Imports
-\href{https://www.w3.org/TR/html-imports/}{(w3c)}}{HTML Imports (w3c)}}\label{html-imports-w3c}
-
-Importing the \texttt{HelloWorld} component is a one-liner:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"Hello.html"}\OtherTok{ async}\KeywordTok{>}
-\end{Highlighting}
-\end{Shaded}
-
-The \texttt{async} flag is optional but like in any other fetching
-event, strongly recommended. Once the imported HTML document comes into
-scope, activation follows a very similar pattern like the aforementioned
-\emph{HTML templates}:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  \textbf{Querying the link node}
-\item
-  \textbf{Parsing the content and preparing the render}\\
-  const content = linknode.import; -\textgreater{} Unlike the \emph{HTML
-  template} the content a fully equipped document object.
-\item
-  \textbf{Optional: Cloning some nodes for multiple use}
-\item
-  \textbf{Appending the clone/original to destination}
-\end{enumerate}
-
-This again is the imperative way to handle a generic \emph{HTML Import}.
-In the declarative world of \emph{web components} a component is
-activated, parsed and anchored solely by its' tag name
-\texttt{\textless{}hello-world\textgreater{}\textless{}/hello-world\textgreater{}}.
-Preliminary, the component needs proper configuration. The next section
-will elaborate the right configuration and composition of a component to
-work out-of-the-box.
-
-Despite from being just a practical document importer \emph{HTML
-imports} acts like a fully fledged dependency manager for the browser.
-Multiple resources, ranging from stylesheets, scripts, documents, media
-files and even other \texttt{imports} can be grouped together in a
-logical \texttt{import} statement. Internally, the browser engine keeps
-track for every imported resource so it won't be loaded twice. The
-inherent complexity is in fact a stumbling block for wider browser
-adoption. Currently only Googles blink web engine supports \emph{HTML
-Imports} as they are the driving force behind the \emph{web components}
-spec in general. Mozilla and Apple imposed distaste for \emph{HTML
-Imports} as a whole. One reason for this can be found in the
-incompatibility of the spec with the upcoming \emph{ES6 module
-loader}.\footnote{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}
-
-Despite the discrepancies among browser vendors \emph{HTML Imports}
-should still be part of the paper as no other native browser technology
-can bundle up CSS, JS and HTML that efficient.
-
-\subsection{\texorpdfstring{Appendix A: Custom Events
-\href{https://dom.spec.whatwg.org/\#interface-customevent}{(whatwg)}}{Appendix A: Custom Events (whatwg)}}\label{appendix-a-custom-events-whatwg}
-
-Events are first-class citizens in the browser world and \emph{Custom
-Events} are no exception. The \emph{Custom Elements} interface is part
-of the DOM since years but with the rise of \emph{Custom Elements} they
-will most likely become an indispensable building block of \emph{web
-components}.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{> index.html}
-\KeywordTok{<hello-world>}
- \KeywordTok{<button>}\NormalTok{Launch CustomEvent}\KeywordTok{</button>}
-\KeywordTok{</hello-world>}
-\CommentTok{<!-- COMPONENT STARTS HERE -->}
-\KeywordTok{<script>}
- \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} 
- \KeywordTok{class} \KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
-  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
-    \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
-   \CommentTok{// Craft a CustomEvent e}
-    \KeywordTok{const} \NormalTok{e }\OperatorTok{=} \KeywordTok{new} \AttributeTok{CustomEvent}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \OperatorTok{\{}
-    \DataTypeTok{bubbles}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,} \CommentTok{//important!}
-    \DataTypeTok{detail}\OperatorTok{:} \StringTok{'Contains scalar or object'}
-   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
-   \CommentTok{// Launch e on child button click}
-   \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'click'}\OperatorTok{,} \NormalTok{click }\OperatorTok{=>} \OperatorTok{\{}
-    \KeywordTok{this}\NormalTok{.}\AttributeTok{dispatchEvent}\NormalTok{(e)}
-    \VariableTok{click}\NormalTok{.}\AttributeTok{stopPropagation}\NormalTok{()}\OperatorTok{;}
-   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
-   \CommentTok{// Catch e. Typically done by some parent}
-     \CommentTok{// node. Message in detail property}
-   \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>} \OperatorTok{\{}
-    \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\VariableTok{e}\NormalTok{.}\AttributeTok{detail}\NormalTok{)}
-   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
-  \OperatorTok{\}}
- \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
-\OperatorTok{<}\SpecialStringTok{/script>}
-\end{Highlighting}
-\end{Shaded}
-
 Naming events after the emitting tag makes the API almost
 self-explanatory. The \texttt{detail} property can be loaded with
-primitives as well as objects. The further sections will elaborate a
-feasible architecture for building a scaling microservice architecture.
-
-Chaining and aggregating events from child nodes should be practiced and
-exercised quiet frequently. As mentioned earlier in the \emph{Custom
-Elements} section, the pattern of \textbf{extending native elements}
-should be somewhat dismissed as it may never implemented outside of
-Chrome. Nevertheless, it is possible to create own kind of quasi native
-buttons when chaining a \emph{CustomEvent} directly after the native
-click event.
-
-Another typical \emph{custom element} use case can be as an actor on
-(native) child elements. In this case, the \emph{Custom Element} catches
-events from children, buffers or rebuild them and eventually fires an
-event towards the document root.
-
-Unfortunately events only work ``upstream'' towards parent nodes. Still
-the web platform offers plenty of possibilities to talk back to child
-nodes.
-
-\subsection{\texorpdfstring{Appendix B: Web Worker
-\href{https://html.spec.whatwg.org/multipage/workers.html}{(whatwg)}}{Appendix B: Web Worker (whatwg)}}\label{appendix-b-web-worker-whatwg}
+scalars as well as objects. Subsequent parent nodes may catch the custom
+event with a clear understanding about the source node.
+
+Chaining and aggregating events from child nodes can be frequently used
+within web components. One use case could be creating a custom button
+like in the previous example. As mentioned earlier in the Custom
+Elements section at 3.1.3. , the pattern of \textbf{extending native
+elements} should be somewhat dismissed as other browser vendors may
+never implement it. A common workaround to eventually create a own
+version a button could be made with a thin wrapper around the native
+button and chaining a \emph{Custom Event} after the click event.
+
+Another common web components use case can be a \textbf{middleware}
+subscribing to certain child events and acting upon them. This involves
+catching events, buffering, destructuring and creation of an own event
+towards the document root. By design, events only bubble upstream
+towards parent nodes. For handing down information towards the child
+nodes we need to query the child node directly.
+
+\subsection{\texorpdfstring{3.6. Web Worker
+\href{https://html.spec.whatwg.org/multipage/workers.html}{(whatwg)}}{3.6. Web Worker (whatwg)}}\label{web-worker-whatwg}
 
@@
 at around 2009 when discussions about browser performance was still in
-the early days. Nevertheless, the addressed problem of \emph{Web
-Workers} is a fundamental language problem of JS itself.
+the early days but addressing fundamental performance bottleneck of the
+JS language.
 
@@
 browser environment, from handling UI events to query and process larget
-amounts of API data and manipulating the DOM, runs on the same
-thread{[}16{]}. Putting a lot of work to the single main thread can slow
-down the web service significantly. From time to time scripts can block
-or fail for whatever reason which leads to a frozen or crashed UI. A
-worker can overcome the bottleneck of the single-threaded nature with
-spawning new \textbf{background threads} which allows the UI to stay
-responsive even when computation-heavy tasks should be carried out.
-Furthermore, a worker thread adds a performance advantage embracing the
-multi core CPU architecture most devices running on today. To grasp the
-full potential of workers, a reader might dive deeper into the Angular 2
-architecture, where most of the application layer is abstracted from the
-main rendering thread into worker threads.\footnote{\href{https://docs.google.com/document/d/1M9FmT05Q6qpsjgvH1XvCm840yn2eWEg0PMskSQz7k4E}{Angular
+amounts of data or manipulating the DOM runs on a single thread{[}16{]}.
+Putting a lot of work to the single main thread can slow down the web
+service significantly. From time to time scripts can block or fail for
+whatever reason which leads to a frozen or crashed UI. A worker can
+overcome the bottleneck of the single-threaded nature with spawning new
+\textbf{background threads} which allows the UI to stay responsive even
+when computation-heavy tasks needs to perform. Furthermore, a worker
+thread adds a performance advantage embracing the multi core CPU
+architecture most devices running on today. To grasp the full potential
+of workers a reader might dive deeper into the Angular 2 architecture,
+where most of the application layer is abstracted from the main
+rendering thread into worker threads.\footnote{\href{https://docs.google.com/document/d/1M9FmT05Q6qpsjgvH1XvCm840yn2eWEg0PMskSQz7k4E}{Angular
   2 Rendering Architecture}}
@@
 A \emph{Web Worker} spawns a new background thread where scripts can run
-concurrent to the main thread. Usually a worker is loaded from a workers
-dedicated file to embrace this kind of separation.
+concurrent to the main thread. Usually a worker is loaded from dedicated
+file to embrace separation.
 
@@
 
-\section{Building an browsernative
-microservice}\label{building-an-browsernative-microservice}
+\section{4. Building a browsernative
+microservice}\label{building-a-browsernative-microservice}
 
@@
 background the paper should briefly join them to form a
-\emph{browsernative microservices}. Needless to say the described
-example is overall simplified to only illustrate the connection between
-browsernative technologies and microservice patterns. Googles library
-Polymer is a good place for learning about web components in depth and
-make use of their simple command line tools. One of their most famous
-proof of concept is the so-called
+\emph{browsernative microservices}. Needless to say the following
+example is overall simplified towards illustrating the connection
+between browsernative technologies and microservice patterns. Googles
+library Polymer is a good place for learning about web components in
+depth and make use of their toolbox. One of their most famous proof of
+concept is the so-called
 \href{https://shop.polymer-project.org/}{Polymer Shop} which is a
@@
 The current checkout can be found at
-https://shop.polymer-project.org/checkout. Currently the checkout is a
-single, 671 lines of code long Polymer component including all required
-fields for sign in, shipping, billing and summarizing the order. In the
-spirit of microservices we will split up the microservice into fine
-grained components. The shopping cart data is pulled out of a local
-storage JSON entity set up previously by another custom-element.
-
-\subsection{Checkout microservice}\label{checkout-microservice}
-
-By breaking down the service the team defined the business boundaries
-within the checkout process and ended up with following granular service
-blocks:
+https://shop.polymer-project.org/checkout. At the time of writing the
+checkout is a single, 671 lines of code long Polymer component including
+all required fields for sign in, shipping, billing and summarizing the
+order. In the spirit of microservices we will split up the microservice
+into fine grained components. The shopping cart data is pulled out of a
+local storage JSON entity set up previously by another custom-element.
+
+By breaking down the service the \textbf{team defined the business
+boundaries} within the checkout process and ended up with following
+granular service blocks:
 
@@
 Yet already we see the simplicity arouse from web components as they
-persue a clear structure. Each of the child components should act
-independently over other child nodes utilizing the loose coupling
-principle. Each child ships all the HTML, CSS and JS code needed to
-fulfil its work following the high cohesion principle. Each component
-may contain different views to accommodate different bounded contexts
-resulting from different devices. And last but not least, all of them
-communicate over an unobstrusive message bus via the service root
-component \texttt{\textless{}shop-checkout\textgreater{}}.
+persue a clear markup. Each of the child components may act
+independently over other child nodes utilizing the \textbf{loose
+coupling principle}. Each child ships all the HTML, CSS and JS code
+needed to fulfil its work following the \textbf{high cohesion
+principle}. Each component may contain different views to accommodate
+different \textbf{bounded contexts resulting from different devices}.
+And last but not least, all of them communicate over an
+\textbf{unobtrusive message bus} via the service root component
+\texttt{\textless{}shop-checkout\textgreater{}}.
 
@@
 familiar. Dan Abramov, the creator of Redux, once defined a simple
-dichotomous pattern for creating UI components. Firstly, he came up with
-the idea of \textbf{presentational components} only related with the
-concern about \emph{how things look}. This component literally doesn't
-know anything about the service in question which makes the component
-highly flexible and reusable. They are controlled solely from the
-outside, receiving data and dispatching unbiased events on user
-interaction.{[}17{]} Most probably every presentational component
-embodies more HTML/CSS markup and less JS code. It should encapsulate
-its styles from bleeding out and protect its styles being overwritten.
-Furthermore, it may contain several templates to change it's look on
-different demands.
+dichotomous pattern for creating UI components.
+
+Firstly, Abramov defined a pattern around \textbf{presentational
+components} only related with the concern about \emph{how things look}.
+This component literally doesn't know anything about the service in
+question which makes the component highly flexible and reusable. They
+are controlled solely from the outside, receiving data and dispatching
+unbiased events on user interaction.{[}17{]} Most probably every
+presentational component embodies more HTML/CSS markup and less JS code.
+It should encapsulate its styles from bleeding out and protect its
+styles being overwritten. Furthermore, it may contain several templates
+to change it's look on different demands.
 
@@
 contain more JS and less if any HTML markup. We probably don't need to
-utilize ShadowDOM as no styles are involved.
-
-Last but not least, we can expand the model for illustrational purposes
+utilize Shadow DOM as no styles are involved.
+
+Last but not least, the pattern can expanded for illustrational purposes
 to \textbf{native components} which is every build-in HTMLElement like
-the HTMLButtonElement. They are solely controllable and styleable from
-the outside and must therefore wrapped in presentational components
-and/or containers.
+the HTMLButtonElement. Native components are mostly deep nested elements
+providing the actual functionality in the browser UI. They are solely
+controllable and styleable from the outside and are therefore wrapped in
+presentational components and/or containers.
 
@@
 
-\subsubsection{Root container}\label{root-container}
+\subsection{Root container}\label{root-container}
 
@@
 functionalities. Encapsulation of CSS won't be necessary at this point
-as no styling is involved. The simplified root container might look like
-the following code snippet
+as no styling is involved. A simplified root container for the checkout
+might look like the following code snippet.
 
@@
 
-The purpose of this simplified code snippet is to outline the MVC
-threefold in the service root. On runtime the root container acts mostly
-like a \textbf{message dispatcher} implementing the microservice
-principle of smart endpoints and dumb pipes.
-
-In order to interact with the MODEL, every child node must implement the
+The purpose of this simplified code snippet is to outline the
+\textbf{MVC threefold} in the service root. The MODEL is pushed into the
+worker thread, the CONTROLLER in the custom element and the VIEW in the
+HTML structuring. On runtime the root container acts like a
+\textbf{message dispatcher} implementing the microservice principle of
+smart endpoints and dumb pipes.
+
+In order to interact with the MODEL every child node must implement the
 \textbf{dedicated \texttt{checkout}} custom event (which will be
@@
 Effects are yielded by the asynchronous operation of messages and create
-actions returned to sender by the message handlers. Effects can be
-created by external messages, like subscription to an WebSocket, too.
-Effects may be created with additional resources from the server or
-syncing with local storage like in the polymer-shop.
+actions returned to sender. Effects can be created by external messages,
+like subscription to an WebSocket, too. Effects may be created with
+additional resources from the server or syncing with local storage like
+in the polymer-shop.
 
@@
 the message. This communication model offers lots of possibilities for
-\textbf{evolutionary design} as child nodes can be loosely dropped in
-the service root and communication might \textbf{fail graceful} as
-simply nothing happens if one or another switch is not defined.
-
-\subsubsection{Container components}\label{container-components}
+\textbf{evolutionary design} as child nodes can be loosely dropped.
+Every child node can be expanded into another full microservice or split
+up into separate nodes without notice of the service root. Communication
+might \textbf{fail graceful} providing a default console log in case
+switches aren't defined yet.
+
+\subsection{Container components}\label{container-components}
 
@@
 \texttt{\textless{}sign-in\textgreater{}} or
-\texttt{\textless{}shipping-details\textgreater{}} can be extended can
+\texttt{\textless{}shipping-details\textgreater{}} can be extended might
 look like this:
@@
 \begin{Highlighting}[]
-\KeywordTok{class} \NormalTok{CheckoutContainerBase }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
-  \CommentTok{// static methods for testing...}
+\KeywordTok{class} \NormalTok{SimpleContainer }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+  \CommentTok{// constructor, static methods ...}
   \NormalTok{set }\AttributeTok{_dispatch}\NormalTok{(msg) }\OperatorTok{\{}
@@
       \DataTypeTok{detail}\OperatorTok{:} \NormalTok{msg}
-      \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+      \OperatorTok{\}}\NormalTok{)}
     \NormalTok{)}\OperatorTok{;}
@@
 
-Extending the \texttt{CheckoutContainerBase} will equipped every
-container node with the unified message interface. Calling
-\texttt{this.\_dispatch(msg)} within a within the container will trigger
-an event bubbling upstream. The service root will implement a simple
-event listener for \texttt{checkout} events. After receiving an answer
-from the MODEL a service root can \texttt{querySelector} the container
-node in question and push forward the answer over the \texttt{\_receive}
-property.
+Extending the \texttt{SimpleContainer} will equip every container node
+with the unified message interface. Calling
+\texttt{this.\_dispatch(msg)} within the container will trigger an event
+bubbling upstream. The service root will implement a simple event
+listener for \texttt{checkout} events. After receiving an answer from
+the MODEL a service root can query the container child node in question
+and push forward the answer over the \texttt{\_receive} property.
 
 There might be even more middleware containers pulled in between
-presentational leafs and service root component to do some extra work
-without caring about bubbling events altogether. Changing or enhancing
-any functionality requires only to touch the controllers in the
-containers in question and the endpoint section at the model.
-\textbf{Infrastructure automation} might be achieved by dynamically
-evaluating the mounted containers beneath the service root and modeling
-the ``backend'' model accordingly. Due to it's standardized message
-system the containers are testable within standardized tests.
-
-\subsubsection{Presentational
-components}\label{presentational-components}
+presentational leafs and the service root to fulfil some extra work
+either like filters on the event or without caring about checkout events
+altogether. Changing or enhancing any functionality requires only the
+controllers in the container in question and the endpoint section at the
+model. \textbf{Infrastructure automation} might be achieved by
+dynamically evaluating the mounted containers beneath the service root
+and modeling the ``backend'' model accordingly. Due to it's standardized
+message system the containers are testable within standardized tests.
+
+\subsection{Presentational components}\label{presentational-components}
 
@@
 
-\section{Thinking further}\label{thinking-further}
-
-cohesive HTML JS CSS
-
-virtual DOM
-
-declarative event management
+template switch
+
+css
+
+shadow dom
+
+\section{5. Thinking further}\label{thinking-further}
+
+Following a \textbf{browsernative first} principle this paper didn't
+made use of third-party libraries so far. Contrary to the example in the
+last section there are a few reasons why web components shouldn't be
+used without any proprietary technology. At first, from a technical
+perspective web components are relatively young and at the time of
+writing only the newer Chrome and Opera browsers can handle them
+correctly. Most probably it will take a long time until native web
+components can be safely used without additional legacy support.The
+second inevitable question is the notion if they should be used pure
+from a \textbf{simplicity perspective} without any proprietary
+technology now and future.
+
+Talking about web components it is impossible to avoid technologies like
+\textbf{React} altogether. In the last couple years React introduced
+some good concepts to web development like the notion of state, passing
+data as properties and declarative event management just to name a few.
+Many of those design decisions were made to overall \textbf{simplify
+browser development through standardized methodologies}. The principles
+of microservices could be applied to React development maybe even
+further as for example the communication model involves a lot of
+delegation and decomposition.
+
+Many of those features and design decisions in React created much of a
+hype around the library and technologies like the virtual DOM while web
+components never caught up in momentum at a comparable rate. According
+to Github Googles Polymer project is even six months older than React
+but has less than one third of its stars. Other notable web components
+projects like Bosonic\footnote{https://github.com/bosonic/bosonic/} or
+Mozilla's X-tag\footnote{https://github.com/x-tag/core} are more dead
+than alive, too.
+
+It seems like the web developing community in a broader sense demands
+for features and standardization in frameworks like React (and many
+others to be fair). Web components and related libraries clearly missed
+a selling point in providing standardized ways for developing scalable
+components and applications or being to cumbersome to use at all. For
+the future it remains unclear if web components will eventually gain
+wider adoption within the UI developing community or will be adopted as
+low-level technology for framework development emphasized by Sebastian
+Markbage, one of the React creators.{[}18{]}
+
+One interesting approach for building web components is the ``reactish''
+Skate\footnote{https://github.com/skatejs/skatejs/}. Following the
+functional rendering model from React it combines native methods with
+additional functionalities known from React like event management. Given
+the focus on native technologies it only weights around 4kb (minified
+and gzipped) which is ten times less than React. \textbf{Small
+encapsulated micro UI frameworks} like Skate could be certainly the near
+future of web development and will hopefully gain more attention. The
+more mature UI framework Riot already expressed their intention to
+gradually develop towards web components.\footnote{http://riotjs.com/compare/\#web-components}
 
@@
 
+\hypertarget{ref-Rauschmayer2015}{}
+{[}18{]} D. A. Rauschmayer, ``Tree-shaking with webpack 2 and babel 6.''
+2015 {[}Online{]}. Available:
+\url{http://www.2ality.com/2015/12/webpack-tree-shaking.html}
+
 \end{document}

### 2017-02-27 paper.tex
pdf/tex build

@@
 \newcommand{\NormalTok}[1]{{#1}}
-\usepackage[normalem]{ulem}
-% avoid problems with \sout in headers with hyperref:
-\pdfstringdefDisableCommands{\renewcommand{\sout}{}}
+\usepackage{longtable,booktabs}
+% Fix footnotes in tables (requires footnote package)
+\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
 \IfFileExists{parskip.sty}{%
@@
 
-Computers can scale, humans can't. Ever since a program or complex
-system made by humans has been constrained by humans mental
-capabilities. Like in the analogy of juggling balls, our brain can just
-``juggle'' a few things at a time. Rich Hickey, the inventor of the
-programming language Clojure gave an inspirational keynote on the topic
-of \textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
+Computers can scale, humans can't. Ever since a complex system made by
+humans has been constrained by humans mental capabilities. Like in the
+analogy of juggling balls our brain can just ``juggle'' a few things at
+a time. Rich Hickey, the inventor of the programming language Clojure
+gave an inspirational keynote on the topic of
+\textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
   Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} In every sphere
-of a humans life, simplicity aligns perception with our mental
+of a humans life simplicity aligns perception with our mental
 capacities.
 
-Derived from the ancient Latin word \textbf{simplex}, simple can be
-understood as ``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
-  Dictionary}} which points directly to the unidimensional aspect. While
+Derived from the Latin word \textbf{simplex}, simple can be understood
+as ``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
+  Dictionary}} which points towards an unidimensional nature. While
 complexity describes the multilayered und entangled nature of
@@
 
-\section{Microservices}\label{microservices}
+\section{2. Microservices}\label{microservices}
 
@@
 When it comes to standardized deployment guidelines previously mentioned
-Ben Issa, described the ING standard workflow. Every component deserves
-a own \textbf{git repo} containing
+Ben Issa, described the ING standard workflow. Every component comes in
+its own \textbf{git repo} containing:
 
@@
 
-\subsection{Design for failure}\label{design-for-failure}
+\subsection{2.7. Design for failure}\label{design-for-failure}
 
@@
 Combinment of the resources in the browser always demanded for
-optimization to avoid unexpected side-effects like the \emph{flash of
-unstyled content (FOUC)}. Googles Polymer propagates the a
-general-purpose pattern called \textbf{PRLP}\footnote{\href{https://www.polymer-project.org/1.0/toolbox/server}{PRLP
+optimization to avoid unexpected side-effects like \emph{flash of
+unstyled content}. Googles Polymer propagates the a general-purpose
+pattern called \textbf{PRLP}\footnote{\href{https://www.polymer-project.org/1.0/toolbox/server}{PRLP
   pattern}}:
@@
 evergreen browser but won't have a reliable internet connection.{[}9{]}
-While \emph{Progressive Enhancement} was once coined on the principle to
-build websites both for Browsers with JS and HTML only, the new
-\emph{Progressive Enhancement} tends towards an \textbf{offline first}
-principle avoiding network connectivity failures. A \emph{browsernative
+While \textbf{Progressive Enhancement} was once related to build
+websites both for browsers with and without JS support the demands
+changed tends towards an \textbf{offline first} principle avoiding
+network connectivity failures.{[}9{]} A \emph{browsernative
 microservice} therefore not only tries to cache data as much as
@@
 
-\subsection{Evolutionary Design}\label{evolutionary-design}
+\subsection{2.8. Evolutionary Design}\label{evolutionary-design}
 
@@
 
-Being flexible when developing for the web is a selling point as
+Pursuing flexibility for web development is a selling point as
 innovation cycles for browser development is fast paced and technologies
@@
 native technology pursues a strong interopt approach with existing
-systems. Andrew Rota for example came up with the pattern to use small,
+systems. Andrew Rota for example came up with a pattern using small,
 encapsulated and stateless web components as leaves in the tree of React
-components instead of native HTML elements.{[}10{]} Even React code can
-profit from the expressiveness of custom components. As web components
-are native after registration there is no difference in using a native
-\texttt{button} over a \texttt{meaningful-button}. Most likely there
-will be always some cutting edge framework promising advantages over
-native code. Whatever new framework will be on the rise within the next
-years native components can eliminate future uncertainty and allowing
-rapid reassembling towards new architectures.
-
-\section{W3C specifications}\label{w3c-specifications}
+components instead of native HTML elements.{[}10{]} Even React can
+eventually profit from the expressiveness of custom components using a
+custom \texttt{\textless{}meaningful-button\textgreater{}} over a native
+\texttt{\textless{}button\textgreater{}}. Most likely there will be
+always some cutting edge framework promising advantages over native
+code. Whatever new framework will be on the rise within the next years
+native components can eliminate future uncertainty and allowing rapid
+reassembling towards new architectures.
+
+\section{3. W3C specifications}\label{w3c-specifications}
 
@@
 
-\subsection{\texorpdfstring{Custom Elements
-\href{http://w3c.github.io/webcomponents/spec/custom/}{(w3c)}}{Custom Elements (w3c)}}\label{custom-elements-w3c}
-
-Custom elements are the fundamental building blocks for web components
-introducing the \emph{Single Responsibility Principle} to the browser.
-In essence, they provide a way to create \textbf{custom HTML tags}
+\subsection{\texorpdfstring{3.1. Custom Elements
+\href{https://html.spec.whatwg.org/multipage/scripting.html\#custom-elements}{(whatwg)}}{3.1. Custom Elements (whatwg)}}\label{custom-elements-whatwg}
+
+\emph{Custom elements} are the fundamental building blocks for web
+components introducing the \textbf{Single Responsibility Principle} to
+the browser. In short, they provide a way to create custom HTML tags
 subsuming behavior, design and functionality. An obligatory
@@
 \item
-  \texttt{connectedCallback()}\\
-  Called upon the time of \emph{connecting or upgrading the node} which
-  means the moment the node is rendered inside the DOM. Typically this
-  method is called straight after the \texttt{constructor()} on insert.
-  Typically, this method contains setup code, such as fetching resources
-  or rendering elements according to attributes.{[}13{]} For a fast
-  initial render of the page, it is highly preferable to put many
-  proceedings in favor of the constructor.
-\item
-  \texttt{disconnectedCallback()}\\
-  Called upon the time of \emph{node removal}. Cleanup code like
-  removing eventListeners or disconnecting websockets can be put here.
-\item
-  \texttt{attributeChangedCallback(attrName,\ oldVal,\ newVal)}\\
-  This method provides an \emph{Onchange handler} that runs for certain
-  attributes called with three values as defined in the signature. It is
-  meant to control an elements' transition from on \texttt{oldVal} to a
-  \texttt{newVal}. Due to performance issues, this callback is only
-  triggered for attributes registered in an \emph{observedAttributes}
-  array.
-\item
-  \texttt{adoptedCallback()}\\
-  Called when moving the node \emph{between documents}.
+  connectedCallback()\\
+  Called upon the time of \textbf{connecting or upgrading the node}
+  which means the moment the node is rendered inside the DOM. Typically
+  this method is called straight after the constructor if the node is
+  inserted directly. For a faster initial render of the page it is
+  highly preferable to put many proceedings in this method. Usually this
+  method contains setup code such as fetching resources or rendering
+  elements according to attributes.{[}13{]}
+\item
+  disconnectedCallback()\\
+  Called upon the time of \textbf{node removal}. Cleanup code like
+  removing event listeners or disconnecting web sockets can be put here.
+\item
+  attributeChangedCallback(attrName, oldVal, newVal)\\
+  This method provides an \textbf{onchange handler} for certain elements
+  attributes. This method is used to guide elements' transition from an
+  old value to a new state. Due to performance issues this callback is
+  only triggered for attributes registered in a dedicated array shipped
+  with the element.
+\item
+  adoptedCallback()\\
+  Called when moving the node \textbf{between documents}. This method
+  comes handy when using HTML imports described later.
 \end{itemize}
 
-\subsubsection{Custom attributes}\label{custom-attributes}
-
-As previously mentioned, the custom elements must \texttt{extend} the
-\texttt{HTMLElement} interface. Therefore, the new element inherits base
-properties and methods commonly used in all HTML elements like
-\texttt{id,\ class,\ addEventListner} . Additionally, it is possible to
-define custom attributes using the \emph{custom elements'}
-\textbf{getter / setter interface} to steer the behavior of the element.
+\subsubsection{3.1.2. Custom attributes}\label{custom-attributes}
+
+As mentioned earlier any custom elements must extend the
+\texttt{HTMLElement} interface ensuring base properties and methods used
+in throughout all HTML elements like \textbf{id, class, addEventListner
+\ldots{}}. Additionally, it is possible to define custom attributes
+using the \emph{custom elements'} \textbf{getter / setter interface} to
+steer the behavior of the element. Note that the keywords \textbf{get}
+and \textbf{set} as well as the previously used \textbf{constructor} are
+optional!
 
@@
     \KeywordTok{this}\NormalTok{.}\AttributeTok{_hello} \OperatorTok{=} \NormalTok{val}\OperatorTok{;}
-  \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\KeywordTok{this}\NormalTok{.}\AttributeTok{_hello}\NormalTok{)}\OperatorTok{;}
   \OperatorTok{\}}
@@
   \OperatorTok{\}}
-\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+\OperatorTok{\}}
 \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}\OperatorTok{;}
-\CommentTok{// Instantiation}
+\CommentTok{// Instantiation via JS instead of HTML}
 \KeywordTok{var} \NormalTok{el }\OperatorTok{=} \KeywordTok{new} \AttributeTok{HelloWorld}\NormalTok{()}\OperatorTok{;}
-\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello} \OperatorTok{=} \StringTok{"earth"}\OperatorTok{;}
-\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello}\OperatorTok{;}\CommentTok{//"earth"}
+\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello} \OperatorTok{=} \StringTok{"earth"}\OperatorTok{;} \CommentTok{// "Call" the setter}
+\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello}\OperatorTok{;}\CommentTok{// Yields "earth"}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
+\OperatorTok{>} \VariableTok{HelloWorld}\NormalTok{.}\AttributeTok{js}
 \KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
-   \CommentTok{// this.onclick...}
    \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
@@
 The new global method \texttt{attachShadow} adds a new document root to
-the \texttt{HelloWorld} which has the same properties as a normal DOM.
-Therefore, invoking \texttt{innerHTML} method would fill the new
-document (fragment) with some arbitrary content. Note that
-\texttt{shadowRoot} is marked as \textbf{open} which ensures that some
-events can bubble out and outside JS can reach in the new root. Nested
-children nodes and other content in the light DOM are ``shadowed'' by
-the new root and must be invited in by so called \texttt{slots}.
-
-\subsubsection{Slots}\label{slots}
+the \texttt{HelloWorld} which has the same properties as a normal, light
+document object. Note that the \texttt{shadowRoot} object is marked as
+\textbf{open} which ensures that some events can bubble out and outside
+JS can reach in the new root.
+
+Filling the \emph{shadow DOM} with an \texttt{innerHTML} string is
+rather impractical. To fill a \emph{shadow DOM} with life usually it
+invites light DOM child nodes nested under the hosting node using a
+technique called \texttt{slots}.
+
+\subsubsection{3.2.1. Slots}\label{slots}
 
 Contradicting to the simplified \texttt{HelloWorld} example, a
-\emph{shadow DOM} shouldn't contain any \sout{valuable} content. While
-technical possible any change of an element would require deeply nested
-calls from the \emph{light DOM} to the \emph{shadow DOM} to update the
-element in place. That's why \emph{shadow DOM} should be perceived more
-as \textbf{static HTML template} and provide therefore a kind of
-internal frame for the render engine. \texttt{Slots} are placeholders
-for \emph{light DOM} nodes used to mark the endpoints in question.
+\emph{shadow DOM} shouldn't contain dynamic content. Changing or
+interacting with the paragraph node from the example would require
+nested JS calls querying the hosting node, entering the \emph{shadow
+DOM} and applying a function. Imported JS behavior from third-party
+libraries in the \emph{light DOM} can't be used inside the \emph{shadow
+DOM}, too.
+
+That's why the shadowed documents should be more perceived as
+\textbf{static documents} filled and managed solely by the render
+engine. \texttt{Slots} are target areas for \emph{light DOM} nodes used
+to mark the endpoints in question.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> index.html}
+\KeywordTok{<HelloWorld-with-ShadowDOM>}
+  \CommentTok{<!--}
+\CommentTok{    All child nodes will be moved inside the}
+\CommentTok{    shadowRoot if shadowRoot.innerHTML = '<slot></slot>'}
+\CommentTok{    -->}
+  \KeywordTok{<p>}\NormalTok{hello I will be scoped}\KeywordTok{</p>}
+\KeywordTok{</HelloWorld-with-ShadowDOM>}
+\end{Highlighting}
+\end{Shaded}
 
@@
 (the \emph{light DOM}) and not the part of the \emph{shadow DOM}. On the
-other hand there are numerous scenarios where JS is just concerned with
-\textbf{styling or animation of an element}.In this case it might be
-more straightforward to apply JS inside the \emph{shadow DOM} to avoid
-mixing with business logic handlers.
-
-Drilling down to a \emph{light DOM} node from an \emph{shadow DOM}
-context is not possible with querying the node directly with
-\texttt{.querySelector()} or \texttt{.getElementById()} as the node is
-not part of the context. To get a distributed node in question it needs
-the way over the slot node and call \texttt{slot.assinedNodes()} to
-receive an array of distributed node(s) which can be accessed and
-manipulated like any other node. Calling \texttt{.assignedNodes()} on an
-empty \texttt{slot} returns an empty array.
-
-Wrapping up this section, \emph{shadow DOM} provides a non-hacky way to
-create uniform looking \emph{custom elements} and even enhance styling
-possibilities without adding much overhead. Still, for smaller
-components with only one or two child nodes, just a little styling
-and/or no structured redistribution a \emph{shadow DOM} might be to hard
-to reason about. Eventually it all depends on the question of ``how hard
-is it to implement it without shadow DOM'' - which can't be answered
-universally. For a more in-depth guide, Google Engineer Eric Bidelman
-wroten a great primer on \emph{shadow DOM}{[}15{]}.
-
-So far, there is still a missing link between \emph{light DOM} and
-\emph{shadow DOM}. The observant reader may have already noticed the
-weak point in the \texttt{HelloWorld} example: how to ``vitalize'' the
-\emph{shadow DOM}. Recapturing the last \texttt{HelloWorld} example a
-string of markup was assigned to the \texttt{shadowRoot.innerHTML}
-property. While it works perfectly fine in this simple case, a string of
+other hand there are numerous scenarios where JS is used for styling or
+animation of an element. In this case it might be more straightforward
+to \textbf{apply JS inside the \emph{shadow DOM}} to avoid mixing with
+logic handlers with styling.
+
+It seems that a \emph{light DOM} node is in the \emph{shadow DOM}
+context after distribution. Visually this may be true but logically the
+node stays in the normal document. To reach a distributed node from the
+\emph{shadow DOM} to apply some JS behavior for styling it needs the
+extra way over the slot node. Calling \texttt{assinedNodes()} on the
+hosting slot element returns a linkage to the distributed node which can
+be accessed and manipulated like in the \emph{light DOM} context.
+
+Wrapping up this section \emph{shadow DOM} provides a non-hacky way to
+create uniform looking custom elements and even enhance styling
+possibilities without adding overhead. For small components with just a
+little styling a \emph{shadow DOM} might be over engineered. Eventually
+it all depends on the question of ``how hard is it to implement it
+without shadow DOM'' - which can't be answered universally. For a more
+in-depth guide, Google Engineer Eric Bidelman wroten a great primer on
+\emph{shadow DOM}{[}15{]}.
+
+There is still a missing link between \emph{light DOM} and \emph{shadow
+DOM}. The observant reader may have already noticed the weak point in
+the HelloWorld example: how to ``vitalize'' the \emph{shadow DOM} with
+slot or other element structuring instead of \texttt{innerHTML} strings.
+While strings works perfectly fine in this simple case a string of
 markup is rather cumbersome and error-prone and doesn't scale well. When
-putting quotes inside another quotes things break quickly. It makes the
-life hard for developers to work with it because it requires manual
-indentation and is out of syntax highlighting. That's the time templates
-come into play.
-
-\subsection{\texorpdfstring{HTML Templates
-\href{https://www.w3.org/TR/html5/scripting-1.html\#the-template-element}{(w3c)}}{HTML Templates (w3c)}}\label{html-templates-w3c}
-
-Among all other new standards \emph{HTML templates} are the most mature
-and adopted standard in the browser environment. All major browsers,
-except from Internet Explorer, support this standard.
-
-One core concept in templates is efficiency: Whatever dropped inside a
-\texttt{template} tag \sout{bucket} will be parsed on runtime - but not
-constructed into the \emph{content tree}. It remains plain HTML Markup
-sitting somewhere in the document until the time of activation.
-
-Activation typically takes four steps:
+putting quotes inside other quotes things break quickly. Strings make
+development harder because code editor features like indentation or
+syntax highlighting won't be supported. The HTML templates is set to
+fill this gap.
+
+\subsection{\texorpdfstring{3.3. HTML Templates
+\href{https://html.spec.whatwg.org/multipage/scripting.html\#the-template-element}{(whatwg)}}{3.3. HTML Templates (whatwg)}}\label{html-templates-whatwg}
+
+Among all other new specifications \emph{HTML templates} are the most
+mature and adopted standard in the browser environment. All major
+browsers support it since years.
+
+One core concept in templates is browser performance. Elements inside a
+\texttt{template} tag will be parsed on runtime - but not constructed
+and rendered into the content tree. They're remaining plain HTML markup
+sitting in the document until the time of activation.
+
+Activation usually takes four steps:
 
@@
 
+  \CommentTok{<!-- light DOM / typical inserted at index.html -->}
+  \KeywordTok{<p}\OtherTok{ slot=}\StringTok{"placeholder"}\KeywordTok{>}
+    \NormalTok{Hello World Web Component}
+  \KeywordTok{</p>}
+\KeywordTok{</hello-world>}
+
 \KeywordTok{<script>}
- \CommentTok{// Switched to anonymous class notation}
- \CommentTok{// for keeping associated code together.}
- \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,}
-  \KeywordTok{class} \KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+  \KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
     \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
@@
       \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
-    \KeywordTok{const} \NormalTok{helloTemplate }\OperatorTok{=} \VariableTok{document}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'#hello'}\NormalTok{)}\OperatorTok{;}
-    \KeywordTok{this}\NormalTok{.}\VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{helloTemplate}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
-   \OperatorTok{\}}
-  \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+      \KeywordTok{const} \NormalTok{temp }\OperatorTok{=} \KeywordTok{this}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'template#hello'}\NormalTok{)}\OperatorTok{;}
+      \KeywordTok{this}\NormalTok{.}\VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{temp}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
+    \OperatorTok{\}}
+  \OperatorTok{\}}
+  \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}\OperatorTok{;}
 \OperatorTok{<}\SpecialStringTok{/script>}
@@
 The updated \texttt{HelloWorld} component looks already pretty mature.
-It combines all the previous mentioned standards into one blob of HTML.
-\emph{Custom Elements} serves the logic, \emph{Shadow DOM} scopes the
-styles and \emph{HTML Templates} efficiently glues DOM and \emph{Shadow
-DOM} together. This separation of concerns comes with a huge gain in
-flexibility. In a real world scenario \texttt{HelloWorld} would
-contain/reference multiple \emph{HTML Templates} and could switch them
-around without any fuss. A developer might to split up templates into
+It combines all the previous mentioned standards into one HTML file.
+Custom elements serves the logic, shadow DOM scopes the styles and
+\emph{HTML Templates} efficiently glues light DOM and shadow DOM
+together. This separation of concerns comes with a surplus in
+flexibility. In a real world scenario \texttt{HelloWorld} could
+reference multiple \emph{HTML Templates} and switch them around without
+any fuss. Even further a developer might split up templates into named
 \textbf{STYLE} templates and \textbf{CONTENT} templates to increase
@@
 The last standard in the row of four is not concerned with the internals
-of a \emph{web component}. \emph{HTML Imports} serves the need for an
-efficient distribution mechanism of components.
-
-\subsection{\texorpdfstring{HTML Imports
-\href{https://www.w3.org/TR/html-imports/}{(w3c)}}{HTML Imports (w3c)}}\label{html-imports-w3c}
+implementation of a web component. HTML Imports serves the need for an
+efficient distribution mechanism of components and other HTML resources.
+
+\subsection{\texorpdfstring{3.4. HTML Imports
+\href{https://www.w3.org/TR/html-imports/}{(w3c)}}{3.4. HTML Imports (w3c)}}\label{html-imports-w3c}
 
 Importing the \texttt{HelloWorld} component is a one-liner:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"Hello.html"}\OtherTok{ async}\KeywordTok{>}
-\end{Highlighting}
-\end{Shaded}
-
-The \texttt{async} flag is optional but like in any other fetching
-event, strongly recommended. Once the imported HTML document comes into
-scope, activation follows a very similar pattern like the aforementioned
-\emph{HTML templates}:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  \textbf{Querying the link node}
-\item
-  \textbf{Parsing the content and preparing the render}\\
-  const content = linknode.import; -\textgreater{} Unlike the \emph{HTML
-  template} the content a fully equipped document object.
-\item
-  \textbf{Optional: Cloning some nodes for multiple use}
-\item
-  \textbf{Appending the clone/original to destination}
-\end{enumerate}
-
-This again is the imperative way to handle a generic \emph{HTML Import}.
-In the declarative world of \emph{web components} a component is
-activated, parsed and anchored solely by its' tag name
-\texttt{\textless{}hello-world\textgreater{}\textless{}/hello-world\textgreater{}}.
-Preliminary, the component needs proper configuration. The next section
-will elaborate the right configuration and composition of a component to
-work out-of-the-box.
-
-Despite from being just a practical document importer \emph{HTML
-imports} acts like a fully fledged dependency manager for the browser.
-Multiple resources, ranging from stylesheets, scripts, documents, media
-files and even other \texttt{imports} can be grouped together in a
-logical \texttt{import} statement. Internally, the browser engine keeps
-track for every imported resource so it won't be loaded twice. The
-inherent complexity is in fact a stumbling block for wider browser
-adoption. Currently only Googles blink web engine supports \emph{HTML
-Imports} as they are the driving force behind the \emph{web components}
-spec in general. Mozilla and Apple imposed distaste for \emph{HTML
-Imports} as a whole. One reason for this can be found in the
-incompatibility of the spec with the upcoming \emph{ES6 module
-loader}.\footnote{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}
-
-Despite the discrepancies among browser vendors \emph{HTML Imports}
-should still be part of the paper as no other native browser technology
-can bundle up CSS, JS and HTML that efficient.
-
-\subsection{\texorpdfstring{Appendix A: Custom Events
-\href{https://dom.spec.whatwg.org/\#interface-customevent}{(whatwg)}}{Appendix A: Custom Events (whatwg)}}\label{appendix-a-custom-events-whatwg}
-
-Events are first-class citizens in the browser world and \emph{Custom
-Events} are no exception. The \emph{Custom Elements} interface is part
-of the DOM since years but with the rise of \emph{Custom Elements} they
-will most likely become an indispensable building block of \emph{web
-components}.
 
@@
 \NormalTok{> index.html}
+\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"Hello.html"}\OtherTok{ async}\KeywordTok{>}
+
+\KeywordTok{<hello-world>}
+  \KeywordTok{<p}\OtherTok{ slot=}\StringTok{"placeholder"}\KeywordTok{>}
+    \NormalTok{Hello World Web Component}
+  \KeywordTok{</p>}
+\KeywordTok{</hello-world>}
+\end{Highlighting}
+\end{Shaded}
+
+The \texttt{async} flag is optional but recommended like any other
+fetching event. Once the imported HTML document comes into scope
+activation follows a very similar process compared to the aforementioned
+HTML templates:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  \textbf{Querying the link node}
+\item
+  \textbf{Parsing the content and preparing the render}\\
+  const content = linknode.import; -\textgreater{} Unlike the \emph{HTML
+  template} a complete document object is constructed.
+\item
+  \textbf{Optional: Cloning some nodes for multiple use}
+\item
+  \textbf{Appending the clone/original to destination}
+\end{enumerate}
+
+Again this is the imperative way to handle a generic \emph{HTML Import}.
+In the declarative world of web components a component is
+\textbf{parsed, auto-activated and anchored} solely by its' tag name
+\texttt{\textless{}hello-world\textgreater{}\textless{}/hello-world\textgreater{}}
+on purpose. The very own lifecycle method \textbf{adoptedCallback()} in
+custom elements shows the strong interconnection between those
+standards.
+
+\emph{HTML imports} can import everything which is wrappable with HTML
+markup. Stylesheets, scripts, documents, media files and even further
+imports statements can form a semantic \emph{HTML import} statement. The
+far reaching possibilities of a single standard has its downsides.
+According to Mozilla \emph{HTML imports} are not fully compatible with
+the dependency model of the new ES6 modules.\footnote{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}
+Under the current situation Mozilla and Apple and Microsoft won't
+implement \emph{HTML Imports} soon if any. Currently only Googles blink
+web engine supports \emph{HTML Imports} as they are the driving force
+behind the web components specs in general. Despite the discrepancies
+among browser vendors \emph{HTML Imports} are part of this paper as no
+other native standard can ship bundled HTML, CSS and JS which is a core
+concept behind web components.
+
+\subsection{\texorpdfstring{3.5. Custom Events
+\href{https://dom.spec.whatwg.org/\#interface-customevent}{(whatwg)}}{3.5. Custom Events (whatwg)}}\label{custom-events-whatwg}
+
+Events are first-class citizens in the browser. After all they provide a
+neat communication channel for dynamic interactions. \emph{Custom
+Events} are part of the DOM since years but with the rise of web
+components they will most likely become an indispensable building block
+of web components.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> hello-world-component.html}
+
 \KeywordTok{<hello-world>}
@@
 \KeywordTok{</hello-world>}
-\CommentTok{<!-- COMPONENT STARTS HERE -->}
+
 \KeywordTok{<script>}
@@
     \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
-   \CommentTok{// Craft a CustomEvent e}
-    \KeywordTok{const} \NormalTok{e }\OperatorTok{=} \KeywordTok{new} \AttributeTok{CustomEvent}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \OperatorTok{\{}
-    \DataTypeTok{bubbles}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,} \CommentTok{//important!}
+
+   \CommentTok{// Craft a helloevent}
+    \KeywordTok{const} \NormalTok{helloevent }\OperatorTok{=} \KeywordTok{new} \AttributeTok{CustomEvent}\NormalTok{(}\StringTok{'helloevent'}\OperatorTok{,} \OperatorTok{\{}
+    \DataTypeTok{bubbles}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,}
     \DataTypeTok{detail}\OperatorTok{:} \StringTok{'Contains scalar or object'}
    \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
-   \CommentTok{// Launch e on child button click}
+
+   \CommentTok{// Launch helloevent if child button clicks}
    \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'click'}\OperatorTok{,} \NormalTok{click }\OperatorTok{=>} \OperatorTok{\{}
-    \KeywordTok{this}\NormalTok{.}\AttributeTok{dispatchEvent}\NormalTok{(e)}
+    \KeywordTok{this}\NormalTok{.}\AttributeTok{dispatchEvent}\NormalTok{(helloevent)}
     \VariableTok{click}\NormalTok{.}\AttributeTok{stopPropagation}\NormalTok{()}\OperatorTok{;}
-   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
-   \CommentTok{// Catch e. Typically done by some parent}
-     \CommentTok{// node. Message in detail property}
-   \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>} \OperatorTok{\{}
-    \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\VariableTok{e}\NormalTok{.}\AttributeTok{detail}\NormalTok{)}
-   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+   \OperatorTok{\}}\NormalTok{)}
   \OperatorTok{\}}
@@
 self-explanatory. The \texttt{detail} property can be loaded with
-primitives as well as objects. The further sections will elaborate a
-feasible architecture for building a scaling microservice architecture.
-
-Chaining and aggregating events from child nodes should be practiced and
-exercised quiet frequently. As mentioned earlier in the \emph{Custom
-Elements} section, the pattern of \textbf{extending native elements}
-should be somewhat dismissed as it may never implemented outside of
-Chrome. Nevertheless, it is possible to create own kind of quasi native
-buttons when chaining a \emph{CustomEvent} directly after the native
-click event.
-
-Another typical \emph{custom element} use case can be as an actor on
-(native) child elements. In this case, the \emph{Custom Element} catches
-events from children, buffers or rebuild them and eventually fires an
-event towards the document root.
-
-Unfortunately events only work ``upstream'' towards parent nodes. Still
-the web platform offers plenty of possibilities to talk back to child
-nodes.
-
-\subsection{\texorpdfstring{Appendix B: Web Worker
-\href{https://html.spec.whatwg.org/multipage/workers.html}{(whatwg)}}{Appendix B: Web Worker (whatwg)}}\label{appendix-b-web-worker-whatwg}
+scalars as well as objects. Subsequent parent nodes may catch the custom
+event with a clear understanding about the source node.
+
+Chaining and aggregating events from child nodes can be frequently used
+within web components. One use case could be creating a custom button
+like in the previous example. As mentioned earlier in the Custom
+Elements section at 3.1.3. , the pattern of \textbf{extending native
+elements} should be somewhat dismissed as other browser vendors may
+never implement it. A common workaround to eventually create a own
+version a button could be made with a thin wrapper around the native
+button and chaining a \emph{Custom Event} after the click event.
+
+Another common web components use case can be a \textbf{middleware}
+subscribing to certain child events and acting upon them. This involves
+catching events, buffering, destructuring and creation of an own event
+towards the document root. By design, events only bubble upstream
+towards parent nodes. For handing down information towards the child
+nodes we need to query the child node directly.
+
+\subsection{\texorpdfstring{3.6. Web Worker
+\href{https://html.spec.whatwg.org/multipage/workers.html}{(whatwg)}}{3.6. Web Worker (whatwg)}}\label{web-worker-whatwg}
 
@@
 background the paper should briefly join them to form a
-\emph{browsernative microservices}. Needless to say the described
-example is overall simplified to only illustrate the connection between
-browsernative technologies and microservice patterns. Googles library
-Polymer is a good place for learning about web components in depth and
-make use of their simple command line tools. One of their most famous
-proof of concept is the so-called
+\emph{browsernative microservices}. Needless to say the following
+example is overall simplified towards illustrating the connection
+between browsernative technologies and microservice patterns.
+Furthermore it is highly opinionated in development and shouldn't
+perceived as ``single source of truth''.
+
+Googles library Polymer is a good place for learning about web
+components in depth and make use of their toolbox. One of their most
+famous proof of concept is the so-called
 \href{https://shop.polymer-project.org/}{Polymer Shop} which is a
@@
 contain more JS and less if any HTML markup. We probably don't need to
-utilize ShadowDOM as no styles are involved.
-
-Last but not least, we can expand the model for illustrational purposes
+utilize Shadow DOM as no styles are involved.
+
+Last but not least, the pattern can expanded for illustrational purposes
 to \textbf{native components} which is every build-in HTMLElement like
-the HTMLButtonElement. They are solely controllable and styleable from
-the outside and must therefore wrapped in presentational components
-and/or containers.
+the HTMLButtonElement. Native components are mostly deep nested elements
+providing the actual functionality in the browser UI. They are solely
+controllable and styleable from the outside and are therefore wrapped in
+presentational components and/or containers.
 
 Lets start the service description top-down beginning with the
-\textbf{root container} managing the overall service.
-
-\subsubsection{Root container}\label{root-container}
+\textbf{service root container} managing the overall service.
+
+\subsection{4.1. Service root}\label{service-root}
 
@@
 the message. This communication model offers lots of possibilities for
-\textbf{evolutionary design} as child nodes can be loosely dropped in
-the service root and communication might \textbf{fail graceful} as
-simply nothing happens if one or another switch is not defined.
-
-\subsubsection{Container components}\label{container-components}
+\textbf{evolutionary design} as child nodes can be loosely dropped.
+Every child node can be expanded into another full microservice or split
+up into separate nodes without notice of the service root. Communication
+might \textbf{fail graceful} providing a default console log in case
+switches aren't defined yet.
+
+\subsection{4.2. Container components}\label{container-components}
 
@@
 \texttt{\textless{}sign-in\textgreater{}} or
-\texttt{\textless{}shipping-details\textgreater{}} still contain a lot
-of logic and no or less styling. Their job is to control their
-underlying presentational leaf components, aggregating events and to
-define a \textbf{set of actions towards the model} . Every container
-mounted directly under the service root may have a dedicated area inside
-the worker thread where it's fulfil his duties. For example, a typical
+\texttt{\textless{}shipping-details\textgreater{}} still contain mostly
+logic and no or less styling. Their job is to control their underlying
+presentational leaf components, aggregating events and to define a
+\textbf{set of actions towards the model} . Every container mounted
+directly under the service root may have a dedicated area inside the
+worker thread where it's fulfil his duties. For example, a typical
 \texttt{\textless{}sign-in\textgreater{}} contains merely two fields
@@
     \CommentTok{// switch to methods}
-    \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\KeywordTok{this}\NormalTok{.}\AttributeTok{localName} \OperatorTok{+} \StringTok{"received msg"}\NormalTok{)}\OperatorTok{;}
-    \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(msg)}\OperatorTok{;}
+    \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\VerbatimStringTok{`}\SpecialCharTok{$\{}\KeywordTok{this}\NormalTok{.}\AttributeTok{localName}\SpecialCharTok{\}}\VerbatimStringTok{ received }\SpecialCharTok{$\{}\NormalTok{msg}\SpecialCharTok{\}}\VerbatimStringTok{`}\NormalTok{)}\OperatorTok{;}
   \OperatorTok{\}}
@@
 
-Extending the \texttt{CheckoutContainerBase} will equipped every
-container node with the unified message interface. Calling
-\texttt{this.\_dispatch(msg)} within a within the container will trigger
-an event bubbling upstream. The service root will implement a simple
-event listener for \texttt{checkout} events. After receiving an answer
-from the MODEL a service root can \texttt{querySelector} the container
-node in question and push forward the answer over the \texttt{\_receive}
-property.
+Extending the \texttt{SimpleContainer} will equip every container node
+with the unified message interface. Calling
+\texttt{this.\_dispatch(msg)} within the container will trigger an event
+bubbling upstream. The service root will implement a simple event
+listener for \texttt{checkout} events. After receiving an answer from
+the MODEL a service root can query the container child node in question
+and push forward the answer over the \texttt{\_receive} property.
 
 There might be even more middleware containers pulled in between
-presentational leafs and service root component to do some extra work
-without caring about bubbling events altogether. Changing or enhancing
-any functionality requires only to touch the controllers in the
-containers in question and the endpoint section at the model.
-\textbf{Infrastructure automation} might be achieved by dynamically
-evaluating the mounted containers beneath the service root and modeling
-the ``backend'' model accordingly. Due to it's standardized message
-system the containers are testable within standardized tests.
-
-\subsubsection{Presentational
+presentational leafs and the service root to fulfil some extra work
+either like filters on the event or without caring about checkout events
+altogether. Changing or enhancing any functionality requires only the
+controllers in the container in question and the endpoint section at the
+model. \textbf{Infrastructure automation} might be achieved by
+dynamically evaluating the mounted containers beneath the service root
+and modeling the ``backend'' model accordingly. Due to it's standardized
+message system the containers are testable within standardized tests.
+
+\subsection{4.3. Presentational
 components}\label{presentational-components}
 
-In comparison to the former presentational component is build mostly
-
-\section{Thinking further}\label{thinking-further}
-
-cohesive HTML JS CSS
-
-virtual DOM
-
-declarative event management
+In comparison to the former container the presentational component is
+build mostly around views and styling and shipping less or no logic
+towards handling the component. To extend the last
+\texttt{\textless{}sign-in\textgreater{}} container example a typical
+presentational component structure could look like the following
+top-level structure
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{<sign-in-container>}
+  \KeywordTok{<sign-in-presentational>}
+  \KeywordTok{</sign-in-presentational>}
+\KeywordTok{</sign-in-container>}
+\end{Highlighting}
+\end{Shaded}
+
+The \texttt{\textless{}sign-in-presentational\textgreater{}} contains
+all required input fields and oauth connectors to Google or Facebook but
+\textbf{will not care about events they create}. A presentation
+component is usually steered via HTML attributes. Contrary to the former
+containers the presentational component highly utilizes \textbf{shadow
+DOM and HTML templates}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> sign-in-presentational.html}
+\CommentTok{<!-- IMPORTS -->}
+\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"facebook-oauth.html"}\OtherTok{ async}\KeywordTok{>}
+\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"high-res-template.html"}\OtherTok{ async}\KeywordTok{>}
+\NormalTok{...}
+
+\CommentTok{<!-- VIEW -->}
+\KeywordTok{<sign-in-presentational>}
+  \KeywordTok{<high-res-template>}
+    \KeywordTok{<facebook-oauth></facebook-oauth>}
+    \KeywordTok{<slot}\OtherTok{ name=}\StringTok{"form"}\KeywordTok{></slot>}
+    \KeywordTok{<slot}\OtherTok{ name=}\StringTok{"button"}\KeywordTok{></slot>}
+  \KeywordTok{</high-res-template>}
+
+  \KeywordTok{<template}\OtherTok{ id=}\StringTok{"lowres"}\KeywordTok{>}\NormalTok{...}\KeywordTok{</template>}
+  \KeywordTok{<template}\OtherTok{ id=}\StringTok{"mobile"}\KeywordTok{>}\NormalTok{...}\KeywordTok{</template>}
+\KeywordTok{</sign-in-presentational>}
+
+\KeywordTok{<script>}
+  \KeywordTok{class} \NormalTok{SignIn }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+    \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
+      \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
+      \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+      \CommentTok{// append templates according to window.innerWidth}
+      \CommentTok{// run this.getAttribute('htmlprop') to control element}
+    \OperatorTok{\}}
+    \AttributeTok{attributeChangedCallback}\NormalTok{(attrName}\OperatorTok{,} \NormalTok{oldVal}\OperatorTok{,} \NormalTok{newVal) }\OperatorTok{\{}
+      \CommentTok{// react on changing attributes}
+    \OperatorTok{\}}
+  \OperatorTok{\}}
+  \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'sign-in-presentational'}\OperatorTok{,} \NormalTok{SignIn)}\OperatorTok{;}
+\OperatorTok{<}\SpecialStringTok{/script>}
+\end{Highlighting}
+\end{Shaded}
+
+Presentational components require configuration and usually ship a long
+list of CSS selectors. Whereas it's possible to expand templates in
+place it could be also possible to define \textbf{template components}
+to atomize the component further and increase code readability.
+
+\section{5. Thinking further}\label{thinking-further}
+
+Following a \textbf{browsernative first} principle this paper didn't
+made use of third-party libraries so far. Contrary to the example in the
+last section there are a few reasons why web components shouldn't be
+used without any proprietary technology. At first, from a technical
+perspective web components are relatively young and at the time of
+writing only the newer Chrome and Opera browsers can handle them
+correctly. Most probably it will take a long time until native web
+components can be safely used without additional legacy support.The
+second inevitable question is the notion whether proprietary technology
+promotes \textbf{simplicity} or complects the implementation.
+
+Talking about web components it is impossible to avoid technologies like
+\textbf{React} altogether. In the last couple years React introduced
+some good concepts to web development like the idea of state, passing
+data as properties and declarative event management just to name a few.
+Many of those design decisions were made to overall \textbf{simplify
+browser development through standardized methodologies}. The principles
+of microservices could be applied to React development maybe even
+further as for example the communication model involves a lot of
+delegation and decomposition.
+
+Many of those features and design decisions in React created much of a
+hype around the library and technologies like the virtual DOM while web
+components never caught up in momentum at a comparable rate. According
+to Github Googles Polymer project is even six months older than React
+but has less than one third of its stars. Other notable web components
+projects like Bosonic\footnote{https://github.com/bosonic/bosonic/} or
+Mozilla's X-tag\footnote{https://github.com/x-tag/core} are more dead
+than alive, too.
+
+It seems like the web developing community in a broader sense demands
+for features and standardization in frameworks like React (and many
+others to be fair). Web components and related libraries clearly missed
+a selling point in providing standardized ways for developing scalable
+components and applications or being to cumbersome to use at all. For
+the future it remains unclear if web components will eventually gain
+wider adoption within the UI developing community or will be adopted as
+low-level technology for framework development emphasized by Sebastian
+Markbage, one of the React creators.{[}18{]}
+
+One interesting approach for building web components is the ``reactish''
+Skate\footnote{https://github.com/skatejs/skatejs/}. Following the
+functional rendering model from React it combines native methods with
+additional functionalities known from React like event management. Given
+the focus on native technologies it only weights around 4kb (minified
+and gzipped) which is ten times less than React. \textbf{Small
+encapsulated micro UI frameworks} like Skate could be certainly the near
+future of web development and will hopefully gain more attention. The
+more mature UI framework Riot already expressed their intention to
+gradually develop towards web components.\footnote{http://riotjs.com/compare/\#web-components}
 

### 2017-02-27 build/paper.tex
pdf/tex build

@@
 Building complex web applications nowadays require additional layers of
-abstraction and often heavily depend on proprietary frameworks. New
-specifications build right into the browserengine provide a native
-service API to overcome tricky abstraction constraints.
+abstraction and often heavily depend on indispensable frameworks. While
+they perfectly circumstance the global paradigm of the DOM they remain
+highly proprietary attached to the framework. New w3c specifications
+build right into the browser provide a native service API in the need to
+create standardized web components for the browser. As of 2017 all major
+browsers will ship those technologies in their browsers. The paper not
+only provides an introduction to the specification. As an assessment
+criteria of it will match those specifications against the concept of
+microservices which is an modular system architecture to build scalable
+applications.
 \end{abstract}
@@
 
-Derived from the Latin word \textbf{simplex}, simple can be understood
-as ``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
-  Dictionary}} which points towards an unidimensional nature. While
-complexity describes the multilayered und entangled nature of
-conditions, simplicity empowers the human brain to reason about issues
-in a straightforward manner. It certainly has some overlapping's with
-easy, but while easy is more of a relative spirit, simple can be laid
-out as a objective manner and therefore universally applicable.
+Derived from the Latin word \textbf{simplex} simple can be understood as
+``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
+  Dictionary}} which points towards an unidimensional state. While
+complexity describes the multilayered und entangled nature of conditions
+simplicity empowers the human brain to reason about issues
+straightforward. It certainly has some overlapping's with easy, but
+while easy is more of a relative spirit, simple can be laid out as a
+objective manner and therefore universally applicable.
 
@@
   Trends}} Leaving images, fonts or other content aside the size of
-HTML, CSS and JS sums up to a total average of 550 kb. One character
+HTML, CSS and JS sums up to an total average of 550 KB. One character
 weights around 1 byte which means an average webpage is delivering
@@
 and design rules into maintainable components. While the DOM can't be
-scoped, JS can. Many frameworks, like ReactJS, AngularJS or VueJS just
-to name a few, ditched the old rule of separated HTML, CSS and JS in
-favor of an additional layer of abstracted JS components (containing
-structure, behavior and sometimes styling). Quiet often those frameworks
-mimic a MVC pattern on top of the browser engine which is a reasonable
-simple design pattern to build graphical user interfaces. While
-frameworks are a valid approach for building scalable web applications
-they remain highly opinionated, embody inherent complexity themselves
-and can change and break over time. Another downside is code inflation
-which is a crucial point for performance. All of those bottlenecks in
-the web demand for a new standardized way for creating and evolving
-complex web services.
+scoped, JS can. Many frameworks, like React, Angular or Vue just to name
+a few, ditched the old rule of separated HTML, CSS and JS in favor of an
+additional layer of abstracted JS components (containing structure,
+behavior and sometimes styling). Quiet often those frameworks mimic a
+MVC pattern on top of the browser engine which is a reasonable simple
+design pattern to build UIs. While frameworks are a valid approach for
+building scalable web applications they remain highly opinionated embody
+inherent complexity themselves and can change and break over time.
+Another downside is code inflation which is a crucial point for
+performance. All of those bottlenecks in the web demand for a new
+standardized way for creating and evolving complex web services.
 
@@
 principle}: changing and deploying one service shouldn't result in
-changing other parts of the system.{[}3, p. 30{]}. Mutations or shadowed
-variables, which is happening a lot in CSS, making it hard to keep
-changes ought to only affect one place in the application. A
-\emph{browsernative microservice} therefore pushing encapsulation and
-avoiding variable mutations outside its scope as much as possible.
-Practically, CSS will be scoped and JS fosters immutable JS entities and
-avoids variables leaking into the global namespace.
+changing other parts of the system.{[}3, p. 30{]} Picking CSS for
+example, where variables tend to be shadowed by higher specific values
+making it increasingly difficult to keep changes ought to only affect
+one place in the application. A \emph{browsernative microservice}
+therefore pushing encapsulation and avoiding variable mutations outside
+its scope as much as possible. One solution would be scoped CSS / JS to
+avoid variables leaking into the global namespace as well as hidden
+implementation details to avoid mutating values.
 
@@
 principle}: Whether designing a microservice or it's components we want
-related behavior sit together, and unrelated behavior to sit
+related behavior sit together and unrelated behavior to sit
 elsewhere.{[}3, p. 30{]} High cohesion can be enhanced towards the more
@@
 \emph{elm}\footnote{\href{http://elm-lang.org/}{elm lang}} where MODEL,
-VIEW and CONTROLLER sit in one place. A browsernative service aims for a
-combinement of the web native trinity HTML, JS and CSS.
+VIEW and CONTROLLER sit in one place. A \emph{browsernative service}
+sought for an combination of the web native trinity HTML, JS and CSS.
 
 Traditional web development relies on \textbf{libraries} to enhance the
-service capabilities of the web platform. Compared to libraries, a
+service capabilities of the web platform. Compared to libraries a
 component service offers multiple advantages for building, deployment
@@
 example is overall simplified towards illustrating the connection
-between browsernative technologies and microservice patterns. Googles
-library Polymer is a good place for learning about web components in
-depth and make use of their toolbox. One of their most famous proof of
-concept is the so-called
+between browsernative technologies and microservice patterns.
+Furthermore it is highly opinionated in development and shouldn't
+perceived as ``single source of truth''.
+
+Googles library Polymer is a good place for learning about web
+components in depth and make use of their toolbox. One of their most
+famous proof of concept is the so-called
 \href{https://shop.polymer-project.org/}{Polymer Shop} which is a
@@
 Lets start the service description top-down beginning with the
-\textbf{root container} managing the overall service.
-
-\subsection{Root container}\label{root-container}
+\textbf{service root container} managing the overall service.
+
+\subsection{4.1. Service root}\label{service-root}
 
@@
 
-\subsection{Container components}\label{container-components}
+\subsection{4.2. Container components}\label{container-components}
 
@@
 
-\subsection{Presentational components}\label{presentational-components}
-
-In comparison to the former presentational component is build mostly
-
-template switch
-
-css
-
-shadow dom
+\subsection{4.3. Presentational
+components}\label{presentational-components}
+
+In comparison to the former container the presentational component is
+build mostly around views and styling containing less or no logic
+towards handling the component. To extend the last
+\texttt{\textless{}sign-in\textgreater{}} container example a typical
+presentational component structure could look like the following
+top-level structure
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{<sign-in-container>}
+  \KeywordTok{<sign-in-presentational>}
+  \KeywordTok{</sign-in-presentational>}
+\KeywordTok{</sign-in-container>}
+\end{Highlighting}
+\end{Shaded}
+
+The \texttt{\textless{}sign-in-presentational\textgreater{}} contains
+all required input fields and oauth connectors to Google or Facebook but
+\textbf{will not care about events they create}. A presentation
+component is usually steered via HTML attributes. Contrary to the former
+containers the presentational component highly utilizes \textbf{shadow
+DOM and HTML templates}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{> sign-in-presentational.html}
+\CommentTok{<!-- IMPORTS -->}
+\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"facebook-oauth.html"}\OtherTok{ async}\KeywordTok{>}
+\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"high-res-template.html"}\OtherTok{ async}\KeywordTok{>}
+\NormalTok{...}
+
+\CommentTok{<!-- VIEW -->}
+\KeywordTok{<sign-in-presentational>}
+  \KeywordTok{<high-res-template>}
+    \KeywordTok{<facebook-oauth></facebook-oauth>}
+    \KeywordTok{<slot}\OtherTok{ name=}\StringTok{"form"}\KeywordTok{></slot>}
+    \KeywordTok{<slot}\OtherTok{ name=}\StringTok{"button"}\KeywordTok{></slot>}
+  \KeywordTok{</high-res-template>}
+
+  \KeywordTok{<template}\OtherTok{ id=}\StringTok{"lowres"}\KeywordTok{>}\NormalTok{...}\KeywordTok{</template>}
+  \KeywordTok{<template}\OtherTok{ id=}\StringTok{"mobile"}\KeywordTok{>}\NormalTok{...}\KeywordTok{</template>}
+\KeywordTok{</sign-in-presentational>}
+
+\KeywordTok{<script>}
+  \KeywordTok{class} \NormalTok{SignIn }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+    \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
+      \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
+      \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+      \CommentTok{// append templates according to window.innerWidth}
+      \CommentTok{// run this.getAttribute('htmlprop') to control element}
+    \OperatorTok{\}}
+    \AttributeTok{attributeChangedCallback}\NormalTok{(attrName}\OperatorTok{,} \NormalTok{oldVal}\OperatorTok{,} \NormalTok{newVal) }\OperatorTok{\{}
+      \CommentTok{// react on changing attributes}
+    \OperatorTok{\}}
+  \OperatorTok{\}}
+  \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'sign-in-presentational'}\OperatorTok{,} \NormalTok{SignIn)}\OperatorTok{;}
+\OperatorTok{<}\SpecialStringTok{/script>}
+\end{Highlighting}
+\end{Shaded}
+
+Presentational components require configuration and usually ship a long
+list of CSS selectors. Whereas it's possible to expand templates in
+place it could be also possible to define \textbf{template components}
+to atomize the component further and increase code readability.
+
+tests
 
@@
 
-Following a \textbf{browsernative first} principle this paper didn't
-made use of third-party libraries so far. Contrary to the example in the
-last section there are a few reasons why web components shouldn't be
-used without any proprietary technology. At first, from a technical
-perspective web components are relatively young and at the time of
-writing only the newer Chrome and Opera browsers can handle them
-correctly. Most probably it will take a long time until native web
-components can be safely used without additional legacy support.The
-second inevitable question is the notion if they should be used pure
-from a \textbf{simplicity perspective} without any proprietary
-technology now and future.
-
-Talking about web components it is impossible to avoid technologies like
-\textbf{React} altogether. In the last couple years React introduced
-some good concepts to web development like the notion of state, passing
-data as properties and declarative event management just to name a few.
-Many of those design decisions were made to overall \textbf{simplify
-browser development through standardized methodologies}. The principles
-of microservices could be applied to React development maybe even
-further as for example the communication model involves a lot of
-delegation and decomposition.
-
-Many of those features and design decisions in React created much of a
-hype around the library and technologies like the virtual DOM while web
+Following a \textbf{browsernative first} dogma this paper didn't put
+third-party libraries into view so far in the search of an independent
+approach. Nevertheless, it should be clear to this point that the
+example in the last section is far away from being usable in production.
+At first, from a technical perspective web components are relatively
+young and at the time of writing only usable in \textbf{newer Chrome and
+Opera browsers}. Most probably it will take a long time until native web
+components can be safely used without additional legacy support. The
+troublesome \emph{HTML imports} spec will certainly need even longer
+which inevitable opens up the question if web components are useful from
+today on. And second, do web components align with the notion of
+simplicity if too many downsides and possible pitfalls require
+attention?!
+
+Any profound discussion about web components would be incomplete without
+touching component alternatives like \textbf{React}. In the last couple
+years React introduced new concepts to web development like the idea of
+state, passing data as properties and declarative event management just
+to name a few. Those design decisions were made to overall simplify
+browser development with full browser support right from the start.
+Overall, Reacts perception about UI development created much of a hype
+around the library and technologies like the virtual DOM while web
 components never caught up in momentum at a comparable rate. According
@@
 
-It seems like the web developing community in a broader sense demands
-for features and standardization in frameworks like React (and many
-others to be fair). Web components and related libraries clearly missed
-a selling point in providing standardized ways for developing scalable
-components and applications or being to cumbersome to use at all. For
-the future it remains unclear if web components will eventually gain
-wider adoption within the UI developing community or will be adopted as
-low-level technology for framework development emphasized by Sebastian
-Markbage, one of the React creators.{[}18{]}
-
-One interesting approach for building web components is the ``reactish''
-Skate\footnote{https://github.com/skatejs/skatejs/}. Following the
-functional rendering model from React it combines native methods with
-additional functionalities known from React like event management. Given
-the focus on native technologies it only weights around 4kb (minified
-and gzipped) which is ten times less than React. \textbf{Small
-encapsulated micro UI frameworks} like Skate could be certainly the near
-future of web development and will hopefully gain more attention. The
-more mature UI framework Riot already expressed their intention to
-gradually develop towards web components.\footnote{http://riotjs.com/compare/\#web-components}
+It seems like the web developing community in a broader sense values
+\textbf{standardized procedures, tooling and browser support} in
+frameworks like React (and many others to be fair). Web components and
+related libraries clearly missed a real selling point towards React and
+Angular. Providing guidelines, simpler syntax and features missed out or
+being to cumbersome to use at all. For the future it remains unclear if
+web components will eventually gain wider adoption within the UI
+developing community or will be adopted as low-level technology for
+framework development emphasized by Sebastian Markbage, one of the React
+creators.{[}18{]}
+
+Even though the downsides seems heavy weight there are notable attempts
+to bring web components into production. A very good example is the
+``reactish'' library Skate\footnote{https://github.com/skatejs/skatejs/}.
+Following the functional rendering model from React it combines native
+methods with additional functionalities known from React like
+declarative event management. Given the focus on native technologies it
+only weights around 4kb (minified and gzipped) which is ten times less
+than React. \textbf{Small encapsulated micro UI frameworks} like Skate
+could certainly be the near future of web development and will hopefully
+pave the way of web components.
+
+The role of web components may not result in an isolated framework but
+move towards a common denominator for future Reacts, Angulars,
+Vues\ldots{} which makes it worth considering it as recyclable
+technology foundation. The more mature UI framework Riot already
+expressed their intention to gradually develop towards web
+components.\footnote{http://riotjs.com/compare/\#web-components}
 

### 2017-02-28 paper.tex
pdf/tex build

@@
 
-Computers can scale, humans can't. Ever since a complex system made by
-humans has been constrained by humans mental capabilities. Like in the
-analogy of juggling balls our brain can just ``juggle'' a few things at
-a time. Rich Hickey, the inventor of the programming language Clojure
-gave an inspirational keynote on the topic of
-\textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
-  Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} In every sphere
-of a humans life simplicity aligns perception with our mental
+Computers can scale, humans can't. Ever since complex systems made by
+humans have been constrained by humans mental capabilities. Like in the
+analogy of juggling balls our brain can just ``juggle'' a few items at a
+time. Rich Hickey, the inventor of the programming language Clojure gave
+an inspirational keynote on the topic of \textbf{simplicity}.{[}1{]} In
+every sphere of a human's life simplicity aligns perception with mental
 capacities.
 
-Derived from the Latin word \textbf{simplex}, simple can be understood
-as ``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
-  Dictionary}} which points towards an unidimensional nature. While
-complexity describes the multilayered und entangled nature of
-conditions, simplicity empowers the human brain to reason about issues
-in a straightforward manner. It certainly has some overlapping's with
-easy, but while easy is more of a relative spirit, simple can be laid
-out as a objective manner and therefore universally applicable.
+Derived from the Latin word \textbf{simplex}, \emph{simple} can be
+understood as ``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
+  Dictionary}} which points towards an unidimensional state. While
+complexity describes the multilayered und entangled nature of conditions
+simplicity empowers the human brain to reason about issues
+straightforward. It certainly has some overlappings with \emph{easy},
+but while \emph{easy} is more of a relative spirit, \emph{simple} can be
+laid out as an objective manner and therefore universally applicable.
 
@@
 evolves in one or another opinionated direction over time. Layers of
-abstractions wrestling with aged legacy code requiring layers of
+abstractions wrestle against aged legacy code requiring additional
 middleware. Mutating assets create subtle bugs and so forth. Eventually
@@
 Opaqueness of the system will slow down innovation to a minimum in the
-need of ``keeping the lights on''.
+need of `keeping the lights on'.
 
@@
 \textbf{impression of extending, substitution, moving, combining and
-repurposing}. The ability to reason about the program at any given time
-is crucial for future decisions and implementations. Recalling again the
-unidimensional nature of simplicity.
-
-Simplicity in ``the web'', read as a loose generalization of
-``everything that runs in the browser'' is certainly a story full of
-misconceptions. While simplicity in the backend is mostly a matter of
-principles and patterns, any browser-based frontend is restricted on the
-highly deterministic nature of the browser platform.
+repurposing}.{[}1{]} The ability to reason about the program at any
+given time is crucial for future decisions and implementations,
+recalling the unidimensional nature of simplicity.
+
+Simplicity in `the web', read as a loose generalization of `everything
+that runs in the browser' is certainly a story full of misconceptions.
+While simplicity in the backend is mostly a matter of principles and
+patterns, any browser-based frontend is restricted on the highly
+deterministic nature of the browser platform.
 
@@
 currently around 2.5 MB.\footnote{\href{http://httparchive.org/trends.php}{HTTPArchive
-  Trends}} Leaving images, fonts or other content aside the size of
-HTML, CSS and JS sums up to a total average of 550 kb. One character
+  Trends}} Leaving images, fonts or other content aside, the size of
+HTML, CSS and JS sums up to a total average of 550 KB. One character
 weights around 1 byte which means an average webpage is delivering
-550.000 character or around 125 pages of single-spaced text. Frederic
+550.000 characters or around 125 pages of single-spaced text. Frederic
 Filloux analyzed the payload on different newspaper websites and came to
-the conclusion, that only round about 5-6 \% of the transferred
-characters made for human consumption.{[}1{]}
-
-Having an 95 \% overhead is rather undesirable for both the consumers
-and creators of the website. Since it's a widespread problem without a
-single point of failure one can argue the platform itself is the
+the conclusion, that only roundabout 5-6 \% of the transferred
+characters are made for human consumption.{[}2{]}
+
+Having a 95 \% overhead is rather undesirable for both the consumers and
+the creators of the website. Since it is a widespread problem without a
+definable point of failure one can argue `the platform itself' is the
 failure. By design, every pageload results in a monolithic DOM tree
 managed by the browser engine. Whether rendering just a bunch of static
-text nodes or an ever changing webapp the underlying global nature of
+text nodes or an ever changing webapp, the underlying global nature of
 the DOM tree remains the same. Every additional piece of code added to
@@
 
-In an non-deterministic runtime environment, encapsulation and
-modularization is a typical pattern to make complexity manageable and
-accommodate future uncertainty.{[}2, p. 1{]} Since years the average JS
+In a non-deterministic runtime environment, encapsulation and
+modularization is a typical pattern to make complexity manageable and to
+accommodate future uncertainty.{[}3, p. 1{]} Since years the average JS
 payload is steadily rising which can be interpreted as a trend towards
-more dynamic websites. The demands to the browser platform changed from
-a static page renderer to a \textbf{dynamic UI machine} without changing
-the underlying architecture significantly. Under the current situation
-only additional layers of abstraction can wrestle complexity.
+more dynamic websites. The demands to the browser platform have changed
+from a static page renderer to a \textbf{dynamic UI machine} without
+significantly changing the underlying architecture. Under the current
+situation only additional layers of abstraction can handle complexity.
 
@@
 methodologies approached the global nature of the DOM by scoping assets
-and design rules into maintainable components. While the DOM can't be
-scoped, JS can. Many frameworks, like ReactJS, AngularJS or VueJS just
-to name a few, ditched the old rule of separated HTML, CSS and JS in
-favor of an additional layer of abstracted JS components (containing
-structure, behavior and sometimes styling). Quiet often those frameworks
-mimic a MVC pattern on top of the browser engine which is a reasonable
-simple design pattern to build graphical user interfaces. While
-frameworks are a valid approach for building scalable web applications
-they remain highly opinionated, embody inherent complexity themselves
-and can change and break over time. Another downside is code inflation
-which is a crucial point for performance. All of those bottlenecks in
-the web demand for a new standardized way for creating and evolving
-complex web services.
+and design rules into maintainable components. While the DOM cannot be
+scoped, JS can. Many frameworks like React, Angular or Vue, just to name
+a few, ditched the old rule of separated HTML, CSS and JS in favor of an
+additional layer of abstracted JS components. Quite often those
+frameworks mimic a MVC pattern on top of the browser engine which is a
+reasonably simple pattern to build UIs. While frameworks are a valid
+approach for building scalable web applications they remain highly
+opinionated, embody inherent complexity themselves and can change and
+break over time. Another drawback is code inflation which is a crucial
+point for performance. All of those bottlenecks in the web demand for a
+new standardized ways for creating and evolving complex web services.
 
@@
 to propose \emph{The Extensible Web Manifesto}.\footnote{\href{https://extensiblewebmanifesto.org/}{The
-  Extensible Web Manifesto}} The manifestos claim was to enhance the
-current web platforms with new low-level capabilities. Those features
-should empower creators of the web to write more declarative code and
-abandon known problems and artificial abstractions. Four years later,
-the enhancement of JavaScript leapfrogged and many new low-level APIs
-brought to life. With this new APIs at hand a vivid web developer can
-create robust websites with less code and less additional libraries.
-This paper is an approach to unfold these \textbf{browsernative}
-technologies to create overall simple and resilient
-\textbf{microservices} for the browser
+  Extensible Web Manifesto}} The claim of the manifesto was to enhance
+the current web platforms with new low-level capabilities. Those
+capabilities aimed to empower creators of the web to write more
+declarative code and to abandon artificial abstractions. Four years
+later, the enhancement of JavaScript leapfrogged and many new low-level
+APIs were brought to life. With this new APIs at hand, a progressive web
+developer can create robust websites with less code and less additional
+libraries. This paper is an approach to unfold these
+\textbf{browsernative} technologies to create overall simple and
+resilient \textbf{microservices} for the browser
 
@@
 In search of a better, simpler web architecture one might look on
-already established pattern that proofed to fulfill enterprise needs.
+already established patters proofed to fulfill enterprise needs.
 Microservices are a good approach for tearing big monolithic systems
@@
 nutshell a microservice is a small, autonomous service that works
-together with other services seamlessly.{[}3, p. 2{]} Or with the words
+together with other services seamlessly.{[}4, p. 2{]} Or with the words
 of Fowler and Lewis: ``\ldots{} the microservice architectural style is
@@
 services, each running in its own process and communicating with
-lightweight mechanisms, often an HTTP resource API.''{[}4{]} Yet at this
+lightweight mechanisms, often an HTTP resource API.''{[}5{]} Yet at this
 point a reader might spot some similarities with microservices and the
-browser-based development: Both wrestling the problem of monolithic
-architecture and both using lightweight communication mechanisms. In
-fact, many big companies of ``the web'' like Amazon or Netflix
-successfully transformed their monolithic system into a service based
-system which gives a taste of the power behind microservices.{[}4{]}
+browser-based development: Both wrestle with the problem of monolithic
+architecture and both use lightweight communication mechanisms. In fact,
+many big companies of `the web' like Amazon or Netflix successfully
+transformed their monolithic system into a service based system which
+gives a taste of the power behind microservices.{[}5{]}
 
 Microservices incorporate a wide array of ideas from developing scalable
-software, like \textbf{domain-driven design} where it pursues the
-incorporation of real world structure in the code.{[}3, p. 2{]} Or
-making use of \textbf{continuous delivery} for pushing software rapidly
-through \textbf{automated deployment} mechanisms into production.{[}4{]}
-Furthermore, microservices transcendent the technical perspective and
-reaches into the team organization. As a primary source of truth this
-paper relies on the work of Sam Newman, who has written a comprehensive
-guide towards microservices{[}3{]} and the work of Fowler and
-Lewis{[}4{]}. The purpose of this section is to gain confidence about
-the microservices architecture in the context of the browser platform.
+software like \textbf{domain-driven design} to pursuing the real world
+structure in the code.{[}4, p. 2{]} Another concept focuses on
+\textbf{continuous delivery} for pushing software rapidly through
+\textbf{automated deployment} mechanisms into production.{[}5{]}
+Furthermore, microservices transcend the technical perspective and reach
+into team organization.
+
+As a primary source of comprehensive information this paper relies on
+the work of Sam Newman{[}4{]} and the work of Fowler and Lewis{[}5{]}.
+The purpose of this section is to gain confidence about the architecture
+of microservices in the context of the browser platform.
 
@@
 ``A~\textbf{component}~is a unit of software that is independently
-replaceable and upgradeable.'' {[}4{]} Components are the building
+replaceable and upgradeable.'' {[}5{]} Components are the building
 blocks of microservices. And microservices are the building blocks of
@@
 microservice or a much more generic component, both share a similar set
-of principles. Therefore this paper referring to both parts when talking
+of principles. Therefore this paper refers to both parts when talking
 about \textbf{services}.
@@
 The first principle of services is the \textbf{loose coupling
-principle}: changing and deploying one service shouldn't result in
-changing other parts of the system.{[}3, p. 30{]}. Mutations or shadowed
-variables, which is happening a lot in CSS, making it hard to keep
-changes ought to only affect one place in the application. A
-\emph{browsernative microservice} therefore pushing encapsulation and
-avoiding variable mutations outside its scope as much as possible.
-Practically, CSS will be scoped and JS fosters immutable JS entities and
-avoids variables leaking into the global namespace.
+principle}: changing and deploying one service should not result in
+changing other parts of the system.{[}4, p. 30{]} Picking CSS for
+example, where variables tend to be shadowed by higher specific values
+making it increasingly difficult to keep changes, ought to merely affect
+one place in the application. Therefore a \emph{browsernative
+microservice} is expected to push encapsulation and to avoid variable
+mutations outside its scope as much as possible. One solution would be
+scoped CSS / JS to avoid variables leaking into the global namespace, as
+well as to hide implementation details in order to avoid mutating
+values.
 
 The second principle of services is the \textbf{high cohesion
-principle}: Whether designing a microservice or it's components we want
-related behavior sit together, and unrelated behavior to sit
-elsewhere.{[}3, p. 30{]} High cohesion can be enhanced towards the more
+principle}: Whether designing a microservice or its components, we want
+related behavior sit together and unrelated behavior to sit
+elsewhere.{[}4, p. 30{]} High cohesion can be enhanced towards the more
 dynamic \textbf{Single Responsibility Principle}: ``Gather together
 those things that change for the same reason and separate those things
-that change for different reasons.''{[}5{]} In a very quick and dirty
+that change for different reasons.''{[}6{]} In a very quick and dirty
 code quality analysis, the quality can be measured just by counting the
-places changes in the code occur in order to implement a functionality.
-An arbitrary threefold MVC system should require a maximum of three
-changes to implement or change functionalities. The problem in browser
-based development is not only the global paradigm which makes changes
+code changes which occur in order to implement a new functionality. An
+arbitrary threefold MVC system should require a maximum of three changes
+to implement or change functionalities. The problem in browser based
+development is not only the global paradigm which makes changes
 deliberately unpredictable. The high cohesion principle is violated by
-the traditional separation along the siloed entities \textbf{HTML, JS
-and CSS}. Understanding the relation of HTML markup and another CSS file
-adds incidental complexity. Different approaches emerged over the recent
-years to join the forces. Most notable the dedicated frontend language
-\emph{elm}\footnote{\href{http://elm-lang.org/}{elm lang}} where MODEL,
-VIEW and CONTROLLER sit in one place. A browsernative service aims for a
-combinement of the web native trinity HTML, JS and CSS.
-
-Traditional web development relies on \textbf{libraries} to enhance the
-service capabilities of the web platform. Compared to libraries, a
-component service offers multiple advantages for building, deployment
-and shipping. As expressed earlier a \emph{components for the web}, or
-web component, is \textbf{self-contained} which means it embodies all
-needed functionality to get it's job done. Therefore it has a much
-better evolution mechanism in the service contracts. Changing
-functionality won't break other services. A component can progressively
-enhanced which guarantees functionality throughout different versions. A
+the traditional separation along the siloed entities HTML, JS and CSS.
+Understanding the relation of HTML markup towards another CSS file
+generates incidental complexity. Different approaches emerged over the
+recent years to join these entities. A \emph{browsernative service}
+sought for a combination of the web native trinity HTML, JS and CSS.
+
+Traditional web development relies on libraries to enhance the service
+capabilities of the web platform. Compared to libraries a component
+service offers multiple advantages for building, deployment and
+shipping. As expressed earlier \emph{components for the web}, or web
+components, are self-contained which means they embody all needed
+functionality to get their job done. Therefore they have a much better
+evolution mechanism in the service contracts. Changing functionality
+will not break other services. A component can be progressively enhanced
+which guarantees functionality throughout different versions whereas a
 library is only loosely coupled to the implementation and therefore hard
@@
 see websites embodying different versions of the same library to
-guarantee functionality but lowers page performance significantly.
+guarantee functionality which lowers page performance significantly.
 
@@
 and especially the critical first page rendering. Libraries for the
-browsers are traditionally ``shipped'' as non static immediately-invoked
-functions run immediately on page load. Following the \emph{Google RAIL}
-model, a user-centric performance measurement, a page load ought be less
-than 1 second to catch up user's attention.{[}6{]} There are many ways
-for optimizing the critical first render but as a rule of thumb a
-build-in web component might be always superior to libraries in terms of
-first rendering. The fine-grained lifecycle methods which will be
-described in the technical section of this paper give the developer far
-reaching optimization opportunities.
+browsers are traditionally `shipped' as non static immediately-invoked
+functions. Following the Google RAIL model, a user-centric performance
+measurement, a page load ought to take less than 1 second to catch up
+users' attention.{[}7{]} There are many ways to optimize the critical
+first render but as a rule of thumb a build-in web component might be
+always superior to libraries in terms of first rendering. The
+fine-grained lifecycle methods which will be described in the technical
+section of this paper give the developer far reaching optimization
+opportunities.
 
 The last argument in favor of components over libraries is the more
-explicit interface.{[}4{]} While the functionality of a library needs
-documentation to be accessible a component functionality is exposed via
-the components' signature. The markup language HTML is by design
-equipped with an expressive syntax and steering a web component using
-attributes and values should be most straightforward even to
-unexperienced web developers.
+explicit interface.{[}5{]} While the functionality of a library needs
+documentation to be accessible, a component functionality is exposed via
+the components' signature. HTML markup is an expressive syntax and
+therefore convenient for steering a web component using merely
+attributes and values.
 
@@
 designs which are copies of the communication~structures~of these
-organizations''. {[}7{]}
+organizations''. {[}8{]}
 \end{quote}
@@
 Emphasizing the human factor in microservices is a key feature.
-Microservices are a product of real-world usage.{[}3, p. 1{]} Instead of
-splitting team structures along the technology stack (UI Experts
+Microservices are a product of real-world usage.{[}4, p. 1{]} Instead of
+splitting the team along the technology stack (UI Experts
 -\textgreater{} Middleware -\textgreater{} Database) a microservice
-approach model teams around \textbf{business capabilities}.{[}4{]}
+approach models teams around \textbf{business capabilities}.{[}5{]}
 Consequently every team is capable of planning, designing, implementing,
 testing and maintaining their very own microservice. Along the
-technology stack every member gains high competence about the the
-service architecture which can be critical for evolving the service over
-its whole lifecycle.
+technology stack every member gains high competence in the service
+architecture which can be positive for evolving the service over its
+whole lifecycle.
 
 Real-world domains tend to be complex and multifaceted. To unfold their
-complexity, domains can subdivided into \textbf{bounded context}.{[}3,
-p. 31{]} For example, customer service is a business domain but with
-varying bounded context. One context can be sales, another context could
-be support. Every context makes different assumptions about the
-underlying model and draws an explicit interface where it decides what
-to share with other contexts.{[}3, p. 30{]} Evaluating each bounded
-context within each business domain will eventually shape the data
-persistency model likewise the interface to the service. This
-methodology can be iterated over and over again. A sales service for
-example might be evaluated to different sales context resulting in
-differing interfaces on differing devices shaped by browsernative
-microservices. By focusing the service into clear defined business
-boundaries, it is easier to define a smart API of the service.
+complexity, domains can be subdivided into \textbf{bounded
+context}.{[}4, p. 31{]} For example, customer service is a business
+domain but with varying bounded contexts. One context can be sales,
+another context could be support etc. Every context makes different
+assumptions about the underlying model and draws an explicit interface
+where it decides what to share with other contexts.{[}4, p. 30{]}
+Evaluating each bounded context within each business domain will
+eventually shape the data persistency model likewise the interface to
+the service. This methodology can be iterated over and over again. A
+sales service for example might be evaluated to different sales contexts
+resulting in differing interfaces on differing devices shaped by
+browsernative microservices. By dividing the service into clear defined
+business boundaries it becomes easier to define a smart API of the
+service.
 
 Assigning service responsibility to a team, the so called
-\textbf{Definition of Done} (DoD) shifts from ``accomplishing projects''
-to ``accomplishing products''. This new paradigm not only changes the
-administrative overhead like budgeting or resource allocation. It
-creates a kind of responsibility connection from the team to the service
-which can be best described as \textbf{shared governance} model. ``Each
-team collectively share responsibility for evolving the technical vision
-of the system.''{[}3, p. 247{]} Expectedly, those teams are more
-motivated within their very own service and exhibit a more sophisticated
-iteration time.{[}4{]}
-
-For many companies working in the spheres of the internet the client
+\textbf{Definition of Done} shifts from `accomplishing projects' to
+`accomplishing products'. This new paradigm not only changes the
+administrative overhead like budgeting or resource allocation.
+Furthermore, it creates a kind of responsibility connection from the
+team to the service which can be best described as \textbf{shared
+governance} model. ``Each team collectively share responsibility for
+evolving the technical vision of the system.''{[}4, p. 247{]}
+Expectedly, those teams are more motivated within their very own service
+and exhibit a more sophisticated iteration time.{[}5{]}
+
+For many companies working in the spheres of the internet, the client
 side is highly important for their business. In fact, business goals and
 capabilities can be derived from frontend needs. The state of the web is
-not only a story of numerous artifacts, it is also a story of an highly
+not only a story of numerous artifacts, it is also a story of a highly
 fragmented market along devices, operating systems, differing sizes and
@@
 about the technology stack. Splitting teams along the stack results in
-an slow paced back and forth negotiation for every change to be made
-make iteration time expensive. As browser technologies, design
-guidelines and devices change frequently it makes absolutely sense to
-shift responsibility towards the teams altogether.
+an slow paced back and forth negotiation about every change to be made
+can result in prolonged and therefor expensive iteration. As browser
+technologies, design guidelines and devices change rapidly it makes
+sense to shift responsibility towards the team altogether.
 
@@
 can be derived from patterns known from traditional postal services. A
-physical letter has only two smart endpoints entitled to read and
-process the message while packaging likewise the connection itself is
-maximized towards unification. \emph{Smart endpoints and dumb pipes} is
-coined to the approach of designing communication mostly decoupled and
-as cohesive as possible.{[}4{]} Applying this rules to the web platform
+physical letter has only two smart endpoints, entitled to read and
+process the message while packaging as well as the connection itself is
+mostly standardized. \textbf{Smart endpoints and dumb pipes} is coined
+to the approach of designing communication mostly decoupled and as
+cohesive as possible.{[}5{]} Applying these rules to the web platform
 can result into building unified JSON message objects passed along
-``dumb'' middleware components. A typical browser event comes close to
+`dumb' middleware components. A typical browser event comes close to
 this definition and suits arguably well for in-memory communication
@@
 Microservices heavily rely on simple HTTP request-response with resource
-APIs and lightweight messaging.{[}4{]} Newman puts his recommendation on
+APIs and lightweight messaging.{[}5{]} Newman recommends
 technologic-agnostic REST APIs to free data persistence from
-implementation constraints.{[}3, p. 247{]} The advantage of this overall
+implementation constraints.{[}4, p. 247{]} The advantage of this overall
 simple communication model is the suitability for both frontend-backend
 likewise backend-backend communication. A service therefore can evolve
-from an heavy backend with a lot of network roundtrips to a leaner
+from a heavy backend with a lot of network roundtrips to a leaner
 backend seamlessly. The browser build-in \textbf{fetch API} which is
@@
 ensure autonomy. This paper already described fragmented services
-bounded to singular business context choreographed by simple
+bounded to singular business context, choreographed by simple
 communication protocols developed and evolved by autonomous teams.
@@
 stack, tools and services designed in the spirit of language- and
-platform independence and share their knowledge with other
-parties.{[}4{]} In the recent years many big companies like Facebook,
+platform independence and to share their knowledge across other
+parties.{[}5{]} In the recent years many big companies like Facebook,
 Google, Netflix and others followed that spirit and published their
-ideas and implementations open source. The previously mentioned ReactJS
-for example is a product of Facebooks need to ensure a consistent
+ideas and implementations open source. The previously mentioned React
+for example is a product of Facebook's need to ensure a consistent
 frontend experience. In fact, many tools and techniques are byproduct of
-vital interaction of concrete domain problems and their implementations.
-
-The spirit of freedom can't be applied universally to
+vital interaction of specific domain problems and their solutions.
+
+The spirit of freedom cannot be applied universally to
 \emph{browsernative microservices} as the browser and its underlying DOM
-will be the limitation factor to a certain degree. Talking about the
-browser, a reader might be tempted to narrowly thinking of the obvious
-VIEW layer only - which is not true anymore. In the recent years the
-major browser engines grow to to a fully-fledged app deployment platform
-offering connectors to build-in databases, multithreading support and
+will, to a certain degree, be the limiting factor. Talking about the
+browser, a reader might be falsely tempted to narrow his perception
+towards the obvious VIEW layer only. In recent years the major browser
+engines have grown to fully-fledged app deployment platforms offering
+connectors to build-in databases, multithreading support and
 ever-growing JS build-ins like speech synthesis or push notifications.
@@
 achieve a similar look and feel like native apps. And last but not least
-services like NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}}
-effectively compiling ``the web'' to native machine code lowering the
-boundary between native and browser code even further.
+services like NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}
+  \#\# 2.5. Decentralized Data Management} effectively compiling `the
+web' to native machine code lower the boundary between native and
+browser code even further.
 
@@
 critical mass there will be most likely some library support or foreign
-function interface towards ES6 modules (which are mandatory for the new
-specifications).
+function interface towards ES6 modules, which are mandatory in the new
+specifications.
 
@@
 perceived as frontend packages with an HTML interface instead of a JS
-signature. Webcomponents.org\footnote{\href{https://webcomponents.org}{Webcomponents.org}}
-is a registry for ready-to-use components of every scale and purpose
-where for even Google shares a lot of their material design elements.
-
-\subsection{2.5. Decentralized Data
-Management}\label{decentralized-data-management}
-
-Data Management in a microservice follows the same modular philosophy
-like the service implementation. As mentioned earlier different bounded
-contexts make different assumptions of the underlying models. A
-\emph{browsernative microservices} takes this idea even further and
-expands it to the fragmented world of electronic devices. Decentralized
-decisions about conceptual models demand for decentralized data storage
-decisions.{[}4{]} Todays web architectures aims to leverage an
-increasing amout of processing to the client to avoid time-consuming
-roundtrips especially in mobiles networks.\footnote{Latency numbers:
-  https://gist.github.com/jboner/2841832} Since network roundtrips are
-costly it is a good advice to only query as much data as needed and
-cache as much as possible. The build-in LocalStorage or its successor
-IndexedDB are mature persistence technologies and libraries like
-PouchDB\footnote{\href{https://pouchdb.com/}{PouchDB}} even offer
-backend adapters for syncing out of the box.
+signature. Webcomponents.org is a registry for ready-to-use components
+of every scale and purpose where even Google shares a lot of their
+material design elements.
+
+Data Management in a microservice follows the same modular philosophy as
+the service implementation. As mentioned before, different bounded
+contexts make different assumptions of the underlying models.
+Decentralized decisions about conceptual models demand decentralized
+data storage decisions.{[}5{]} Todays web architectures aim to leverage
+an increasing amout of \textbf{processing to the client} to avoid
+time-consuming roundtrips especially in mobile networks.{[}9{]} Since
+network roundtrips are costly it is a good advice to only query as much
+data as needed and cache as much as possible. The build-in LocalStorage
+or its successor IndexedDB are mature persistence technologies and
+libraries like PouchDB\footnote{\href{https://pouchdb.com/}{PouchDB}}
+even offer adapters for syncing to the server out of the box.
 
@@
 emphasizes this pragmatism on APIs in a conference talk. At ING the
-frontend demands tailor the backend APIs, APIs may be produced
-automatically and not even Issa knows how many APIs exists.{[}8{]} They
-are using a pattern called \textbf{backend for frontends} empowering the
-team working to craft their UI and backend in a one-to-one
-relationship.{[}3, p. 72{]}
-
-To see this pattern in the field a reader might have a look at Facebooks
-GraphQL\footnote{\href{http://graphql.org/}{GraphQL}}. GraphQL is a
-query language for the frontend. The backend solely replies on the
-frontend needs. Another well documented example in the field is
+frontend demands tailor the backend APIs. APIs may be produced
+automatically and not even Issa knows how many APIs exists.{[}10{]} At
+ING, they are using a pattern called \textbf{backends for frontends}
+empowering the team to craft their UI and backend in a one-to-one
+relationship.{[}4, p. 72{]}
+
+To see this pattern in the field a reader might have a look at
+Facebook's GraphQL\footnote{\href{http://graphql.org/}{GraphQL}}.
+GraphQL is a query language for the frontend. The backend solely replies
+to the frontend needs. Another well documented example in the field is
 Cognitects Datomic\footnote{\href{http://www.datomic.com/}{Datomic}},
@@
 microservice is designed holistically containing both front- and
-backends. For the sake of simplicity data management won't be explored
-into depth throughout this paper.
+backends. For the sake of simplicity data management will not be
+explored into depth throughout this paper.
 
@@
 Microservices tend to increase complexity as this model adds a sheer
-number of moving parts to the system whereas requires proper
-orchestration.{[}3, p. 246{]} Arguably every more sophisticated web
-developer already came across build tools like Webpack or infrastructure
-automation tools like Gulp. Testing and deploying web components
-shouldn't be an obstacle in development.
-
-In the global nature of web development the development couldn't
+number of moving parts to the system which requires proper
+orchestration.{[}4, p. 246{]} Arguably every sophisticated web developer
+already came across build tools like Webpack or infrastructure
+automation tools like Gulp. Therefore, testing and deploying web
+components should not be an obstacle in development.
+
+In the global nature of web development the development could not be
 completely decoupled from the production environment. This circumstance
@@
 \textbf{immediate feedback principle} for developing user
-interfaces.\footnote{\href{https://vimeo.com/36579366}{Bret Victor -
-  Inventing on Principle}} In his talk he emphasizes the importance of
-an immediate connection between the creator of a product and product
-itself. Any change must results in an immediate visible feedback. Web
+interfaces.{[}11{]} In his talk he emphasizes the importance of an
+immediate connection between the creator of a product and the product
+itself. Any change must result in an immediate visible feedback. Web
 components catch up with this principle as they allow isolated
-development within a single file containing all bits and pieces of the
-web component. Every major browser devtool offers a direct file
-manipulation functionality so development can be even in place.
-
-When it comes to standardized deployment guidelines previously mentioned
-Ben Issa, described the ING standard workflow. Every component comes in
-its own \textbf{git repo} containing:
+development within a single file containing all bits and pieces of a
+single service. Every major browser devtool offers a direct file
+manipulation functionality, so development can be even in place.
+
+When it comes to standardized development previously mentioned, Ben Issa
+described the ING standard workflow as follows. Every component is
+packed in its own \textbf{git repo} containing:
 
@@
 Even though this example is an opinionated perception it gives a sense
-of a mature component build for the web. This example should made clear
-that all parts of the component put together in one place. Every
-check-in is handled as release candidate and can be independently tested
-and deployed by a fully automated machinery.{[}8{]} Due to an exhaustive
-amount of testing and deployment tools for JS an automated
-infrastructure shouldn't be a problem.
+of a mature component built for the web. This example should illustrate
+that all parts of the component put together in one place as well as
+tests, demos and blueprints are part of the component from day one.
+Every check-in is handled as a release candidate and can be
+independently tested and deployed by a fully automated
+machinery.{[}10{]} Due to an exhaustive amount of testing and deployment
+tools for JS an automated infrastructure should not be an obstacle.
 
@@
 
-In theory a microservice is designed with focus on monitoring for both
-the architectural elements and business relevant metrics.{[}4{]} Due to
-the modular structure weak points can occur in the orchestration of the
+In theory a microservice is designed with focus on monitoring of both
+the architectural elements and business relevant metrics.{[}5{]} Due to
+the modular structure, weak points can occur in the orchestration of the
 services. A microservice should track down every communication flow and
-provide defaults and meaningful error messages where communication
+provide defaults and meaningful error messages where communication might
 stuck. Testing every single component with predefined synthetic events
@@
 
-Combinment of the resources in the browser always demanded for
+Combination of different resources in the browser always demand
 optimization to avoid unexpected side-effects like \emph{flash of
-unstyled content}. Googles Polymer propagates the a general-purpose
-pattern called \textbf{PRLP}\footnote{\href{https://www.polymer-project.org/1.0/toolbox/server}{PRLP
+unstyled content}. Googles Polymer propagates a general-purpose pattern
+called \textbf{PRLP}\footnote{\href{https://www.polymer-project.org/1.0/toolbox/server}{PRLP
   pattern}}:
@@
 solution instead of the latest browser optimized version. After the
-initial paint, critical resources like top-level microservices or other
-app logic can be loaded and registered.
-
-Regarding the evolution of the web, the ``next billion'' internet users
-most likely using Android, have decent specs mobile phones, use an
-evergreen browser but won't have a reliable internet connection.{[}9{]}
+initial paint, critical resources like top-level microservices or
+related parts can be loaded and registered.
+
+Regarding the evolution of the web, the `next billion' internet users
+will most likely use Android, have decent specs mobile phones, use an
+evergreen browser but won't have a reliable internet connection.{[}12{]}
 While \textbf{Progressive Enhancement} was once related to build
-websites both for browsers with and without JS support the demands
-changed tends towards an \textbf{offline first} principle avoiding
-network connectivity failures.{[}9{]} A \emph{browsernative
-microservice} therefore not only tries to cache data as much as
-possible, it should also bring in a lot of program logic as described in
-the previous chapters.
-
-\subsection{2.8. Evolutionary Design}\label{evolutionary-design}
+websites both for browsers with and without JS support, the demands have
+changed towards an \textbf{offline first} principle avoiding network
+connectivity failures.{[}12{]} A \emph{browsernative microservice}
+therefore not only tries to cache data as much as possible, it should
+also bring in a lot of program logic as described in the previous
+chapters.
+
+\subsection{2.8. Evolutionary design}\label{evolutionary-design}
 
 Microservices tend to become smaller over time. An evolutionary design
-approach puts emphasizes on decomposition and scrapping the service.
-``The key property of a component is the notion of independent
-replacement and upgradeability.''{[}4{]} Therefore we can safely change
-and chop services. Lazy components of the system which won't change
-often should be separated from parts undergoing a lot of churn.{[}4{]}
-And services which change for the same reason might be moved together or
+approach emphasizes decomposition and scrapping the service. ``The key
+property of a component is the notion of independent replacement and
+upgradeability.''{[}5{]} Therefore we can safely change and chop
+services. Lazy components of the system which will not change often
+should be separated from parts undergoing a lot of churn.{[}5{]}
+Services which change for the same reason might be moved together or
 even could be merged.
 
-Pursuing flexibility for web development is a selling point as
-innovation cycles for browser development is fast paced and technologies
-can change quickly. Frontend related hardware, software and
-methodologies innovate rapidly over time.
+Pursuing flexibility in web development is a selling point as innovation
+cycles in browser development are fast and technologies can change
+quickly. Frontend related hardware, software and methodologies innovate
+rapidly over time.
 
 \emph{Browsernative microservices} should be perceived as complementary
-technology in contrast to full-service frameworks like Angular. Being
-native technology pursues a strong interopt approach with existing
-systems. Andrew Rota for example came up with a pattern using small,
-encapsulated and stateless web components as leaves in the tree of React
-components instead of native HTML elements.{[}10{]} Even React can
-eventually profit from the expressiveness of custom components using a
-custom \texttt{\textless{}meaningful-button\textgreater{}} over a native
-\texttt{\textless{}button\textgreater{}}. Most likely there will be
-always some cutting edge framework promising advantages over native
-code. Whatever new framework will be on the rise within the next years
-native components can eliminate future uncertainty and allowing rapid
+technology in contrast to full-service frameworks like Angular. Being a
+native technology, they pursue a strong interoperability approach to
+existing systems. Andrew Rota, for example, came up with a pattern using
+small, encapsulated and stateless web components as leaves in the tree
+of React components instead of native HTML elements.{[}13{]} Even React
+can eventually profit from the expressiveness of custom components using
+a custom \texttt{\textless{}meaningful-button\textgreater{}} over a
+native \texttt{\textless{}button\textgreater{}}. Most likely there will
+always be some cutting edge framework promising advantages over native
+code. Whatever new framework will be on the rise within the next years,
+native components can eliminate future uncertainty, allowing rapid
 reassembling towards new architectures.
@@
 
-For building a native microservice running on the ``bare-metal'' browser
+Building a native microservice running on the `bare-metal' browser
 engine requires a bunch of new specifications and assumptions. Most
-importantly the quasi specification \textbf{Web Components} is needed.
-Web Components is not a real standard. It's an amalgam of APIs from
-multiple w3c specs which can be used independently. A web developer may
-choose one spec and embrace the freedom in architecture which can be
-combined with other frameworks/libraries.
+importantly multiple \textbf{Web Components} specifications are needed.
+Web Components is not a single standard. They are a kind of amalgam of
+combined APIs from multiple w3c specs which can be used independently. A
+web developer may choose one spec which can be combined with other
+frameworks.
 
 Depending on the context, some people argue for only two specs which
-essentially make it possible to create a scoped component but not caring
-too much on it's distribution{[}11{]}. Some people prefer the three
-specs {[}12{]}, but the majority advocating the four specs variant,
-which is listed on the quasi-official
+essentially make it possible to create a scoped component but do not
+care about its distribution{[}14{]}. Some people prefer three specs
+{[}15{]}, but the majority advocate the four specs variant, which is
+listed on the quasi-official
 \href{http://webcomponents.org}{webcomponents.org} website. For the
-purpose of this article, all four specs will be discussed briefly to
-provide a rough understanding. It is not meant to cover all bits and
-pieces.
+purpose of this article, the four specs variant will be discussed
+briefly to provide a rough understanding.
 
 \emph{Disclaimer:} This paper introduces many new browser build-ins with
-the focus on testability. As the time of writing, many examples can be
-tried frictionless in the console of the latest versions of
-\textbf{Google Chrome, Opera and Apple Safari}.\footnote{\href{http://jonrimmer.github.io/are-we-componentized-yet/}{Are
-  we componentized yet?}} On Mozilla Firefox technologies work behind a
-flag and Microsoft Edge implementation is unfortunately far behind. But
-Browser implementation changes quickly and soon technology adoption
-won't be an issue. Meanwhile new standards can be used through
-\textbf{polyfills} even on legacy browsers.
-
-\subsection{\texorpdfstring{3.1. Custom Elements
-\href{https://html.spec.whatwg.org/multipage/scripting.html\#custom-elements}{(whatwg)}}{3.1. Custom Elements (whatwg)}}\label{custom-elements-whatwg}
+the focus on accessibility. At the time of writing, all examples can be
+tested in the console of the latest versions of \textbf{Google Chrome,
+Opera and Apple Safari}.\footnote{\href{http://jonrimmer.github.io/are-we-componentized-yet/}{Are
+  we componentized yet?}} On Mozilla Firefox they could be manually
+enabled. However, browser implementation changes quickly and soon
+technology adoption will not be an issue in all major browsers.
+Meanwhile all new standards can be used through \textbf{polyfills} even
+on legacy browsers.
+
+\subsection{\texorpdfstring{3.1. Custom elements
+\href{https://html.spec.whatwg.org/multipage/scripting.html\#custom-elements}{(whatwg)}}{3.1. Custom elements (whatwg)}}\label{custom-elements-whatwg}
 
@@
 This example should be almost self-explanatory in functionality.
-\emph{Custom elements} come in the fashion of \textbf{ES6
-Classes}\footnote{\href{https://developer.mozilla.org/en/docs/Web/JavaScript/Reference/Classes}{JavaScript
-  Classes}} in favor of the normal JavaScript prototype-based
-inheritance model which was part of an older specification. Every valid
-element must extend the base \texttt{HTMLElement} interface
-which~``ensures the newly created element inherits the entire DOM API
-and any properties/methods that you add to the class become part of the
-element's DOM interface.''{[}13{]} Like any other ES6 class any
-\emph{custom element} can be specialized further using the typical
-inheritance model allowing higher levels of abstraction.
+\emph{Custom elements} come in the fashion of ES6 Classes in favor of
+the JS prototype-based inheritance model which was part of an older
+specification. Every valid element must \textbf{extend the base
+\texttt{HTMLElement} interface} which~``ensures the newly created
+element inherits the entire DOM API and any properties/methods that you
+add to the class become part of the element's DOM interface.''{[}16{]}
+Like any other ES6 class any \emph{Custom element} can be specialized
+further using the typical inheritance model allowing higher levels of
+abstraction.
 
@@
 After declaration the new HTML element needs to be registered in the
-global build-in \texttt{customElements} object with an dedicated tag
-name acting as key to the element. Mind the dash inside the tag name to
+global build-in \texttt{customElements} object with a dedicated tag name
+acting as key to the element. Mind the dash inside the tag name to
 conform the spec. Finally, the new element can be mounted inside the
@@
 In addition to the constructor which runs procedures on initialization,
-the spec defines \textbf{lifecycle callbacks} for controlling elements
+the spec defines \textbf{lifecycle callbacks} for controlling elements'
 behavior towards DOM interaction. Many popular frameworks like React or
@@
   connectedCallback()\\
-  Called upon the time of \textbf{connecting or upgrading the node}
+  called upon the time of \textbf{connecting or upgrading the node}
   which means the moment the node is rendered inside the DOM. Typically
-  this method is called straight after the constructor if the node is
+  this method is called straight after the constructor, if the node is
   inserted directly. For a faster initial render of the page it is
@@
   method contains setup code such as fetching resources or rendering
-  elements according to attributes.{[}13{]}
+  elements according to attributes.{[}16{]}
 \item
   disconnectedCallback()\\
-  Called upon the time of \textbf{node removal}. Cleanup code like
+  called upon the time of \textbf{node removal}. Cleanup code like
   removing event listeners or disconnecting web sockets can be put here.
@@
   attributeChangedCallback(attrName, oldVal, newVal)\\
-  This method provides an \textbf{onchange handler} for certain elements
-  attributes. This method is used to guide elements' transition from an
-  old value to a new state. Due to performance issues this callback is
-  only triggered for attributes registered in a dedicated array shipped
-  with the element.
+  This method provides an \textbf{onchange handler} for certain
+  elements' attributes. This method is used to guide elements'
+  transition from an old value to a new state. Due to performance issues
+  this callback is only triggered for attributes registered in a
+  dedicated array shipped with the element.
 \item
   adoptedCallback()\\
-  Called when moving the node \textbf{between documents}. This method
-  comes handy when using HTML imports described later.
+  called when moving the node \textbf{between documents}. This method
+  comes handy when using HTML Imports described later.
 \end{itemize}
@@
 
-As mentioned earlier any custom elements must extend the
+As mentioned before any custom element must extend the
 \texttt{HTMLElement} interface ensuring base properties and methods used
-in throughout all HTML elements like \textbf{id, class, addEventListner
-\ldots{}}. Additionally, it is possible to define custom attributes
-using the \emph{custom elements'} \textbf{getter / setter interface} to
-steer the behavior of the element. Note that the keywords \textbf{get}
-and \textbf{set} as well as the previously used \textbf{constructor} are
-optional!
+throughout all HTML elements like id, class, addEventListner etc.
+Additionally, it is possible to define custom attributes using the
+\emph{custom elements'} \textbf{getter / setter interface} to steer the
+behavior of the element. Note that the get/set keywords as well as the
+previously used constructor are optional!
 
@@
 Native DOM properties always reflect their values between HTML and
-JS.{[}14, Para. 2.6.1{]} Declaring
+JS.{[}17, Para. 2.6.1{]} Declaring
 \texttt{\textless{}hello-world\ id="hello"\textgreater{}} equals to the
@@
 
-This behavior won't work out-of-the-box with methods defined with
-setters as they are strictly JS only . Mounting
-\texttt{\textless{}hello-world\ sayhello="mars"\textgreater{}} would't
-call the \texttt{sayhello} method in the previous setup. Value
-reflection can be implemented inside the \emph{custom elements} using
+This behavior will not work out-of-the-box with methods defined by
+setters as they are strictly JS. Mounting
+\texttt{\textless{}hello-world\ sayhello="mars"\textgreater{}} would not
+result in calling the \texttt{sayhello} method in the previous setup.
+Value reflection can be implemented inside \emph{custom elements} using
 the native methods \texttt{getAttributes} and \texttt{setAttributes}.
-Using them exhaustively throughout lifecycles methods the new components
-can configured to read and listen to HTML attributes accordingly.
+Using them exhaustively throughout lifecycle methods the new components
+can be configured to read and listen to HTML attributes accordingly.
 
@@
 
-One aspect didn't mentioned yet is the possibility of extending other
+One aspect not mentioned yet is the possibility of extending other
 build-in elements by extending other interfaces instead of the
@@
 spec'd it is strongly rejected by some browser vendors.\footnote{https://github.com/w3c/webcomponents/issues/509}
-Most likely the spec will change in future in one or other way on this
-issue and therefore customized build-in elements left out of this paper
-intentionally.
+Most likely the spec will change in future in one or other direction
+concerning this issue and therefore customized build-in elements are
+left out in this paper intentionally.
 
@@
 
-A \emph{shadow DOM} is just an isolated DOM tree living inside an
-another DOM tree. The spec refers the hosting tree as \textbf{light DOM
-tree} and the attached DOM as \textbf{shadow DOM tree}. Conceptually
-\emph{shadow DOM} issues a single important issue for building scalable
-software which is namely \textbf{encapsulation}. While custom elements
-provide a good way to encapsulate JS behavior \emph{shadow DOM} tends
-strongly to the direction of style and event encapsulation.
+A \emph{shadow DOM} is just an isolated DOM tree living inside another
+DOM tree. The spec refers the hosting tree as \textbf{light DOM tree}
+and the attached DOM as \textbf{shadow DOM tree}. Conceptually
+\emph{shadow DOM} issues a single important function for building
+scalable software which is namely \textbf{encapsulation}. While custom
+elements provide a good way to encapsulate JS behavior \emph{shadow DOM}
+tends strongly to the direction of style and event encapsulation.
 
@@
 organization and leads over times to highly fragmented bits of CSS and
-obscured CSS selectors. Of course this situation lowers code clarity and
-reusability dramatically. The only solution which won't break with the
-existing global paradigm of the DOM effectively is to allow separate
-pieces of encapsulated code sit on top of the global DOM - introducing
-the shadowed DOM approach.
+obscured CSS selectors. Of course this situation dramatically lowers
+code clarity and reusability. The only solution which will not break
+with the existing global paradigm of the DOM is to allow separate pieces
+of encapsulated code sit on top of the global DOM - introducing the
+shadowed DOM approach.
 
 Enhancing the previous example the new encapsulated \texttt{HelloWorld}
-would like this:
+would look like the following code snippet:
 
@@
 the \texttt{HelloWorld} which has the same properties as a normal, light
-document object. Note that the \texttt{shadowRoot} object is marked as
-\textbf{open} which ensures that some events can bubble out and outside
-JS can reach in the new root.
+document object. Note that the \texttt{shadowRoot} object is
+\textbf{marked as open} which ensures that some events can bubble out
+and outer JS can capture the new root.
 
 Filling the \emph{shadow DOM} with an \texttt{innerHTML} string is
-rather impractical. To fill a \emph{shadow DOM} with life usually it
-invites light DOM child nodes nested under the hosting node using a
+rather impractical. To fill a \emph{shadow DOM} with life, it usually
+pulls light DOM child nodes nested under the hosting node using a
 technique called \texttt{slots}.
@@
 
-Contradicting to the simplified \texttt{HelloWorld} example, a
-\emph{shadow DOM} shouldn't contain dynamic content. Changing or
-interacting with the paragraph node from the example would require
-nested JS calls querying the hosting node, entering the \emph{shadow
-DOM} and applying a function. Imported JS behavior from third-party
-libraries in the \emph{light DOM} can't be used inside the \emph{shadow
-DOM}, too.
-
-That's why the shadowed documents should be more perceived as
-\textbf{static documents} filled and managed solely by the render
-engine. \texttt{Slots} are target areas for \emph{light DOM} nodes used
-to mark the endpoints in question.
+Contradicting the simplified \texttt{HelloWorld} example, a \emph{shadow
+DOM} should not contain dynamic content. Changing or interacting with
+the paragraph node from the previous example would require nested JS
+calls querying the hosting node, entering the \emph{shadow DOM} and
+applying a function. Imported JS behavior from third-party libraries in
+the \emph{light DOM} cannot easily reach inside the \emph{shadow DOM},
+too.
+
+This is the reason why the shadowed document root should rather be
+perceived as \textbf{static document} filled and managed solely by the
+render engine. \texttt{Slots} are target areas for \emph{light DOM}
+nodes used to mark the endpoints in question.
 
@@
 
-Technically, the \emph{light DOM} nodes are not moved inside the
-\emph{shadow DOM}. Their just rendered in place. It's an subtle but
-important difference towards handling a node. All JS behavior and CSS
-styles applied in the \emph{light DOM} will be valid in the \emph{shadow
-DOM}. The render engine literally taking the nodes and drop them inside
-the \texttt{slot} tag. This procedure is commonly referred as
-\textbf{flattening} of the DOM trees.
-
-It's possible to add semantics to the \emph{shadow DOM} in naming the
-slots which frees the \emph{light DOM} from the responsiblity to deliver
+From a technical perspective, the \emph{light DOM} nodes are not moved
+inside the \emph{shadow DOM}. They are just rendered in place. This is a
+subtle but important difference towards handling a node. All JS behavior
+and CSS styles applied in the \emph{light DOM} will be valid in the
+\emph{shadow DOM}. The render engine literally takes the nodes and drops
+them inside the \texttt{slot} tag. This procedure is commonly referred
+as \textbf{flattening} of the DOM trees.
+
+It is possible to add semantics to the \emph{shadow DOM} in naming the
+slots which free the \emph{light DOM} from the responsiblity to deliver
 nodes in a correct top-to-bottom order. Combining \emph{shadow DOM} with
-HTML templates provides the web developer with a flexible \textbf{HTML
+HTML templates equips the web developer with a flexible \textbf{HTML
 template engine}.
@@
 \texttt{\textless{}slot\textgreater{}} tag is considered as a good
-practice as it provides the developer with visual clues what nodes must
-be delivered. This functionality makes a \emph{shadow DOM} pretty much
-self-explanatory. Inside a default slot tag the render engine expands
-all \emph{light DOM} children without a named \texttt{slot} attribution.
+practice as it provides the developer with visual clues about what nodes
+must be delivered. This functionality makes a \emph{shadow DOM} pretty
+much self-explanatory. Inside a default slot tag the render engine
+expands all \emph{light DOM} children without a named \texttt{slot}
+attribution.
 
@@
 
-As mentioned in the last section, there is a distinct difference about
-the nature of nodes. Nodes declared and rendered exclusively in the
-\emph{shadow DOM} are not affected by any styling from outside. Nodes
-which are distributed will be styled in the \emph{light DOM} and can be
+As mentioned in the last section, there is a distinct difference between
+nodes. Nodes declared and rendered exclusively in the \emph{shadow DOM}
+are not affected by any styling from outside. Nodes which are
+distributed will be styled in the \emph{light DOM} and can be
 additionally painted in the \emph{shadow DOM}.
 
-Note that styles from the outside have an higher specify than styles
+Note that styles from the outside have an higher specificity than styles
 assigned after distribution. Therefore it is generally a good advice to
-minimize the global stylings to some base styling for uniformity of the
+minimize the global stylings to some base stylings for uniformity of the
 web site while leaving the specific stylings to the component. It is
@@
 
-Regarding the importance style encapsulation, a couple of \textbf{new
+Regarding the importance of style encapsulation, a couple of \textbf{new
 CSS rules} emerged that are exclusively targeting the \emph{shadow DOM}.
-The table below outlines styling possibilities for the use INSIDE the
+The points below outline styling possibilities for the use INSIDE the
 \emph{shadow DOM}:
@@
   ::slotted(selector)\\
-  Applies to distributed nodes and repaints them after distribution.
-  Slotted won't override outsides styles but can complement them
+  applies to distributed nodes and repaints them after distribution.
+  Slotted will not override outsides styles but can complement
   previously unset style rules.
@@
 \texttt{host-context()} allows the creation of context-aware custom
-elements. A possible usecase would be ``theming'' a component.
+elements. A possible usecase would be `theming' a component.
 
@@
 
-As mentioned earlier any logic applied to \emph{light DOM} nodes stays
-with the node even after redistribution. For the sake of separation of
-concerns the business logic should be part of the \emph{custom element}
-(the \emph{light DOM}) and not the part of the \emph{shadow DOM}. On the
+As mentioned before, any logic applied to \emph{light DOM} nodes stays
+within the node even after redistribution. For the sake of separation of
+concerns the business logic should be part of the custom element (the
+\emph{light DOM}) and not the part of the \emph{shadow DOM}. On the
 other hand there are numerous scenarios where JS is used for styling or
 animation of an element. In this case it might be more straightforward
-to \textbf{apply JS inside the \emph{shadow DOM}} to avoid mixing with
-logic handlers with styling.
+to \textbf{apply JS inside the \emph{shadow DOM}} to avoid mixing logic
+handlers with styling.
 
@@
 node stays in the normal document. To reach a distributed node from the
-\emph{shadow DOM} to apply some JS behavior for styling it needs the
+\emph{shadow DOM} to apply some JS behavior for styling, it needs the
 extra way over the slot node. Calling \texttt{assinedNodes()} on the
@@
 
-Wrapping up this section \emph{shadow DOM} provides a non-hacky way to
+Wrapping up this section, \emph{shadow DOM} provides a non-hacky way to
 create uniform looking custom elements and even enhance styling
 possibilities without adding overhead. For small components with just a
-little styling a \emph{shadow DOM} might be over engineered. Eventually
-it all depends on the question of ``how hard is it to implement it
-without shadow DOM'' - which can't be answered universally. For a more
-in-depth guide, Google Engineer Eric Bidelman wroten a great primer on
-\emph{shadow DOM}{[}15{]}.
+little styling, \emph{shadow DOM} might be over engineered. Eventually
+it all depends on the question of `how hard is it to implement it
+without shadow DOM' - which cannot be answered in general. For a more
+in-depth guide, Google Engineer Eric Bidelman wrote a primer on
+\emph{shadow DOM}{[}18{]}.
 
 There is still a missing link between \emph{light DOM} and \emph{shadow
-DOM}. The observant reader may have already noticed the weak point in
-the HelloWorld example: how to ``vitalize'' the \emph{shadow DOM} with
-slot or other element structuring instead of \texttt{innerHTML} strings.
-While strings works perfectly fine in this simple case a string of
-markup is rather cumbersome and error-prone and doesn't scale well. When
-putting quotes inside other quotes things break quickly. Strings make
+DOM}. The attentive reader may have already noticed the weak point in
+the HelloWorld example: how to `vitalize' the \emph{shadow DOM} with
+slot tags or other element structuring instead of markup strings. While
+strings work perfectly fine in this simple case a string of markup is
+rather cumbersome and error-prone and does not scale well. When putting
+quotes inside other quotes things break quickly. Strings make
 development harder because code editor features like indentation or
-syntax highlighting won't be supported. The HTML templates is set to
+syntax highlighting will not be supported. The HTML templates are set to
 fill this gap.
 
-\subsection{\texorpdfstring{3.3. HTML Templates
-\href{https://html.spec.whatwg.org/multipage/scripting.html\#the-template-element}{(whatwg)}}{3.3. HTML Templates (whatwg)}}\label{html-templates-whatwg}
+\subsection{\texorpdfstring{3.3. HTML templates
+\href{https://html.spec.whatwg.org/multipage/scripting.html\#the-template-element}{(whatwg)}}{3.3. HTML templates (whatwg)}}\label{html-templates-whatwg}
 
@@
 
-One core concept in templates is browser performance. Elements inside a
+A core concept in templates is browser performance. Elements inside a
 \texttt{template} tag will be parsed on runtime - but not constructed
-and rendered into the content tree. They're remaining plain HTML markup
-sitting in the document until the time of activation.
+and rendered into the content tree. They are remaining plain HTML
+markups sitting in the document until the time of activation.
 
@@
 
-As easy and minimal \emph{HTML templates} are they're missing out a
-crucial feature other template implementations used to have. As
-templates are basically just containers for HTML markup there is no
-idiomatic way to define \textbf{placeholders} for dynamic content.
-Templates could be mock up this way using JS for altering the content
-but the much cleaner way leverages the previously described slot
-technique from shadow DOM.
+As easy and minimal as \emph{HTML templates} are, they skip a crucial
+feature other template implementations used to have. As template tags
+are basically just containers for HTML markup there is no idiomatic way
+to define \textbf{placeholders} for dynamic content. Templates could be
+mocked up this way using JS for altering the content but a much cleaner
+way leverages the previously described slot technique from shadow DOM.
 
@@
 The updated \texttt{HelloWorld} component looks already pretty mature.
-It combines all the previous mentioned standards into one HTML file.
-Custom elements serves the logic, shadow DOM scopes the styles and
-\emph{HTML Templates} efficiently glues light DOM and shadow DOM
+It combines all the previously mentioned standards into one HTML file.
+Custom elements serve the logic, shadow DOM scopes the styles and
+\emph{HTML templates} efficiently glue light DOM and shadow DOM
 together. This separation of concerns comes with a surplus in
 flexibility. In a real world scenario \texttt{HelloWorld} could
-reference multiple \emph{HTML Templates} and switch them around without
+reference multiple \emph{HTML templates} and switch them around without
 any fuss. Even further a developer might split up templates into named
-\textbf{STYLE} templates and \textbf{CONTENT} templates to increase
-reusability even further.
-
-The last standard in the row of four is not concerned with the internals
-implementation of a web component. HTML Imports serves the need for an
+\textbf{STYLE templates and CONTENT templates} to increase reusability
+even further.
+
+The last standard in the row of four is not concerned with the
+implementation of a web component. HTML imports serves the need for an
 efficient distribution mechanism of components and other HTML resources.
 
-\subsection{\texorpdfstring{3.4. HTML Imports
-\href{https://www.w3.org/TR/html-imports/}{(w3c)}}{3.4. HTML Imports (w3c)}}\label{html-imports-w3c}
+\subsection{\texorpdfstring{3.4. HTML imports
+\href{https://www.w3.org/TR/html-imports/}{(w3c)}}{3.4. HTML imports (w3c)}}\label{html-imports-w3c}
 
@@
 
-The \texttt{async} flag is optional but recommended like any other
-fetching event. Once the imported HTML document comes into scope
+The \texttt{async} flag is optional but recommended like in any other
+fetching event. Once the imported HTML document comes into scope,
 activation follows a very similar process compared to the aforementioned
@@
   \textbf{Parsing the content and preparing the render}\\
-  const content = linknode.import; -\textgreater{} Unlike the \emph{HTML
-  template} a complete document object is constructed.
+  const content = linknode.import; -\textgreater{} Contrary to HTML
+  templates a complete document object is constructed.
 \item
@@
 
-Again this is the imperative way to handle a generic \emph{HTML Import}.
+Again this is the imperative way to handle a generic \emph{HTML import}.
 In the declarative world of web components a component is
-\textbf{parsed, auto-activated and anchored} solely by its' tag name
-\texttt{\textless{}hello-world\textgreater{}\textless{}/hello-world\textgreater{}}
-on purpose. The very own lifecycle method \textbf{adoptedCallback()} in
-custom elements shows the strong interconnection between those
-standards.
-
-\emph{HTML imports} can import everything which is wrappable with HTML
-markup. Stylesheets, scripts, documents, media files and even further
-imports statements can form a semantic \emph{HTML import} statement. The
-far reaching possibilities of a single standard has its downsides.
+\textbf{parsed, auto-activated and anchored} solely by its tag name
+\texttt{\textless{}hello-world\textgreater{}} on purpose. The very own
+lifecycle method \textbf{adoptedCallback()} in custom elements shows the
+strong interconnection between those standards.
+
+\emph{HTML imports} can import everything wrappable in HTML markup.
+Stylesheets, scripts, documents, media files and even further import
+statements can form a semantic \emph{HTML import} statement.
+
+The far reaching possibilities of a single standard has its drawbacks.
 According to Mozilla \emph{HTML imports} are not fully compatible with
 the dependency model of the new ES6 modules.\footnote{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}
-Under the current situation Mozilla and Apple and Microsoft won't
-implement \emph{HTML Imports} soon if any. Currently only Googles blink
-web engine supports \emph{HTML Imports} as they are the driving force
-behind the web components specs in general. Despite the discrepancies
-among browser vendors \emph{HTML Imports} are part of this paper as no
-other native standard can ship bundled HTML, CSS and JS which is a core
-concept behind web components.
-
-\subsection{\texorpdfstring{3.5. Custom Events
-\href{https://dom.spec.whatwg.org/\#interface-customevent}{(whatwg)}}{3.5. Custom Events (whatwg)}}\label{custom-events-whatwg}
-
-Events are first-class citizens in the browser. After all they provide a
-neat communication channel for dynamic interactions. \emph{Custom
-Events} are part of the DOM since years but with the rise of web
-components they will most likely become an indispensable building block
-of web components.
+In the current situation Mozilla, Apple and Microsoft will not implement
+\emph{HTML imports} soon if any. Only Google's blink web engine supports
+\emph{HTML imports} as they are the driving force behind the web
+components specs in general. Despite the discrepancies among browser
+vendors, \emph{HTML imports} are part of this paper as no other native
+standard can ship bundled HTML, CSS and JS which is a core concept
+behind web components.
+
+\subsection{\texorpdfstring{3.5. CustomEvent
+\href{https://dom.spec.whatwg.org/\#interface-customevent}{(whatwg)}}{3.5. CustomEvent (whatwg)}}\label{customevent-whatwg}
+
+Events are first-class citizens in the browser providing a neat
+communication channel for dynamic interactions. \emph{CustomEvent} is
+part of the DOM since years but with the rise of web components it will
+most likely become an indispensable building block of web components.
 
@@
 
-Naming events after the emitting tag makes the API almost
-self-explanatory. The \texttt{detail} property can be loaded with
+Naming events after the emitting tag makes the \textbf{API almost
+self-explanatory}. The \texttt{detail} property can be loaded with
 scalars as well as objects. Subsequent parent nodes may catch the custom
@@
 Elements section at 3.1.3. , the pattern of \textbf{extending native
-elements} should be somewhat dismissed as other browser vendors may
-never implement it. A common workaround to eventually create a own
-version a button could be made with a thin wrapper around the native
-button and chaining a \emph{Custom Event} after the click event.
-
-Another common web components use case can be a \textbf{middleware}
-subscribing to certain child events and acting upon them. This involves
-catching events, buffering, destructuring and creation of an own event
-towards the document root. By design, events only bubble upstream
-towards parent nodes. For handing down information towards the child
-nodes we need to query the child node directly.
-
-\subsection{\texorpdfstring{3.6. Web Worker
-\href{https://html.spec.whatwg.org/multipage/workers.html}{(whatwg)}}{3.6. Web Worker (whatwg)}}\label{web-worker-whatwg}
-
-Like \emph{Custom Events}, \emph{Web Workers} had been around for a long
-time and therefore enjoy full support among major browsers. They emerged
-at around 2009 when discussions about browser performance was still in
-the early days but addressing fundamental performance bottleneck of the
-JS language.
+elements} should be dismissed as certain browser vendors imposed
+distaste towards this functionality. A common workaround to an eventual
+creation of an extended native element could be made with a thin wrapper
+around the native element and chaining a \emph{CustomEvent} after the
+click event.
+
+Another common web components' use case can be a \textbf{middleware}
+subscribing to certain child events and acting upon them. For example
+embedding another third-party web component or widget which can be
+wrapped in a middleware component to align it to system conventions.
+This involves catching events, buffering, destructuring and creation of
+own events. By design, events only bubble upstream towards parent nodes.
+For handing down information towards the child nodes we need to query
+the child node directly.
+
+\subsection{\texorpdfstring{3.6. Web worker
+\href{https://html.spec.whatwg.org/multipage/workers.html}{(whatwg)}}{3.6. Web worker (whatwg)}}\label{web-worker-whatwg}
+
+Like CustomEvent, \emph{web workers} have been around for a long time
+and therefore enjoy full browser support. They emerged at around 2009
+when discussions about browser performance were still in the early days.
+\emph{Web workers} however addressed a fundamental performance
+bottleneck of the JS language.
 
 JS runs in a single-threaded language environment. Every script in the
-browser environment, from handling UI events to query and process larget
-amounts of data or manipulating the DOM runs on a single thread{[}16{]}.
-Putting a lot of work to the single main thread can slow down the web
+browser environment, from handling UI events to query and process large
+amounts of data or manipulating the DOM runs on a single thread{[}19{]}.
+Putting a lot of work into the single main thread can slow down the web
 service significantly. From time to time scripts can block or fail for
@@
 \textbf{background threads} which allows the UI to stay responsive even
-when computation-heavy tasks needs to perform. Furthermore, a worker
-thread adds a performance advantage embracing the multi core CPU
-architecture most devices running on today. To grasp the full potential
-of workers a reader might dive deeper into the Angular 2 architecture,
-where most of the application layer is abstracted from the main
-rendering thread into worker threads.\footnote{\href{https://docs.google.com/document/d/1M9FmT05Q6qpsjgvH1XvCm840yn2eWEg0PMskSQz7k4E}{Angular
+when computation-heavy tasks need to perform. Furthermore, a worker adds
+a performance advantage embracing the multi core CPU architecture most
+devices are running on today. To grasp the full potential of workers, a
+reader might dive deeper into the Angular 2 architecture, where most of
+the application layer is abstracted from the main rendering thread into
+worker threads.\footnote{\href{https://docs.google.com/document/d/1M9FmT05Q6qpsjgvH1XvCm840yn2eWEg0PMskSQz7k4E}{Angular
   2 Rendering Architecture}}
-
-A \emph{Web Worker} spawns a new background thread where scripts can run
-concurrent to the main thread. Usually a worker is loaded from dedicated
-file to embrace separation.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{const} \NormalTok{worker }\OperatorTok{=} \KeywordTok{new} \AttributeTok{Worker}\NormalTok{(}\StringTok{'worker.js'}\NormalTok{)}\OperatorTok{;}
-\end{Highlighting}
-\end{Shaded}
-
-After initialization a worker communicates over a simple \textbf{message
-based interface} with the main thread.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
-\CommentTok{// Send to worker}
-\VariableTok{worker}\NormalTok{.}\AttributeTok{postMessage}\NormalTok{(}\StringTok{'Hello World'}\NormalTok{)}\OperatorTok{;}
-\CommentTok{// Receive msg from worker}
-\VariableTok{worker}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'message'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>}
- \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\StringTok{'Worker said: '}\OperatorTok{,} \VariableTok{e}\NormalTok{.}\AttributeTok{data}\NormalTok{))}\OperatorTok{;}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\OperatorTok{>} \VariableTok{worker}\NormalTok{.}\AttributeTok{js}
-\CommentTok{// Receive msg and echo back}
-\KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'message'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>}
- \KeywordTok{this}\NormalTok{.}\AttributeTok{postMessage}\NormalTok{(}\StringTok{"Echo "} \OperatorTok{+} \VariableTok{e}\NormalTok{.}\AttributeTok{data}\NormalTok{))}\OperatorTok{;}
-\end{Highlighting}
-\end{Shaded}
 
@@
 background the paper should briefly join them to form a
-\emph{browsernative microservices}. Needless to say the following
+\emph{browsernative microservice}. Needless to say that the following
 example is overall simplified towards illustrating the connection
 between browsernative technologies and microservice patterns.
-Furthermore it is highly opinionated in development and shouldn't
-perceived as ``single source of truth''.
-
-Googles library Polymer is a good place for learning about web
-components in depth and make use of their toolbox. One of their most
-famous proof of concept is the so-called
-\href{https://shop.polymer-project.org/}{Polymer Shop} which is a
-fully-fledged online shop nested within a single root element
-\texttt{\textless{}shop-app\textgreater{}}. This app made of several
-main views and many more invisible wrapper elements for routing, service
-worker caching, theming, etc. The whole shop runs as a single
-application fetching and updating remote resources and switching views.
-Let's assume we work in a sales engineering team of the Polymer Shop and
-need to rebuild the checkout microservice.
+Furthermore it is an opinionated approach and should not be perceived as
+a `single source of truth'.
+
+Google's Polymer project is a good place for learning about web
+components in depth and make use of their toolbox. One of their proof of
+concept is the so-called \textbf{Polymer Shop}\footnote{\href{https://shop.polymer-project.org/}{Polymer
+  Shop}} which is a fully-fledged online shop nested within a single
+root element \texttt{\textless{}shop-app\textgreater{}}. This app is
+made of several main views and many more invisible wrapper elements for
+routing, service worker caching, theming, etc. The whole shop runs as a
+single application fetching and updating remote resources and switching
+views. Let us assume we work in a sales engineering team of the Polymer
+Shop and we need to rebuild the checkout microservice.
 
 The current checkout can be found at
-https://shop.polymer-project.org/checkout. At the time of writing the
+https://shop.polymer-project.org/checkout. At the time of writing, the
 checkout is a single, 671 lines of code long Polymer component including
-all required fields for sign in, shipping, billing and summarizing the
-order. In the spirit of microservices we will split up the microservice
-into fine grained components. The shopping cart data is pulled out of a
-local storage JSON entity set up previously by another custom-element.
+all required fields for signing in, shipping, billing and summarizing
+the order. In the spirit of microservices we will split up the
+microservice into independent components. The shopping cart data is
+pulled out of a local storage JSON entity, set up previously by another
+custom element.
 
 By breaking down the service the \textbf{team defined the business
-boundaries} within the checkout process and ended up with following
+boundaries} within the checkout process and came up with following
 granular service blocks:
@@
 \item
-  Sign in or Sign up
+  Sign in
 \item
@@
 Yet already we see the simplicity arouse from web components as they
-persue a clear markup. Each of the child components may act
+persue a clean markup. Each of the child components may act
 independently over other child nodes utilizing the \textbf{loose
@@
 
-Before diving deeper into implementation, its worth to clarify an
-\textbf{architectural pattern} behind components. Any reader of the
-paper came across ReactJS / Redux, the concept of components may look
+Before diving deeper into implementation, it is worth to clarify an
+\textbf{architectural pattern} behind components. To any reader of the
+paper who already came across React, the concept of components may look
 familiar. Dan Abramov, the creator of Redux, once defined a simple
@@
 components} only related with the concern about \emph{how things look}.
-This component literally doesn't know anything about the service in
-question which makes the component highly flexible and reusable. They
-are controlled solely from the outside, receiving data and dispatching
-unbiased events on user interaction.{[}17{]} Most probably every
+This component literally does not know anything about the service in
+question which makes the component highly flexible and reusable. It is
+controlled solely from the outside, receiving data and dispatching
+unbiased events on user interaction.{[}20{]} Most probably every
 presentational component embodies more HTML/CSS markup and less JS code.
 It should encapsulate its styles from bleeding out and protect its
-styles being overwritten. Furthermore, it may contain several templates
-to change it's look on different demands.
-
-Secondly, Abramov described components he refers as \textbf{containers}.
-A container component is concerned with \emph{how things work}.{[}17{]}
-Containers acts as invisible wrappers around presentational components
-acting in the sense of UNIX filters. Their job is to fetch data from
-child nodes, aggregating events, interacting with the model and push
-state back to the presentational components. Consequently they might
-contain more JS and less if any HTML markup. We probably don't need to
-utilize Shadow DOM as no styles are involved.
-
-Last but not least, the pattern can expanded for illustrational purposes
-to \textbf{native components} which is every build-in HTMLElement like
-the HTMLButtonElement. Native components are mostly deep nested elements
-providing the actual functionality in the browser UI. They are solely
-controllable and styleable from the outside and are therefore wrapped in
-presentational components and/or containers.
+styles from being overwritten. Furthermore, it may contain several
+templates to change its look on different demands.
+
+Secondly, Abramov described components he refers to as
+\textbf{containers}. A container component is concerned with \emph{how
+things work}.{[}20{]} Containers act as invisible wrappers around
+presentational components acting in the sense of UNIX filters. Their job
+is to fetch data from child nodes, aggregating events, interacting with
+the model and push state back to the presentational components.
+Consequently they might contain more JS and less if any HTML markup. We
+probably do not need to utilize shadow DOM as no styles are involved.
+
+Last but not least, the pattern can be expanded for illustrational
+purposes to \textbf{native components} such as every build-in
+HTMLElement like an ordinary HTMLButtonElement. Native components are
+mostly deeply nested elements providing the actual functionality in the
+browser UI. They are solely controllable and styleable from the outside
+and are therefore wrapped in presentational components and/or
+containers.
 
@@
 basically just an encapsulation layer in terms of service
-functionalities. Encapsulation of CSS won't be necessary at this point
-as no styling is involved. A simplified root container for the checkout
-might look like the following code snippet.
+functionalities. Encapsulation of CSS will not be an issue at this point
+as no styling is involved. A simplified \emph{root container} for the
+checkout might look like the following code snippet:
 
@@
 \OperatorTok{\}}    
-\VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'shop-checkout'}\OperatorTok{,} \NormalTok{shopCheckout)}\OperatorTok{;}
+\VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'shop-checkout'}\OperatorTok{,} \NormalTok{ShopCheckout)}\OperatorTok{;}
 \OperatorTok{<}\SpecialStringTok{/script>}
@@
 The purpose of this simplified code snippet is to outline the
-\textbf{MVC threefold} in the service root. The MODEL is pushed into the
-worker thread, the CONTROLLER in the custom element and the VIEW in the
-HTML structuring. On runtime the root container acts like a
-\textbf{message dispatcher} implementing the microservice principle of
-smart endpoints and dumb pipes.
+\textbf{MVC threefold} in the \emph{service root}. The MODEL is pushed
+into the worker thread, the CONTROLLER in the custom element and the
+VIEW in the HTML structuring. On runtime the \emph{root container} acts
+like a \textbf{message dispatcher} implementing the microservice
+principle of \textbf{smart endpoints and dumb pipes}.
 
 In order to interact with the MODEL every child node must implement the
-\textbf{dedicated \texttt{checkout}} custom event (which will be
-explained further on in the next section). Summarizing the
-unidirectional communication from the service root perspective looks
-like following plot:
+dedicated \texttt{checkout} CustomEvent (which will be explained further
+on in the next section). Illustrating the \textbf{unidirectional
+communication} from the \emph{service root} perspective may look like
+the following plot:
 
@@
 Effects are yielded by the asynchronous operation of messages and create
-actions returned to sender. Effects can be created by external messages,
-like subscription to an WebSocket, too. Effects may be created with
-additional resources from the server or syncing with local storage like
-in the polymer-shop.
-
-While the msg handlers are basic ``dumb'' switch statements the
-\textbf{smartness} solely arouse from intelligent controllers processing
+actions returned to sender. Effects can be created by external
+operations, like subscription to an WebSocket, too. Effects may be
+created with additional resources from the server or syncing with local
+storage like those in the polymer-shop.
+
+While the msg handlers are basic `dumb' switch statements the
+\textbf{smartness} solely arises from intelligent controllers processing
 the message. This communication model offers lots of possibilities for
@@
 Every child node can be expanded into another full microservice or split
-up into separate nodes without notice of the service root. Communication
-might \textbf{fail graceful} providing a default console log in case
-switches aren't defined yet.
+up into separate nodes without notice of the \emph{service root}.
+Communication may fail without harming effects, providing a default
+console log.
 
@@
 \texttt{\textless{}sign-in\textgreater{}} or
-\texttt{\textless{}shipping-details\textgreater{}} still contain mostly
-logic and no or less styling. Their job is to control their underlying
-presentational leaf components, aggregating events and to define a
-\textbf{set of actions towards the model} . Every container mounted
-directly under the service root may have a dedicated area inside the
-worker thread where it's fulfil his duties. For example, a typical
-\texttt{\textless{}sign-in\textgreater{}} contains merely two fields
-username and password and a submit button. The component listens for the
-submit action, aggregating the credentials and might add some semantics
-to it like \emph{action: `SIGNIN SUBMITTED'}. Fields and action message
-will be dispatched towards the model for further processing like
-initiating a authorization process.
-
-Every container component is eligible to aggregate subordinate events
-from their children, buffer them and interact with the model trough via
-a \textbf{unified message system}. A base class, from which
+\texttt{\textless{}shipping-details\textgreater{}} still contains mostly
+logic and little or less styling. The job of a \emph{container
+component} is to control its underlying presentational components and to
+define a \textbf{set of ACTIONS towards the model}. Every
+\emph{container} mounted directly under the service root may have a
+dedicated area inside the worker thread reserved for its duties. For
+example, a typical \texttt{\textless{}sign-in\textgreater{}} contains
+merely two fields, username and password and a submit button. The
+\emph{container component} waits for a submit action, aggregating the
+credentials, and might add some semantics to them. Field values and the
+action message will be dispatched towards the model for further
+processing like initiating an authorization process.
+
+Every \emph{container component} is eligible to aggregate subordinate
+events from its children, buffer them and interact with the model via a
+\textbf{unified message system}. A base class, from which
 \texttt{\textless{}sign-in\textgreater{}} or
-\texttt{\textless{}shipping-details\textgreater{}} can be extended might
-look like this:
+\texttt{\textless{}shipping-details\textgreater{}} can be extended,
+might look like this:
 
@@
 
-Extending the \texttt{SimpleContainer} will equip every container node
-with the unified message interface. Calling
-\texttt{this.\_dispatch(msg)} within the container will trigger an event
-bubbling upstream. The service root will implement a simple event
-listener for \texttt{checkout} events. After receiving an answer from
-the MODEL a service root can query the container child node in question
-and push forward the answer over the \texttt{\_receive} property.
-
-There might be even more middleware containers pulled in between
-presentational leafs and the service root to fulfil some extra work
-either like filters on the event or without caring about checkout events
+Extending the \texttt{SimpleContainer} will equip every \emph{container
+node} with the unified message interface. Calling
+\texttt{this.\_dispatch(msg)} within the \emph{container} will trigger
+an event. The service root will implement a simple event listener for
+\texttt{checkout} events. After receiving an answer from the MODEL a
+service root can query the \emph{container child node} in question and
+push forward the answer over the \texttt{\_receive} property.
+
+There might be even \textbf{more middleware containers} pulled in
+between presentational leafs and the service root to fulfil some extra
+work like filters on the event or without caring about checkout events
 altogether. Changing or enhancing any functionality requires only the
-controllers in the container in question and the endpoint section at the
-model. \textbf{Infrastructure automation} might be achieved by
-dynamically evaluating the mounted containers beneath the service root
-and modeling the ``backend'' model accordingly. Due to it's standardized
-message system the containers are testable within standardized tests.
+controller in the \emph{container} in question and the endpoint section
+at the model. \textbf{Infrastructure automation} might be achieved by
+dynamically evaluating the mounted \emph{containers} beneath the service
+root and modeling the `backend' model accordingly. Due to its
+standardized message system the \emph{containers} are testable within
+standardized tests exposing them to different synthetic events.
 
@@
 
-In comparison to the former container the presentational component is
-build mostly around views and styling and shipping less or no logic
-towards handling the component. To extend the last
-\texttt{\textless{}sign-in\textgreater{}} container example a typical
-presentational component structure could look like the following
-top-level structure
+In comparison to the former container the \emph{presentational
+component} is build mostly around views and styling with just a little
+logic to switch templates around or to alter styling accordingly. As
+much as the container can be perceived as a wrapper for logic, the
+\emph{presentational component} can be perceived as \textbf{wrapper for
+styling} needs. To extend the last
+\texttt{\textless{}sign-in\textgreater{}} example a \emph{presentational
+component} structure could look like the following top-level structure:
 
@@
 \begin{Highlighting}[]
-\KeywordTok{<sign-in-container>}
-  \KeywordTok{<sign-in-presentational>}
-  \KeywordTok{</sign-in-presentational>}
-\KeywordTok{</sign-in-container>}
+\KeywordTok{<sign-in>}
+  \KeywordTok{<sign-in-styling}\OtherTok{ theme=}\StringTok{"dark"} \ErrorTok{...}\KeywordTok{>}
+    \CommentTok{<!-- Native components // form button ... -->}
+  \KeywordTok{</sign-in-styling>}
+\KeywordTok{</sign-in>}
 \end{Highlighting}
@@
 
-The \texttt{\textless{}sign-in-presentational\textgreater{}} contains
-all required input fields and oauth connectors to Google or Facebook but
-\textbf{will not care about events they create}. A presentation
-component is usually steered via HTML attributes. Contrary to the former
-containers the presentational component highly utilizes \textbf{shadow
-DOM and HTML templates}.
+The \texttt{\textless{}sign-in-styling\textgreater{}} contains all
+required input fields and Oauth connectors to Google or Facebook but
+\textbf{will not care about events they create}. A \emph{presentation
+component} is usually steered via HTML attributes. Contrary to the
+former containers, the \emph{presentational component} highly utilizes
+\textbf{shadow DOM and HTML templates}.
 
@@
 \begin{Highlighting}[]
-\NormalTok{> sign-in-presentational.html}
+\NormalTok{> sign-in-styling.html}
 \CommentTok{<!-- IMPORTS -->}
 \KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"facebook-oauth.html"}\OtherTok{ async}\KeywordTok{>}
-\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"high-res-template.html"}\OtherTok{ async}\KeywordTok{>}
+\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"mobile-template.html"}\OtherTok{ async}\KeywordTok{>}
 \NormalTok{...}
@@
 \CommentTok{<!-- VIEW -->}
-\KeywordTok{<sign-in-presentational>}
-  \KeywordTok{<high-res-template>}
-    \KeywordTok{<facebook-oauth></facebook-oauth>}
+\KeywordTok{<sign-in-styling>}
+  \KeywordTok{<template}\OtherTok{ id=}\StringTok{"highres"}\KeywordTok{>}
+    \KeywordTok{<style>}\FloatTok{...}\KeywordTok{</style>}
+    \KeywordTok{<div}\OtherTok{ id=}\StringTok{"content"}\KeywordTok{>}
+      \KeywordTok{<facebook-oauth>}\NormalTok{Yet another element}\KeywordTok{</facebook-oauth>}
       \KeywordTok{<slot}\OtherTok{ name=}\StringTok{"form"}\KeywordTok{></slot>}
       \KeywordTok{<slot}\OtherTok{ name=}\StringTok{"button"}\KeywordTok{></slot>}
-  \KeywordTok{</high-res-template>}
+    \KeywordTok{</div>}
+  \KeywordTok{</template>}
 
   \KeywordTok{<template}\OtherTok{ id=}\StringTok{"lowres"}\KeywordTok{>}\NormalTok{...}\KeywordTok{</template>}
-  \KeywordTok{<template}\OtherTok{ id=}\StringTok{"mobile"}\KeywordTok{>}\NormalTok{...}\KeywordTok{</template>}
-\KeywordTok{</sign-in-presentational>}
+  \KeywordTok{<mobile-template>}\NormalTok{Template custom element}\KeywordTok{</template>}
+\KeywordTok{</sign-in-styling>}
 
 \KeywordTok{<script>}
-  \KeywordTok{class} \NormalTok{SignIn }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
+  \KeywordTok{class} \NormalTok{SignInStyling }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
     \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
@@
       \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
-      \CommentTok{// append templates according to window.innerWidth}
-      \CommentTok{// run this.getAttribute('htmlprop') to control element}
+      \KeywordTok{this}\NormalTok{.}\VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{temp}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
+      \CommentTok{// f.e. append templates according to window.innerWidth}
+      \CommentTok{// style related JS altering}
+      \CommentTok{// this.getAttribute('theme') for configuration}
     \OperatorTok{\}}
@@
   \OperatorTok{\}}
-  \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'sign-in-presentational'}\OperatorTok{,} \NormalTok{SignIn)}\OperatorTok{;}
+  \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'sign-in-styling'}\OperatorTok{,} \NormalTok{SignInStyling)}\OperatorTok{;}
 \OperatorTok{<}\SpecialStringTok{/script>}
@@
 
-Presentational components require configuration and usually ship a long
-list of CSS selectors. Whereas it's possible to expand templates in
-place it could be also possible to define \textbf{template components}
-to atomize the component further and increase code readability.
+As mentioned before, HTML templates provide an opportunity to define
+\textbf{CSS modules}. Tags like \texttt{\textless{}style\textgreater{}}
+and \texttt{\textless{}slot\textgreater{}} create a kind of stencil
+filled by content from the \texttt{index.html}. As it is possible to
+expand templates in place it could also be possible to define
+\textbf{template components} like the \texttt{mobile-template} to
+atomize the component further and to increase code brevity.
+
+Overall, the \texttt{\textless{}shop-checkout\textgreater{}} should
+provide a taste of web components and their ingredients in practical
+usage. On a bigger scale things would certainly look different and less
+static than in the given example. Still, the microservice approach is
+intended to be somewhat visible in this example.
 
@@
 
-Following a \textbf{browsernative first} principle this paper didn't
-made use of third-party libraries so far. Contrary to the example in the
-last section there are a few reasons why web components shouldn't be
-used without any proprietary technology. At first, from a technical
-perspective web components are relatively young and at the time of
-writing only the newer Chrome and Opera browsers can handle them
-correctly. Most probably it will take a long time until native web
-components can be safely used without additional legacy support.The
-second inevitable question is the notion whether proprietary technology
-promotes \textbf{simplicity} or complects the implementation.
-
-Talking about web components it is impossible to avoid technologies like
-\textbf{React} altogether. In the last couple years React introduced
-some good concepts to web development like the idea of state, passing
-data as properties and declarative event management just to name a few.
-Many of those design decisions were made to overall \textbf{simplify
-browser development through standardized methodologies}. The principles
-of microservices could be applied to React development maybe even
-further as for example the communication model involves a lot of
-delegation and decomposition.
-
-Many of those features and design decisions in React created much of a
-hype around the library and technologies like the virtual DOM while web
-components never caught up in momentum at a comparable rate. According
-to Github Googles Polymer project is even six months older than React
-but has less than one third of its stars. Other notable web components
-projects like Bosonic\footnote{https://github.com/bosonic/bosonic/} or
-Mozilla's X-tag\footnote{https://github.com/x-tag/core} are more dead
-than alive, too.
-
-It seems like the web developing community in a broader sense demands
-for features and standardization in frameworks like React (and many
-others to be fair). Web components and related libraries clearly missed
-a selling point in providing standardized ways for developing scalable
-components and applications or being to cumbersome to use at all. For
-the future it remains unclear if web components will eventually gain
-wider adoption within the UI developing community or will be adopted as
-low-level technology for framework development emphasized by Sebastian
-Markbage, one of the React creators.{[}18{]}
-
-One interesting approach for building web components is the ``reactish''
-Skate\footnote{https://github.com/skatejs/skatejs/}. Following the
-functional rendering model from React it combines native methods with
-additional functionalities known from React like event management. Given
-the focus on native technologies it only weights around 4kb (minified
-and gzipped) which is ten times less than React. \textbf{Small
-encapsulated micro UI frameworks} like Skate could be certainly the near
-future of web development and will hopefully gain more attention. The
-more mature UI framework Riot already expressed their intention to
-gradually develop towards web components.\footnote{http://riotjs.com/compare/\#web-components}
+In the search for an independent \textbf{browsernative} approach this
+paper left out third-party libraries so far. Nevertheless, it should be
+clear to this point that the example in the last section is far away
+from being usable in production. At first, from a technical perspective
+web components are relatively young and at the time of writing only
+usable in \textbf{newer Chrome and Opera browsers}. Most probably it
+will take a long time until native web components can be safely used
+without additional legacy support. The troublesome \emph{HTML imports}
+spec will certainly need even longer which inevitably opens up the
+question if web components are useful from today onwards. And second, do
+web components align with the notion of simplicity if too many downsides
+and possible pitfalls require attention?!
+
+Any profound discussion about web components would be incomplete without
+touching component alternatives like \textbf{React}. In the last couple
+years React introduced new concepts to web development like the idea of
+state, passing data as properties and declarative event management just
+to name a few. Those design decisions were made to simplify browser
+development with full browser support right from the start. Overall,
+React's perception about UI development created much of a hype around
+the library and technologies like the virtual DOM while web components
+never caught up in momentum at a comparable rate. According to Github,
+Google's Polymer project is even six months older than React but has
+less than one third of its stars. Other notable web component projects
+like Bosonic\footnote{https://github.com/bosonic/bosonic/} or Mozilla's
+X-tag\footnote{https://github.com/x-tag/core} are more dead than alive,
+too.
+
+It seems like the web developing community in a broader sense values
+\textbf{standardized procedures, tooling and browser support} in
+frameworks like React (and many others to be fair). Previously mentioned
+web component libraries clearly missed a real selling point compared to
+React or Angular. Providing guidelines, simpler syntax and features are
+missed out or are too cumbersome to use at all. For the future it
+remains unclear if web components eventually gain wider adoption within
+the UI developing community or will be adopted as low-level technology
+for framework development emphasized by Sebastian Markbage, one of the
+React creators.{[}21{]}
+
+Even though the drawbacks seems heavy weight, there are notable attempts
+to bring web components into production. A very good example is the
+``reactish'' library Skate\footnote{https://github.com/skatejs/skatejs/}.
+Following the functional rendering model from React, it combines native
+methods with additional functionalities known from React like
+declarative event management. Given the focus on native technologies it
+only weights around 4kb (minified and gzipped) which is ten times less
+than React. \textbf{Micro UI frameworks} like Skate could certainly be
+the near future of web development, offering a real selling point.
+Previously mentioned specs could lead to a modular, pluggable building
+pipelines offering the same features and tooling as React already does.
+After all, web components fit better with the microservice
+\textbf{decentralization aspect} than `walled garden' frameworks.
+
+The role of web components may not result in an isolated implementation
+but move towards a common denominator for future Reacts, Angulars, Vues
+etc. This makes \textbf{web components worth considering them as
+recyclable technology}. The more mature UI framework Riot already
+expressed its intention to gradually develop towards web
+components.\footnote{http://riotjs.com/compare/\#web-components}
 
 \hypertarget{refs}{}
+\hypertarget{ref-Hickey2012}{}
+{[}1{]} R. Hickey, ``Rails conf 2012 keynote: Simplicity matters by rich
+hickey.'' 2012 {[}Online{]}. Available:
+\url{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}
+
 \hypertarget{ref-Filloux2016}{}
-{[}1{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016
+{[}2{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016
 {[}Online{]}. Available:
@@
 \hypertarget{ref-Baldwin2006}{}
-{[}2{]} C. Y. Baldwin and K. B. Clark, ``Modularity in the Design of
+{[}3{]} C. Y. Baldwin and K. B. Clark, ``Modularity in the Design of
 Complex Engineering Systems,'' in \emph{Complex engineered systems:
@@
 \hypertarget{ref-Newman2015}{}
-{[}3{]} S. Newman, \emph{Building microservices}. O'Reilly Media, Inc.,
+{[}4{]} S. Newman, \emph{Building microservices}. O'Reilly Media, Inc.,
 2016.
@@
 \hypertarget{ref-Fowler2014}{}
-{[}4{]} M. Fowler and J. Lewis, ``Microservices: A definition of this
+{[}5{]} M. Fowler and J. Lewis, ``Microservices: A definition of this
 new architectural term,'' Jan. 2014 {[}Online{]}. Available:
@@
 \hypertarget{ref-Martin}{}
-{[}5{]} R. C. Martin, ``The single responsibility principle.''
+{[}6{]} R. C. Martin, ``The single responsibility principle.''
 {[}Online{]}. Available:
@@
 \hypertarget{ref-Kearney2016}{}
-{[}6{]} M. Kearney, ``Measure performance with the rail model.'' 2016
+{[}7{]} M. Kearney, ``Measure performance with the rail model.'' 2016
 {[}Online{]}. Available:
@@
 \hypertarget{ref-Conway1968}{}
-{[}7{]} M. E. Conway, ``How do committees invent?'' 1968 {[}Online{]}.
+{[}8{]} M. E. Conway, ``How do committees invent?'' 1968 {[}Online{]}.
 Available: \url{http://www.melconway.com/Home/Committees_Paper.html}
 
+\hypertarget{ref-Boner2012}{}
+{[}9{]} J. Bonér, ``Latency numbers every programmer should know.'' 2012
+{[}Online{]}. Available: \url{https://gist.github.com/jboner/2841832}
+
 \hypertarget{ref-Issa2016}{}
-{[}8{]} B. Issa, ``The way of the web.'' Polymer Summit 2016, Oct-2016
+{[}10{]} B. Issa, ``The way of the web.'' Polymer Summit 2016, Oct-2016
 {[}Online{]}. Available:
@@
 
+\hypertarget{ref-Victor2012}{}
+{[}11{]} B. Victor, ``Inventing on principle.'' 2012 {[}Online{]}.
+Available: \url{https://vimeo.com/36579366}
+
 \hypertarget{ref-Lawson2016}{}
-{[}9{]} N. Lawson, ``Progressive enhancement isn't dead, but it smells
+{[}12{]} N. Lawson, ``Progressive enhancement isn't dead, but it smells
 funny.'' 2016 {[}Online{]}. Available:
@@
 \hypertarget{ref-Rota2015}{}
-{[}10{]} A. Rota, ``React.js conf 2015 - the complementarity of react
+{[}13{]} A. Rota, ``React.js conf 2015 - the complementarity of react
 and web components.'' 2015 {[}Online{]}. Available:
@@
 \hypertarget{ref-Buchner2016}{}
-{[}11{]} D. Buchner, ``Demythstifying web components,'' 2016
+{[}14{]} D. Buchner, ``Demythstifying web components,'' 2016
 {[}Online{]}. Available:
@@
 \hypertarget{ref-vanKesteren2014}{}
-{[}12{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
+{[}15{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
 {[}Online{]}. Available:
@@
 \hypertarget{ref-Bidelman2016}{}
-{[}13{]} E. Bidelman, ``Custom elements v1: reusable web components.''
+{[}16{]} E. Bidelman, ``Custom elements v1: reusable web components.''
 2016 {[}Online{]}. Available:
@@
 \hypertarget{ref-HTML}{}
-{[}14{]} \emph{HTML living standard --- last updated 11 january 2017}.
+{[}17{]} \emph{HTML living standard --- last updated 11 january 2017}.
 {[}Online{]}. Available: \url{https://html.spec.whatwg.org/multipage/}
@@
 \hypertarget{ref-Bidelman2016shadow}{}
-{[}15{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
+{[}18{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
 2016 {[}Online{]}. Available:
@@
 \hypertarget{ref-Bidelman2010}{}
-{[}16{]} E. Bidelman, ``The basics of web workers.'' 2010 {[}Online{]}.
+{[}19{]} E. Bidelman, ``The basics of web workers.'' 2010 {[}Online{]}.
 Available: \url{https://www.html5rocks.com/en/tutorials/workers/basics/}
@@
 \hypertarget{ref-Abramov2015}{}
-{[}17{]} D. Abramov, ``Presentational and Container Components --
+{[}20{]} D. Abramov, ``Presentational and Container Components --
 Medium.'' 2015 {[}Online{]}. Available:
@@
 \hypertarget{ref-Rauschmayer2015}{}
-{[}18{]} D. A. Rauschmayer, ``Tree-shaking with webpack 2 and babel 6.''
+{[}21{]} D. A. Rauschmayer, ``Tree-shaking with webpack 2 and babel 6.''
 2015 {[}Online{]}. Available:

### 2017-02-28 paper.tex
pdf/tex build

@@
 the work of Sam Newman{[}4{]} and the work of Fowler and Lewis{[}5{]}.
-The purpose of this section is to gain confidence about the architecture
-of microservices in the context of the browser platform.
+The purpose of this section is to develop confidence about the
+architecture of microservices in the context of the browser platform.
 
@@
 global nature of the DOM creates a daunting situation for code
-organization and leads over times to highly fragmented bits of CSS and
+organization and leads over times to highly fragmented bits of CSS or
 obscured CSS selectors. Of course this situation dramatically lowers
@@
 the application layer is abstracted from the main rendering thread into
-worker threads.\footnote{\href{https://docs.google.com/document/d/1M9FmT05Q6qpsjgvH1XvCm840yn2eWEg0PMskSQz7k4E}{Angular
-  2 Rendering Architecture}}
+worker threads.{[}\^{}Jbanov{]}
 
@@
       \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
+      \ControlFlowTok{if} \NormalTok{(}\VariableTok{window}\NormalTok{.}\AttributeTok{innerWidth} \OperatorTok{>} \DecValTok{1280}\NormalTok{) }\OperatorTok{\{}
+        \KeywordTok{const} \NormalTok{temp }\OperatorTok{=} \KeywordTok{this}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'#highres'}\NormalTok{)}\OperatorTok{;}
         \KeywordTok{this}\NormalTok{.}\VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{temp}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
-      \CommentTok{// f.e. append templates according to window.innerWidth}
-      \CommentTok{// style related JS altering}
+      \OperatorTok{\}}  
       \CommentTok{// this.getAttribute('theme') for configuration}
+      \CommentTok{// selecting nodes for altering}
     \OperatorTok{\}}
@@
 expand templates in place it could also be possible to define
-\textbf{template components} like the \texttt{mobile-template} to
+\textbf{template components} like the \texttt{mobile-template.html} to
 atomize the component further and to increase code brevity.

### 2017-03-01 build/paper.tex
pdf/tex build

@@
 \author{Jan Peteler, FH Würzburg-Schweinfurt, jan.peteler@student.fhws.de}
-\date{Januar 2017}
+\date{March 2017}
 
@@
 abstraction and often heavily depend on indispensable frameworks. While
-they perfectly circumstance the global paradigm of the DOM they remain
-highly proprietary attached to the framework. New w3c specifications
-build right into the browser provide a native service API in the need to
-create standardized web components for the browser. As of 2017 all major
+they elegantly circuit the global paradigm of the DOM they remain highly
+proprietary pieces of technology. New w3c specifications build right
+into the browser provide a native service API in the need to create
+standardized web components for the browser. As of 2017 all major
 browsers will ship those technologies in their browsers. The paper not
-only provides an introduction to the specification. As an assessment
-criteria of it will match those specifications against the concept of
-microservices which is an modular system architecture to build scalable
-applications.
+only provides an brief and practical introduction to the specification.
+As an assessment criteria this paper will match those specifications
+against the concept of microservices, which is an modular system
+architecture to build scalable applications.
 \end{abstract}
@@
 
-Computers can scale, humans can't. Ever since a complex system made by
-humans has been constrained by humans mental capabilities. Like in the
-analogy of juggling balls our brain can just ``juggle'' a few things at
-a time. Rich Hickey, the inventor of the programming language Clojure
-gave an inspirational keynote on the topic of
-\textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
-  Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} In every sphere
-of a humans life simplicity aligns perception with our mental
+Computers can scale, humans can't. Ever since complex systems made by
+humans have been constrained by humans mental capabilities. Like in the
+analogy of juggling balls our brain can just ``juggle'' a few items at a
+time. Rich Hickey, the inventor of the programming language Clojure gave
+an inspirational keynote on the topic of \textbf{simplicity}.{[}1{]} In
+every sphere of a human's life simplicity aligns perception with mental
 capacities.
 
-Derived from the Latin word \textbf{simplex} simple can be understood as
-``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
+Derived from the Latin word \textbf{simplex}, \emph{simple} can be
+understood as ``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
   Dictionary}} which points towards an unidimensional state. While
@@
 simplicity empowers the human brain to reason about issues
-straightforward. It certainly has some overlapping's with easy, but
-while easy is more of a relative spirit, simple can be laid out as a
-objective manner and therefore universally applicable.
+straightforward. It certainly has some overlappings with \emph{easy},
+but while \emph{easy} is more of a relative spirit, \emph{simple} can be
+laid out as an objective manner and therefore universally applicable.
 
@@
 currently around 2.5 MB.\footnote{\href{http://httparchive.org/trends.php}{HTTPArchive
-  Trends}} Leaving images, fonts or other content aside the size of
-HTML, CSS and JS sums up to an total average of 550 KB. One character
+  Trends}} Leaving images, fonts or other content aside, the size of
+HTML, CSS and JS sums up to a total average of 550 KB. One character
 weights around 1 byte which means an average webpage is delivering
-550.000 character or around 125 pages of single-spaced text. Frederic
+550.000 characters or around 125 pages of single-spaced text. Frederic
 Filloux analyzed the payload on different newspaper websites and came to
-the conclusion, that only round about 5-6 \% of the transferred
-characters made for human consumption.{[}1{]}
-
-Having an 95 \% overhead is rather undesirable for both the consumers
-and creators of the website. Since it's a widespread problem without a
-single point of failure one can argue the platform itself is the
+the conclusion, that only roundabout 5-6 \% of the transferred
+characters are made for human consumption.{[}2{]}
+
+Having a 95 \% overhead is rather undesirable for both the consumers and
+the creators of the website. Since it is a widespread problem without a
+definable point of failure one can argue `the platform itself' is the
 failure. By design, every pageload results in a monolithic DOM tree
 managed by the browser engine. Whether rendering just a bunch of static
-text nodes or an ever changing webapp the underlying global nature of
+text nodes or an ever changing webapp, the underlying global nature of
 the DOM tree remains the same. Every additional piece of code added to
@@
 methodologies approached the global nature of the DOM by scoping assets
-and design rules into maintainable components. While the DOM can't be
-scoped, JS can. Many frameworks, like React, Angular or Vue just to name
+and design rules into maintainable components. While the DOM cannot be
+scoped, JS can. Many frameworks like React, Angular or Vue, just to name
 a few, ditched the old rule of separated HTML, CSS and JS in favor of an
-additional layer of abstracted JS components (containing structure,
-behavior and sometimes styling). Quiet often those frameworks mimic a
-MVC pattern on top of the browser engine which is a reasonable simple
-design pattern to build UIs. While frameworks are a valid approach for
-building scalable web applications they remain highly opinionated embody
-inherent complexity themselves and can change and break over time.
-Another downside is code inflation which is a crucial point for
-performance. All of those bottlenecks in the web demand for a new
-standardized way for creating and evolving complex web services.
+additional layer of abstracted JS components. Quite often those
+frameworks mimic a MVC pattern on top of the browser engine which is a
+reasonably simple pattern to build UIs. While frameworks are a valid
+approach for building scalable web applications they remain highly
+opinionated, embody inherent complexity themselves and can change and
+break over time. Another drawback is code inflation which is a crucial
+point for performance. All of those bottlenecks in the web demand for a
+new standardized ways for creating and evolving complex web services.
 
@@
 services, each running in its own process and communicating with
-lightweight mechanisms, often an HTTP resource API.''{[}4{]} Yet at this
+lightweight mechanisms, often an HTTP resource API.''{[}5{]} Yet at this
 point a reader might spot some similarities with microservices and the
-browser-based development: Both wrestling the problem of monolithic
-architecture and both using lightweight communication mechanisms. In
-fact, many big companies of ``the web'' like Amazon or Netflix
-successfully transformed their monolithic system into a service based
-system which gives a taste of the power behind microservices.{[}4{]}
+browser-based development: Both wrestle with the problem of monolithic
+architecture and both use lightweight communication mechanisms. In fact,
+many big companies of `the web' like Amazon or Netflix successfully
+transformed their monolithic system into a service based system which
+gives a taste of the power behind microservices.{[}5{]}
 
 Microservices incorporate a wide array of ideas from developing scalable
-software, like \textbf{domain-driven design} where it pursues the
-incorporation of real world structure in the code.{[}3, p. 2{]} Or
-making use of \textbf{continuous delivery} for pushing software rapidly
-through \textbf{automated deployment} mechanisms into production.{[}4{]}
-Furthermore, microservices transcendent the technical perspective and
-reaches into the team organization. As a primary source of truth this
-paper relies on the work of Sam Newman, who has written a comprehensive
-guide towards microservices{[}3{]} and the work of Fowler and
-Lewis{[}4{]}. The purpose of this section is to gain confidence about
-the microservices architecture in the context of the browser platform.
+software like \textbf{domain-driven design} to pursuing the real world
+structure in the code.{[}4, p. 2{]} Another concept focuses on
+\textbf{continuous delivery} for pushing software rapidly through
+\textbf{automated deployment} mechanisms into production.{[}5{]}
+Furthermore, microservices transcend the technical perspective and reach
+into team organization.
+
+As a primary source of comprehensive information this paper relies on
+the work of Sam Newman{[}4{]} and the work of Fowler and Lewis{[}5{]}.
+The purpose of this section is to develop confidence about the
+architecture of microservices in the context of the browser platform.
 
 \subsection{2.1. Componentization via
-Services}\label{componentization-via-services}
+services}\label{componentization-via-services}
 
 ``A~\textbf{component}~is a unit of software that is independently
-replaceable and upgradeable.'' {[}4{]} Components are the building
+replaceable and upgradeable.'' {[}5{]} Components are the building
 blocks of microservices. And microservices are the building blocks of
@@
 The first principle of services is the \textbf{loose coupling
-principle}: changing and deploying one service shouldn't result in
-changing other parts of the system.{[}3, p. 30{]} Picking CSS for
+principle}: changing and deploying one service should not result in
+changing other parts of the system.{[}4, p. 30{]} Picking CSS for
 example, where variables tend to be shadowed by higher specific values
-making it increasingly difficult to keep changes ought to only affect
-one place in the application. A \emph{browsernative microservice}
-therefore pushing encapsulation and avoiding variable mutations outside
-its scope as much as possible. One solution would be scoped CSS / JS to
-avoid variables leaking into the global namespace as well as hidden
-implementation details to avoid mutating values.
+making it increasingly difficult to keep changes, ought to merely affect
+one place in the application. Therefore a \emph{browsernative
+microservice} is expected to push encapsulation and to avoid variable
+mutations outside its scope as much as possible. One solution would be
+scoped CSS / JS to avoid variables leaking into the global namespace, as
+well as to hide implementation details in order to avoid mutating
+values.
 
 The second principle of services is the \textbf{high cohesion
-principle}: Whether designing a microservice or it's components we want
+principle}: Whether designing a microservice or its components, we want
 related behavior sit together and unrelated behavior to sit
-elsewhere.{[}3, p. 30{]} High cohesion can be enhanced towards the more
+elsewhere.{[}4, p. 30{]} High cohesion can be enhanced towards the more
 dynamic \textbf{Single Responsibility Principle}: ``Gather together
 those things that change for the same reason and separate those things
-that change for different reasons.''{[}5{]} In a very quick and dirty
+that change for different reasons.''{[}6{]} In a very quick and dirty
 code quality analysis, the quality can be measured just by counting the
-places changes in the code occur in order to implement a functionality.
-An arbitrary threefold MVC system should require a maximum of three
-changes to implement or change functionalities. The problem in browser
-based development is not only the global paradigm which makes changes
+code changes which occur in order to implement a new functionality. An
+arbitrary threefold MVC system should require a maximum of three changes
+to implement or change functionalities. The problem in browser based
+development is not only the global paradigm which makes changes
 deliberately unpredictable. The high cohesion principle is violated by
-the traditional separation along the siloed entities \textbf{HTML, JS
-and CSS}. Understanding the relation of HTML markup and another CSS file
-adds incidental complexity. Different approaches emerged over the recent
-years to join the forces. Most notable the dedicated frontend language
-\emph{elm}\footnote{\href{http://elm-lang.org/}{elm lang}} where MODEL,
-VIEW and CONTROLLER sit in one place. A \emph{browsernative service}
-sought for an combination of the web native trinity HTML, JS and CSS.
-
-Traditional web development relies on \textbf{libraries} to enhance the
-service capabilities of the web platform. Compared to libraries a
-component service offers multiple advantages for building, deployment
-and shipping. As expressed earlier a \emph{components for the web}, or
-web component, is \textbf{self-contained} which means it embodies all
-needed functionality to get it's job done. Therefore it has a much
-better evolution mechanism in the service contracts. Changing
-functionality won't break other services. A component can progressively
-enhanced which guarantees functionality throughout different versions. A
+the traditional separation along the siloed entities HTML, JS and CSS.
+Understanding the relation of HTML markup towards another CSS file
+generates incidental complexity. Different approaches emerged over the
+recent years to join these entities. A \emph{browsernative service}
+sought for a combination of the web native trinity HTML, JS and CSS.
+
+Traditional web development relies on libraries to enhance the service
+capabilities of the web platform. Compared to libraries a component
+service offers multiple advantages for building, deployment and
+shipping. As expressed earlier \emph{components for the web}, or web
+components, are self-contained which means they embody all needed
+functionality to get their job done. Therefore they have a much better
+evolution mechanism in the service contracts. Changing functionality
+will not break other services. A component can be progressively enhanced
+which guarantees functionality throughout different versions whereas a
 library is only loosely coupled to the implementation and therefore hard
@@
 and especially the critical first page rendering. Libraries for the
-browsers are traditionally ``shipped'' as non static immediately-invoked
-functions run immediately on page load. Following the \emph{Google RAIL}
-model, a user-centric performance measurement, a page load ought be less
-than 1 second to catch up user's attention.{[}6{]} There are many ways
-for optimizing the critical first render but as a rule of thumb a
-build-in web component might be always superior to libraries in terms of
-first rendering. The fine-grained lifecycle methods which will be
-described in the technical section of this paper give the developer far
-reaching optimization opportunities.
+browsers are traditionally `shipped' as non static immediately-invoked
+functions. Following the Google RAIL model, a user-centric performance
+measurement, a page load ought to take less than 1 second to catch up
+users' attention.{[}7{]} There are many ways to optimize the critical
+first render but as a rule of thumb a build-in web component might be
+always superior to libraries in terms of first rendering. The
+fine-grained lifecycle methods which will be described in the technical
+section of this paper give the developer far reaching optimization
+opportunities.
 
 The last argument in favor of components over libraries is the more
-explicit interface.{[}4{]} While the functionality of a library needs
-documentation to be accessible a component functionality is exposed via
-the components' signature. The markup language HTML is by design
-equipped with an expressive syntax and steering a web component using
-attributes and values should be most straightforward even to
-unexperienced web developers.
-
-\subsection{2.2. Organized around Business
-Capabilities}\label{organized-around-business-capabilities}
+explicit interface.{[}5{]} While the functionality of a library needs
+documentation to be accessible, a component functionality is exposed via
+the components' signature. HTML markup is an expressive syntax and
+therefore convenient for steering a web component using merely
+attributes and values.
+
+\subsection{2.2. Organized around business
+capabilities}\label{organized-around-business-capabilities}
 
@@
 \subsection{2.4. Decentralized
-Governance}\label{decentralized-governance}
+governance}\label{decentralized-governance}
 
@@
 achieve a similar look and feel like native apps. And last but not least
-services like NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}}
-effectively compiling ``the web'' to native machine code lowering the
-boundary between native and browser code even further.
+services like NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}
+  \#\# 2.5. Decentralized data management} effectively compiling `the
+web' to native machine code lower the boundary between native and
+browser code even further.
 
@@
 global nature of the DOM creates a daunting situation for code
-organization and leads over times to highly fragmented bits of CSS and
-obscured CSS selectors. Of course this situation lowers code clarity and
-reusability dramatically. The only solution which won't break with the
-existing global paradigm of the DOM effectively is to allow separate
-pieces of encapsulated code sit on top of the global DOM - introducing
-the shadowed DOM approach.
+organization and leads over times to highly fragmented bits of CSS or
+obscured CSS selectors. Of course this situation dramatically lowers
+code clarity and reusability. The only solution which will not break
+with the existing global paradigm of the DOM is to allow separate pieces
+of encapsulated code sit on top of the global DOM - introducing the
+shadowed DOM approach.
 
 Enhancing the previous example the new encapsulated \texttt{HelloWorld}
-would like this:
+would look like the following code snippet:
 
@@
 \textbf{background threads} which allows the UI to stay responsive even
-when computation-heavy tasks needs to perform. Furthermore, a worker
-thread adds a performance advantage embracing the multi core CPU
-architecture most devices running on today. To grasp the full potential
-of workers a reader might dive deeper into the Angular 2 architecture,
-where most of the application layer is abstracted from the main
-rendering thread into worker threads.\footnote{\href{https://docs.google.com/document/d/1M9FmT05Q6qpsjgvH1XvCm840yn2eWEg0PMskSQz7k4E}{Angular
-  2 Rendering Architecture}}
-
-A \emph{Web Worker} spawns a new background thread where scripts can run
-concurrent to the main thread. Usually a worker is loaded from dedicated
-file to embrace separation.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{const} \NormalTok{worker }\OperatorTok{=} \KeywordTok{new} \AttributeTok{Worker}\NormalTok{(}\StringTok{'worker.js'}\NormalTok{)}\OperatorTok{;}
-\end{Highlighting}
-\end{Shaded}
-
-After initialization a worker communicates over a simple \textbf{message
-based interface} with the main thread.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
-\CommentTok{// Send to worker}
-\VariableTok{worker}\NormalTok{.}\AttributeTok{postMessage}\NormalTok{(}\StringTok{'Hello World'}\NormalTok{)}\OperatorTok{;}
-\CommentTok{// Receive msg from worker}
-\VariableTok{worker}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'message'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>}
- \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\StringTok{'Worker said: '}\OperatorTok{,} \VariableTok{e}\NormalTok{.}\AttributeTok{data}\NormalTok{))}\OperatorTok{;}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\OperatorTok{>} \VariableTok{worker}\NormalTok{.}\AttributeTok{js}
-\CommentTok{// Receive msg and echo back}
-\KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'message'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>}
- \KeywordTok{this}\NormalTok{.}\AttributeTok{postMessage}\NormalTok{(}\StringTok{"Echo "} \OperatorTok{+} \VariableTok{e}\NormalTok{.}\AttributeTok{data}\NormalTok{))}\OperatorTok{;}
-\end{Highlighting}
-\end{Shaded}
+when computation-heavy tasks need to perform. Furthermore, a worker adds
+a performance advantage embracing the multi core CPU architecture most
+devices are running on today. To grasp the full potential of workers, a
+reader might dive deeper into the Angular 2 architecture, where most of
+the application layer is abstracted from the main rendering thread into
+worker threads.{[}\^{}Jbanov{]}
 
@@
 Firstly, Abramov defined a pattern around \textbf{presentational
-components} only related with the concern about \emph{how things look}.
-This component literally doesn't know anything about the service in
-question which makes the component highly flexible and reusable. They
-are controlled solely from the outside, receiving data and dispatching
-unbiased events on user interaction.{[}17{]} Most probably every
+components} only related with the concern about \textbf{how things
+look}. This component literally does not know anything about the service
+in question which makes the component highly flexible and reusable. It
+is controlled solely from the outside, receiving data and dispatching
+unbiased events on user interaction.{[}20{]} Most probably every
 presentational component embodies more HTML/CSS markup and less JS code.
 It should encapsulate its styles from bleeding out and protect its
-styles being overwritten. Furthermore, it may contain several templates
-to change it's look on different demands.
-
-Secondly, Abramov described components he refers as \textbf{containers}.
-A container component is concerned with \emph{how things work}.{[}17{]}
-Containers acts as invisible wrappers around presentational components
-acting in the sense of UNIX filters. Their job is to fetch data from
-child nodes, aggregating events, interacting with the model and push
-state back to the presentational components. Consequently they might
-contain more JS and less if any HTML markup. We probably don't need to
-utilize Shadow DOM as no styles are involved.
-
-Last but not least, the pattern can expanded for illustrational purposes
-to \textbf{native components} which is every build-in HTMLElement like
-the HTMLButtonElement. Native components are mostly deep nested elements
-providing the actual functionality in the browser UI. They are solely
-controllable and styleable from the outside and are therefore wrapped in
-presentational components and/or containers.
+styles from being overwritten. Furthermore, it may contain several
+templates to change its look on different demands.
+
+Secondly, Abramov described components he refers to as
+\textbf{containers}. A container component is concerned with \textbf{how
+things work}.{[}20{]} Containers act as invisible wrappers around
+presentational components acting in the sense of UNIX filters. Their job
+is to fetch data from child nodes, aggregating events, interacting with
+the model and push state back to the presentational components.
+Consequently they might contain more JS and less if any HTML markup. We
+probably do not need to utilize shadow DOM as no styles are involved.
+
+Last but not least, the pattern can be expanded for illustrational
+purposes to \textbf{native components} such as every build-in
+HTMLElement like an ordinary HTMLButtonElement. Native components are
+mostly deeply nested elements providing the actual functionality in the
+browser UI. They are solely controllable and styleable from the outside
+and are therefore wrapped in presentational components and/or
+containers.
 
@@
 
-The root container \texttt{\textless{}shop-checkout\textgreater{}} is
-basically just an encapsulation layer in terms of service
-functionalities. Encapsulation of CSS won't be necessary at this point
-as no styling is involved. A simplified root container for the checkout
-might look like the following code snippet.
+The \emph{root container}
+\texttt{\textless{}shop-checkout\textgreater{}} is basically just an
+encapsulation layer in terms of service functionalities. Encapsulation
+of CSS will not be an issue at this point as no styling is involved. A
+simplified \emph{root container} for the checkout might look like the
+following code snippet:
 
@@
 \texttt{\textless{}sign-in\textgreater{}} or
-\texttt{\textless{}shipping-details\textgreater{}} still contain mostly
-logic and no or less styling. Their job is to control their underlying
-presentational leaf components, aggregating events and to define a
-\textbf{set of actions towards the model} . Every container mounted
-directly under the service root may have a dedicated area inside the
-worker thread where it's fulfil his duties. For example, a typical
-\texttt{\textless{}sign-in\textgreater{}} contains merely two fields
-username and password and a submit button. The component listens for the
-submit action, aggregating the credentials and might add some semantics
-to it like \emph{action: `SIGNIN SUBMITTED'}. Fields and action message
-will be dispatched towards the model for further processing like
-initiating a authorization process.
-
-Every container component is eligible to aggregate subordinate events
-from their children, buffer them and interact with the model trough via
-a \textbf{unified message system}. A base class, from which
+\texttt{\textless{}shipping-details\textgreater{}} still contains mostly
+logic and little or less styling. The job of a \emph{container
+component} is to control its underlying presentational components and to
+define and spell a \textbf{set of ACTIONS towards the model}. Every
+\emph{container} mounted directly under the service root may have a
+dedicated area inside the worker thread reserved for its duties. For
+example, a typical \texttt{\textless{}sign-in\textgreater{}} contains
+merely two fields, username and password and a submit button. The
+\emph{container component} waits for a submit action, aggregating the
+credentials, and might add some semantics to them. Field values and the
+action message will be dispatched towards the model for further
+processing like initiating an authorization process.
+
+Every \emph{container component} is eligible to aggregate subordinate
+events from its children, buffer them and interact with the model via a
+\textbf{unified message system}. A base class, from which
 \texttt{\textless{}sign-in\textgreater{}} or
-\texttt{\textless{}shipping-details\textgreater{}} can be extended might
-look like this:
+\texttt{\textless{}shipping-details\textgreater{}} can be extended,
+might look like this:
 
@@
 
-In comparison to the former container the presentational component is
-build mostly around views and styling containing less or no logic
-towards handling the component. To extend the last
-\texttt{\textless{}sign-in\textgreater{}} container example a typical
-presentational component structure could look like the following
-top-level structure
+In comparison to the former container the \emph{presentational
+component} is build mostly around views and styling with just a little
+logic to switch templates around or to alter styling accordingly. As
+much as the container can be perceived as a wrapper for logic, the
+\emph{presentational component} can be perceived as \textbf{wrapper for
+styling} needs. To extend the last
+\texttt{\textless{}sign-in\textgreater{}} example a \emph{presentational
+component} structure could look like the following top-level structure:
 
@@
       \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
-      \CommentTok{// append templates according to window.innerWidth}
-      \CommentTok{// run this.getAttribute('htmlprop') to control element}
+      \ControlFlowTok{if} \NormalTok{(}\VariableTok{window}\NormalTok{.}\AttributeTok{innerWidth} \OperatorTok{>} \DecValTok{1280}\NormalTok{) }\OperatorTok{\{}
+        \KeywordTok{const} \NormalTok{temp }\OperatorTok{=} \KeywordTok{this}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'#highres'}\NormalTok{)}\OperatorTok{;}
+        \KeywordTok{this}\NormalTok{.}\VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{temp}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
+      \OperatorTok{\}}  
+      \CommentTok{// this.getAttribute('theme') for configuration}
+      \CommentTok{// selecting nodes for altering}
     \OperatorTok{\}}
@@
 
-Presentational components require configuration and usually ship a long
-list of CSS selectors. Whereas it's possible to expand templates in
-place it could be also possible to define \textbf{template components}
-to atomize the component further and increase code readability.
-
-tests
+As mentioned before, HTML templates provide an opportunity to define
+\textbf{CSS modules}. Tags like \texttt{\textless{}style\textgreater{}}
+and \texttt{\textless{}slot\textgreater{}} create a kind of stencil
+filled by content from the \texttt{index.html}. As it is possible to
+expand templates in place it could also be possible to define
+\textbf{template components} like the \texttt{mobile-template.html} to
+atomize the component further and to increase code brevity.
+
+Overall, the \texttt{\textless{}shop-checkout\textgreater{}} should
+provide a taste of web components and their ingredients in practical
+usage. On a bigger scale things would certainly look different and less
+static than in the given example. Still, the microservice approach is
+intended to be somewhat visible in this example.
 
@@
 
-Following a \textbf{browsernative first} dogma this paper didn't put
-third-party libraries into view so far in the search of an independent
-approach. Nevertheless, it should be clear to this point that the
-example in the last section is far away from being usable in production.
-At first, from a technical perspective web components are relatively
-young and at the time of writing only usable in \textbf{newer Chrome and
-Opera browsers}. Most probably it will take a long time until native web
-components can be safely used without additional legacy support. The
-troublesome \emph{HTML imports} spec will certainly need even longer
-which inevitable opens up the question if web components are useful from
-today on. And second, do web components align with the notion of
-simplicity if too many downsides and possible pitfalls require
-attention?!
+In the search for an independent \textbf{browsernative} approach this
+paper left out third-party libraries so far. Nevertheless, it should be
+clear to this point that the example in the last section is far away
+from being usable in production. At first, from a technical perspective
+web components are relatively young and at the time of writing only
+usable in \textbf{newer Chrome and Opera browsers}. Most probably it
+will take a long time until native web components can be safely used
+without additional legacy support. The troublesome \emph{HTML imports}
+spec will certainly need even longer which inevitably opens up the
+question if web components are useful from today onwards. And second, do
+web components align with the notion of simplicity if too many downsides
+and possible pitfalls require attention?!
 
@@
 state, passing data as properties and declarative event management just
-to name a few. Those design decisions were made to overall simplify
-browser development with full browser support right from the start.
-Overall, Reacts perception about UI development created much of a hype
-around the library and technologies like the virtual DOM while web
-components never caught up in momentum at a comparable rate. According
-to Github Googles Polymer project is even six months older than React
-but has less than one third of its stars. Other notable web components
-projects like Bosonic\footnote{https://github.com/bosonic/bosonic/} or
-Mozilla's X-tag\footnote{https://github.com/x-tag/core} are more dead
-than alive, too.
+to name a few. Those design decisions were made to simplify browser
+development with full browser support right from the start. Overall,
+React's perception about UI development created much of a hype around
+the library and technologies like the virtual DOM while web components
+never caught up in momentum at a comparable rate. According to Github,
+Google's Polymer project is even six months older than React but has
+less than one third of its stars. Other notable web component projects
+like Bosonic\footnote{https://github.com/bosonic/bosonic/} or Mozilla's
+X-tag\footnote{https://github.com/x-tag/core} are more dead than alive,
+too.
 
@@
 \textbf{standardized procedures, tooling and browser support} in
-frameworks like React (and many others to be fair). Web components and
-related libraries clearly missed a real selling point towards React and
-Angular. Providing guidelines, simpler syntax and features missed out or
-being to cumbersome to use at all. For the future it remains unclear if
-web components will eventually gain wider adoption within the UI
-developing community or will be adopted as low-level technology for
-framework development emphasized by Sebastian Markbage, one of the React
-creators.{[}18{]}
-
-Even though the downsides seems heavy weight there are notable attempts
+frameworks like React (and many others to be fair). Previously mentioned
+web component libraries clearly missed a real selling point compared to
+React or Angular. Providing guidelines, simpler syntax and features are
+missed out or are too cumbersome to use at all. For the future it
+remains unclear if web components eventually gain wider adoption within
+the UI developing community or will be adopted as low-level technology
+for framework development emphasized by Sebastian Markbage, one of the
+React creators.{[}21{]}
+
+Even though the drawbacks seems heavy weight, there are notable attempts
 to bring web components into production. A very good example is the
 ``reactish'' library Skate\footnote{https://github.com/skatejs/skatejs/}.
-Following the functional rendering model from React it combines native
+Following the functional rendering model from React, it combines native
 methods with additional functionalities known from React like
@@
 only weights around 4kb (minified and gzipped) which is ten times less
-than React. \textbf{Small encapsulated micro UI frameworks} like Skate
-could certainly be the near future of web development and will hopefully
-pave the way of web components.
-
-The role of web components may not result in an isolated framework but
-move towards a common denominator for future Reacts, Angulars,
-Vues\ldots{} which makes it worth considering it as recyclable
-technology foundation. The more mature UI framework Riot already
-expressed their intention to gradually develop towards web
+than React. \textbf{Micro UI frameworks} like Skate could certainly be
+the near future of web development, offering a real selling point.
+Previously mentioned specs could lead to modular, pluggable building
+pipelines offering the same features and tooling as React already does.
+After all, web components fit better with the microservice
+\textbf{decentralization aspect} than `walled garden' frameworks.
+
+The role of web components may not result in an isolated implementation
+but move towards a common denominator for future Reacts, Angulars, Vues
+etc. This makes \textbf{web components worth considering them as
+recyclable technology}. The more mature UI framework Riot already
+expressed its intention to gradually develop towards web
 components.\footnote{http://riotjs.com/compare/\#web-components}
@@
 \hypertarget{refs}{}
+\hypertarget{ref-Hickey2012}{}
+{[}1{]} R. Hickey, ``Rails conf 2012 keynote: Simplicity matters by rich
+hickey.'' 2012 {[}Online{]}. Available:
+\url{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}
+
 \hypertarget{ref-Filloux2016}{}
-{[}1{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016
+{[}2{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016
 {[}Online{]}. Available:

### 2017-03-01 paper.tex
pdf/tex build

@@
 \subsection{2.1. Componentization via
-Services}\label{componentization-via-services}
+services}\label{componentization-via-services}
 
@@
 
-\subsection{2.2. Organized around Business
-Capabilities}\label{organized-around-business-capabilities}
+\subsection{2.2. Organized around business
+capabilities}\label{organized-around-business-capabilities}
 
@@
 services like NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}
-  \#\# 2.5. Decentralized Data Management} effectively compiling `the
+  \#\# 2.5. Decentralized data management} effectively compiling `the
 web' to native machine code lower the boundary between native and
@@
 Firstly, Abramov defined a pattern around \textbf{presentational
-components} only related with the concern about \emph{how things look}.
-This component literally does not know anything about the service in
-question which makes the component highly flexible and reusable. It is
-controlled solely from the outside, receiving data and dispatching
+components} only related with the concern about \textbf{how things
+look}. This component literally does not know anything about the service
+in question which makes the component highly flexible and reusable. It
+is controlled solely from the outside, receiving data and dispatching
 unbiased events on user interaction.{[}20{]} Most probably every
@@
 Secondly, Abramov described components he refers to as
-\textbf{containers}. A container component is concerned with \emph{how
+\textbf{containers}. A container component is concerned with \textbf{how
 things work}.{[}20{]} Containers act as invisible wrappers around
@@
 
-The root container \texttt{\textless{}shop-checkout\textgreater{}} is
-basically just an encapsulation layer in terms of service
-functionalities. Encapsulation of CSS will not be an issue at this point
-as no styling is involved. A simplified \emph{root container} for the
-checkout might look like the following code snippet:
+The \emph{root container}
+\texttt{\textless{}shop-checkout\textgreater{}} is basically just an
+encapsulation layer in terms of service functionalities. Encapsulation
+of CSS will not be an issue at this point as no styling is involved. A
+simplified \emph{root container} for the checkout might look like the
+following code snippet:
 
@@
 component} is to control its underlying presentational components and to
-define a \textbf{set of ACTIONS towards the model}. Every
+define and spell a \textbf{set of ACTIONS towards the model}. Every
 \emph{container} mounted directly under the service root may have a
@@
 
-\section{5. Thinking further}\label{thinking-further}
-
-In the search for an independent \textbf{browsernative} approach this
-paper left out third-party libraries so far. Nevertheless, it should be
-clear to this point that the example in the last section is far away
-from being usable in production. At first, from a technical perspective
-web components are relatively young and at the time of writing only
-usable in \textbf{newer Chrome and Opera browsers}. Most probably it
-will take a long time until native web components can be safely used
-without additional legacy support. The troublesome \emph{HTML imports}
-spec will certainly need even longer which inevitably opens up the
-question if web components are useful from today onwards. And second, do
-web components align with the notion of simplicity if too many downsides
-and possible pitfalls require attention?!
+\section{5. Outlook}\label{outlook}
+
+In the search for an independent browsernative approach this paper left
+out \textbf{third-party libraries} purposely. It has to be pointed out
+that the example given in chapter 4 of this paper is far from being
+usable in production. At first, from a technical perspective web
+components are relatively young and at the time of writing only usable
+in \textbf{newer Chrome and Opera browsers}. Most probably it will take
+time until native web components can be safely used without additional
+legacy support. The troublesome \emph{HTML imports} spec will certainly
+need even longer which inevitably opens up the question if web
+components are useful from today onwards. And second, do web components
+align with the notion of simplicity if too many drawbacks and possible
+pitfalls require attention?!
 
 Any profound discussion about web components would be incomplete without
-touching component alternatives like \textbf{React}. In the last couple
-years React introduced new concepts to web development like the idea of
-state, passing data as properties and declarative event management just
-to name a few. Those design decisions were made to simplify browser
-development with full browser support right from the start. Overall,
-React's perception about UI development created much of a hype around
-the library and technologies like the virtual DOM while web components
-never caught up in momentum at a comparable rate. According to Github,
-Google's Polymer project is even six months older than React but has
-less than one third of its stars. Other notable web component projects
-like Bosonic\footnote{https://github.com/bosonic/bosonic/} or Mozilla's
-X-tag\footnote{https://github.com/x-tag/core} are more dead than alive,
-too.
-
-It seems like the web developing community in a broader sense values
-\textbf{standardized procedures, tooling and browser support} in
-frameworks like React (and many others to be fair). Previously mentioned
-web component libraries clearly missed a real selling point compared to
-React or Angular. Providing guidelines, simpler syntax and features are
-missed out or are too cumbersome to use at all. For the future it
-remains unclear if web components eventually gain wider adoption within
-the UI developing community or will be adopted as low-level technology
-for framework development emphasized by Sebastian Markbage, one of the
-React creators.{[}21{]}
-
-Even though the drawbacks seems heavy weight, there are notable attempts
-to bring web components into production. A very good example is the
-``reactish'' library Skate\footnote{https://github.com/skatejs/skatejs/}.
-Following the functional rendering model from React, it combines native
-methods with additional functionalities known from React like
+ever touching \textbf{component alternatives} like React or Angular. In
+the last couple of years both introduced new concepts to web development
+like the idea of state, passing data as properties and declarative event
+management just to name a few. Even though they use the notion of
+components too, they differentiate much from the philosophy of web
+components. Web components are more about the DOM whereas others focus
+on JS. Implicitly this approach opens up a case for \textbf{`standards
+vs.~opinions'}.
+
+Overall, React's perception about UI development created much of a hype
+around the library and technologies like the virtual DOM while web
+components have never caught up in momentum at a comparable rate.
+Certainly the biggest drawbacks of web components were frequently
+changing specs and the uneasiness of using an underperforming polyfilled
+version. In the highly competitive field of UI frameworks, \textbf{web
+components clearly missed a real selling point} compared to opinionated
+frameworks so far. Notable web component projects like Bosonic\footnote{https://github.com/bosonic/bosonic/}
+or Mozilla's X-tag\footnote{https://github.com/x-tag/core} are more dead
+than alive. For the future it remains unclear if web components
+eventually gain wider adoption within the UI developing community or
+will be adopted as low-level technology for framework development
+emphasized by Sebastian Markbage, one of the React creators.{[}21{]}
+
+Even though the drawbacks seems heavy weight for the moment things will
+certainly change for good with flourishing browser adoption. A good
+example of recent efforts is the ``reactish'' library Skate\footnote{https://github.com/skatejs/skatejs/}.
+Following the functional rendering model from React, Skate combines
+native methods with additional functionalities known from React like
 declarative event management. Given the focus on native technologies it
-only weights around 4kb (minified and gzipped) which is ten times less
+only weights around 4 KB (minified and gzipped) which is ten times less
 than React. \textbf{Micro UI frameworks} like Skate could certainly be
 the near future of web development, offering a real selling point.
-Previously mentioned specs could lead to a modular, pluggable building
+Previously mentioned specs could lead to modular, pluggable building
 pipelines offering the same features and tooling as React already does.
